### 2022-01-03 src/plot-tikz_maup/maup2.tex
Add code to programmatically produce TikZ plots demonstrating MAUP

\documentclass[tikz]{standalone}

\usetikzlibrary{shapes, decorations, arrows ,calc, arrows.meta,
                fit, positioning, automata, shapes.geometric, patterns}

\usepackage{tikz}

\begin{document}

\begin{tikzpicture}
\filldraw[pattern color=gray, pattern=north east lines, opacity=0.9] (0,-1) -- (0,1) -- (1,1) -- (1,-1) -- cycle;
\node[] at (-0.5, 0) {1};
\node[] at (0.5, 0) {2};
\draw (-1,1) -- (1,1);
\draw (-1,-1) -- (1,-1);
\draw (-1,-1) -- (-1,1);
\draw (1,-1) -- (1,1);
\draw (0,-1) -- (0,1);

\filldraw[pattern color=gray, pattern=north east lines, opacity=0.9] (3,-1) -- (3,1) -- (4,1) -- (4,-1) -- cycle;
\node[] at (2.5, 0) {1};
\node[] at (3.5, 0.5) {2};
\node[] at (3.5, -0.5) {3};
\draw (2,1) -- (4,1);
\draw (2,-1) -- (4,-1);
\draw (2,-1) -- (2,1);
\draw (4,-1) -- (4,1);
\draw (3,-1) -- (3,1);
\draw (3,0) -- (4,0);

\filldraw[pattern color=gray, pattern=north east lines, opacity=0.9] (6,-1) -- (6,1) -- (7,1) -- (7,-1) -- cycle;
\node[] at (5.5, 0) {1};
\node[] at (6.5, 0.666) {2};
\node[] at (6.5, 0) {3};
\node[] at (6.5, -0.666) {4};
\draw (5,1) -- (7,1);
\draw (5,-1) -- (7,-1);
\draw (5,-1) -- (5,1);
\draw (7,-1) -- (7,1);
\draw (6,-1) -- (6,1);
\draw (6,-0.333) -- (7,-0.333);
\draw (6,0.333) -- (7,0.333);

\draw [->,>=stealth] (1.25,0) -- (1.75,0);
\draw [->,>=stealth] (4.25,0) -- (4.75,0);
\draw [->,>=stealth] (7.25,0) -- (7.75,0);

\node[] at (8.25, 0) {$\cdots$};
\end{tikzpicture}

\end{document}

### 2022-03-28 src/0_demo_areal-kernels/areal-kernels.Rmd
Updating naming of folders

# Comparison

```{r}
fit_centroid <- readRDS("centroid.rds")
fit_integrated <- readRDS("integrated.rds")
```

## Sampling speed

```{r}
samples_per_second <- function(fit) {
  # Outside of warm-up phase
  times <- rstan::get_elapsed_time(fit)
  # Number of samples in first chain
  nsim_iter <- fit@stan_args[[1]]$iter
  nsim_iter / mean(times[, 2])
}
samples_per_second(fit_centroid)
samples_per_second(fit_integrated)
```

## Convergence analysis

Trace-plots:

```{r}
bayesplot::mcmc_trace(fit_centroid, pars = "l")
bayesplot::mcmc_trace(fit_centroid, pars = "beta_0")
bayesplot::mcmc_trace(fit_centroid, pars = "rho[28]") # For example
bayesplot::mcmc_trace(fit_integrated, pars = "l")
bayesplot::mcmc_trace(fit_integrated, pars = "beta_0")
bayesplot::mcmc_trace(fit_integrated, pars = "rho[28]")
```

Rank histograms:

```{r}
bayesplot::mcmc_rank_overlay(fit_centroid, pars = "l")
bayesplot::mcmc_rank_overlay(fit_centroid, pars = "beta_0")
bayesplot::mcmc_rank_overlay(fit_centroid, pars = "rho[28]")
bayesplot::mcmc_rank_overlay(fit_integrated, pars = "l")
bayesplot::mcmc_rank_overlay(fit_integrated, pars = "beta_0")
bayesplot::mcmc_rank_overlay(fit_integrated, pars = "rho[28]")
```

All Rhat < 1.1 necessary but not sufficient:

```{r}
which(rstan::summary(fit_centroid)[["summary"]][, "Rhat"] > 1.1)
which(rstan::summary(fit_integrated)[["summary"]][, "Rhat"] > 1.1)
```

Fitted values comparison:

```{r}
fitted_centroid <- summary(fit_centroid)$summary[paste0("rho[", 1:28, "]"), "mean"]
fitted_integrated <- summary(fit_integrated)$summary[paste0("rho[", 1:28, "]"), "mean"]
plot(fitted_centroid, fitted_integrated)
```

### 2022-03-28 src/0_plot-tikz_maup/maup3.tex
Updating naming of folders

--- 2022-01-03 src/plot-tikz_maup/maup2.tex
@@
 \begin{tikzpicture}
-\filldraw[pattern color=gray, pattern=north east lines, opacity=0.9] (0,-1) -- (0,1) -- (1,1) -- (1,-1) -- cycle;
+\node[] at (-1, 1.5) {(i)};nonumber
 \node[] at (-0.5, 0) {1};
 \node[] at (0.5, 0) {2};
-\draw (-1,1) -- (1,1);
-\draw (-1,-1) -- (1,-1);
+\node[] at (1.5, 0) {3};
+\node[] at (2.5, 0) {4};
+\draw (-1,1) -- (3,1);
+\draw (-1,-1) -- (3,-1);
 \draw (-1,-1) -- (-1,1);
+\draw (0,-1) -- (0,1);
 \draw (1,-1) -- (1,1);
-\draw (0,-1) -- (0,1);
+\draw (2,-1) -- (2,1);
+\draw (3,-1) -- (3,1);
+\filldraw[pattern color=gray, pattern=north east lines, opacity=0.9] (1,-1) -- (1,1) -- (3,1) -- (3,-1) -- cycle;
 
-\filldraw[pattern color=gray, pattern=north east lines, opacity=0.9] (3,-1) -- (3,1) -- (4,1) -- (4,-1) -- cycle;
-\node[] at (2.5, 0) {1};
-\node[] at (3.5, 0.5) {2};
-\node[] at (3.5, -0.5) {3};
-\draw (2,1) -- (4,1);
-\draw (2,-1) -- (4,-1);
-\draw (2,-1) -- (2,1);
+\node[] at (4, 1.5) {(ii)};nonumber
+\node[] at (4.5, 0) {1};
+\node[] at (5.5, 0) {2};
+\node[] at (6.25, 0) {3};
+\node[] at (6.75, 0) {4};
+\node[] at (7.25, 0) {5};
+\node[] at (7.75, 0) {6};
+\draw (4,1) -- (8,1);
+\draw (4,-1) -- (8,-1);
 \draw (4,-1) -- (4,1);
-\draw (3,-1) -- (3,1);
-\draw (3,0) -- (4,0);
+\draw (5,-1) -- (5,1);
+\draw (6,-1) -- (6,1);
+\draw (7,-1) -- (7,1);
+\draw (8,-1) -- (8,1);
 
-\filldraw[pattern color=gray, pattern=north east lines, opacity=0.9] (6,-1) -- (6,1) -- (7,1) -- (7,-1) -- cycle;
-\node[] at (5.5, 0) {1};
-\node[] at (6.5, 0.666) {2};
-\node[] at (6.5, 0) {3};
-\node[] at (6.5, -0.666) {4};
-\draw (5,1) -- (7,1);
-\draw (5,-1) -- (7,-1);
-\draw (5,-1) -- (5,1);
-\draw (7,-1) -- (7,1);
-\draw (6,-1) -- (6,1);
-\draw (6,-0.333) -- (7,-0.333);
-\draw (6,0.333) -- (7,0.333);
+% Additional lines down
+\draw (6.5,-1) -- (6.5,1);
+\draw (7.5,-1) -- (7.5,1);
 
-\draw [->,>=stealth] (1.25,0) -- (1.75,0);
-\draw [->,>=stealth] (4.25,0) -- (4.75,0);
-\draw [->,>=stealth] (7.25,0) -- (7.75,0);
+% Shading
+\filldraw[pattern color=gray, pattern=north east lines, opacity=0.9] (6,-1) -- (6,1) -- (8,1) -- (8,-1) -- cycle;
 
-\node[] at (8.25, 0) {$\cdots$};
 \end{tikzpicture}

### 2022-03-28 src/0_check_bin-gaussian/functions.R

@@ -1,0 +1,27 @@
+experiment <- function(nsim, n, phi_mean, phi_sd, y_sample_size) {
+  phi <- rnorm(n, phi_mean, phi_sd)
+  rho <- plogis(phi)
+  m <- rep(y_sample_size, n)
+
+  #' Without aggregation
+  y <- replicate(nsim, rbinom(n, m, rho), simplify = FALSE)
+  y <- lapply(y, sum)
+
+  #' With aggregation (at the level of the latent field)
+  y_agg <- replicate(nsim, rbinom(1, sum(m), weighted.mean(rho)), simplify = FALSE)
+
+  return(list(y = unlist(y), y_agg = unlist(y_agg), phi = phi, rho = rho))
+}
+
+plot_hist <- function(result) {
+  data.frame(
+    type = c(rep("No aggregation (truth)", nsim), rep("Aggregation (approximation)", nsim)),
+    y = c(result$y, result$y_agg)
+  ) %>%
+    ggplot() +
+    geom_histogram(aes(x = y, fill = type), alpha = 0.5, position = "identity") +
+    theme_minimal() +
+    theme(
+      legend.position = "bottom"
+    )
+}

### 2022-03-28 src/0_check_bin-gaussian/script.R

@@ -1,0 +1,47 @@
+# orderly::orderly_develop_start("check_bin-gaussian")
+# setwd("src/check_bin-gaussian")
+
+#' Binomial of sum of Gaussians versus sum of binomial of Gaussians
+#' IID latent field
+#' Without aggregation:
+#' `y = \sum_{i = 1}^n y_i$, $y_i \sim \text{Bin}(m, \rho_i)$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`
+#' With aggregation:
+#' `y \sim \text{Bin}(nm, \bar \rho)$, $\bar \rho = \frac{1}{n} \sum_{i = 1}^n \rho_i$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`
+
+nsim <- 500
+n <- 36
+
+#' Low prevalence setting (disease)
+result <- experiment(nsim, n, -2.5, 0.5, 30)
+plot(result$rho) #' This is what the underlying latent field looks like for these settings
+
+pdf("histogram-disease.pdf", h = 4, w = 6.25)
+
+plot_hist(result) #' And then the difference between the truth and the aggregated approximation
+
+dev.off()
+
+sd(result$y)
+sd(result$y_agg)
+sd(result$y_agg) / sd(result$y) #' Ratio
+
+#' Probability approximately 1/2 setting (elections)
+#' As before but with different `phi_mean` settings:
+result_half <- experiment(nsim, n, 0, 0.5, 30)
+plot_hist(result_half$rho)
+
+pdf("histogram-election.pdf", h = 4, w = 6.25)
+
+hist_fun(result_half)
+
+dev.off()
+
+sd(result_half$y)
+sd(result_half$y_agg)
+sd(result_half$y_agg) / sd(result_half$y) #' Ratio
+
+#' Conclusions, things to add:
+#' * Standard deviation of aggregation tends to overestimate true standard deviation,
+#' * Effect is larger in the small `\rho` setting than the large `\rho` setting
+#' * Should test latent field with spatial structure e.g. ICAR latent field and so on
+#' * What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?

### 2022-03-28 src/0_check_ik-compute/script.R

@@ -1,0 +1,72 @@
+# orderly::orderly_develop_start("check_ik-compute")
+# setwd("src/check_ik-compute")
+
+#' Checking that the Stan code to calculate the integrated kernel is correct
+
+sf <- mw
+L <- 10
+type = "random"
+
+n <- nrow(sf)
+samples <- sf::st_sample(sf, type = type, exact = TRUE, size = rep(L, n))
+S <- sf::st_distance(samples, samples)
+sample_index <- sf::st_intersects(sf, samples)
+sample_lengths <- lengths(sample_index)
+start_index <- sapply(sample_index, function(x) x[1])
+
+#' First calculate it directly using R only
+D <- sf::st_distance(samples, samples)
+kD <- matern(D, l = 1)
+K <- matrix(nrow = n, ncol = n)
+
+for(i in 1:n) {
+  K[i, i] <- mean(kD[sample_index[[i]], sample_index[[i]]])
+}
+
+for(i in 1:(n - 1)) {
+  for(j in (i + 1):n) {
+    K[i, j] <- mean(kD[sample_index[[i]], sample_index[[j]]]) #' Fill the upper triangle
+    K[j, i] <- K[i, j] #' Fill the lower triangle
+  }
+}
+
+cov_r <- K
+
+#' Now calculate it in Stan via two different methods
+#' rstan::expose_stan_functions puts the C++ functions into the R local workspace to use
+rstan::expose_stan_functions("cov_sample_average.stan")
+
+#' 1. Newer method
+cov_stan <- cov_sample_average(
+  S = S,
+  l = 1,
+  n = n,
+  start_index = start_index,
+  sample_lengths = sample_lengths,
+  total_samples = sum(sample_lengths)
+)
+
+#' 2. Old method
+cov_stan_old <- cov_sample_average_old(
+  n = n,
+  L = 10,
+  l = 1,
+  S = S
+)
+
+#' Comparison (these statements should all be TRUE)
+all.equal(cov_r, cov_stan)
+all.equal(cov_r, cov_stan_old)
+all.equal(cov_stan, cov_stan_old)
+
+#' Hard to tell from an image if something is identical but anyway
+pdf("cov-comparison.pdf", h = 5, w = 6.25)
+
+cowplot::plot_grid(
+  plot_matrix(cov_r) + labs(title = "R"),
+  plot_matrix(cov_stan_old) + labs(title = "Stan old"),
[…]

### 2022-03-28 src/0_check_ik-converge/functions.R

@@ -1,0 +1,42 @@
+adapted_integrated_covariance <- function(samples) {
+  sample_index <- sf::st_intersects(grid, samples)
+  D <- sf::st_distance(samples, samples)
+  kD <- matern(D, l = 1)
+  K <- matrix(nrow = n, ncol = n)
+  #' Diagonal entries
+  for(i in 1:(n - 1)) {
+    K[i, i] <- mean(kD[sample_index[[i]], sample_index[[i]]])
+    for(j in (i + 1):n) {
+      #' Off-diagonal entries
+      K[i, j] <- mean(kD[sample_index[[i]], sample_index[[j]]]) #' Fill the upper triangle
+      K[j, i] <- K[i, j] #' Fill the lower triangle
+    }
+  }
+  K[n, n] <- mean(kD[sample_index[[n]], sample_index[[n]]])
+  return(K)
+}
+
+plot_samples <- function(samples){
+  ggplot(grid) +
+    geom_sf(fill = "lightgrey") +
+    geom_sf(data = samples, alpha = 0.5, shape = 4) +
+    labs(x = "Longitude", y = "Latitude") +
+    theme_minimal() +
+    labs(subtitle = paste0(length(samples$geom), " samples"), fill = "") +
+    theme(axis.title = element_blank(),
+          axis.text = element_blank(),
+          axis.ticks = element_blank())
+}
+
+#' Simple metric between matrices of identical dimension
+frobenius <- function(M1, M2) {
+  diff <- M1 - M2
+  sqrt(sum(diff^2))
+}
+
+int_convergence_experiment <- function(L, n_sim = 100) {
+  sample_sets <- lapply(1:n_sim, function(i) sf::st_sample(grid, type = type, exact = TRUE, size = rep(L, n)))
+  covs <- lapply(sample_sets, adapted_integrated_covariance)
+  frobs <- sapply(covs, function(cov) frobenius(gt_cov, cov))
+  return(frobs)
+}

### 2022-03-28 src/0_check_xbinomial-sampling/script.R

@@ -1,0 +1,29 @@
+# orderly::orderly_develop_start("check_xbinomial-sampling")
+# setwd("src/check_xbinomial-sampling")
+
+#' Test of sampling from the distribution produced by custom likelihood: xbinomial_lpdf
+temp1 <- rstan::stan(file = "xbinom.stan", data = list(m = 10, rho = 0.5), warmup = 400, iter = 600, verbose = TRUE)
+rstan::summary(temp1)$summary
+out1 <- rstan::extract(temp1)
+plot(out1$y)
+hist(out1$y)
+
+#' Test of sampling from the distribution produced by custom likelihood: xbinomial_logit_lpdf
+temp2 <- rstan::stan(file = "xbinom_logit.stan", data = list(m = 10, eta = 0), warmup = 100, iter = 900, verbose = TRUE)
+rstan::summary(temp2)$summary
+out2 <- rstan::extract(temp2)
+plot(out2$y)
+hist(out2$y)
+
+df1 <- data.frame(x = out1$y)
+df2 <- data.frame(x = out2$y)
+
+ks.test(out1$y, out2$y)
+
+pdf("xbinom-ecdf-comparison.pdf", h = 4, w = 6.25)
+
+ggplot() +
+  stat_ecdf(data = df1, aes(x = x), geom = "step") +
+  stat_ecdf(data = df2, aes(x = x), geom = "step", col = "red")
+
+dev.off()

### 2022-03-28 src/0_demo_areal-kernels/centroid.R

@@ -1,0 +1,28 @@
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+sf <- mw
+
+centroid_distance <- function(sf) {
+  cent <- sf::st_centroid(sf)
+  D <- sf::st_distance(cent, cent)
+  return(D)
+}
+
+D <- centroid_distance(sf)
+
+dat <- list(n = nrow(sf),
+            y = sf$y,
+            m = sf$n_obs,
+            mu = rep(0, nrow(sf)),
+            D = D)
+
+nsim_warm <- 100
+nsim_iter <- 1000
+
+fit <- rstan::stan("centroid.stan",
+                   data = dat,
+                   warmup = nsim_warm,
+                   iter = nsim_iter)
+
+saveRDS(fit, file = "centroid.rds")

### 2022-03-28 src/0_explore_inla-spde/script.R

@@ -1,0 +1,4 @@
+# orderly::orderly_develop_start("explore_inla-spde")
+# setwd("src/explore_inla-spde")
+
+rmarkdown::render("inla-spde.Rmd")

### 2022-03-28 src/0_explore_variogram-method/script.R

@@ -1,0 +1,48 @@
+# orderly::orderly_develop_start("explore_variogram-method")
+# setwd("src/explore_variogram-method")
+
+#' Development of variogram methods
+#' Aim to automatically determine length-scale
+
+cent_mw <- sf::st_centroid(mw)
+
+#' Size of point corresponds to estimate
+
+pdf("bubble-plot.pdf", h = 7, w = 6.25)
+
+ggplot(mw) +
+  geom_sf(fill = "lightgrey") +
+  geom_sf(data = cent_mw, aes(size = est)) +
+  labs(x = "Longitude", y = "Latitude", fill = "Estimate") +
+  theme_minimal() +
+  labs(fill = "") +
+  theme_minimal()
+
+dev.off()
+
+vario <- variogram(est ~ 1, cent_mw)
+
+pdf("variogram.pdf", h = 4, w = 6.25)
+
+plot(vario)
+
+dev.off()
+
+#' * np: the number of point pairs for this estimate
+#' * dist: the average distance of all point pairs considered for this estimate
+#' * gamma: the actual sample variogram estimate
+
+fit <- fit.variogram(
+  vario,
+  model = vgm(
+    psill = 1,
+    model = "Mat",
+    range = 300,
+    kappa = 1.5
+  )
+)
+
+#' * psill: (partial) sill of the variogram model component
+#' * model: model type, e.g. "Exp", "Sph", "Gau", "Mat"
+#' * range: range parameter of the variogram model component; in case of anisotropy: major range
+#' * kappa: smoothness parameter for the Matern class of variogram models

### 2022-03-28 src/0_explore_wilson-pointless/functions.R

@@ -1,0 +1,22 @@
+# Make a smaller version of the A matrix so that predicting the risk surface
+#   is less computationally involved
+#
+# Args:
+#   mesh.true: The mesh
+#   adm0: Spatial polygons dataframe that is for the boundary of Kenya
+#
+# Return: list with prediction matrix
+#   and points where predictions will be made
+MakePredictionAMatrix <- function(mesh.true, adm0, dims = c(120, 120)) {
+  smallproj <- inla.mesh.projector(mesh.true, dims=dims)
+  x <- smallproj$x
+  y <- smallproj$y
+  all.temp <- (expand.grid(x, y , KEEP.OUT.ATTRS=F))
+  names(all.temp) <- c("long", "lat")
+  coordinates(all.temp) <- ~ long + lat
+  all.temp@proj4string <- adm0@proj4string
+  rel.points <- all.temp[adm0, ]
+  rel.points.clip <- as.data.frame(rel.points)
+  A.pred <- inla.spde.make.A(mesh.true, loc=as.matrix(rel.points.clip))
+  return(list(A.pred=A.pred, rel.points.clip=rel.points.clip))
+}

### 2022-03-28 src/0_plot-tikz_besag/script.R

@@ -1,0 +1,5 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("plot-tikz_besag")
+# setwd("src/plot-tikz_besag")
+
+tools::texi2dvi("besag.tex", pdf = TRUE, clean = TRUE)

### 2022-03-28 src/0_plot_areal-data/script.R

@@ -1,0 +1,36 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("plot_areal-data")
+# setwd("src/plot_areal-data")
+
+sf_lightgrey <- "#E6E6E6"
+lightgrey <- "#D3D3D3"
+
+indicator <- c(0, 0, 1, rep(0, 25))
+
+a <- ggplot() +
+  geom_sf(data = mw, aes(fill = as.factor(indicator))) +
+  labs(x = "", y = "", title = "Country: Malawi", fill = "") +
+  scale_fill_manual(values = c(sf_lightgrey, lightgrey)) +
+  theme_void() +
+  theme(legend.position = "none")
+
+lilongwe <- filter(mw, name_1 == "Lilongwe")
+points <- st_sample(lilongwe, size = 5)
+
+v <- points %>%
+  st_union() %>%
+  st_voronoi()
+
+dat <- st_intersection(st_cast(v), st_union(lilongwe))
+
+b <- ggplot(dat) +
+  geom_sf(fill = lightgrey) +
+  stat_sf_coordinates(size = 0.5) +
+  labs(x = "", y = "", title = "District: Lilongwe") +
+  theme_void()
+
+pdf("areal-data.pdf", h = 2.5, w = 6.25)
+
+cowplot::plot_grid(b, a, rel_widths = c(1.5, 1), align = "v", ncol = 2)
+
+dev.off()

### 2022-03-28 src/0_plot_geometry-graph/script.R

@@ -1,0 +1,15 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("plot_geometry-graph")
+# setwd("src/plot_geometry-graph")
+
+pdf("geometry-graph-mwi.pdf", h = 2.5, w = 6.25)
+
+bsae::plot_graph(mw, add_geography = TRUE)
+
+dev.off()
+
+pdf("geometry-graph-zwe.pdf", h = 2.5, w = 6.25)
+
+bsae::plot_graph(zw, add_geography = TRUE)
+
+dev.off()

### 2022-03-28 src/0_plot_integration-method/script.R

@@ -1,0 +1,28 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("plot_integration-method")
+# setwd("src/plot_integration-method")
+
+#' The number of points in each area
+L <- 10
+
+#' The number of areas
+n <- nrow(mw)
+
+#' The different integration strategies
+#' Note: sampling methods from package spatstat are interfaced!
+#' https://cran.r-project.org/web/packages/spatstat/spatstat.pdf
+
+random <- sf::st_sample(mw, size = rep(L, n))
+hexagonal <- sf::st_sample(mw, size = rep(L, n), type = "hexagonal", exact = TRUE)
+regular <- sf::st_sample(mw, size = rep(L, n), type = "regular", exact = TRUE)
+
+pdf("integration-method.pdf", h = 2.5, w = 6.25)
+
+cowplot::plot_grid(
+  plot_samples(random, title = "Random"),
+  plot_samples(hexagonal, title = "Hexagonal"),
+  plot_samples(regular, title = "Regular"),
+  ncol = 3
+)
+
+dev.off()

### 2022-03-28 src/1_analyze_assessment-results/script.R

@@ -1,0 +1,3 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("analyze_assessment-results")
+# setwd("src/analyze_assessment-results")

### 2022-03-28 src/1_assess_intercept-marginal/functions.R

@@ -1,0 +1,55 @@
+marginal_intercept <- function(fit, ...) {
+  UseMethod("marginal_intercept")
+}
+
+marginal_intercept.inla <- function(fit, S = 1000) {
+  df <- data.frame(fit$marginals.fixed$"(Intercept)")
+  samples_list <- INLA::inla.posterior.sample(n = S, fit)
+  samples <- sapply(samples_list, function(x) x$latent["(Intercept):1", ])
+  summary <- fit$summary.fixed["(Intercept)", ]
+  return(list(
+    df = df, samples = samples, mean = summary$mean, mode = summary$mode,
+    lower = summary$"0.025quant", upper = summary$"0.975quant"
+  ))
+}
+
+marginal_intercept.inlax <- function(fit, S = 1000) {
+  df <- data.frame(fit$marginals.fixed$"(Intercept)")
+  samples <- sapply(fit[["samples"]], function(x) x$latent["(Intercept):1", ])
+  summary <- fit$summary.fixed["(Intercept)", ]
+  return(list(
+    df = df, samples = samples, mean = summary$mean, mode = summary$mode,
+    lower = summary$"0.025quant", upper = summary$"0.975quant"
+  ))
+}
+
+marginal_intercept.stanfit <- function(fit, S = 1000) {
+  samples <- rstan::extract(fit, pars = "beta_0")$beta_0
+  kde <- density(samples)
+  df <- data.frame(x = kde$x, y = kde$y)
+  summary <- rstan::summary(fit)$summary["beta_0", ]
+  return(list(
+    df = df, samples = samples, mean = summary[["mean"]], mode = summary[["50%"]],
+    lower = summary[["2.5%"]], upper = summary[["97.5%"]]
+  ))
+}
+
+assess_marginal_intercept <- function(intercept = -2, fit) {
+  marginal <- marginal_intercept(fit)
+  obs <- intercept
+  error_samples <- (marginal$samples - obs)
+  f <- approxfun(marginal$df$x, marginal$df$y, yleft = 0, yright = 0)
+
+  return(list(
+    obs = obs,
+    mean = marginal$mean,
+    mode = marginal$mode,
+    lower = marginal$lower,
+    upper = marginal$upper,
+    mse = mean(error_samples^2),
+    mae = mean(abs(error_samples)),
+    crps = bsae::crps(marginal$samples, obs),
+    lds = log(f(obs)),
+    quantile = cubature::cubintegrate(f, lower = -Inf, upper = obs, method = "pcubature")$integral
+  ))
+}

### 2022-03-28 src/1_assess_lengthscale-marginal/benchmark.R

@@ -1,0 +1,11 @@
+#' Testing
+fits <- readRDS("depends/fits_icar_1_ck_stan.rds")
+
+#' Just the first one
+fit <- fits[[1]]
+
+#' The lengthscale marginal
+marginal_lengthscale(fit)
+
+#' And its assessment
+assess_marginal_lengthscale(lengthscale = 5/2, fit)

### 2022-03-28 src/1_assess_lengthscale-marginal/functions.R

@@ -1,0 +1,58 @@
+#' Most models don't have lengthscales
+#' The behaviour I'm aiming for is that lengthscale_df eventually just contains rows for the ones that do
+#' Perhaps returning NULL is a good way to do this, likely there are better solutions
+
+marginal_lengthscale <- function(fit, ...) {
+  UseMethod("marginal_lengthscale")
+}
+
+marginal_lengthscale.inla <- function(fit, S = 1000) {
+  warning("R-INLA models don't have a length-scale!")
+  return(NULL)
+}
+
+marginal_lengthscale.inlax <- function(fit, S = 1000) {
+  warning("R-INLA models don't have a length-scale!")
+  return(NULL)
+}
+
+marginal_lengthscale.stanfit <- function(fit, S = 1000) {
+  if(!("l" %in% rownames(rstan::summary(fit)$summary))){
+    # For other use-cases need more general approach to deal with e.g. rho[1]
+    warning("This Stan model doesn't have a length-scale!")
+    return(NULL)
+  }
+  samples <- rstan::extract(fit, pars = "l")$l
+  kde <- density(samples)
+  df <- data.frame(x = kde$x, y = kde$y)
+  summary <- rstan::summary(fit)$summary["l", ]
+  return(list(
+    df = df, samples = samples, mean = summary[["mean"]], mode = summary[["50%"]],
+    lower = summary[["2.5%"]], upper = summary[["97.5%"]]
+  ))
+}
+
+assess_marginal_lengthscale <- function(lengthscale = 5/2, fit) {
+  marginal <- marginal_lengthscale(fit)
+
+  if(is.null(marginal)){
+    return(NULL)
+  }
+
+  obs <- lengthscale
+  error_samples <- (marginal$samples - obs)
+  f <- approxfun(marginal$df$x, marginal$df$y, yleft = 0, yright = 0)
+
+  return(list(
+    obs = obs,
+    mean = marginal$mean,
+    mode = marginal$mode,
+    lower = marginal$lower,
+    upper = marginal$upper,
+    mse = mean(error_samples^2),
+    mae = mean(abs(error_samples)),
+    crps = bsae::crps(marginal$samples, obs),
+    lds = log(f(obs)),
+    quantile = cubature::cubintegrate(f, lower = 0, upper = obs, method = "pcubature")$integral
+  ))
+}

### 2022-03-28 src/1_assess_lengthscale-marginal/script.R

@@ -1,0 +1,56 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("assess_lengthscale-marginal")
+# setwd("src/assess_lengthscale-marginal")
+
+geometries <- c("grid", "civ", "tex", as.character(1:4))
+sim_models <- c("ik")
+
+inf_models <- c("ck_stan")
+# inf_models <- c("ck_stan", "ik_stan")
+
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models,
+  "inf_model" = inf_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to process fits
+process_fits <- function(geometry, sim_model, inf_model) {
+  pb$tick()$print()
+
+  #' CK models can't handle concentric circles
+  if(geometry == "2" & inf_model %in% c("fck_inla", "ck_stan")) return(NULL)
+
+  #' The fitted models
+  fits_loc <- paste0("depends/fits_", sim_model, "_", geometry, "_", inf_model, ".rds")
+  fits <- readRDS(fits_loc)
+
+  #' Assess the lengthscale marginal
+  df <- lapply(fits, assess_marginal_lengthscale, lengthscale = 5/2) %>%
+    bsae::list_to_df() %>%
+    #' Add columns for meta data
+    mutate(
+      geometry = geometry,
+      sim_model = sim_model,
+      inf_model = inf_model,
+      .before = replicate
+    )
+
+  return(df)
+}
+
+#' Eventually this will be iterated over geometries, sim_models and inf_models
+df <- purrr::pmap_df(pars, process_fits)
+
+df <- df %>%
+  mutate(
+    mse_mean = (obs - mean)^2,
+    mae_mean = abs(obs - mean),
+    mse_mode = (obs - mode)^2,
+    mae_mode = abs(obs - mode)
+  )
+
+saveRDS(df, "df.rds")

### 2022-03-28 src/1_assess_rho-marginals/script.R

@@ -1,0 +1,63 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("assess_rho-marginals")
+# setwd("src/assess_rho-marginals")
+
+geometries <- c("grid", "civ", "tex", as.character(1:4))
+sim_models <- c("iid", "icar", "ik")
+
+#' Currently we're testing with a subset, but in future it will seven different models
+inf_models <- c("constant_inla", "iid_inla", "besag_inla", "bym2_inla", "fck_inla", "fik_inla")
+# inf_models <- c("constant_inla", "iid_inla", "besag_inla", "bym2_inla", "fck_inla", "fik_inla", "ck_stan", "ik_stan")
+
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models,
+  "inf_model" = inf_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to process fits
+process_fits <- function(geometry, sim_model, inf_model) {
+  pb$tick()$print()
+
+  #' CK models can't handle concentric circles
+  if(geometry == "2" & inf_model %in% c("fck_inla", "ck_stan")) return(NULL)
+
+  #' The underlying truth
+  data_loc <- paste0("depends/data_", sim_model, "_", geometry, ".rds")
+  data <- readRDS(data_loc)
+
+  #' The fitted models
+  fits_loc <- paste0("depends/fits_", sim_model, "_", geometry, "_", inf_model, ".rds")
+  fits <- readRDS(fits_loc)
+
+  #' Assess the rho marginals
+  rhos <- lapply(data, "[[", "rho")
+  df <- Map(assess_marginals_rho, rhos, fits) %>%
+    bsae::list_to_df() %>%
+    #' Add columns for meta data
+    mutate(
+      geometry = geometry,
+      sim_model = sim_model,
+      inf_model = inf_model,
+      .before = replicate
+    )
+
+  return(df)
+}
+
+#' Eventually this will be iterated over geometries, sim_models and inf_models
+df <- purrr::pmap_df(pars, process_fits)
+
+#' Calculated in post (vectorised) rather than in the purrr:: call
+df <- df %>%
+  mutate(
+    mse_mean = (obs - mean)^2,
+    mae_mean = abs(obs - mean),
+    mse_mode = (obs - mode)^2,
+    mae_mode = abs(obs - mode)
+  )
+
+saveRDS(df, "df.rds")

### 2022-03-28 src/1_assess_time-taken/script.R

@@ -1,0 +1,46 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("assess_time-taken")
+# setwd("src/assess_time-taken")
+
+geometries <- c("grid", "civ", "tex", as.character(1:4))
+sim_models <- c("iid", "icar", "ik")
+
+#' Currently we're testing with a subset, but in future it will seven different models
+inf_models <- c("constant_inla", "iid_inla", "besag_inla", "bym2_inla", "fck_inla", "fik_inla")
+# inf_models <- c("constant_inla", "iid_inla", "besag_inla", "bym2_inla", "fck_inla", "fik_inla", "ck_stan", "ik_stan")
+
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models,
+  "inf_model" = inf_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to process fits
+process_fits <- function(geometry, sim_model, inf_model) {
+  pb$tick()$print()
+
+  #' The fitted models
+  fits_loc <- paste0("depends/fits_", sim_model, "_", geometry, "_", inf_model, ".rds")
+  fits <- readRDS(fits_loc)
+
+  #' Assess the intercept marginal
+  df <- lapply(fits, function(fit) list(t = get_time(fit))) %>%
+    bsae::list_to_df() %>%
+    #' Add columns for meta data
+    mutate(
+      geometry = geometry,
+      sim_model = sim_model,
+      inf_model = inf_model,
+      .before = replicate
+    )
+
+  return(df)
+}
+
+#' Eventually this will be iterated over geometries, sim_models and inf_models
+df <- purrr::pmap_df(pars, process_fits)
+
+saveRDS(df, "df.rds")

### 2022-03-28 src/1_check_lengthscale-recovery/script.R

@@ -1,0 +1,88 @@
+# orderly::orderly_develop_start("check_lengthscale-recovery")
+# setwd("src/check_lengthscale-recovery")
+
+#' Simulating data from l = 1 or 3 and seeing if it can be recovered
+
+grid <- readRDS("depends/grid.rds")
+
+#' l = 1
+ck_one <- sim_ck(grid, nsim = 10, l = 1, save = FALSE)
+
+fits_one <- lapply(ck_one, FUN = function(x) bsae::ck_stan(x$sf))
+lengthscale_summary_one <- lapply(fits_one, FUN = function(fit) summary(fit)$summary["l", ])
+df_lengthscale_one <- list_to_df(lengthscale_summary_one)
+
+pdf("1-recovery-lengthscale.pdf", h = 3, w = 6.25)
+
+ggplot(df_lengthscale_one, aes(x = replicate, y = mean)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
+  geom_hline(yintercept = 1, size = 0.75, col = "grey", alpha = 0.75) +
+  labs(x = "Replicate", y = "Posterior CrI", title = "Is l = 1 recovered?") +
+  theme_minimal()
+
+dev.off()
+
+#' Ratio of lengthscale to noise
+sigma_summary_one <- lapply(fits_one, FUN = function(fit) summary(fit)$summary["sigma_phi", ])
+df_sigma_one <- list_to_df(sigma_summary_one)
+
+pdf("1-recovery-sigma.pdf", h = 3, w = 6.25)
+
+ggplot(df_sigma_one, aes(x = replicate, y = mean)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
+  geom_hline(yintercept = 1, size = 0.75, col = "grey", alpha = 0.75) +
+  labs(x = "Replicate", y = "Posterior CrI", title = "Is sigma = 1 recovered?") +
+  theme_minimal()
+
+dev.off()
+
+pdf("1-lengthscale-sigma-relationship-mean.pdf", h = 3, w = 3)
+
+plot(df_sigma_one$mean, df_lengthscale_one$mean)
+
+dev.off()
+
+lengthscale_draws_one <- lapply(fits_one, FUN = function(fit) as.data.frame(rstan::extract(fit)$l))
[…]

### 2022-03-28 src/1_check_plausible-lengthscale/functions.R

@@ -1,0 +1,18 @@
+graph_length_plot <- function(g) {
+  nb <- sf_to_nb(g)
+
+  nb_sf <- spdep::nb2lines(nb, coords = sp::coordinates(as(g, "Spatial"))) %>%
+    as("sf") %>%
+    st_set_crs(st_crs(g))
+
+  line_sf <- nb_sf %>%
+    mutate(length = round(sf::st_length(nb_sf), 2),
+           geometry = sf::st_centroid(geometry))
+
+  ggplot(g) +
+    geom_sf() +
+    geom_sf(data = nb_sf) +
+    geom_sf_label(data = line_sf, aes(label = length)) +
+    theme_minimal() +
+    theme_void()
+}

### 2022-03-28 src/1_check_prior-sensitivity/script.R

@@ -1,0 +1,117 @@
+# orderly::orderly_develop_start("check_prior-sensitivity")
+# setwd("src/check_prior-sensitivity")
+
+#' Investigating the sensitivity of the length-scale posterior to prior specification
+
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+data <- readRDS("depends/data_iid_grid.rds")
+sf <- data[[1]]$sf
+
+D <- centroid_distance(sf)
+dat <- list(n = nrow(sf), y = sf$y, m = sf$n_obs, mu = rep(0, nrow(sf)), D = D)
+
+# No prior on l (uniform): this shouldn't be a good idea
+fit_non <- rstan::stan("prior-sensitivity/non-informative.stan", data = dat, warmup = 100, iter = 1000)
+# Nice visual checking tool--a lot of between chain posterior variability would be a bad sign
+bayesplot::mcmc_hist_by_chain(fit_non, pars = "l")
+
+# l ~ Gamma(1, 1)
+fit_gamma <- rstan::stan("prior-sensitivity/gamma_.stan", data = dat, warmup = 100, iter = 1000)
+bayesplot::mcmc_hist_by_chain(fit_non, pars = "l")
+
+# l ~ N+(0, (ub - lb) / 3): Half-normal with standard deviation based on range of distances
+# https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html
+dist <- as.vector(D)
+lb <- min(dist) # Could do dist[dist > 0] here but results in lb = 1
+lb <- 0.1
+ub <- max(dist)
+(ub - lb) / 3 # Approximately 2.3
+fit_normal_data <- rstan::stan("prior-sensitivity/normal-data.stan", data = dat, warmup = 100, iter = 1000)
+bayesplot::mcmc_hist_by_chain(fit_normal_data, pars = "l")
+
+# l ~ N(1, 0.5): Informative prior at the true value
+fit_normal_inform <- rstan::stan("prior-sensitivity/informative.stan", data = dat, warmup = 100, iter = 1000)
+bayesplot::mcmc_hist_by_chain(fit_normal_inform, pars = "l")
+
+# l ~ Inverse-Gamma(a, b):
+# https://warwick.ac.uk/fac/sci/statistics/crism/workshops/masterclassapril/presentation4.pdf
+
[…]

### 2022-03-28 src/1_fit_besag-inla/script.R

@@ -1,0 +1,51 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("fit_besag-inla")
+# setwd("src/fit_besag-inla")
+
+vignette_geometries <- as.character(1:4)
+realistic_geometries <- c("grid", "civ", "tex")
+
+geometries <- c()
+if(vignette) geometries <- c(geometries, vignette_geometries)
+if(realistic) geometries <- c(geometries, realistic_geometries)
+if(length(geometries) == 0) stop("Either vignette or realistic must be TRUE")
+
+sim_models <- c("iid", "icar", "ik")
+
+#' Prevent orderly complaining if I don't produce all simulations all of the time
+lapply(
+  expand.grid(
+    "sim_model" = sim_models,
+    "geometry" = c(vignette_geometries, realistic_geometries)
+  ) %>%
+    as.data.frame() %>%
+    mutate(file = paste0("fits_", sim_model, "_", geometry, ".rds")) %>%
+    pull(file),
+  function(file) saveRDS(NULL, file)
+)
+
+#' Geometry and simulation model
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to run bsae::besag_inla
+run_models <- function(geometry, sim_model) {
+  pb$tick()$print()
+  data <- readRDS(paste0("depends/data_", sim_model, "_", geometry, ".rds"))
+  fits <- lapply(data, function(x) {
+    fit <- bsae::besag_inla(x$sf)
+    samples <- INLA::inla.posterior.sample(n = 1000, fit)
+    fit[["samples"]] <- samples
+    class(fit) <- "inlax"
+    return(fit)
+  })
+  saveRDS(fits, file = paste0("fits_", sim_model, "_", geometry, ".rds"))
+}
+
+# Run models and save for each row of pars
+purrr::pmap_df(pars, run_models)

### 2022-03-28 src/1_fit_bym2-inla/script.R

@@ -1,0 +1,51 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("fit_bym2-inla")
+# setwd("src/fit_bym2-inla")
+
+vignette_geometries <- as.character(1:4)
+realistic_geometries <- c("grid", "civ", "tex")
+
+geometries <- c()
+if(vignette) geometries <- c(geometries, vignette_geometries)
+if(realistic) geometries <- c(geometries, realistic_geometries)
+if(length(geometries) == 0) stop("Either vignette or realistic must be TRUE")
+
+sim_models <- c("iid", "icar", "ik")
+
+#' Prevent orderly complaining if I don't produce all simulations all of the time
+lapply(
+  expand.grid(
+    "sim_model" = sim_models,
+    "geometry" = c(vignette_geometries, realistic_geometries)
+  ) %>%
+    as.data.frame() %>%
+    mutate(file = paste0("fits_", sim_model, "_", geometry, ".rds")) %>%
+    pull(file),
+  function(file) saveRDS(NULL, file)
+)
+
+#' Geometry and simulation model
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to run bsae::bym2_inla
+run_models <- function(geometry, sim_model) {
+  pb$tick()$print()
+  data <- readRDS(paste0("depends/data_", sim_model, "_", geometry, ".rds"))
+  fits <- lapply(data, function(x) {
+    fit <- bsae::bym2_inla(x$sf)
+    samples <- INLA::inla.posterior.sample(n = 1000, fit)
+    fit[["samples"]] <- samples
+    class(fit) <- "inlax"
+    return(fit)
+  })
+  saveRDS(fits, file = paste0("fits_", sim_model, "_", geometry, ".rds"))
+}
+
+# Run models and save for each row of pars
+purrr::pmap_df(pars, run_models)

### 2022-03-28 src/1_fit_ck-stan/script.R

@@ -1,0 +1,49 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("fit_ck-stan")
+# setwd("src/fit_ck-stan")
+
+vignette_geometries <- as.character(1:4)
+realistic_geometries <- c("grid", "civ", "tex")
+
+geometries <- c()
+if(vignette) geometries <- c(geometries, vignette_geometries)
+if(realistic) geometries <- c(geometries, realistic_geometries)
+if(length(geometries) == 0) stop("Either vignette or realistic must be TRUE")
+
+sim_models <- c("iid", "icar", "ik")
+
+#' Prevent orderly complaining if I don't produce all simulations all of the time
+lapply(
+  expand.grid(
+    "sim_model" = sim_models,
+    "geometry" = c(vignette_geometries, realistic_geometries)
+  ) %>%
+    as.data.frame() %>%
+    mutate(file = paste0("fits_", sim_model, "_", geometry, ".rds")) %>%
+    pull(file),
+  function(file) saveRDS(NULL, file)
+)
+
+#' Geometry and simulation model
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to run bsae::ck_stan
+run_models <- function(geometry, sim_model) {
+  pb$tick()$print()
+
+  #' This model can't handle the concentric circles case!
+  if(geometry == "2") return(NULL)
+
+  data <- readRDS(paste0("depends/data_", sim_model, "_", geometry, ".rds"))
+  fits <- lapply(data, function(x) bsae::ck_stan(x$sf))
+  saveRDS(fits, file = paste0("fits_", sim_model, "_", geometry, ".rds"))
+}
+
+# Run models and save for each row of pars
+purrr::pmap_df(pars, run_models)

### 2022-03-28 src/1_fit_fck-inla/script.R

@@ -1,0 +1,55 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("fit_fck-inla")
+# setwd("src/fit_fck-inla")
+
+vignette_geometries <- as.character(1:4)
+realistic_geometries <- c("grid", "civ", "tex")
+
+geometries <- c()
+if(vignette) geometries <- c(geometries, vignette_geometries)
+if(realistic) geometries <- c(geometries, realistic_geometries)
+if(length(geometries) == 0) stop("Either vignette or realistic must be TRUE")
+
+sim_models <- c("iid", "icar", "ik")
+
+#' Prevent orderly complaining if I don't produce all simulations all of the time
+lapply(
+  expand.grid(
+    "sim_model" = sim_models,
+    "geometry" = c(vignette_geometries, realistic_geometries)
+  ) %>%
+    as.data.frame() %>%
+    mutate(file = paste0("fits_", sim_model, "_", geometry, ".rds")) %>%
+    pull(file),
+  function(file) saveRDS(NULL, file)
+)
+
+#' Geometry and simulation model
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to run bsae::fck_inla
+run_models <- function(geometry, sim_model) {
+  pb$tick()$print()
+
+  #' This model can't handle the concentric circles case!
+  if(geometry == "2") return(NULL)
+
+  data <- readRDS(paste0("depends/data_", sim_model, "_", geometry, ".rds"))
+  fits <- lapply(data, function(x) {
+    fit <- bsae::fck_inla(x$sf)
+    samples <- INLA::inla.posterior.sample(n = 1000, fit)
+    fit[["samples"]] <- samples
+    class(fit) <- "inlax"
+    return(fit)
+  })
+  saveRDS(fits, file = paste0("fits_", sim_model, "_", geometry, ".rds"))
+}
+
+# Run models and save for each row of pars
+purrr::pmap_df(pars, run_models)

### 2022-03-28 src/1_fit_fik-inla/script.R

@@ -1,0 +1,51 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("fit_fik-inla")
+# setwd("src/fit_fik-inla")
+
+vignette_geometries <- as.character(1:4)
+realistic_geometries <- c("grid", "civ", "tex")
+
+geometries <- c()
+if(vignette) geometries <- c(geometries, vignette_geometries)
+if(realistic) geometries <- c(geometries, realistic_geometries)
+if(length(geometries) == 0) stop("Either vignette or realistic must be TRUE")
+
+sim_models <- c("iid", "icar", "ik")
+
+#' Prevent orderly complaining if I don't produce all simulations all of the time
+lapply(
+  expand.grid(
+    "sim_model" = sim_models,
+    "geometry" = c(vignette_geometries, realistic_geometries)
+  ) %>%
+    as.data.frame() %>%
+    mutate(file = paste0("fits_", sim_model, "_", geometry, ".rds")) %>%
+    pull(file),
+  function(file) saveRDS(NULL, file)
+)
+
+#' Geometry and simulation model
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to run bsae::fik_inla
+run_models <- function(geometry, sim_model) {
+  pb$tick()$print()
+  data <- readRDS(paste0("depends/data_", sim_model, "_", geometry, ".rds"))
+  fits <- lapply(data, function(x) {
+    fit <- bsae::fik_inla(x$sf)
+    samples <- INLA::inla.posterior.sample(n = 1000, fit)
+    fit[["samples"]] <- samples
+    class(fit) <- "inlax"
+    return(fit)
+  })
+  saveRDS(fits, file = paste0("fits_", sim_model, "_", geometry, ".rds"))
+}
+
+# Run models and save for each row of pars
+purrr::pmap_df(pars, run_models)

### 2022-03-28 src/1_fit_ik-stan/script.R

@@ -1,0 +1,45 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("fit_ik-stan")
+# setwd("src/fit_ik-stan")
+
+vignette_geometries <- as.character(1:4)
+realistic_geometries <- c("grid", "civ", "tex")
+
+geometries <- c()
+if(vignette) geometries <- c(geometries, vignette_geometries)
+if(realistic) geometries <- c(geometries, realistic_geometries)
+if(length(geometries) == 0) stop("Either vignette or realistic must be TRUE")
+
+sim_models <- c("iid", "icar", "ik")
+
+#' Prevent orderly complaining if I don't produce all simulations all of the time
+lapply(
+  expand.grid(
+    "sim_model" = sim_models,
+    "geometry" = c(vignette_geometries, realistic_geometries)
+  ) %>%
+    as.data.frame() %>%
+    mutate(file = paste0("fits_", sim_model, "_", geometry, ".rds")) %>%
+    pull(file),
+  function(file) saveRDS(NULL, file)
+)
+
+#' Geometry and simulation model
+pars <- expand.grid(
+  "geometry" = geometries,
+  "sim_model" = sim_models
+)
+
+#' Initialise progress bar
+pb <- progress_estimated(nrow(pars))
+
+#' Function to run bsae::ik_stan
+run_models <- function(geometry, sim_model) {
+  pb$tick()$print()
+  data <- readRDS(paste0("depends/data_", sim_model, "_", geometry, ".rds"))
+  fits <- lapply(data, function(x) bsae::ik_stan(x$sf))
+  saveRDS(fits, file = paste0("fits_", sim_model, "_", geometry, ".rds"))
+}
+
+# Run models and save for each row of pars
+purrr::pmap_df(pars, run_models)

### 2022-03-28 src/1_plot_covariance-matrix/script.R

@@ -1,0 +1,3 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("plot_covariance-matrix")
+# setwd("src/plot_covariance-matrix")

### 2022-03-28 src/1_plot_coverage/script.R

@@ -1,0 +1,74 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("plot_coverage")
+# setwd("src/plot_coverage")
+
+df <- readRDS("depends/df_rho.rds") %>%
+  mutate(replicate = as.numeric(replicate)) %>%
+  bsae::update_naming() %>%
+  filter(inf_model != "Constant")
+
+geometries <- unique(df$geometry)
+
+pdf("histogram-ecdf-no-constant.pdf", h = 8, w = 6.25)
+
+lapply(geometries, function(x) {
+  df <- filter(df, geometry == as.character(x))
+  S <- max(df$replicate) * max(df$id) #' Number of Monte Carlo samples
+  bins <- 20
+  alpha <- 0.05
+
+  ci <- qbinom(
+    p = c(alpha / 2, 0.5, (1 - alpha / 2)),
+    size = S,
+    prob = 1 / bins
+  )
+
+  polygon_data <- data.frame(
+    x = c(-0.05, 0, 1, 0, -0.05, 1.05, 1, 1.05, -0.05),
+    y = c(ci[1], ci[2], ci[2], ci[2], ci[3], ci[3], ci[2], ci[1], ci[1]) / S
+  )
+
+  plotA <- ggplot(df, aes(x = quantile)) +
+    facet_grid(sim_model ~ inf_model, drop = TRUE, scales = "free") +
+    geom_histogram(aes(y = (..count..) / tapply(..count..,..PANEL..,sum)[..PANEL..]),
+                   breaks = seq(0, 1, length.out = bins + 1), fill = "#009E73", col = "black", alpha = 0.9) +
+    geom_polygon(data = polygon_data, aes(x = x, y = y), fill = "grey75", color = "grey50", alpha = 0.6) +
+    labs(x = "", y = "", title = paste0(x)) +
+    scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c(0, 0.25, 0.5, 0.75, 1)) +
+    theme_minimal()
+
+  lims <- get_lims(n = S, alpha, K = 100)
+
+  plotB <- df %>%
+    filter(!is.na(quantile)) %>%
+    split(~sim_model + inf_model) %>%
+    lapply(function(y) {
+      empirical_coverage <- purrr::map_dbl(seq(0, 1, by = 0.01), ~ multi.utils::empirical_coverage(y$quantile, .x))
[…]

### 2022-03-28 src/1_plot_lengthscale-recovery/functions.R

@@ -1,0 +1,28 @@
+#' Values of the length-scale recovered by the best_average() function, without any data
+calc_best_average <- function(file) {
+  readRDS(paste0("depends/", file)) %>%
+    centroid_distance() %>%
+    bsae::best_average()
+}
+
+lengthscale_plot <- function(df, inf_model, geometry, best = NA, subtitle = NA) {
+
+  overall_mean <- df %>%
+    filter(sim_model == "IK", inf_model == !!inf_model, geometry == !!geometry) %>%
+    summarise(overall_mean = mean(mean)) %>%
+    as.numeric()
+
+  df %>%
+    filter(sim_model == "IK", inf_model == !!inf_model, geometry == !!geometry) %>%
+    select(mean, upper, lower) %>%
+    tibble::rownames_to_column(var = "id") %>%
+    ggplot(aes(x = id, y = mean)) +
+    geom_point(alpha = 0.5) +
+    geom_errorbar(aes(ymin = lower, ymax = upper), alpha = 0.5) +
+    geom_hline(yintercept = 2.5, col = lightblue, size = 1.5) +
+    geom_hline(yintercept = overall_mean, col = lightgreen, size = 1.5) +
+    geom_hline(yintercept = best, col = "#E69F00", size = 1.5) +
+    labs(x = "Simulation number", y = "Lengthscale", subtitle = subtitle) +
+    theme(axis.text.x=element_blank(),
+          axis.ticks.x=element_blank())
+}

### 2022-03-28 src/1_plot_lengthscale-recovery/script.R

@@ -1,0 +1,59 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("plot_lengthscale-recovery")
+# setwd("src/plot_lengthscale-recovery")
+
+df_lengthscale <- readRDS("depends/df_lengthscale.rds")
+
+best_lengthscale <- c(
+  "Grid" = calc_best_average("grid.rds"),
+  "Cote d'Ivoire" = calc_best_average("civ.rds"),
+  "Texas" = calc_best_average("tex.rds"),
+  "1" = calc_best_average("geometry-1.rds"),
+  "2" = calc_best_average("geometry-2.rds"),
+  "3" = calc_best_average("geometry-3.rds"),
+  "4" = calc_best_average("geometry-4.rds")
+)
+
+lengthscale_mean <- df_lengthscale %>%
+  group_by(inf_model, geometry) %>%
+  summarise(overall_mean = mean(mean))
+
+cbpalette <- multi.utils::cbpalette()
+
+pdf("lengthscale-recovery.pdf", h = 4, w = 6.25)
+
+df_lengthscale %>%
+  left_join(
+    lengthscale_mean,
+    by = c("inf_model", "geometry")
+  ) %>%
+  bsae::update_naming() %>%
+  split(.$geometry) %>%
+  lapply(function(x) {
+    x_max <- max(as.numeric(x$replicate))
+    obs <- mean(x$obs)
+    overall_mean <- mean(x$overall_mean)
+    best <- best_lengthscale[x$geometry[1]]
+
+    x %>%
+      select(replicate, obs, mean, upper, lower, overall_mean) %>%
+      ggplot(aes(x = replicate, y = mean)) +
+      geom_point(alpha = 0.5) +
+      geom_errorbar(aes(ymin = lower, ymax = upper), alpha = 0.5) +
+      geom_hline(yintercept = obs, col = cbpalette[1], size = 1, alpha = 0.7) +
+      annotate("text", label = "Truth", col = cbpalette[1], x = x_max + 1.75, y = obs + 0.1, hjust = 0) +
+      geom_hline(yintercept = overall_mean, col = cbpalette[2], size = 1, alpha = 0.7) +
+      annotate("text", label = "Mean", col = cbpalette[2], x = x_max + 1.75, y = overall_mean + 0.1, hjust = 0) +
+      geom_hline(yintercept = best, col = cbpalette[3], size = 1, alpha = 0.7) +
[…]

### 2022-03-28 src/1_process_metric-tables/functions.R

@@ -1,0 +1,116 @@
+group_mean_and_se <- function(df, group_variables) {
+  df %>%
+    select(-c(obs, mean, mode, lower, upper)) %>%
+    group_by(.dots = lapply(group_variables, as.symbol)) %>%
+    summarise(n = n(), across(mse:lds, list(mean = mean, se = ~ sd(.x) / sqrt(length(.x)))))
+}
+
+#' Create table for particular model assessment metric.
+#'
+#' @param df A dataframe of marginal assessments.
+#' @param metric A metric name as a string.
+#' @param title Title for the table as a string, defaults to `NULL`.
+#' @param subtitle Subtitle for the table as a string, defaults to `NULL`.
+#' @param latex Should the table be returned in LaTeX (`TRUE`) or Markdown (`FALSE`, default)?
+#' @param scale A value to multiply all values in table by, to facilitate easier reading, defaults to `1`.
+#' @param figures How many significant figures should be used, defaults to `2`. Passed to `signif`.
+#' @return A table in either Markdown or LaTeX format.
+metric_table <- function(df, metric, title = NULL, subtitle = NULL, latex = FALSE, scale = 1, figures = 2) {
+
+  metric_mean <- paste0(metric, "_mean")
+  metric_se <- paste0(metric, "_se")
+
+  df_wide <- df %>%
+    select(geometry, sim_model, inf_model, !!metric_mean, !!metric_se) %>%
+    rename(mean = !!metric_mean, se = !!metric_se) %>%
+    mutate(mean = scale * signif(mean, figures),
+           se = scale * signif(se, figures),
+           val = paste0(mean, " (", se, ")")) %>%
+    select(-mean, -se) %>%
+    tidyr::spread(inf_model, val)
+
+  table <- df_wide %>%
+    gt(rowname_col = "sim_model", groupname_col = "geometry") %>%
+    tab_style(style = cell_text(weight = "bold"), locations = cells_row_groups()) %>%
+    tab_header(title = title, subtitle = subtitle) %>%
+    tab_spanner(
+      label = "Inferential model",
+      columns = everything()
+    ) %>%
+    tab_stubhead("Simulation model")
+
+  #' The column with the minimum value of the metric
+  min_ind <- pmap_int(
+    df %>%
+      select(geometry, sim_model, inf_model, !!metric_mean) %>%
+      tidyr::spread(inf_model, !!metric_mean) %>%
[…]

### 2022-03-28 src/1_process_metric-tables/script.R

@@ -1,0 +1,55 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("process_metric-tables")
+# setwd("src/process_metric-tables")
+
+df_rho <- readRDS("depends/df_rho.rds")
+df_intercept <- readRDS("depends/df_intercept.rds")
+df_time <- readRDS("depends/df_time.rds")
+
+#' Some bug happening currently with latex = TRUE here (not with latex = FALSE, so nothing fundamental)
+
+#' CRPS table for rho
+sink("crps-table-rho.txt")
+
+df_rho %>%
+  group_mean_and_se(group_variables = c("geometry", "sim_model", "inf_model")) %>%
+  bsae::update_naming() %>%
+  metric_table(
+    metric = "crps",
+    title = "Continuous Ranked Probability Score",
+    latex = TRUE,
+    scale = 10^3,
+    figures = 3
+  )
+
+sink()
+
+#' CRPS table for the intercept
+sink("crps-table-intercept.txt")
+
+df_intercept %>%
+  group_mean_and_se(group_variables = c("geometry", "sim_model", "inf_model")) %>%
+  bsae::update_naming() %>%
+  metric_table(
+    metric = "crps",
+    title = "Continuous Ranked Probability Score",
+    latex = TRUE,
+    scale = 10^3,
+    figures = 3
+  )
+
+sink()
+
+#' Time table
+sink("time-table.txt")
+
+df_time %>%
+  group_by(geometry, sim_model, inf_model) %>%
+  bsae::update_naming() %>%
+  summarise(n = n(), across(t, list(mean = mean, se = ~ sd(.x) / sqrt(length(.x))))) %>%
+  metric_table(
+    metric = "t",
+    latex = FALSE
+  )
+
+sink()

### 2022-03-28 src/2_assess_cv/script.R

@@ -1,0 +1,80 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("assess_cv")
+# setwd("src/assess_cv")
+
+inf_models <- c("constant_inla", "iid_inla", "besag_inla", "bym2_inla")
+# inf_models <- c("constant_inla", "iid_inla", "besag_inla", "bym2_inla", "fck_inla", "fik_inla", "ck_stan", "ik_stan")
+surveys <- c("civ2017phia", "mwi2016phia", "tza2017phia", "zwe2016phia")
+types <- c("loo", "sloo")
+
+#' Direct model fit measures (DIC, WAIC and CPO)
+pars <- expand.grid(
+  "inf_model" = inf_models,
+  "survey" = surveys
+)
+
+direct <- purrr::pmap_df(pars, function(inf_model, survey) {
+  fit <- readRDS(paste0("depends/fit_", substr(survey, 1, 3), "_", inf_model, ".rds"))
+  data.frame(dic = fit$dic$local.dic, waic = fit$waic$local.waic, cpo = fit$cpo$cpo) %>%
+    tibble::rownames_to_column("id") %>%
+    mutate(
+      survey = survey,
+      inf_model = inf_model,
+      .before = id
+    ) %>%
+    mutate(id = as.numeric(id))
+})
+
+saveRDS(direct, "direct.rds")
+
+#' Bind all of the direct fits together
+
+df_models <- purrr::pmap_df(pars, function(inf_model, survey) {
+  fit <- readRDS(paste0("depends/fit_", substr(survey, 1, 3), "_", inf_model, ".rds"))
+  st_read(paste0("depends/", survey, ".geojson")) %>%
+    select(survey_id, area_id, area_name, geometry) %>%
+    mutate(
+      inf_model = inf_model,
+      estimate = fit$summary.fitted.values$mean,
+      ci_lower = fit$summary.fitted.values$`0.025quant`,
+      ci_upper = fit$summary.fitted.values$`0.975quant`,
+      .before = geometry
+    )
+})
+
+df_raw <- purrr::pmap_df(list(surveys), function(survey) {
+  st_read(paste0("depends/", survey, ".geojson")) %>%
+    select(survey_id, area_id, area_name, estimate, ci_lower, ci_upper, geometry) %>%
+    mutate(
+      inf_model = "raw",
+      .before = estimate
+    )
+})
+
[…]

### 2022-03-28 src/2_cv_constant-inla/script.R

@@ -1,0 +1,41 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("cv_constant-inla")
+# setwd("src/cv_constant-inla")
+
+surveys <- c("civ2017phia", "mwi2016phia", "tza2017phia", "zwe2016phia")
+types <- c("loo", "sloo")
+
+pars <- expand.grid(
+  "survey" = surveys,
+  "type" = types
+)
+
+#' Function to run bsae::constant_inla for each cross-validation fold
+run_cv_models <- function(survey, type) {
+  sf <- st_read(paste0("depends/", survey, ".geojson"))
+  training_sets <- create_folds(sf, type = toupper(type))
+
+  print(paste0("Begin ", toupper(type), " cross-valdiation of ", toupper(survey)))
+  pb <- progress_estimated(nrow(sf)) #' Initialise progress bar
+
+  training_sets <- lapply(training_sets, function(training_set) {
+    training_set$fit <- bsae::constant_inla(training_set$data)
+    pb$tick()$print()
+    return(training_set)
+  })
+
+  saveRDS(training_sets, file = paste0("fits_", substr(survey, 1, 3), "_", tolower(type), ".rds"))
+}
+
+#' Run models and save for each row of pars
+purrr::pmap_df(pars, run_cv_models)
+
+#' Function to run bsae::constant_inla for the whole dataset
+#' This will be useful for comparing R-INLA and Stan inbuilt model comparison metrics to manual CV
+run_models <- function(survey, type) {
+  sf <- st_read(paste0("depends/", survey, ".geojson"))
+  fit <- bsae::constant_inla(sf)
+  saveRDS(fit, file = paste0("fit_", substr(survey, 1, 3), ".rds"))
+}
+
+purrr::pmap_df(expand.grid("survey" = surveys), run_models)

### 2022-03-28 src/2_cv_ik-stan/script.R

@@ -1,0 +1,41 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("cv_ik-stan")
+# setwd("src/cv_ik-stan")
+
+surveys <- c("civ2017phia", "mwi2016phia", "tza2017phia", "zwe2016phia")
+types <- c("loo", "sloo")
+
+pars <- expand.grid(
+  "survey" = surveys,
+  "type" = types
+)
+
+#' Function to run bsae::ik_stan for each cross-validation fold
+run_cv_models <- function(survey, type) {
+  sf <- st_read(paste0("depends/", survey, ".geojson"))
+  training_sets <- create_folds(sf, type = toupper(type))
+
+  print(paste0("Begin ", toupper(type), " cross-valdiation of ", toupper(survey)))
+  pb <- progress_estimated(nrow(sf)) #' Initialise progress bar
+
+  training_sets <- lapply(training_sets, function(training_set) {
+    training_set$fit <- bsae::ik_stan(training_set$data)
+    pb$tick()$print()
+    return(training_set)
+  })
+
+  saveRDS(training_sets, file = paste0("fits_", substr(survey, 1, 3), "_", tolower(type), ".rds"))
+}
+
+#' Run models and save for each row of pars
+purrr::pmap_df(pars, run_cv_models)
+
+#' Function to run bsae::ik_stan for the whole dataset
+#' This will be useful for comparing R-INLA and Stan inbuilt model comparison metrics to manual CV
+run_models <- function(survey, type) {
+  sf <- st_read(paste0("depends/", survey, ".geojson"))
+  fit <- bsae::ik_stan(sf)
+  saveRDS(fit, file = paste0("fit_", substr(survey, 1, 3), ".rds"))
+}
+
+purrr::pmap_df(expand.grid("survey" = surveys), run_models)

### 2022-03-28 src/2_plot_metric-maps/script.R

@@ -1,0 +1,51 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("plot_hiv-metric-maps")
+# setwd("src/plot_hiv-metric-maps")
+
+surveys <- c("civ2017phia", "mwi2016phia", "tza2017phia", "zwe2016phia")
+
+hiv_surveys <- lapply(surveys, function(x) {
+  st_read(paste0("depends/", x, ".geojson")) %>%
+    tibble::rownames_to_column("id") %>%
+    mutate(id = as.numeric(id))
+  }) %>%
+  bind_rows() %>%
+  select(survey = survey_id, id, geometry)
+
+sf <- readRDS("depends/df.rds") %>%
+  mutate(
+    survey = toupper(survey),
+    id = as.numeric(id),
+    inf_model = recode_factor(inf_model,
+      "constant_inla" = "Constant",
+      "iid_inla" = "IID",
+      "besag_inla" = "Besag",
+      "bym2_inla" = "BYM2",
+      "fck_inla" = "FCK",
+      "ck_stan" = "CK",
+      "fik_inla" = "FIK",
+      "ik_stan" = "IK")
+  ) %>%
+  left_join(
+    hiv_surveys,
+    by = c("survey", "id")
+  ) %>%
+  st_as_sf()
+
+pdf("crps-map-hiv-surveys-no-constant.pdf", h = 5, w = 6.25)
+
+metric_map(sf, metric = "crps")
+
+dev.off()
+
+pdf("mse-map-hiv-surveys-no-constant.pdf", h = 5, w = 6.25)
+
+metric_map(sf, metric = "mse")
+
+dev.off()
+
+pdf("mae-map-hiv-surveys-no-constant.pdf", h = 5, w = 6.25)
+
+metric_map(sf, metric = "mae")
+
+dev.off()

### 2022-03-28 src/2_process_hiv-surveys/functions.R

@@ -1,0 +1,29 @@
+extract_survey <- function(iso3) {
+  df <- read_csv(file = paste0("depends/", iso3, "_survey_hiv_indicators.csv"))
+  areas <- read_sf(paste0("depends/", tolower(iso3), "_areas.geojson"))
+
+  sf <- df  %>%
+    mutate(iso3 = substr(area_id, 1, 3)) %>%
+    filter(
+      survey_id == survey_name[toupper(iso3)],
+      indicator == "prevalence",
+      age_group == "Y015_049",
+      sex == "both"
+    ) %>%
+    right_join(
+      areas %>%
+        filter(area_level == analysis_level[toupper(iso3)]),
+      by = c("area_id", "area_name")
+    ) %>%
+    mutate(
+      y = estimate * n_eff_kish,
+      n_obs = n_eff_kish,
+      .after = ci_upper
+    ) %>%
+    st_as_sf() %>%
+    st_transform(4326)
+
+  st_write(sf, paste0(tolower(survey_name[toupper(iso3)]), ".geojson"))
+
+  return(sf)
+}

### 2022-03-28 src/2_process_metric-tables/functions.R

@@ -1,0 +1,46 @@
+#' Only works for LOO and SLOO metrics
+metric_table <- function(df, metric, title = NULL, subtitle = NULL, latex = FALSE, scale = 1) {
+
+  metric_mean.LOO <- paste0(metric, "_mean.LOO")
+  metric_se.LOO <- paste0(metric, "_se.LOO")
+  metric_mean.SLOO <- paste0(metric, "_mean.SLOO")
+  metric_se.SLOO <- paste0(metric, "_se.SLOO")
+
+  df <- df %>%
+    select(geometry, inf_model,
+           !!metric_mean.LOO, !!metric_se.LOO,
+           !!metric_mean.SLOO, !!metric_se.SLOO) %>%
+    rename(mean.LOO = !!metric_mean.LOO, se.LOO = !!metric_se.LOO,
+           mean.SLOO = !!metric_mean.SLOO, se.SLOO = !!metric_se.SLOO) %>%
+    mutate(mean.LOO = scale * signif(mean.LOO, 3),
+           se.LOO = scale * signif(se.LOO, 3),
+           mean.SLOO = scale * signif(mean.SLOO, 3),
+           se.SLOO = scale * signif(se.SLOO, 3),
+           val.LOO = paste0(mean.LOO, " (", se.LOO, ")"),
+           val.SLOO = paste0(mean.SLOO, " (", se.SLOO, ")")) %>%
+    select(-mean.LOO, -se.LOO, -mean.SLOO, -se.SLOO)
+
+  table <- df %>%
+    arrange(match(inf_model, c("Constant", "IID", "Besag", "BYM2", "FCK", "CK", "FIK", "IK"))) %>%
+    arrange_at(1) %>%
+    gt(rowname_col = "inf_model", groupname_col = "geometry") %>%
+    tab_spanner_delim(delim = ".") %>%
+    tab_style(style = cell_text(weight = "bold"), locations = cells_row_groups()) %>%
+    tab_header(title = title, subtitle = subtitle) %>%
+    tab_spanner(
+      label = toupper(metric),
+      columns = everything()
+    ) %>%
+    tab_stubhead("Inferential model")
+
+  if(latex){
+    return(
+      table %>%
+        as_latex() %>%
+        as.character() %>%
+        cat()
+    )
+  } else {
+    return(table)
+  }
+}

### 2022-03-28 src/2_process_metric-tables/script.R

@@ -1,0 +1,50 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("process_hiv-metric-tables")
+# setwd("src/process_hiv-metric-tables")
+
+manual <- readRDS("depends/manual.rds")
+
+metric <- "mse"
+
+metric_mean.LOO <- paste0(metric, "_mean.LOO")
+metric_se.LOO <- paste0(metric, "_se.LOO")
+metric_mean.SLOO <- paste0(metric, "_mean.SLOO")
+metric_se.SLOO <- paste0(metric, "_se.SLOO")
+
+df <- manual %>%
+  select(survey, inf_model,
+         !!metric_mean.LOO, !!metric_se.LOO,
+         !!metric_mean.SLOO, !!metric_se.SLOO) %>%
+  rename(mean.LOO = !!metric_mean.LOO, se.LOO = !!metric_se.LOO,
+         mean.SLOO = !!metric_mean.SLOO, se.SLOO = !!metric_se.SLOO) %>%
+  mutate(mean.LOO = scale * signif(mean.LOO, 3),
+         se.LOO = scale * signif(se.LOO, 3),
+         mean.SLOO = scale * signif(mean.SLOO, 3),
+         se.SLOO = scale * signif(se.SLOO, 3),
+         val.LOO = paste0(mean.LOO, " (", se.LOO, ")"),
+         val.SLOO = paste0(mean.SLOO, " (", se.SLOO, ")")) %>%
+  select(-mean.LOO, -se.LOO, -mean.SLOO, -se.SLOO)
+
+table <- df %>%
+  arrange(match(inf_model, c("Constant", "IID", "Besag", "BYM2", "FCK", "CK", "FIK", "IK"))) %>%
+  arrange_at(1) %>%
+  gt(rowname_col = "inf_model", groupname_col = "geometry") %>%
+  tab_spanner_delim(delim = ".") %>%
+  tab_style(style = cell_text(weight = "bold"), locations = cells_row_groups()) %>%
+  tab_header(title = title, subtitle = subtitle) %>%
+  tab_spanner(
+    label = toupper(metric),
+    columns = everything()
+  ) %>%
+  tab_stubhead("Inferential model")
+
+if(latex){
+  return(
+    table %>%
+      as_latex() %>%
+      as.character() %>%
+      cat()
+  )
+} else {
+  return(table)
+}

### 2022-07-12 src/0_explore_inla-spde/inla-spde.Rmd
Large upload

---
title: "SPDE approach notes"
author: "Adam Howes"
output: pdf_document
bibliography: citations.bib
---

## List of possible sources

* @miller2020understanding (recommended by Samir)
* [Chapter 9](https://www.paulamoraga.com/book-geospatial/) of Geospatial Health Data: Modeling and Visualization with R-INLA and Shiny by Paula Moraga
* [Most of](https://becarioprecario.bitbucket.io/spde-gitbook/index.html) Advanced Spatial Modeling with Stochastic Partial Differential Equations Using R and INLA by Elias T. Krainski and coauthors
* [Comment](https://groups.google.com/g/r-inla-discussion-group/c/4tKXemUciGU) on Google group by Finn
* [Markdown document](https://rstudio-pubs-static.s3.amazonaws.com/302639_3ccd091c277d4d6c9ad9c2cf524250b6.html) by Zhe Sha

## Background

The solution to certain stochastic partial differential equations (SPDEs) is a Gaussian field with Matérn covariance [@whittle1954stationary].
@lindgren2010explicit propose to represent a Gaussian field with Matérn covariance as a GMRF by representing the solution of the SPDE using the finite element method.
The Gaussian field $f(x)$ with the Matérn covariance function is a solution to the linear fractional (since $\alpha \neq 2$) SPDE
$$
(\kappa^2 - \Delta)^{\alpha/2} f(x) = \epsilon(x), \quad x \in \mathbb{R}^d, \quad \alpha = \nu + d/2, \quad \kappa > 0, \quad \nu > 0,
$$
where $(\kappa^2 - \Delta)^{\alpha/2}$ is a pseduo-differential operator, the innovation process $\epsilon$ is spatial Gaussian white noise with unit variance and $\Delta$ is the Laplacian
$$
\Delta = \sum_{i=1}^d \frac{\delta^2}{\delta x_i^2},
$$
and the marginal variance is
$$
\sigma^2 = \frac{\Gamma(\nu)}{\Gamma(\nu + d/2)(4\pi)^{d/2}\kappa^{2\nu}}
$$

## @miller2020understanding explanation

SPDE equation:

$$Df = \epsilon$$

where $D$ is a differential operator, $f$ is a funciton and $\epsilon$ is a stochastic process (often white noise).
This corresponds to the integral form:

$$Df = \epsilon \iff \int Df(x) \phi(x) \text{d}x = \int \epsilon(x) \phi(x) \text{d}x \, \forall \phi$$

where $\phi$ is known as a test function.
Using the inner product notation $\langle f, g \rangle = \int f(x) g(x) \text{d}x$ this is equivalent to:

$$\langle Df, \phi \rangle = \langle \epsilon, \phi \rangle$$

Solve for points on a mesh $j = 1, \ldots, M$ and $f(x) = \sum_{j = 1}^M \beta_j \psi_j (x)$ then:

$$
\langle D \sum_{j = 1}^M \beta_j \psi_j, \phi \rangle 
= \sum_{j = 1}^M \beta_j \langle D \psi_j, \phi \rangle 
= \langle \epsilon, \phi \rangle
$$

Test only functions $\phi$ of the form $\sum_{j = 1}^M \beta_j \psi_j (x)$, equivalent to (showing it's equal for all of the basis shows it's equal for any linear combination of them) showing that:

$$
\sum_{j = 1}^M \beta_j \langle D \psi_j, \psi_i \rangle = \langle \epsilon, \psi_i \rangle \, \forall \psi_i
$$

This can be written in matrix form as $\mathbf{P} \mathbf{\beta} = \mathbf{e}$ where $P_{i, j}$ is $\langle D \psi_i, \psi_j \rangle$ and $e_j = \langle \epsilon, \psi_j \rangle$.
The random vector $\mathbf{e} \sim \mathcal{N}(\mathbf{0}, \mathbf{Q}_e ^{-1})$ where the precision has $(i, j)$th entry $\langle \psi_i, \psi_j \rangle$.
If $\mathbf{P} \mathbf{\beta} \sim \mathcal{N}(\mathbf{0},  \mathbf{Q}_e ^{-1})$ then $\mathbf{\beta} \sim \mathcal{N}(\mathbf{0},  \mathbf{Q})$ where $\mathbf{Q}$ is such that $\mathbf{Q}^{-1} = \mathbf{P} \mathbf{Q}_e^{-1} \mathbf{P}^\top \implies \mathbf{Q} = \mathbf{P}^\top \mathbf{Q}_e \mathbf{P}$ (does $\mathbf{P}^\top \mathbf{P} = I$?).

The steps are to use finite element method to compute $\mathbf{Q}$, then simulate $\tilde \beta$ from the Gaussian, then you have a sample $\tilde f$ from the stochastic process which is a solution to the SPDE.

Need to choose a mesh, grid, triangulation to do the FEM.
`R-INLA` uses regular grid in 1D, or constrained Delaunay triangulation in 2D to produce a mesh.
Piecewise linear basis functions (linear B-spline).

### Connection to basis-penalty smoothing

Assume $f$ to have the form $f(x) = \sum_{j = 1}^M \beta_j \psi_j$.
Observations give a log-likelihood of $\ell(\beta)$.
Point estimate $\hat \beta$ leads to function $\hat f$.
Smoothing penalty $J(\beta, \lambda)$ to give penalised log-likelihood $\ell_p(\beta, \lambda) = \ell(\beta) - J(\beta, \lambda)$.
Estimate both $\lambda$ and $\beta$ using REML.

Lots of choices for the penalty!

* Second derivative
* Thin-plate spline (total variation in the gradient of $f$ including the interaction between coordinates). Ends up being $J(\beta, \lambda) = \lambda \int (Df)^2 \text{d} x$ for some $D$ (can also write using inner product). Can also write this in matrix form

You can interpret penalised likelihood as just Bayesian.
This gets you to the following (improper) prior:

$$
p(\beta) \propto \text{exp}( - \lambda \beta^\top S \beta)
$$

This is just a Gaussian with precision $\lambda S$.

Approximate precision matrix for SPDE $Df = \epsilon$ is the same as $\lambda S$ computed using the smoothing penalty $\langle Df, Df, \rangle$.

1. Any SPDE of the form $Df = \epsilon$ can be interpreted instead as a smoothing penalty and vice-versa
2. Differences in the approaches are differences in implementation only

What smoothing penalty does the Matérn correspond to?
This one:

$$\tau \int (\kappa^2 f - \Delta f)^2 \text{d} x$$

Trade off between the value of $f$ and the second derivative in each direction.
Higher $\kappa$ means that $\Delta f$ can be higher without incurring penalty.
The other parameter $\tau$ controls the overall smoothness.
Similar question about how distinuishable these are as with lengthscale and noise (probably not very).

## A short introduction on how to fit a SPDE model with INLA (Krainski and Rue)

```{r message=FALSE}
library(INLA)
```

Generate fake data:

```{r}
n <- 200
coo <- matrix(runif(2 * n), n)

plot(coo[,1], coo[,2])
```

```{r}
s2u <- 0.5 # "sigma squared u"
k <- 10
r <- 2/k

R <- s2u * exp(-k * as.matrix(dist(coo)))

image(R)
```

Sample one multivariate Gaussian observation (note that `rnorm(n) %*% chol(R)` is equivalent to `MASS::mvrnorm`):

```{r}
u <- drop(rnorm(n) %*% chol(R))
u2 <- MASS::mvrnorm(1, rep(0, n), Sigma = R)

plot(u)
points(u2, add = TRUE, col = "red")
```

Add a covariate effect and noise (nugget) terms:

```{r}
x <- runif(n)
beta <- 1:2
s2e <- 0.2 # "sigma squared e"
lin.pred <- beta[1] + beta[2] * x + u # "linear predictor"
e <- rnorm(n, 0, sqrt(s2e)) # "noise"
y <- lin.pred + e
```

Mesh:

```{r}
mesh <- inla.mesh.2d(
  coo, 
  cutoff = r/10, 
  max.edge = c(r/4, r/2),
  offset = c(r/2, r)
)

plot(mesh, asp = 1)
points(coo, col = "red")
```

Question as to if all of the points lie at verticies?
It doesn't look like it, but most of them are.

Create the projection matrix `A`:

```{r}
A <- inla.spde.make.A(mesh = mesh, loc = coo)
class(A)
```

`A` is a $200 \times 1740$ sparse Matrix.
Where do the dimensions come from?
200 is the number of data points, what about the 1740?
I'd guess the number of mesh points but it's not exactly right:

```{r}
mesh$n
```

Set-up the SPDE:

```{r}
spde <- inla.spde2.pcmatern(
  mesh = mesh,
  alpha = 1.5,
  prior.range = c(0.2, 0.5), # P(range < 0.2) = 0.5
  prior.sigma = c(1, 0.5)    # P(sigma > 1) = 0.5
)
```

```{r}
stk.e <- inla.stack(
  tag = "est",
  data = list(y = y),
  A = list(1, A),
  effects = list(
    data.frame(b0 = 1, x = x),
    idx.u = 1:spde$n.spde)
)

pcprec <- list(
  hyper = list(theta = list(prior = "pc.prec", param = c(1, 0.1)))
)

mf <- y ~ 0 + b0 + x + f(idx.u, model = spde)

res <- inla(
  mf, 
  control.family = pcprec,
  data = inla.stack.data(stk.e),
  control.predictor = list(compute = TRUE, A = inla.stack.A(stk.e))
)

round(res$summary.fixed, 4)

pmd.s2e <- inla.tmarginal(
  function(x) sqrt(1/x), 
  res$marginals.hyperpar$`Precision for the Gaussian observations`
)

plot(pmd.s2e, type = "l", ylab = "Density", xlab = expression(sigma[e]))
```

## @krainski2018advanced notes

### First result

Consider a regular 2D lattice with number of sites tending to infinity then
$$
\mathbb{E}(u_{i, j} \, | \, u_{-i, j}) = \frac{1}{a}(u_{i-1, j} + u_{i + 1, j} + u_{i, j - 1} + u_{i, j + 1})
$$
and $\mathbb{V}(u_{i, j} \, | \, u_{-i, j}) = 1/a$ with $|a| > 4$.
For $\nu = 1$ and $\nu = 2$ the GMRF representation is a convolution of processes with... (missing diagram here but it's a Besag precision).
As the smoothness $\nu$ increases the precision matrix in GMRF representation becomes less sparse, the conditional distributions depend on a wider neighbourhood.

```{r}
q1 <- INLA:::inla.rw1(n = 5)
q1

crossprod(q1)

q2 <- INLA:::inla.rw2(n = 5)
q2
```

### Second result

$$
u(s) = \sum_{k = 1}^m \psi_k(s) w_k
$$
where $\psi_k$ are basis functions and $w_k$ are Gaussian weights, $m$ the number of points in the triangulation.
The distribution of weights determines the full distribution in the continuous domain.

Projector matrix $A$ works using barycentric coordinates of the point with respect to the coordinates of the triangle vertices.

### Toy example

```{r}
data(SPDEtoy)
str(SPDEtoy)

plot(x = SPDEtoy$s1, y = SPDEtoy$s2, cex = SPDEtoy$y / 10)

pl.dom <- cbind(c(0, 1, 1, 0.7, 0), c(0, 0, 0.7, 1, 1))
mesh5 <- inla.mesh.2d(loc.domain = pl.dom, max.e = c(0.092, 0.2))
plot(mesh5)

spde5 <- inla.spde2.pcmatern(
  mesh = mesh5, alpha = 2,     # Mesh and smoothness parameter
  prior.range = c(0.3, 0.5),   # P(practic.range < 0.3) = 0.5
  prior.sigma = c(10, 0.01)    # P(sigma > 10) = 0.01
) 
```

Projector matrix is sparse (maximum three elements in each row are non-zero, and each row sums to one).

```{r}
coords <- as.matrix(SPDEtoy[, 1:2])
A5 <- inla.spde.make.A(mesh5, loc = coords)
```

```{r}
dim(A5)

table(rowSums(A5 > 0)) # Three non-zero elements
table(rowSums(A5))     # Which add to one

table(colSums(A5) > 0) # Columns which correspond to triangles with no points inside can be dropped
# This is done automatically by inla.stack()
```

`inla.stack()` organises the data, covariates, indicies and projector matrices

```{r}
stk5 <- inla.stack(
  data = list(resp = SPDEtoy$y),
  A = list(A5, 1), 
  effects = list(i = 1:spde5$n.spde, beta0 = rep(1, nrow(SPDEtoy))),
  tag = 'est'
)
```

```{r}
dim(inla.stack.A(stk5)) # Number of columns in A5 which have non-zero entries, plus one column for the intercept
```

```{r}
res5 <- inla(
  resp ~ 0 + beta0 + f(i, model = spde5),
  data = inla.stack.data(stk5),
  control.predictor = list(A = inla.stack.A(stk5))
)

res5$summary.fixed
```

Now for predicting at new points.

```{r}
pts3 <- rbind(c(0.1, 0.1), c(0.5, 0.55), c(0.7, 0.9))
A5pts3 <- inla.spde.make.A(mesh5, loc = pts3)
dim(A5pts3)

jj3 <- which(colSums(A5pts3) > 0)
A5pts3[, jj3]
```

Interpolation of the posterior mean:

```{r}
drop(A5pts3 %*% res5$summary.random$i$mean)
```

Projection onto a grid:

```{r}
pgrid0 <- inla.mesh.projector(mesh5, xlim = 0:1, ylim = 0:1, dims = c(101, 101))

prd0.m <- inla.mesh.project(pgrid0,  res5$summary.random$i$mean)
prd0.s <- inla.mesh.project(pgrid0,  res5$summary.random$i$sd)

image(prd0.m)
image(prd0.s)
```

Prediction of new data:

```{r}
# Create a new stack
stk5.pmu <- inla.stack(
  data = list(resp = NA),
  A = list(A5pts3, 1), 
  effects = list(i = 1:spde5$n.spde, beta0 = rep(1, 3)), 
  tag = 'prd5.mu'
)

# Append it
stk5.full <- inla.stack(stk5, stk5.pmu)

r5pmu <- inla(
  resp ~ 0 + beta0 + f(i, model = spde5), 
  data = inla.stack.data(stk5.full), 
  control.mode = list(theta = res5$mode$theta, restart = FALSE), 
  control.predictor = list(A = inla.stack.A(stk5.full), compute = TRUE)
)
```

```{r}
indd3r <- inla.stack.index(stk5.full, 'prd5.mu')$data
indd3r

r5pmu$summary.fitted.values[indd3r, c(1:3, 5)]
```

## Bibliography {-}

### 2022-07-12 src/docs_paper/appendix.Rmd
Large upload

---
title: Appendix to "Models for spatial structure in small-area estimation"
subtitle: |
  Corresponding author: Adam Howes (`ath19@ic.ac.uk`)
output:
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    includes:
      in_header: preamble.tex
bibliography: citations.bib
---

<!-- So that the tables and figures in this appendix are prefixed by "A." -->
\setcounter{table}{0}  
\renewcommand{\thetable}{A.\arabic{table}} 
\setcounter{figure}{0} 
\renewcommand{\thefigure}{A.\arabic{figure}}

\newpage

# Implementation details for the Besag model \label{app:freni}

Here we briefly review three best practices for using the Besag model, scaling, singletons, and constraints, as recommended by @freni2018note:

1. \textit{Scaling}
The structure matrix $\mathbf{R}$ should be rescaled to have generalised variance, defined by the geometric mean of the diagonal elements of its generalised inverse
\begin{equation}
    \sigma^2_{\text{GV}}(\mathbf{R}) = \prod_{i = 1}^n (R^-_{ii})^{1/n} = \exp \left( \frac{1}{n} \sum_{i = 1}^n \log (R^-_{ii}) \right),
\end{equation}
equal to one, by replacing $\mathbf{R}$ with $\mathbf{R}^\star = \mathbf{R} / \sigma^2_{\text{GV}}(\mathbf{R})$.
As the diagonal elements $R^-_{ii}$ correspond to marginal variances, the generalised variance gives a measure of the average marginal variance.
However, this measure, introduced by @sorbye2014scaling, ignores off-diagonal entries and more broadly any measure of typical variance could be used.
Scaling mitigates the influence of the adjacency graph on the variance of $\bphi$.
Allowing the variance to be controlled by $\tau_\phi$ alone is important as it allows for consistent, interpretable prior selection.
When the adjacency graph is disconnected it is not appropriate to scale the structure matrix $\mathbf{R}$ uniformly since for a given precision $\tau_\phi$, local smoothing operates on each connected component independently.
As such, each connected component should be scaled independently to have generalised variance one giving $\mathbf{R}^\star_I = \mathbf{R}_I / \sigma^2_{\text{GV}}(\mathbf{R}_I)$ where $\mathbf{R}_I$ is the sub-matrix of the structure matrix corresponding to index set $I$.

2.
\textit{Singletons}
When one of the connected components is a single area, known either as a singleton or an island, the probability density $\exp \left( -\frac{\tau_\phi}{2} \sum_{i \sim j} (\phi_i - \phi_j)^2 \right)$ has no dependence on $\phi_i$.
This is equivalent to using an improper prior $p(\phi_i) \propto 1$ and can be avoided by setting each singleton to have independent Gaussian noise $p(\phi_i) \sim \mathcal{N}(0, 1)$.

3.
\textit{Constraints}
To avoid confounding of the spatial random effects with the intercept, it is recommended to place a sum-to-zero constraint on each non-singleton connected component.
In other words, for each $|I| > 1$ that $\sum_{i \in I} \phi_i = 0$.

# References {#references .unnumbered}

### 2022-07-12 src/docs_paper/paper.Rmd
Large upload

---
title: Models for spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    corresponding: ath19@ic.ac.uk
  # - name: Harrison Zhu
  #   affiliation: A
  # - name: Yidan Xu
  #   affiliation: C
  # - name: Samir Bhatt
  #   affiliation: D
  - name: Jeff W. Eaton
    affiliation: B
  - name: Seth R. Flaxman
    affiliation: E
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  # - code: C
  #   address: Department of Statistics, University of Michigan
  # - code: D
  #   address: Department of Public Health, University of Copenhagen
  - code: E
    address: Department of Computer Science, University of Oxford
abstract: |
  **Background:**

  **Methods:**

  **Results:**

  **Conclusions:**

author_summary: |
  Author summary

bibliography: citations.bib
output:
  rticles::elsevier_article:
    number_sections: true
    includes:
      in_header: preamble.tex
# plos.csl can be download and used locally
csl: http://www.zotero.org/styles/plos
---

# Introduction

Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
Gaussian Markov random fields (GMRFs) [@rue2005gaussian], which combine a Gaussian distribution with Markov conditional independence assumptions between areas, are the most common class of models used to specify spatial random effects.
Using a GMRF often reflects an assumption that observations made in areas close together are correlated, and more distant relationships may be ignored.
Arguably the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
This model requires minimal additional modelling choices and is widely implemented, including as a part of the `R-INLA` package [@rue2017bayesian], popular within spatial statistics.
Use of this model is widespread, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
However, for irregular geometries like the administrative divisions of a country, the assumptions made about space are unrealistic.
We hypothesised that incorporating a spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models.
We set out to test this hypothesis, and in doing so provide practical recommendations for modelling spatial structure.
<!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)? -->

Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
@duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model^[This conclusion was specific to the (often-analysed) Scottish lip cancer example, and like the majority of the literature, based on the deviance information criteria (DIC). The DIC is based on summarising the posterior distribution by a point estimate, and is recommended against by @vehtari2017practical.].
Another approach is to take into account that areal data is typically produced by aggregating higher resolution data, so as to be a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change] and dating to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially], `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages.
Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].

We also hypothesised that improvements to the Besag model may be transferable to improving convolution models.
Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models.
Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive.
These convolutions use a Besag model as the spatially structured component, though any choice of spatial random effect model is possible.
The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and more broadly both models are of substantive practical interest.

<!-- @bradley2015spatio develop a framework for spatio-temporal change of support (STCOS) modelling, and a review of its implementation in \textsc{R} is provided by @raim2021spatio. -->
<!-- Lee find that Leroux model of @leroux2000estimation achieves consistently good performance across a range of spatial correlation settings. -->
<!-- @riebler2016intuitive reach a similar conclusion about both the Leroux and BYM2 model. -->

The remainder of this paper is organised as follows.
Section \ref{sec:background} provides background on areal data and the Bayesian hierarchical modelling approach to small-area estimation.
In Section \ref{sec:ab}, we review developments in specifying spatial random effects based on adjacency, before presenting an alternative kernel-based approach in Section \ref{sec:ks}.
In Section \ref{sec:simulation} we compare models using simulated data, before applying them to mapping HIV prevalence in Section \ref{sec:hiv}.
Finally, in Section \ref{sec:conclusion} we discuss our conclusions and directions for future research.

# Background \label{sec:background}

## Areal and point spatial data

Let $\Sc \subset \mathbb{R}^2$ be the study region and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$.
Areal data are a type of spatial data where observations $y_i$ are associated to areas $A_i$.
Examples include the colour of a pixel, the minimum wage in a state, and the number of disease cases in a region.
Point data are another type of spatial data where instead observations $y(s)$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.
Often areal may be conceptualised as arising from aggregation of point data so that $y_i = \int_{A_i} y(s) \text{d}s$.
There are some exceptions, such as the minimum wage or disease control measures, which are determined at an administrative level and have a genuinely discrete spatial structure\footnote{For data of this sort, there is less clear justification for using spatial smoothing. Realistically, data relate to many covariates, some number of which are likely have discrete spatial structure and others of which continuous spatial structure, such that the result is often an intermediate. It may be possible to model this explicitly, though the result might look similar to what is achieved using e.g. BYM2.}.

If the areal data we observe are indeed an aggregation of point data, why not use the higher resolution point data in our models rather than areal data?
Explainations include that (a) the underlying point data are unobserved, (b) although point data are collected, they may not be accessible due to privacy constraints, administrative practicality or storage capacity, and (c) the point data are available, but we decide not to use them in the model.
In the final case (c), researchers may prefer to use area-level models because models for point data are typically more complex, require more data, including covariate data which may not exist, and have higher computational burden.
Choosing to model the underlying data generating mechanism as either areal or point is often a matter of pragmatism.

## Small-area estimation

Auto-correlation is an important property of most spatial processes.
Usually, outcomes for locations close together in space tend to be more similar than those far apart.
This fact is known as "Tobler's first law of geography" [@tobler1970computer] and, from a statistical point of view, is both a challenge and a benefit of working with spatial data.
Each observation provides less information than it would have had the samples been independent, making it more difficult to estimate global parameters.
However, spatial correlation can be used to improve local parameter estimates, particularly in parts of the study region where little to no information is available^[We use local parameters to refer to those parameters those indexed by a particular spatial (or spatio-temporal) location. These parameters corresponds to the latent field, as defined in Section \ref{sec:lgm}.].

The latter benefit is basis for the statistical task of small-area estimation, which aims to produce reliable local estimates where small sample sizes lead to noisy data [@pfeffermann2013new].
In a spatial setting, this is often in small geographic areas, though the phrase "small-area" is not restricted to geographic areas and may be interpreted more broadly to mean any area where data are insufficient to make accurate local parameter inferences^[For example, in the context of multilevel regression and post-stratification [@gelman1997poststratification] small-areas are generated by the intersection of demographic variables like age, gender and race, alongside geographic variables.].
For example, due to the cost of gathering samples, a survey may be designed to give reliable estimates at a national level but not at a district level.
Although direct estimators of local parameters are unbiased, when data are sparse the total error may be reduced by accepting some bias in exchange for reduced variance using so-called indirect estimators.
Smoothing approaches use information from similar units to "borrow strength" from one parameter to another, with determining precisely what is meant by "similar" a central challenge.
@wakefield2020small review design-based approaches and model-based approaches to spatial analysis of health-indicators, including both area and unit-level analysis.

## Bayesian model-based approach and latent Gaussian models \label{sec:lgm}

Bayesian hierarchical models are an attractive framework for small-area estimation [@rao2015small].
To model areal observations $\y$ corresponding to the areas $\{A_i\}_{i = 1}^n$ of a study region $\Sc$, a three-layer structure [@berliner1996hierarchical; @cressie2015statistics] given by
\begin{alignat}{2}
&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, g^{-1}(x_i), \btheta), \quad i = 1, \ldots, n, \label{eq:data} \\
&\text{(Latent field)}  &        \x &\sim p(\x \, | \, \btheta), \label{eq:process} \\
&\text{(Parameters)}    & \qquad \btheta &\sim p(\btheta), \label{eq:parameters}
\end{alignat}
may be used, where $\x$ is the $n$-dimensional latent field, $\btheta$ are the $m$-dimensional\footnote{Ordinarily $m$ is much smaller than $n$ as few parameters are required in order to control the properties of the latent field and likelihood.} parameters and $g$ is a link function.
Spatial structure is modelled at the level of the latent field, allowing the observation model, often either binomial or Poisson, to be conditionally independent and identically distributed $p(\y \, | \, \x, \btheta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \btheta)$.
As small-area estimation models are usually descriptive rather than mechanistic, it is natural [@best2005comparison] to use a Gaussian distribution $\x \sim \mathcal{N}(\bm{\mu}, \mathbf{\Sigma})$ for the latent field^[The Gaussian distribution is the maximal entropy distribution which retains finite mean and variance, as well as being computationally convenient.].
Three-layer models with Gaussian latent fields comprise a wide array of popular models, including generalised linear mixed models (GLMMs), and have been collectively studied under the title latent Gaussian models (LGMs) [@rue2009approximate].

## Spatial random effects

Decomposing the latent field in a LGM gives the sum of a constant mean $\bm{\mu}$, known also as the fixed effects, and (spatial) random effects $\bphi \sim \mathcal{N}(\mathbf{0}, \mathbf{\Sigma})$.
The mean may be modelled using a linear predictor $\mu_i = \beta_0 + \z_i^\top \bm{\beta}$, where $(\beta_0, \bm{\beta})$ are regression parameters corresponding to areal covariates^[To denote length $p$ vectors of areal and point covariates, we use $\z_i = (z_{i,1}, \ldots, z_{i,p})^\top$ and $\z(s) = (z_1(s), \ldots, z_p(s))^\top$ respectively.} $(1, \z_i)^\top$].
Our focus is on the spatial random effects $\bphi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
Specifying $\bphi$ amounts to specifying the entries of a precision or covariance matrix, which we discuss below.

If no spatial structure remains after inclusion of covariates in the mean, independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ should be used, where $\tau_\phi$ is a shared precision.
Exploratory data analysis techniques such as Moran's $I$ coefficient [@cliff1981spatial], visual inspection, may be used to determine if there remains any spatial auto-correlation in the data.
<!-- TODO: Add calculation of Moran's $I$ to my HIV survey data-sets. -->

# Spatial random effects based on adjacency \label{sec:ab}

\begin{figure}
\centering
\includegraphics{depends/geometry-graph-zwe.pdf}
\caption{A map of the districts of Zimbabwe together with the corresponding adjacency graph structure $\mathcal{G}$.}
\label{fig:zi}
\end{figure}

## Besag \label{sec:besag}

One way to encode spatial structure between areas is by using a symmetric relation, where $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring, as often defined by a shared border^[Any choice of adjacency definition is possible. For example, one might chose to allow inclusion of second degree neighbours resulting in so-called higher-order models [@paciorek2013spatial].].
The Besag model [@besag1991bayesian] is a type of improper conditional auto-regressive (ICAR) model where the full conditional distribution of the $i$th spatial random effect is given by
\begin{equation}
    \phi_i \, | \, \bphi_{-i} \sim \N \left(\frac{1}{n_{\delta i}} \sum_{j: j \sim i} \phi_j, \frac{1}{n_{\delta i}\tau_\phi}\right), \label{eq:besag}
\end{equation}
where $\delta i$ is the set of neighbours of $A_i$ with cardinality $n_{\delta i} = |\delta i|$ and $\bphi_{-i} = (\phi_1, \ldots, \phi_{i - 1}, \phi_{i + 1}, \ldots, \phi_n)^\top$ is the vector of spatial random effects with the $i$th entry removed.
The conditional mean of the random effect $\phi_i$ is the average of its neighbours $\{\phi_j\}_{j \sim i}$ and the precision $n_{\delta i}\tau_\phi$ is proportional to the number of neighbours $n_{\delta i}$.
By Brook's lemma \cite[Chapter 2]{rue2005gaussian} the set of full conditionals of the Besag model are equivalent to the Gaussian Markov random field (GMRF) given by
\begin{equation}
    \bphi \sim \mathcal{N}(\mathbf{0}, \tau_\phi^{-1} \mathbf{R}^{-}), \label{eq:gmrf}
\end{equation}
where $\mathbf{R}^{-}$ is the generalised inverse of the rank-deficient structure matrix\footnote{So-called because it defines the \textit{structure} of the precision matrix $\mathbf{Q} = \tau_\phi \mathbf{R}$.} $\mathbf{R}$ with entries
\begin{equation}
    R_{ij} =
    \begin{cases}
        n_{\delta i}, & i = j \\
        -1, & i \sim j \\
        0, & \text{otherwise.}
    \end{cases}
\end{equation}
The Markov property arises due to the conditional independence structure $p(\phi_i \, | \, \bphi_{-i}) = p(\phi_i \, | \, \bphi_{\delta i})$ whereby each area only depends on its neighbours.
This is reflected in the sparsity of $\mathbf{R}$ whereby $\phi_i \perp \phi_j \, | \, \phi_{-ij}$ if and only if $R_{ij} = 0$.
The structure matrix $\mathbf{R}$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
Figure \ref{fig:zi} shows the adjacency graph for the districts of Zimbabwe alongside the geometry.
Shortly, in Section \ref{sec:concerns} we discuss the appropriateness of using the graph rather than the full geometry.

Rewriting Equation \ref{eq:gmrf}, the probability density function of $\bphi$ is
\begin{equation}
    p(\bphi)
    \propto \exp \left( -\frac{\tau_\phi}{2} \bphi^\top \mathbf{R} \bphi \right)
    \propto \exp \left( -\frac{\tau_\phi}{2} \sum_{i \sim j} (\phi_i - \phi_j)^2 \right). \label{eq:pdfphi}
\end{equation}
This density is a function of the pairwise differences $\phi_i - \phi_j$ and so is invariant to the addition of a constant $c$ to each entry $p(\bphi) = p(\bphi + c\mathbf{1})$, leading to an improper uniform distribution on the average of the $\phi_i$.
If $\mathcal{G}$ is connected\footnote{In that by traversing the edges, any vertex can be reached from any other vertex.} then there is only one impropriety in the model and $\text{rank}(\mathbf{R}) = n - 1$, while if $\mathcal{G}$ is disconnected, and composed of $n_c \geq 2$ connected components with index sets $I_1, \ldots, I_{n_c}$, then the corresponding structure matrix $\mathbf{R}$ has rank $n - n_c$ and the density is invariant to the addition of a constant to each of the connected components $p(\bphi_{I}) = p(\bphi_{I} + c \mathbf{1})$ where $I = I_1, \ldots, I_{n_c}$.

<!-- ## Concerns about the Besag model's representation of space \label{sec:concerns} -->

<!-- \begin{figure} -->
<!-- \centering -->
<!-- \includegraphics{depends/maup1.pdf} -->
<!-- \caption{Each of these four geometries corresponds to the same adjacency graph.} -->
<!-- \label{fig:tikz1} -->
<!-- \end{figure} -->

<!-- \begin{figure} -->
<!-- \centering -->
<!-- \includegraphics{depends/maup2.pdf} -->
<!-- \caption{A sequence of geometries where the number of neighbours of area one grows by one at each iteration.} -->
<!-- \label{fig:tikz2} -->
<!-- \end{figure} -->

<!-- \begin{figure} -->
<!-- \centering -->
<!-- \includegraphics{depends/maup3.pdf} -->
<!-- \caption{Each of the shaded areas are split into two moving from geometry (i) to (ii).} -->
<!-- \label{fig:tikz3} -->
<!-- \end{figure} -->

<!-- The Besag model was originally proposed for use in image analysis, where areas correspond to pixels arranged in a regular lattice structure. -->
<!-- Since then, it has seen wider use, including in situations where the spatial structure is less regular. -->
<!-- This leads to a number of concerns about the model's applicability, which we highlight below. -->
<!-- Closely linked are the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups. -->
<!-- <!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. --> -->

<!-- ### Adjacency compression -->

<!-- Summarising all spatial information with an adjacency graph represents a loss of information. -->
<!-- Many geometries share the same adjacency graph, and therefore the same probability model, as illustrated by Figure \ref{fig:tikz1}. -->
<!-- This fact is not concerning in itself, though it prompts consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.\footnote{The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be. In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.} -->
<!-- It is intuitive that the more regular the spatial structure the less information is lost in compression to an adjacency graph. -->
<!-- In image analysis, very little spatial information is lost in compression a lattice structure to an adjacency graph. -->
<!-- On the other hand, the regions of a country, determined by political and geographic forces, tend to display greater irregularity. -->
<!-- As such, the appropriateness of adjacency compression varies by the type of geometry common to the application setting. -->

<!-- ### Mean structure -->

<!-- A feature specific to the Besag model is that all adjacent areas count equally. -->
<!-- This assumption is unsatisfying: for most geometries, we expect some areas to be more highly correlated to some neighbours than to others. -->
<!-- Figure \ref{fig:tikz1} illustrates a number of heuristic features for neighbour importance, including length of shared border in geometry (iii), and the proximity of centers of mass in geometry (iv). -->
<!-- This may be remedied using more general weighted ICAR models, as we outline in Section \ref{sec:icar}. -->

<!-- ### Variance structure -->

<!-- A striking feature of Equation~\ref{eq:besag} is that the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.  -->
<!-- It follow that as $n_{\delta i} \to \infty$ then $\text{Var}(\phi_i) \to 0$. -->
<!-- This is illustrated by Figure \ref{fig:tikz2} where the area on the right is repeatedly divided such that its number of neighbours increases by one at each step. -->
<!-- This property is a consequence of averaging the conditional mean over a greater number of areas, which, in certain situations, can correspond to a greater amount of information. -->
<!-- However, if the amount of information in the shaded area remains fixed, it is inappropriate that $\text{Var}(\phi_1)$ should tend to zero just as a result of drawing additional, arbitrary, boundaries. -->
<!-- In the image analysis setting this modelling assumption is reasonable: each pixel represents a fixed amount of information and a higher pixel density represents a greater amount of information. -->
<!-- On the other hand, in public health and epidemiology, drawing boundaries to create additional regional areas cannot be expected to correspond to a greater amount of information and this assumption is not appropriate. -->

<!-- Suppose we fit a Besag model upon identical data using each of the two geometries in Figure \ref{fig:tikz3}. -->
<!-- If the spatial variation is relatively smooth, dividing the shaded areas into two will result in a lower estimated variance $\sigma^2_\phi$ in geometry (ii) as compared with geometry (i) because there will appear to be less variation between neighbouring areas. -->
<!-- This problem does not only apply locally: since the effect of $\sigma^2_\phi$ applies everywhere, the smoothing will change even in unaltered parts of the study region. -->

<!-- ## Weighted ICAR \label{sec:icar} -->

<!-- The Besag model is a special case of a more general class of (zero-mean) ICAR models in which each neighbour may not be weighted equally. -->
<!-- These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\bm{\kappa} = (\kappa_1, \ldots, \kappa_n)^\top$ such that $\phi_i \, | \, \bphi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$. -->
<!-- The structure matrix $\mathbf{R}$ corresponding to the above full conditionals is given by $\mathbf{R} = \mathbf{D}_\kappa(\mathbf{I} - \mathbf{B})$, where $\mathbf{B}$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $\mathbf{D}_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$. -->
<!-- As the structure matrix is symmetric we must have the condition $-b_{ij}\kappa_i = -b_{ji}\kappa_j$. -->
<!-- To meet this condition, it is often simpler to use an unscaled weights matrix $\mathbf{W} = \mathbf{D}_\kappa \mathbf{B}$ such that $\mathbf{R} = \mathbf{D}_\kappa - \mathbf{W}$. -->
<!-- For example, in the Besag model $\mathbf{W}$ corresponds to the adjacency matrix. -->
<!-- The scaled weights can then be recovered by $b_{ij} = w_{ij} / \kappa_i$ where $\kappa_i = \sum_{k: k \sim i} w_{ik}$. -->
<!-- A thorough discussion of methods for specifying $\mathbf{W}$ is provided by @duncan2017spatial. -->
<!-- Much of the work in this area focuses on the case where the geometry is a lattice. -->

<!-- ## BYM2 \label{sec:bym2} -->

<!-- Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model. -->
<!-- The Besag-York-Molli\'{e} (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\bphi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$. -->
<!-- The resulting distribution is -->
<!-- \begin{equation} -->
<!--     \bphi \sim \N(\mathbf{0}, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} \mathbf{R}^{-}) \label{eq:bym}. -->
<!-- \end{equation} -->
<!-- Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$. -->
<!-- However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise. -->
<!-- In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale. -->
<!-- Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation. -->
<!-- This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal. -->
<!-- Building on the models of @leroux2000estimation and @dean2001detecting which tackle this the identifiability problem, but do not scale the spatially structured noise, @simpson2017penalising propose a reparameterisation $(\tau_v, \tau_u) \mapsto (\tau_\phi, \pi)$ of the BYM model known as the BYM2 model and given by -->
<!-- \begin{align} -->
<!-- \bphi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2} -->
<!-- \end{align} -->
<!-- where $\tau_\phi$ is the marginal precision of $\bphi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $\bu$ with precision matrix given by the scaled structure matrix $\mathbf{R}^\star$ (see~\ref{app:freni}). -->
<!-- When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model. -->
<!-- To borrow an analogy, the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately [@rue2020comment]. -->

<!-- \section{Areal kernel spatial random effects} \label{sec:ks} -->

<!-- % Cast the ICAR models as an example of a kernel? -->

<!-- \subsection{Areal kernels} -->

<!-- Section \ref{sec:ab} reviewed ways to construct the precision matrix of the spatial random effects using an adjacency relation. -->
<!-- Another approach is to define the covariance matrix directly using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$, where $\mathcal{P}$ denotes the power set such that $\mathcal{P}(\Sc)$ is the space of subsets of the study region, which gives a measure of similarity between two areas. -->
<!-- If $K$ is positive semi-definite, then we define areal kernel spatial random effects by -->
<!-- \begin{equation} -->
<!--     \bphi \sim \mathcal{N}(\mathbf{0}, \frac{1}{\tau_\phi} \mathbf{K}), \label{eq:arealkernel} -->
<!-- \end{equation} -->
<!-- where the $n \times n$ Gram matrix $\mathbf{K}$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix. -->
<!-- We keep $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices. -->

<!-- Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$. -->
<!-- To construct $K$ we use a simple method, namely averaging the kernel $k$ computed on some collection of points from within each area. -->
<!-- In the multiple instance learning literature, $K$ is described as a type of kernels on sets [@gartner2002multi]. -->
<!-- Although it may be possible to learn the method for combining kernels [@gonen2011multiple] from data, we do believe that this is feasible in the small-area estimation setting, and instead hope to construct areal kernels which represent our prior beliefs about the spatial process. -->

<!-- \subsection{Centroid kernel} -->

<!-- The simplest approach is to use a single point, which one might hope would be representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude. -->
<!-- This results in the centroid kernel -->
<!-- \begin{equation} -->
<!--     K(A_i, A_j) = k(c_i, c_j). -->
<!-- \end{equation} -->
<!-- % More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
<!-- Usage of the centroid kernel includes @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both using the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$. -->
<!-- The former fix the length-scale $l$ using a semi-variogram and the later place a prior on the inverse length-scale. -->
<!-- In their simulation experiment, \citet[Section 3]{best2005comparison} used data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas (though it is unsurprising that a stationary covariance structure would struggle to recover non-stationary spatial structure). -->
<!-- More recently, @teh2021efficient used a centroid kernel as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19. -->

<!-- \subsection{Integrated kernel} \label{sec:ik} -->

<!-- \begin{figure} -->
<!-- \centering -->
<!-- \includegraphics[width=\textwidth]{depends/integration-method.pdf} -->
<!-- \caption{Three ways to choose integration points with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample} [@pebesma2018sf].} -->
<!-- \label{fig:41} -->
<!-- \end{figure} -->

<!-- Rather than attempting to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest. -->
<!-- This results in the integrated kernel -->
<!-- \begin{equation} -->
<!-- K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s', \label{eq:ik1} -->
<!-- \end{equation} -->
<!-- which is equivalent to the covariance structure obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition. -->
<!-- Vaious authors [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi] in the machine learning literature have recently contributed to the development of these aggregated Gaussian process models. -->
<!-- % For example, @law2018variational develop a variational Bayesian inference method for aggregated observation models and @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region. -->
<!-- Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out. -->

<!-- Considering the aggregation of point data is particularly natural within the log-Gaussian Cox Process framework [@diggle2013spatial]. -->
<!-- A Cox process is an inhomogeneous Poisson process with a continuous stochastic intensity function $\{ x(s), s \in \Sc \}$ such that conditional on the realisation of $x(s)$ the number of points in any area $A_i$ follows a Poisson distribution. -->
<!-- The rate parameter of this Poisson distribution is explicitly aggregated as follows -->
<!-- \begin{equation} -->
<!-- y_i \, | \, x(s) \sim \text{Poisson} \left(\int_{s \in A_i} x(s) \text{d}s \right). -->
<!-- \end{equation} -->
<!-- In a LGCP the log intensity $\log x(s) = \eta(s)$ is modelled using a Gaussian process prior $\eta(s) \sim \mathcal{GP}(\mu(s), k(s, s'))$. -->
<!-- @johnson2019spatially obtain Equation~\ref{eq:ik2} by considering a discrete Poisson log-linear mixed model approximation to a continuous LGCP, whereby $\eta(s)$ is approximated by a piecewise constant $\eta_i = \mu_i + \phi_i$ in each area $A_i$.\footnote{@johnson2019spatially use points generated from a class of inhibition processes which combine sequential inhibition with rejection sampling to evaluate the integral.} -->
<!-- The $i$th discrete spatial random effect is then $\phi_i = \int_{A_i} w_i(s) \phi(s) \text{d}s$, -->
<!-- with covariance structure -->
<!-- \begin{equation} -->
<!-- \text{Cov} \left( \int_{A_i} w_i(s) \phi(s) \text{d}s, \int_{A_j} w_j(s') \phi(s') \text{d}s' \right) -->
<!-- = \int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d}s\text{d}s', -->
<!-- \end{equation} -->
<!-- corresponding to an areal integrated kernel with a logarithmic link function and Poisson likelihood. -->

<!-- One benefit of integration is that additional information accounting for heterogeneity over $A_i$ may be incorperated into the model if it is available. -->
<!-- This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure. -->
<!-- The weighted integrated kernel is given by -->
<!-- \begin{equation} -->
<!-- K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2} -->
<!-- \end{equation} -->
<!-- This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, and utilising a high resolution measure of population density as the weighting distribution would account for this. -->

<!-- If areal kernel spatial random effects are created using the integrated kernel, then aggregation occurs at the level of the latent field in the three stage model rather than at the level of the data. -->
<!-- This corresponds to a generative model of the form $y_i \sim p(y_i \, | \,  g^{-1}(x_i))$, with $x_i = |A_i|^{-1}\int_{A_i} x(s) \text{d} s$. -->
<!-- If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data. -->
<!-- On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model mismatches the proposed data generating process $y_i = |A_i|^{-1}\int_{A_i} y(s) \text{d}s$.  -->

<!-- Most of the time we do not expect to be able to calculate the integral in Equation~\ref{eq:ik2} analytically. -->
<!-- Instead, given $n$ collections of $L_i$ samples $\{s^{(i)}_l \}_{l = 1}^{L_i} \sim \mathcal{U}(A_i)$ drawn uniformly from each area then the integral may be approximated using Monte Carlo integration by the double sum -->
<!-- \begin{equation} -->
<!-- K(A_i, A_j) \approx \frac{1}{L_i L_j} \sum_{l = 1}^{L_i} \sum_{m = 1}^{L_j}  -->
<!-- w_i \left( s^{(i)}_l \right) w_j \left( s^{(j)}_m \right)  -->
<!-- k \left( s^{(i)}_l, s^{(j)}_m \right). \label{eq:mcintegratedkernel} -->
<!-- \end{equation} -->
<!-- Equivalently, samples drawn from $W_i$ may be used unweighted by $w_i(s)$. -->
<!-- Integration points may also be selected by some deterministic process, and used in Equation~\ref{eq:mcintegratedkernel} to give a numerical integration estimate of the kernel. -->
<!-- As with learning the way in which the kernels should be combined, in more data rich settings it may be possible to learn the integration points [@campbell2019automated]. -->
<!-- Taking any of these approaches requires $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $\mathbf{K}$. -->
<!-- This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or integration points it is feasible to use. -->

<!-- Another closely linked approach is disaggregation regression (also known as downscaling or interpolation) which, rather than using the aggregate nature of areal observations as just a route towards better area-level estimates as is our focus, aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation]. -->
<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates.  -->

<!-- <!-- Disaggregation is challenging without auxiliary information. --> -->
<!-- <!-- Kelsall and Wakefield They use the cubic correlation function and, to reduce the number of Gram matrix constructions and inversions, learn a range parameter by placing a discrete prior. --> -->
<!-- <!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. --> -->
<!-- <!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. --> -->
<!-- <!-- Explain under what circumstances we believe this approximation to be acceptable. --> -->

<!-- \clearpage -->

<!-- # Simulation study \label{sec:simulation} -->

<!-- We tested the ability of inferential models with different spatial random effect specifications to accurately recover small-area quantities. -->
<!-- The data and modelling choices were designed with a spatial epidemiology setting in mind, and are similar to those used in Section \ref{sec:hiv} where we considered models for HIV prevalence using data from national household surveys in sub-Saharan Africa. -->
<!-- The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`. -->
<!-- We used the `orderly` package [@orderly] for reproducible research. -->

<!-- ## Synthetic data-sets -->

<!-- \begin{table}[] -->
<!-- \small -->
<!-- \begin{tabularx}{\textwidth}{@{}lX@{}} -->
<!-- \toprule -->
<!-- Simulation model & Additional details \\  -->
<!--  \midrule -->
<!-- \hypertarget{S1}{S1}. \textit{IID} & $\bphi \sim \mathcal{N}(0, I_n)$ \\ -->
<!-- \hypertarget{S2}{S2}. \textit{Besag} & $\bphi \sim \mathcal{N}(0, (\mathbf{R}^\star)^{-})$ as described in Section \ref{sec:besag} where $\mathbf{R}^\star$ is the scaled structure matrix \\  -->
<!-- \hypertarget{S3}{S3}. \textit{IK} & $\bphi \sim \mathcal{N}(0, \mathbf{K}^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation~\ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area. -->
<!-- Like the Besag model, we scaled the Gram matrix to have generalised variance one \\  -->
<!-- \bottomrule -->
<!-- \end{tabularx} -->
<!-- \caption{The three simulation models from which we generate $\bphi$.} -->
<!-- \label{tab:simulationmodels} -->
<!-- \small -->
<!-- \end{table} -->

<!-- To study the robustness of the inferential models to spatial misspecification, we simulated synthetic data-sets from three known data generating processes. -->
<!-- The spatial random effects $\bphi$ were generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}). -->
<!-- After obtaining spatial random effects $\bphi$ from each simulation model, we generated synthetic data $\y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via -->
<!-- \begin{equation} -->
<!-- \log \left( \frac{\rho_i}{1 - \rho_i} \right) = x_i = \beta_0 + \phi_i, \quad i = 1, \ldots, n. -->
<!-- \end{equation} -->
<!-- We fixed the sample sizes to $m_i = 25$ for all $i$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$. -->
<!-- We used seven different geometries: the four vignette geometries in Figure \ref{fig:tikz1}, all of which have the same adjacency graph, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of C\^{o}te d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}). -->
<!-- These more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test the hypothesis that model performance depends on geometry regularity. -->
<!-- For each combination of spatial random effect and geometry we replicated the simulation process above 200 times to generate a total of 1400 synthetic data-sets. -->

<!-- ## Inferential models -->

<!-- We fit eight inferential models to each simulated data-set. -->
<!-- Each inferential model shared most of its structure with the data generating process, only differing in the spatial random effect specification, which varied across the eight models (Table \ref{tab:inferentialmodels}), allowing us to test sensitivity to misspecification of the spatial structure. -->

<!-- \begin{table} -->
<!-- \small -->
<!-- \begin{tabularx}{\textwidth}{llX} -->
<!-- \toprule -->
<!-- Inferential model & Algorithm & Additional details \\  -->
<!--  \midrule -->
<!-- \hypertarget{I1}{I1}. \textit{Constant} & INLA &  -->
<!-- No spatial random effects \label{row:I1} \\ -->
<!-- \hypertarget{I2}{I2}. \textit{IID} & INLA &  -->
<!-- $\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} I_n)$\\ -->
<!-- \hypertarget{I3}{I3}. \textit{Besag} & INLA &  -->
<!-- $\bphi \sim \mathcal{N}(0, (\tau_\phi \mathbf{R}^\star)^{-})$ following additional recommendations in~\ref{app:freni}  \\  -->
<!-- \hypertarget{I4}{I4}. \textit{BYM2} & INLA &  -->
<!-- $\bphi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\ -->
<!-- \hypertarget{I5}{I5}. \textit{FCK} & INLA & -->
<!-- $\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{K})$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\  -->
<!-- \hypertarget{I6}{I6}. \textit{CK} & NUTS & As in model~\hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\  -->
<!-- \hypertarget{I7}{I7}. \textit{FIK} & INLA & $\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{K})$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\ -->
<!-- \hypertarget{I8}{I8}. \textit{IK} & NUTS & As in model~\hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\  -->
<!-- \bottomrule -->
<!-- \end{tabularx} -->
<!-- \caption{Each inferential model is implemented as a part of the \textsf{R} package \texttt{bsae} available from \url{github.com/athowes/bsae}.} -->
<!-- \label{tab:inferentialmodels} -->
<!-- \small -->
<!-- \end{table} -->

<!-- ### Priors -->

<!-- For the precision, we placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$. -->
<!-- The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one, neither of which we assumed to reflect reality. -->
<!-- A weakly informative $\mathcal{N}(-2, 1)$ prior was placed on $\beta_0$, setting most of the prior probability density for $\rho_i$ within a range $[0, 0.25]$ typical for a disease prevalence. -->

<!-- ### Kernel model details -->

<!-- Following @stein1999interpolation, for the models~\hyperlink{I5}{I5}-\hyperlink{I8}{I8} we used the Mat\'{e}rn kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by -->
<!-- \begin{equation} -->
<!-- k(s, s') =  -->
<!-- \frac{1}{2^{\nu - 1}\Gamma(\nu)} \left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right)^\nu -->
<!-- K_\nu\left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right), \label{eq:matern} -->
<!-- \end{equation} -->
<!-- where $K_\nu$ is the modified Bessel function of the second kind, $|s - s'|$ is the Euclidean distance between $s$ and $s'$, $\nu$ is the smoothness hyperparameter and $l$ is the length-scale hyperparameter on the latitude-longitude scale. -->
<!-- We fixed the smoothness parameter, which is typically unidentifiable from data, to be $3/2$, matching that used in simulation model~\hyperlink{S3}{S3}, and convenient in that it simplifies Equation~\ref{eq:matern} to be the product of a polynomial and an exponential -->
<!-- \begin{equation} -->
<!-- k(s, s') = \left(1 + \sqrt{3} \lvert s - s' \rvert / l \right) \exp \left(- \sqrt{3} \lvert s - s' \rvert / l \right). -->
<!-- \end{equation} -->
<!-- Taking a similar approach to @best1999bayesian, in models~\hyperlink{I5}{I5} and \hyperlink{I7}{I7} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data. -->
<!-- In models~\hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust]. -->
<!-- For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51} (though the true number of integration points may be 1-2 above or below 10, as `sf::st_sample` with `type = "hexagonal"` is not guaranteed to be exact). -->
<!-- We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}. -->
<!-- For Model~\hyperlink{I6}{I6} although it is feasible to use a much larger number of integration points, since the Gram matrix is only computed once, we chose not to for comparability. -->

<!-- ## Inference algorithms -->

<!-- In models~\hyperlink{I1}{I1}-\hyperlink{I5}{I5} choosing a Gaussian distribution for $\beta_0$ ensures that the latent random variables are Gaussian and the model is a LGM. -->
<!-- This facilitated use of the integrated nested Laplace approximation (INLA) algorithm [@rue2009approximate] for inference, as implemented by the \texttt{R-INLA} package. -->
<!-- INLA is a deterministic method for approximate Bayesian inference based upon a combination of numerical integration and the Laplace approximation. -->
<!-- INLA is significantly faster than MCMC and has been shown to have comparable accuracy for LGMs in realistic, pre-asymptotic settings. -->
<!-- <!-- Although there are particular conditions under which INLA is not accurate [@rue2017bayesian] they do not apply here, as we illustrate in Appendix \ref{app:inlamcmc} where we compare INLA to MCMC using a large number of iterations. --> -->

<!-- ## Model assessment -->

<!-- We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, subject to calibration. -->
<!-- In the following two sections we describe our approach to model assessment. -->
<!-- % Analogous approaches may be used to check recovery of the intercept parameter $\beta_0$, precision parameter $\tau_\phi$, and, when relevant, the length-scale $l$. -->

<!-- ### Continuous ranked probability score -->

<!-- The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts. -->
<!-- For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, \y)$ and $\omega_i$ be the true value. -->
<!-- Writing $F$ as the cumulative distribution corresponding to $f$ then -->
<!-- \begin{equation} -->
<!-- \text{CRPS}(f, \omega_i) = \int_{-\infty}^{\infty} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i = \int_{0}^{1} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i, \label{eq:crps} -->
<!-- \end{equation} -->
<!-- where $\mathbbm{1}$ denotes the indicator function and the second equality follows from $0 \leq \rho_i \leq 1$. -->
<!-- The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$. -->
<!-- We use the CRPS because it is a strictly proper scoring rule [@gneiting2007strictly]. -->
<!-- This ensures that the (uniquely) best approach to minimising the score is to truthfully report probabilistic beliefs, which here correspond to the posterior distribution. -->
<!-- The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, but in line with the findings of @kruger2020predictive, using kernel density estimation we found it difficult to compute satisfactorily. -->
<!-- In contrast, given a collection of Monte Carlo samples $\{\rho_i^s\}_{s = 1}^S \sim f$, Equation~\ref{eq:crps} may easily be approximated \citep[Equation 37]{sigrist2015stochastic} by -->
<!-- \begin{equation} -->
<!-- \text{CRPS}(f, \omega_i) \approx \frac{1}{S} \sum_{s = 1}^S | \rho_i^s - \omega_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | \rho_i^s - \rho_i^l |. \label{eq:crpsapprox} -->
<!-- \end{equation} -->
<!-- We obtained these samples using the function \texttt{INLA::inla.posterior.sample} for models fit with \texttt{R-INLA} and directly from the Markov chain using \texttt{rstan::extract} for models fit with Stan. -->

<!-- \subsubsection{Coverage} -->

<!-- To assess the coverage of our estimates, we considered probability integral transform (PIT) histograms and empirical cumulative distribution function (ECDF) difference plots [@sailynoja2021graphical]. -->
<!-- We estimated the quantile $q_i = F(\omega_i)$ of the true value within the marginal posterior using numerical integration. -->
<!-- For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that for any given nominal significance level $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$, should be $1 - \alpha$. -->

<!-- ## Funding -->

<!-- AH, HZ were supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).  -->
<!-- AH, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661). -->
<!-- YX was supported by the Mary Lister McCammon Summer Research Fellowship 2019. -->
<!-- HZ was supported by Cervest Limited. -->
<!-- JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664).  -->
<!-- SF was supported by the EPSRC (EP/V002910/1). -->
<!-- This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth \& Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union. -->

<!-- For the purpose of Open Access, the authors have applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission. -->

<!-- ## Acknowledgements -->

<!-- AH thanks Robert Ashton for help with \textsf{R} package development, and Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for providing feedback on earlier drafts of this manuscript. -->

<!-- ## Disclaimer -->

<!-- The findings and conclusions in this manuscript are those of the authors and do not necessarily represent the official position of the funding agencies. -->

<!-- \clearpage -->

<!-- # References {#references .unnumbered} -->

### 2022-07-12 src/docs_paper/preamble.tex
Large upload

\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage{bm}
\RequirePackage{enumerate}
\RequirePackage{enumitem}
\RequirePackage{tabularx}
\RequirePackage{adjustbox}
\RequirePackage{tikz}
\RequirePackage{bbm}
\RequirePackage{lscape}
\RequirePackage{booktabs}
\RequirePackage{longtable}
\RequirePackage{lscape}
\RequirePackage{pifont}
\usepackage{footnote}

\newcommand{\Sc}{\mathcal{S}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\bb}{\mathbf{b}}
\newcommand{\brho}{\bm{\rho}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bpsi}{\bm{\psi}}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

### 2022-07-12 src/1_process_metric-tables/script.R

@@ -1,6 +1,6 @@
 #' Uncomment and run the two line below to resume development of this script
-# orderly::orderly_develop_start("process_metric-tables")
-# setwd("src/process_metric-tables")
+# orderly::orderly_develop_start("1_process_metric-tables")
+# setwd("src/1_process_metric-tables")
 
 df_rho <- readRDS("depends/df_rho.rds")
 df_intercept <- readRDS("depends/df_intercept.rds")

### 2022-07-12 src/docs_paper/script.R

@@ -1,0 +1,5 @@
+# orderly::orderly_develop_start("docs_paper")
+# setwd("src/docs_paper")
+
+rmarkdown::render("paper.Rmd")
+rmarkdown::render("appendix.Rmd")

### 2022-07-13 src/docs_paper/appendix.Rmd
Work on manuscript and simulation study

@@
 1. \textit{Scaling}
-The structure matrix $\mathbf{R}$ should be rescaled to have generalised variance, defined by the geometric mean of the diagonal elements of its generalised inverse
+The structure matrix $R$ should be rescaled to have generalised variance, defined by the geometric mean of the diagonal elements of its generalised inverse
 \begin{equation}
-    \sigma^2_{\text{GV}}(\mathbf{R}) = \prod_{i = 1}^n (R^-_{ii})^{1/n} = \exp \left( \frac{1}{n} \sum_{i = 1}^n \log (R^-_{ii}) \right),
+    \sigma^2_{\text{GV}}(R) = \prod_{i = 1}^n (R^-_{ii})^{1/n} = \exp \left( \frac{1}{n} \sum_{i = 1}^n \log (R^-_{ii}) \right),
 \end{equation}
-equal to one, by replacing $\mathbf{R}$ with $\mathbf{R}^\star = \mathbf{R} / \sigma^2_{\text{GV}}(\mathbf{R})$.
+equal to one, by replacing $R$ with $R^\star = R / \sigma^2_{\text{GV}}(R)$.
 As the diagonal elements $R^-_{ii}$ correspond to marginal variances, the generalised variance gives a measure of the average marginal variance.
@@
 Allowing the variance to be controlled by $\tau_\phi$ alone is important as it allows for consistent, interpretable prior selection.
-When the adjacency graph is disconnected it is not appropriate to scale the structure matrix $\mathbf{R}$ uniformly since for a given precision $\tau_\phi$, local smoothing operates on each connected component independently.
-As such, each connected component should be scaled independently to have generalised variance one giving $\mathbf{R}^\star_I = \mathbf{R}_I / \sigma^2_{\text{GV}}(\mathbf{R}_I)$ where $\mathbf{R}_I$ is the sub-matrix of the structure matrix corresponding to index set $I$.
+When the adjacency graph is disconnected it is not appropriate to scale the structure matrix $R$ uniformly since for a given precision $\tau_\phi$, local smoothing operates on each connected component independently.
+As such, each connected component should be scaled independently to have generalised variance one giving $R^\star_I = R_I / \sigma^2_{\text{GV}}(R_I)$ where $R_I$ is the sub-matrix of the structure matrix corresponding to index set $I$.
 
@@
 
+\clearpage
+
+# Simulation study
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=1]{depends/crps-boxplot-rho.pdf}}
+\label{fig:crps-boxplot-rho-1}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=2]{depends/crps-boxplot-rho.pdf}}
+\label{fig:crps-boxplot-rho-2}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=3]{depends/crps-boxplot-rho.pdf}}
+\label{fig:crps-boxplot-rho-3}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=4]{depends/crps-boxplot-rho.pdf}}
+\label{fig:crps-boxplot-rho-4}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=5]{depends/crps-boxplot-rho.pdf}}
+\label{fig:crps-boxplot-rho-5}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=6]{depends/crps-boxplot-rho.pdf}}
+\label{fig:crps-boxplot-rho-6}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=7]{depends/crps-boxplot-rho.pdf}}
+\label{fig:crps-boxplot-rho-7}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=1]{depends/histogram-ecdf.pdf}}
+\label{fig:histogram-ecdf-1}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=2]{depends/histogram-ecdf.pdf}}
+\label{fig:histogram-ecdf-2}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=3]{depends/histogram-ecdf.pdf}}
+\label{fig:histogram-ecdf-3}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=4]{depends/histogram-ecdf.pdf}}
+\label{fig:histogram-ecdf-4}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=5]{depends/histogram-ecdf.pdf}}
+\label{fig:histogram-ecdf-5}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=6]{depends/histogram-ecdf.pdf}}
+\label{fig:histogram-ecdf-6}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=7]{depends/histogram-ecdf.pdf}}
+\label{fig:histogram-ecdf-7}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=1]{depends/crps-maps.pdf}}
+\label{fig:crps-maps-1}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=2]{depends/crps-maps.pdf}}
+\label{fig:crps-maps-2}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=3]{depends/crps-maps.pdf}}
+\label{fig:crps-maps-3}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=4]{depends/crps-maps.pdf}}
+\label{fig:crps-maps-4}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=5]{depends/crps-maps.pdf}}
+\label{fig:crps-maps-5}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=6]{depends/crps-maps.pdf}}
+\label{fig:crps-maps-6}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\makebox[\textwidth][c]{\includegraphics[page=7]{depends/crps-maps.pdf}}
+\label{fig:crps-maps-7}
+\end{figure}
+
+\clearpage
+
+# HIV study
+
+\clearpage
+
 # References {#references .unnumbered}

### 2022-07-13 src/docs_paper/paper.Rmd
Work on manuscript and simulation study

@@
     corresponding: ath19@ic.ac.uk
-  # - name: Harrison Zhu
-  #   affiliation: A
-  # - name: Yidan Xu
-  #   affiliation: C
-  # - name: Samir Bhatt
-  #   affiliation: D
-  - name: Jeff W. Eaton
+  - name: Harrison Zhu
+    affiliation: A
+  - name: Yidan Xu
+    affiliation: C
+  - name: Samir Bhatt
+    affiliation: D
+  - name: Jeffrey W. Eaton
     affiliation: B
@@
     address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
-  # - code: C
-  #   address: Department of Statistics, University of Michigan
-  # - code: D
-  #   address: Department of Public Health, University of Copenhagen
+  - code: C
+    address: Department of Statistics, University of Michigan
+  - code: D
+    address: Department of Public Health, University of Copenhagen
   - code: E
@@
 
+```{r echo = FALSE}
+options(scipen = 100)
+knitr::opts_chunk$set(
+  echo = FALSE,
+  warning = FALSE,
+  message = FALSE,
+  dpi = 320,
+  cache = TRUE,
+  out.width = "95%",
+  fig.align = 'center'
+)
+
+comma <- function(x) formatC(x, format = "d", big.mark = ",")
+```
+
 # Introduction
@@
 Gaussian Markov random fields (GMRFs) [@rue2005gaussian], which combine a Gaussian distribution with Markov conditional independence assumptions between areas, are the most common class of models used to specify spatial random effects.
-Using a GMRF often reflects an assumption that observations made in areas close together are correlated, and more distant relationships may be ignored.
-Arguably the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
+Using a GMRF typically reflects the assumption that observations made in areas close together are correlated, and more distant relationships may be ignored.
+The simplest GMRF model is arguably that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
 This model requires minimal additional modelling choices and is widely implemented, including as a part of the `R-INLA` package [@rue2017bayesian], popular within spatial statistics.
@@
 We hypothesised that incorporating a spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models.
-We set out to test this hypothesis, and in doing so provide practical recommendations for modelling spatial structure.
+Here, we set out to test this hypothesis, and in doing so provide practical recommendations for modelling spatial structure.
 <!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)?
@@
 
-Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
-@duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model^[This conclusion was specific to the (often-analysed) Scottish lip cancer example, and like the majority of the literature, based on the deviance information criteria (DIC).
+<!-- Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related. -->
+<!-- @duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model^[This conclusion was specific to the (often-analysed) Scottish lip cancer example, and like the majority of the literature, based on the deviance information criteria (DIC).
 The DIC is based on summarising the posterior distribution by a point estimate, and is recommended against by @vehtari2017practical.].
-Another approach is to take into account that areal data is typically produced by aggregating higher resolution data, so as to be a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change] and dating to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
-Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
-This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially], `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages.
-Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].
+-->
+<!-- Another approach is to take into account that areal data is typically produced by aggregating higher resolution data, so as to be a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change] and dating to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
+-->
+<!-- Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
+-->
+<!-- This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially], `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages.
+-->
+<!-- Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].
+-->
 
-We also hypothesised that improvements to the Besag model may be transferable to improving convolution models.
-Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
-Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models.
-Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive.
-These convolutions use a Besag model as the spatially structured component, though any choice of spatial random effect model is possible.
-The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and more broadly both models are of substantive practical interest.
+<!-- We also hypothesised that improvements to the Besag model may be transferable to improving convolution models. -->
+<!-- Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
+-->
+<!-- Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models. -->
+<!-- Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive. -->
+<!-- These convolutions use a Besag model as the spatially structured component, though any choice of spatial random effect model is possible. -->
+<!-- The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and more broadly both models are of substantive practical interest.
+-->
 
@@
 Section \ref{sec:background} provides background on areal data and the Bayesian hierarchical modelling approach to small-area estimation.
-In Section \ref{sec:ab}, we review developments in specifying spatial random effects based on adjacency, before presenting an alternative kernel-based approach in Section \ref{sec:ks}.
-In Section \ref{sec:simulation} we compare models using simulated data, before applying them to mapping HIV prevalence in Section \ref{sec:hiv}.
-Finally, in Section \ref{sec:conclusion} we discuss our conclusions and directions for future research.
+In Section \ref{sec:ab}, we review developments in specifying spatial random effects based on adjacency, before presenting an alternative approach based on kernels in Section \ref{sec:ks}.
+In Section \ref{sec:simulation} we compare models using simulated data, before applying them to mapping HIV prevalence in sub-Saharan Africa in Section \ref{sec:hiv}.
+Finally, we discuss our conclusions and directions for future research in Section \ref{sec:conclusion}.
 
@@
 
-Let $\Sc \subset \mathbb{R}^2$ be the study region and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$.
+Let $\Sc \subset \mathbb{R}^2$ be the study region, and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$.
 Areal data are a type of spatial data where observations $y_i$ are associated to areas $A_i$.
@@
 Point data are another type of spatial data where instead observations $y(s)$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.
-Often areal may be conceptualised as arising from aggregation of point data so that $y_i = \int_{A_i} y(s) \text{d}s$.
-There are some exceptions, such as the minimum wage or disease control measures, which are determined at an administrative level and have a genuinely discrete spatial structure\footnote{For data of this sort, there is less clear justification for using spatial smoothing.
+
+Usually, areal may be conceptualised as arising from aggregation of point data so that $y_i = \int_{A_i} y(s) \text{d}s$.
+Exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a genuinely discrete spatial structure\footnote{For data of this sort, there is less clear justification for using spatial smoothing.
 Realistically, data relate to many covariates, some number of which are likely have discrete spatial structure and others of which continuous spatial structure, such that the result is often an intermediate.
@@
 BYM2.}.
-
 If the areal data we observe are indeed an aggregation of point data, why not use the higher resolution point data in our models rather than areal data?
@@
 The latter benefit is basis for the statistical task of small-area estimation, which aims to produce reliable local estimates where small sample sizes lead to noisy data [@pfeffermann2013new].
-In a spatial setting, this is often in small geographic areas, though the phrase "small-area" is not restricted to geographic areas and may be interpreted more broadly to mean any area where data are insufficient to make accurate local parameter inferences^[For example, in the context of multilevel regression and post-stratification [@gelman1997poststratification] small-areas are generated by the intersection of demographic variables like age, gender and race, alongside geographic variables.].
-For example, due to the cost of gathering samples, a survey may be designed to give reliable estimates at a national level but not at a district level.
+In a spatial setting, this is often in small geographic areas, though the phrase "small-area" is not restricted to geographic areas and may be interpreted more broadly to mean any area where data are insufficient to make accurate local parameter inferences.
+In the context of multilevel regression and post-stratification [@gelman1997poststratification], small-areas are generated by the intersection of demographic variables like age, gender and race, alongside geographic variables.
+Due to the cost of gathering samples, a survey may be designed to give reliable estimates at at an aggregated level but not at a small-area level.
 Although direct estimators of local parameters are unbiased, when data are sparse the total error may be reduced by accepting some bias in exchange for reduced variance using so-called indirect estimators.
@@
 
-Bayesian hierarchical models are an attractive framework for small-area estimation [@rao2015small].
-To model areal observations $\y$ corresponding to the areas $\{A_i\}_{i = 1}^n$ of a study region $\Sc$, a three-layer structure [@berliner1996hierarchical; @cressie2015statistics] given by
+Bayesian hierarchical models are an attractive framework for small-area estimation, whereby areal data $y$ corresponding to the areas $\{A_i\}_{i = 1}^n$, are modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
 \begin{alignat}{2}
-&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, g^{-1}(x_i), \btheta), \quad i = 1, \ldots, n, \label{eq:data} \\
-&\text{(Latent field)}  &        \x &\sim p(\x \, | \, \btheta), \label{eq:process} \\
-&\text{(Parameters)}    & \qquad \btheta &\sim p(\btheta), \label{eq:parameters}
+&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, g^{-1}(x_i), \theta), \quad i = 1, \ldots, n, \label{eq:data} \\
+&\text{(Latent field)}  &        x &\sim p(x \, | \, \theta), \label{eq:process} \\
+&\text{(Parameters)}    & \qquad \theta &\sim p(\theta), \label{eq:parameters}
 \end{alignat}
-may be used, where $\x$ is the $n$-dimensional latent field, $\btheta$ are the $m$-dimensional\footnote{Ordinarily $m$ is much smaller than $n$ as few parameters are required in order to control the properties of the latent field and likelihood.} parameters and $g$ is a link function.
-Spatial structure is modelled at the level of the latent field, allowing the observation model, often either binomial or Poisson, to be conditionally independent and identically distributed $p(\y \, | \, \x, \btheta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \btheta)$.
-As small-area estimation models are usually descriptive rather than mechanistic, it is natural [@best2005comparison] to use a Gaussian distribution $\x \sim \mathcal{N}(\bm{\mu}, \mathbf{\Sigma})$ for the latent field^[The Gaussian distribution is the maximal entropy distribution which retains finite mean and variance, as well as being computationally convenient.].
+where $x$ is the $n$-dimensional latent field, $\theta$ are the $m$-dimensional parameters, $m < n$, and $g$ is a link function.
+Spatial structure is modelled at the level of the latent field, and the observation model, often either binomial or Poisson, is conditionally independent and identically distributed $p(\y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
+As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $x \sim \mathcal{N}(\mu, \Sigma)$ for the latent field.
 Three-layer models with Gaussian latent fields comprise a wide array of popular models, including generalised linear mixed models (GLMMs), and have been collectively studied under the title latent Gaussian models (LGMs) [@rue2009approximate].
@@
 
-Decomposing the latent field in a LGM gives the sum of a constant mean $\bm{\mu}$, known also as the fixed effects, and (spatial) random effects $\bphi \sim \mathcal{N}(\mathbf{0}, \mathbf{\Sigma})$.
-The mean may be modelled using a linear predictor $\mu_i = \beta_0 + \z_i^\top \bm{\beta}$, where $(\beta_0, \bm{\beta})$ are regression parameters corresponding to areal covariates^[To denote length $p$ vectors of areal and point covariates, we use $\z_i = (z_{i,1}, \ldots, z_{i,p})^\top$ and $\z(s) = (z_1(s), \ldots, z_p(s))^\top$ respectively.} $(1, \z_i)^\top$].
-Our focus is on the spatial random effects $\bphi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
-Specifying $\bphi$ amounts to specifying the entries of a precision or covariance matrix, which we discuss below.
-
+The latent field in a LGM is comprised a constant mean $\mu$, known also as the fixed effects, and (spatial) random effects $\phi \sim \mathcal{N}(0, \Sigma)$.
+The mean may be modelled using a linear predictor $\mu_i = \beta_0 + z_i^\top \beta$, where $(\beta_0, \beta)$ are regression parameters corresponding to areal covariates^[To denote length $p$ vectors of areal and point covariates, we use $z_i = (z_{i,1}, \ldots, z_{i,p})^\top$ and $z(s) = (z_1(s), \ldots, z_p(s))^\top$ respectively.} $(1, \z_i)^\top$].
+Here, we focus on the spatial random effects $\phi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
 If no spatial structure remains after inclusion of covariates in the mean, independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ should be used, where $\tau_\phi$ is a shared precision.
-Exploratory data analysis techniques such as Moran's $I$ coefficient [@cliff1981spatial], visual inspection, may be used to determine if there remains any spatial auto-correlation in the data.
+Exploratory data analysis techniques such as visual inspection or Moran's $I$ coefficient [@cliff1981spatial] may be used to determine if there remains any spatial auto-correlation in the data.
 <!-- TODO: Add calculation of Moran's $I$ to my HIV survey data-sets. -->
+Specifying $\phi$ amounts to specifying the entries of a precision or covariance matrix, as we discuss below.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{depends/geometry-graph-zwe.pdf}
-\caption{A map of the districts of Zimbabwe together with the corresponding adjacency graph structure $\mathcal{G}$.}
-\label{fig:zi}
-\end{figure}
+```{r geometry-graph-zwe, fig.align="center", fig.cap="A map of the districts of Zimbabwe together with the corresponding adjacency graph structure $\\mathcal{G}$. \\label{fig:geometry-graph-zwe}"}
+knitr::include_graphics("depends/geometry-graph-zwe.pdf")
+```
 
@@
 \begin{equation}
-    \phi_i \, | \, \bphi_{-i} \sim \N \left(\frac{1}{n_{\delta i}} \sum_{j: j \sim i} \phi_j, \frac{1}{n_{\delta i}\tau_\phi}\right), \label{eq:besag}
+    \phi_i \, | \, \phi_{-i} \sim \N \left(\frac{1}{n_{\delta i}} \sum_{j: j \sim i} \phi_j, \frac{1}{n_{\delta i}\tau_\phi}\right), \label{eq:besag}
 \end{equation}
-where $\delta i$ is the set of neighbours of $A_i$ with cardinality $n_{\delta i} = |\delta i|$ and $\bphi_{-i} = (\phi_1, \ldots, \phi_{i - 1}, \phi_{i + 1}, \ldots, \phi_n)^\top$ is the vector of spatial random effects with the $i$th entry removed.
+where $\delta i$ is the set of neighbours of $A_i$ with cardinality $n_{\delta i} = |\delta i|$ and $\phi_{-i} = (\phi_1, \ldots, \phi_{i - 1}, \phi_{i + 1}, \ldots, \phi_n)^\top$ is the vector of spatial random effects with the $i$th entry removed.
 The conditional mean of the random effect $\phi_i$ is the average of its neighbours $\{\phi_j\}_{j \sim i}$ and the precision $n_{\delta i}\tau_\phi$ is proportional to the number of neighbours $n_{\delta i}$.
@@
 \begin{equation}
-    \bphi \sim \mathcal{N}(\mathbf{0}, \tau_\phi^{-1} \mathbf{R}^{-}), \label{eq:gmrf}
+    \phi \sim \mathcal{N}(\mathbf{0}, \tau_\phi^{-1} R^{-}), \label{eq:gmrf}
 \end{equation}
-where $\mathbf{R}^{-}$ is the generalised inverse of the rank-deficient structure matrix\footnote{So-called because it defines the \textit{structure} of the precision matrix $\mathbf{Q} = \tau_\phi \mathbf{R}$.} $\mathbf{R}$ with entries
+where $R^{-}$ is the generalised inverse of the rank-deficient structure matrix\footnote{So-called because it defines the \textit{structure} of the precision matrix $\mathbf{Q} = \tau_\phi R$.} $R$ with entries
 \begin{equation}
@@
 \end{equation}
-The Markov property arises due to the conditional independence structure $p(\phi_i \, | \, \bphi_{-i}) = p(\phi_i \, | \, \bphi_{\delta i})$ whereby each area only depends on its neighbours.
-This is reflected in the sparsity of $\mathbf{R}$ whereby $\phi_i \perp \phi_j \, | \, \phi_{-ij}$ if and only if $R_{ij} = 0$.
-The structure matrix $\mathbf{R}$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
+The Markov property arises due to the conditional independence structure $p(\phi_i \, | \, \phi_{-i}) = p(\phi_i \, | \, \phi_{\delta i})$ whereby each area only depends on its neighbours.
+This is reflected in the sparsity of $R$ whereby $\phi_i \perp \phi_j \, | \, \phi_{-ij}$ if and only if $R_{ij} = 0$.
+The structure matrix $R$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
 Figure \ref{fig:zi} shows the adjacency graph for the districts of Zimbabwe alongside the geometry.
@@
 
-Rewriting Equation \ref{eq:gmrf}, the probability density function of $\bphi$ is
+Rewriting Equation \ref{eq:gmrf}, the probability density function of $\phi$ is
 \begin{equation}
-    p(\bphi)
-    \propto \exp \left( -\frac{\tau_\phi}{2} \bphi^\top \mathbf{R} \bphi \right)
+    p(\phi)
+    \propto \exp \left( -\frac{\tau_\phi}{2} \phi^\top R \phi \right)
     \propto \exp \left( -\frac{\tau_\phi}{2} \sum_{i \sim j} (\phi_i - \phi_j)^2 \right). \label{eq:pdfphi}
 \end{equation}
-This density is a function of the pairwise differences $\phi_i - \phi_j$ and so is invariant to the addition of a constant $c$ to each entry $p(\bphi) = p(\bphi + c\mathbf{1})$, leading to an improper uniform distribution on the average of the $\phi_i$.
-If $\mathcal{G}$ is connected\footnote{In that by traversing the edges, any vertex can be reached from any other vertex.} then there is only one impropriety in the model and $\text{rank}(\mathbf{R}) = n - 1$, while if $\mathcal{G}$ is disconnected, and composed of $n_c \geq 2$ connected components with index sets $I_1, \ldots, I_{n_c}$, then the corresponding structure matrix $\mathbf{R}$ has rank $n - n_c$ and the density is invariant to the addition of a constant to each of the connected components $p(\bphi_{I}) = p(\bphi_{I} + c \mathbf{1})$ where $I = I_1, \ldots, I_{n_c}$.
+This density is a function of the pairwise differences $\phi_i - \phi_j$ and so is invariant to the addition of a constant $c$ to each entry $p(\phi) = p(\phi + c\mathbf{1})$, leading to an improper uniform distribution on the average of the $\phi_i$.
+If $\mathcal{G}$ is connected\footnote{In that by traversing the edges, any vertex can be reached from any other vertex.} then there is only one impropriety in the model and $\text{rank}(R) = n - 1$, while if $\mathcal{G}$ is disconnected, and composed of $n_c \geq 2$ connected components with index sets $I_1, \ldots, I_{n_c}$, then the corresponding structure matrix $R$ has rank $n - n_c$ and the density is invariant to the addition of a constant to each of the connected components $p(\phi_{I}) = p(\phi_{I} + c \mathbf{1})$ where $I = I_1, \ldots, I_{n_c}$.
 
-<!-- ## Concerns about the Besag model's representation of space \label{sec:concerns} -->
+## Concerns about the Besag model's representation of space \label{sec:concerns}
 
-<!-- \begin{figure} -->
-<!-- \centering -->
-<!-- \includegraphics{depends/maup1.pdf} -->
-<!-- \caption{Each of these four geometries corresponds to the same adjacency graph.} -->
-<!-- \label{fig:tikz1} -->
-<!-- \end{figure} -->
+\begin{figure}
+\centering
+\includegraphics{depends/maup1.pdf}
+\caption{Each of these four geometries corresponds to the same adjacency graph.}
+\label{fig:tikz1}
+\end{figure}
 
-<!-- \begin{figure} -->
-<!-- \centering -->
-<!-- \includegraphics{depends/maup2.pdf} -->
-<!-- \caption{A sequence of geometries where the number of neighbours of area one grows by one at each iteration.} -->
-<!-- \label{fig:tikz2} -->
-<!-- \end{figure} -->
+\begin{figure}
+\centering
+\includegraphics{depends/maup2.pdf}
+\caption{A sequence of geometries where the number of neighbours of area one grows by one at each iteration.}
+\label{fig:tikz2}
+\end{figure}
 
-<!-- \begin{figure} -->
-<!-- \centering -->
-<!-- \includegraphics{depends/maup3.pdf} -->
-<!-- \caption{Each of the shaded areas are split into two moving from geometry (i) to (ii).} -->
-<!-- \label{fig:tikz3} -->
-<!-- \end{figure} -->
+\begin{figure}
+\centering
+\includegraphics{depends/maup3.pdf}
+\caption{Each of the shaded areas are split into two moving from geometry (i) to (ii).}
+\label{fig:tikz3}
+\end{figure}
 
-<!-- The Besag model was originally proposed for use in image analysis, where areas correspond to pixels arranged in a regular lattice structure. -->
-<!-- Since then, it has seen wider use, including in situations where the spatial structure is less regular. -->
-<!-- This leads to a number of concerns about the model's applicability, which we highlight below. -->
-<!-- Closely linked are the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups.
--->
-<!-- <!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. --> -->
+The Besag model was originally proposed for use in image analysis, where areas correspond to pixels arranged in a regular lattice structure.
+Since then, it has seen wider use, including in situations where the spatial structure is less regular.
+This leads to a number of concerns about the model's applicability, which we highlight below.
+Closely linked are the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups.
+<!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. -->
 
-<!-- ### Adjacency compression -->
+### Adjacency compression
 
-<!-- Summarising all spatial information with an adjacency graph represents a loss of information. -->
-<!-- Many geometries share the same adjacency graph, and therefore the same probability model, as illustrated by Figure \ref{fig:tikz1}. -->
-<!-- This fact is not concerning in itself, though it prompts consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.\footnote{The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be.
-In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.} -->
-<!-- It is intuitive that the more regular the spatial structure the less information is lost in compression to an adjacency graph. -->
-<!-- In image analysis, very little spatial information is lost in compression a lattice structure to an adjacency graph. -->
-<!-- On the other hand, the regions of a country, determined by political and geographic forces, tend to display greater irregularity. -->
-<!-- As such, the appropriateness of adjacency compression varies by the type of geometry common to the application setting. -->
+Summarising all spatial information with an adjacency graph represents a loss of information.
+Many geometries share the same adjacency graph, and therefore the same probability model, as illustrated by Figure \ref{fig:tikz1}.
+This fact is not concerning in itself, though it prompts consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.\footnote{The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be.
+In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.}
+It is intuitive that the more regular the spatial structure the less information is lost in compression to an adjacency graph.
+In image analysis, very little spatial information is lost in compression a lattice structure to an adjacency graph.
+On the other hand, the regions of a country, determined by political and geographic forces, tend to display greater irregularity.
+As such, the appropriateness of adjacency compression varies by the type of geometry common to the application setting.
 
-<!-- ### Mean structure -->
+### Mean structure
 
-<!-- A feature specific to the Besag model is that all adjacent areas count equally. -->
-<!-- This assumption is unsatisfying: for most geometries, we expect some areas to be more highly correlated to some neighbours than to others. -->
-<!-- Figure \ref{fig:tikz1} illustrates a number of heuristic features for neighbour importance, including length of shared border in geometry (iii), and the proximity of centers of mass in geometry (iv).
--->
-<!-- This may be remedied using more general weighted ICAR models, as we outline in Section \ref{sec:icar}. -->
+A feature specific to the Besag model is that all adjacent areas count equally.
+This assumption is unsatisfying: for most geometries, we expect some areas to be more highly correlated to some neighbours than to others.
+Figure \ref{fig:tikz1} illustrates a number of heuristic features for neighbour importance, including length of shared border in geometry (iii), and the proximity of centers of mass in geometry (iv).
+This may be remedied using more general weighted ICAR models, as we outline in Section \ref{sec:icar}.
 
-<!-- ### Variance structure -->
+### Variance structure
 
-<!-- A striking feature of Equation~\ref{eq:besag} is that the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.  -->
-<!-- It follow that as $n_{\delta i} \to \infty$ then $\text{Var}(\phi_i) \to 0$. -->
-<!-- This is illustrated by Figure \ref{fig:tikz2} where the area on the right is repeatedly divided such that its number of neighbours increases by one at each step. -->
-<!-- This property is a consequence of averaging the conditional mean over a greater number of areas, which, in certain situations, can correspond to a greater amount of information. -->
-<!-- However, if the amount of information in the shaded area remains fixed, it is inappropriate that $\text{Var}(\phi_1)$ should tend to zero just as a result of drawing additional, arbitrary, boundaries.
--->
-<!-- In the image analysis setting this modelling assumption is reasonable: each pixel represents a fixed amount of information and a higher pixel density represents a greater amount of information.
--->
-<!-- On the other hand, in public health and epidemiology, drawing boundaries to create additional regional areas cannot be expected to correspond to a greater amount of information and this assumption is not appropriate.
--->
+A striking feature of Equation~\ref{eq:besag} is that the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.
+It follow that as $n_{\delta i} \to \infty$ then $\text{Var}(\phi_i) \to 0$.
+This is illustrated by Figure \ref{fig:tikz2} where the area on the right is repeatedly divided such that its number of neighbours increases by one at each step.
+This property is a consequence of averaging the conditional mean over a greater number of areas, which, in certain situations, can correspond to a greater amount of information.
+However, if the amount of information in the shaded area remains fixed, it is inappropriate that $\text{Var}(\phi_1)$ should tend to zero just as a result of drawing additional, arbitrary, boundaries.
+In the image analysis setting this modelling assumption is reasonable: each pixel represents a fixed amount of information and a higher pixel density represents a greater amount of information.
+On the other hand, in public health and epidemiology, drawing boundaries to create additional regional areas cannot be expected to correspond to a greater amount of information and this assumption is not appropriate.
 
-<!-- Suppose we fit a Besag model upon identical data using each of the two geometries in Figure \ref{fig:tikz3}. -->
-<!-- If the spatial variation is relatively smooth, dividing the shaded areas into two will result in a lower estimated variance $\sigma^2_\phi$ in geometry (ii) as compared with geometry (i) because there will appear to be less variation between neighbouring areas.
--->
-<!-- This problem does not only apply locally: since the effect of $\sigma^2_\phi$ applies everywhere, the smoothing will change even in unaltered parts of the study region. -->
+Suppose we fit a Besag model upon identical data using each of the two geometries in Figure \ref{fig:tikz3}.
+If the spatial variation is relatively smooth, dividing the shaded areas into two will result in a lower estimated variance $\sigma^2_\phi$ in geometry (ii) as compared with geometry (i) because there will appear to be less variation between neighbouring areas.
+This problem does not only apply locally: since the effect of $\sigma^2_\phi$ applies everywhere, the smoothing will change even in unaltered parts of the study region.
 
@@
 <!-- The Besag model is a special case of a more general class of (zero-mean) ICAR models in which each neighbour may not be weighted equally. -->
-<!-- These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\bm{\kappa} = (\kappa_1, \ldots, \kappa_n)^\top$ such that $\phi_i \, | \, \bphi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$.
+<!-- These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\bm{\kappa} = (\kappa_1, \ldots, \kappa_n)^\top$ such that $\phi_i \, | \, \phi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$.
 -->
-<!-- The structure matrix $\mathbf{R}$ corresponding to the above full conditionals is given by $\mathbf{R} = \mathbf{D}_\kappa(\mathbf{I} - \mathbf{B})$, where $\mathbf{B}$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $\mathbf{D}_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$.
+<!-- The structure matrix $R$ corresponding to the above full conditionals is given by $R = \mathbf{D}_\kappa(\mathbf{I} - \mathbf{B})$, where $\mathbf{B}$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $\mathbf{D}_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$.
 -->
 <!-- As the structure matrix is symmetric we must have the condition $-b_{ij}\kappa_i = -b_{ji}\kappa_j$. -->
-<!-- To meet this condition, it is often simpler to use an unscaled weights matrix $\mathbf{W} = \mathbf{D}_\kappa \mathbf{B}$ such that $\mathbf{R} = \mathbf{D}_\kappa - \mathbf{W}$. -->
+<!-- To meet this condition, it is often simpler to use an unscaled weights matrix $\mathbf{W} = \mathbf{D}_\kappa \mathbf{B}$ such that $R = \mathbf{D}_\kappa - \mathbf{W}$. -->
 <!-- For example, in the Besag model $\mathbf{W}$ corresponds to the adjacency matrix. -->
@@
 <!-- Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model. -->
-<!-- The Besag-York-Molli\'{e} (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\bphi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
+<!-- The Besag-York-Molli\'{e} (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\phi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
 -->
@@
 <!-- \begin{equation} -->
-<!--     \bphi \sim \N(\mathbf{0}, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} \mathbf{R}^{-}) \label{eq:bym}. -->
+<!--     \phi \sim \N(\mathbf{0}, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} R^{-}) \label{eq:bym}. -->
 <!-- \end{equation} -->
@@
 <!-- \begin{align} -->
-<!-- \bphi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2} -->
+<!-- \phi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2} -->
 <!-- \end{align} -->
-<!-- where $\tau_\phi$ is the marginal precision of $\bphi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $\bu$ with precision matrix given by the scaled structure matrix $\mathbf{R}^\star$ (see~\ref{app:freni}).
+<!-- where $\tau_\phi$ is the marginal precision of $\phi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $\bu$ with precision matrix given by the scaled structure matrix $R^\star$ (see~\ref{app:freni}).
 -->
@@
 
-<!-- \section{Areal kernel spatial random effects} \label{sec:ks} -->
+# Areal kernel spatial random effects \label{sec:ks}
 
@@
 
-<!-- \subsection{Areal kernels} -->
+## Areal kernels
 
-<!-- Section \ref{sec:ab} reviewed ways to construct the precision matrix of the spatial random effects using an adjacency relation. -->
-<!-- Another approach is to define the covariance matrix directly using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$, where $\mathcal{P}$ denotes the power set such that $\mathcal{P}(\Sc)$ is the space of subsets of the study region, which gives a measure of similarity between two areas.
--->
-<!-- If $K$ is positive semi-definite, then we define areal kernel spatial random effects by -->
-<!-- \begin{equation} -->
-<!--     \bphi \sim \mathcal{N}(\mathbf{0}, \frac{1}{\tau_\phi} \mathbf{K}), \label{eq:arealkernel} -->
-<!-- \end{equation} -->
-<!-- where the $n \times n$ Gram matrix $\mathbf{K}$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix. -->
-<!-- We keep $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices. -->
+Section \ref{sec:ab} reviewed ways to construct spatial random effect precision matrices using an adjacency relation.
+Another approach is to define the covariance matrix using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$, where $\mathcal{P}$ denotes the power set such that $\mathcal{P}(\Sc)$ is the space of subsets of the study region, which gives a measure of similarity between two areas.
+If $K$ is positive semi-definite, then we define areal kernel spatial random effects by
+\begin{equation}
+    \phi \sim \mathcal{N}(\mathbf{0}, \frac{1}{\tau_\phi} K), \label{eq:arealkernel}
+\end{equation}
+where the $n \times n$ Gram matrix $K$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix.
+We keep $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.
 
-<!-- Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$. -->
-<!-- To construct $K$ we use a simple method, namely averaging the kernel $k$ computed on some collection of points from within each area. -->
-<!-- In the multiple instance learning literature, $K$ is described as a type of kernels on sets [@gartner2002multi]. -->
-<!-- Although it may be possible to learn the method for combining kernels [@gonen2011multiple] from data, we do believe that this is feasible in the small-area estimation setting, and instead hope to construct areal kernels which represent our prior beliefs about the spatial process.
--->
+Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$.
+To construct $K$ we use a simple method, namely averaging the kernel $k$ computed on some collection of points from within each area.
+As such, in the multiple instance learning literature $K$ would be described as a type of kernels on sets [@gartner2002multi].
+Although it may be possible to learn the method for combining kernels from data [@gonen2011multiple], we do believe that this is feasible in the small-area estimation setting, and instead hope to construct areal kernels which represent our prior beliefs about the spatial process.
 
-<!-- \subsection{Centroid kernel} -->
+## Centroid kernel
 
-<!-- The simplest approach is to use a single point, which one might hope would be representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude.
--->
-<!-- This results in the centroid kernel -->
-<!-- \begin{equation} -->
-<!--     K(A_i, A_j) = k(c_i, c_j). -->
-<!-- \end{equation} -->
-<!-- % More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
-<!-- Usage of the centroid kernel includes @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both using the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$.
--->
-<!-- The former fix the length-scale $l$ using a semi-variogram and the later place a prior on the inverse length-scale. -->
-<!-- In their simulation experiment, \citet[Section 3]{best2005comparison} used data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas (though it is unsurprising that a stationary covariance structure would struggle to recover non-stationary spatial structure).
--->
-<!-- More recently, @teh2021efficient used a centroid kernel as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19. -->
+The simplest approach is to use a single point, which one might hope would be representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude.
+This results in the centroid kernel
+\begin{equation}
+    K(A_i, A_j) = k(c_i, c_j).
+\end{equation}
+% More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation.
+Usage of the centroid kernel includes @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both using the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$.
+The former fix the length-scale $l$ using a semi-variogram and the later place a prior on the inverse length-scale.
+In their simulation experiment, \citet[Section 3]{best2005comparison} used data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas (though it is unsurprising that a stationary covariance structure would struggle to recover non-stationary spatial structure).
+More recently, @teh2021efficient used a centroid kernel as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
 
-<!-- \subsection{Integrated kernel} \label{sec:ik} -->
+## Integrated kernel \label{sec:ik}
 
@@
 <!-- which is equivalent to the covariance structure obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition. -->
-<!-- Vaious authors [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi] in the machine learning literature have recently contributed to the development of these aggregated Gaussian process models.
+<!-- Various authors [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi] in the machine learning literature have recently contributed to the development of these aggregated Gaussian process models.
 -->
@@
 <!-- As with learning the way in which the kernels should be combined, in more data rich settings it may be possible to learn the integration points [@campbell2019automated]. -->
-<!-- Taking any of these approaches requires $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $\mathbf{K}$. -->
+<!-- Taking any of these approaches requires $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$. -->
 <!-- This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or integration points it is feasible to use.
@@
 
-<!-- <!-- Disaggregation is challenging without auxiliary information. --> -->
-<!-- <!-- Kelsall and Wakefield They use the cubic correlation function and, to reduce the number of Gram matrix constructions and inversions, learn a range parameter by placing a discrete prior.
---> -->
-<!-- <!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. --> -->
-<!-- <!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. --> -->
-<!-- <!-- Explain under what circumstances we believe this approximation to be acceptable. --> -->
+<!-- Disaggregation is challenging without auxiliary information. -->
+<!-- Kelsall and Wakefield They use the cubic correlation function and, to reduce the number of Gram matrix constructions and inversions, learn a range parameter by placing a discrete prior. -->
+<!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. -->
+<!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. -->
+<!-- Explain under what circumstances we believe this approximation to be acceptable. -->
 
-<!-- \clearpage -->
+# Simulation study \label{sec:simulation}
 
-<!-- # Simulation study \label{sec:simulation} -->
+```{r simulation-geometries, fig.align="center", fig.cap="Simulation geometries."}
+knitr::include_graphics("depends/simulation-geometries.pdf")
+```
 
-<!-- We tested the ability of inferential models with different spatial random effect specifications to accurately recover small-area quantities. -->
-<!-- The data and modelling choices were designed with a spatial epidemiology setting in mind, and are similar to those used in Section \ref{sec:hiv} where we considered models for HIV prevalence using data from national household surveys in sub-Saharan Africa.
--->
-<!-- The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`. -->
-<!-- We used the `orderly` package [@orderly] for reproducible research. -->
+We tested the ability of inferential models with varying spatial random effect specifications to accurately recover small-area quantities.
+The data and modelling choices were designed with a spatial epidemiology setting in mind, and are similar to those used in Section \ref{sec:hiv} where we considered models for HIV prevalence using data from national household surveys in sub-Saharan Africa.
+The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`.
+We used the `orderly` package [@orderly] for reproducible research.
 
-<!-- ## Synthetic data-sets -->
+## Synthetic data-sets
 
-<!-- \begin{table}[] -->
-<!-- \small -->
-<!-- \begin{tabularx}{\textwidth}{@{}lX@{}} -->
-<!-- \toprule -->
-<!-- Simulation model & Additional details \\  -->
-<!--  \midrule -->
-<!-- \hypertarget{S1}{S1}. \textit{IID} & $\bphi \sim \mathcal{N}(0, I_n)$ \\ -->
-<!-- \hypertarget{S2}{S2}. \textit{Besag} & $\bphi \sim \mathcal{N}(0, (\mathbf{R}^\star)^{-})$ as described in Section \ref{sec:besag} where $\mathbf{R}^\star$ is the scaled structure matrix \\  -->
-<!-- \hypertarget{S3}{S3}.
-\textit{IK} & $\bphi \sim \mathcal{N}(0, \mathbf{K}^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation~\ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
--->
-<!-- Like the Besag model, we scaled the Gram matrix to have generalised variance one \\  -->
-<!-- \bottomrule -->
-<!-- \end{tabularx} -->
-<!-- \caption{The three simulation models from which we generate $\bphi$.} -->
-<!-- \label{tab:simulationmodels} -->
-<!-- \small -->
-<!-- \end{table} -->
+\begin{table}[]
+\small
+\begin{tabularx}{\textwidth}{@{}lX@{}}
+\toprule
+Simulation model & Additional details \\
+ \midrule
+\hypertarget{S1}{S1}. \textit{IID} & $\phi \sim \mathcal{N}(0, I_n)$ \\
+\hypertarget{S2}{S2}. \textit{Besag} & $\phi \sim \mathcal{N}(0, (R^\star)^{-})$ as described in Section \ref{sec:besag} where $R^\star$ is the scaled structure matrix \\
+\hypertarget{S3}{S3}.
+\textit{IK} & $\phi \sim \mathcal{N}(0, K^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation~\ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
+Like the Besag model, we scaled the Gram matrix to have generalised variance one \\
+\bottomrule
+\end{tabularx}
+\caption{The three simulation models from which we generate $\phi$.}
+\label{tab:simulationmodels}
+\small
+\end{table}
 
-<!-- To study the robustness of the inferential models to spatial misspecification, we simulated synthetic data-sets from three known data generating processes. -->
-<!-- The spatial random effects $\bphi$ were generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}). -->
-<!-- After obtaining spatial random effects $\bphi$ from each simulation model, we generated synthetic data $\y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via -->
-<!-- \begin{equation} -->
-<!-- \log \left( \frac{\rho_i}{1 - \rho_i} \right) = x_i = \beta_0 + \phi_i, \quad i = 1, \ldots, n. -->
-<!-- \end{equation} -->
-<!-- We fixed the sample sizes to $m_i = 25$ for all $i$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$. -->
-<!-- We used seven different geometries: the four vignette geometries in Figure \ref{fig:tikz1}, all of which have the same adjacency graph, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of C\^{o}te d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
--->
-<!-- These more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test the hypothesis that model performance depends on geometry regularity. -->
-<!-- For each combination of spatial random effect and geometry we replicated the simulation process above 200 times to generate a total of 1400 synthetic data-sets. -->
+We studied robustness to spatial misspecification, by simulating synthetic data-sets from three known data generating processes.
+The spatial random effects $\phi$ were generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}).
+We then generated synthetic data $\y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via
+\begin{equation}
+\log \left( \frac{\rho_i}{1 - \rho_i} \right) = x_i = \beta_0 + \phi_i, \quad i = 1, \ldots, n.
+\end{equation}
+We fixed the sample sizes to $m_i = 25$ for all $i$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$.
+We used seven different geometries: the four vignette geometries in Figure \ref{fig:tikz1}, which share an adjacency graph, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of C\^{o}te d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
+These more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test how model performance varies by geometry regularity.
+For each combination of spatial random effect and geometry we replicated the simulation process above 200 times to generate a total of 1400 synthetic data-sets.
 
-<!-- ## Inferential models -->
+## Inferential models
 
-<!-- We fit eight inferential models to each simulated data-set. -->
-<!-- Each inferential model shared most of its structure with the data generating process, only differing in the spatial random effect specification, which varied across the eight models (Table \ref{tab:inferentialmodels}), allowing us to test sensitivity to misspecification of the spatial structure.
+We fit eight inferential models, each of which differing in the spatial random effect specification (Table \ref{tab:inferentialmodels}), allowing us to test sensitivity to spatial structure misspecification.
+
+\begin{table}
+\small
+\begin{tabularx}{\textwidth}{llX}
+\toprule
+Inferential model & Algorithm & Additional details \\
+ \midrule
+\hypertarget{I1}{I1}. \textit{Constant} & INLA &
+No spatial random effects \label{row:I1} \\
+\hypertarget{I2}{I2}. \textit{IID} & INLA &
+$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} I_n)$\\
+\hypertarget{I3}{I3}. \textit{Besag} & INLA &
+$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations in~\ref{app:freni}  \\
+\hypertarget{I4}{I4}. \textit{BYM2} & INLA &
+$\phi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
+\hypertarget{I5}{I5}. \textit{FCK} & INLA &
+$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
+\hypertarget{I6}{I6}. \textit{CK} & NUTS & As in model~\hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
+\hypertarget{I7}{I7}.
+\textit{FIK} & INLA & $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
+\hypertarget{I8}{I8}. \textit{IK} & NUTS & As in model~\hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
+\bottomrule
+\end{tabularx}
+\caption{Each inferential model is implemented as a part of the \textsf{R} package \texttt{bsae} available from \url{github.com/athowes/bsae}.}
+\label{tab:inferentialmodels}
+\small
+\end{table}
+
+### Priors
+
+For the precision, we placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
+The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one, neither of which we assumed to reflect reality.
+A weakly informative $\mathcal{N}(-2, 1)$ prior was placed on $\beta_0$, setting most of the prior probability density for $\rho_i$ within a range $[0, 0.25]$ typical for a disease prevalence.
+
+### Kernel model details
+
+Following @stein1999interpolation, for the models~\hyperlink{I5}{I5}-\hyperlink{I8}{I8} we used the Mat\'{e}rn kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
+\begin{equation}
+k(s, s') =
+\frac{1}{2^{\nu - 1}\Gamma(\nu)} \left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right)^\nu
+K_\nu\left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right), \label{eq:matern}
+\end{equation}
+where $K_\nu$ is the modified Bessel function of the second kind, $|s - s'|$ is the Euclidean distance between $s$ and $s'$, $\nu$ is the smoothness hyperparameter and $l$ is the length-scale hyperparameter on the latitude-longitude scale.
+We fixed the smoothness parameter, which is typically unidentifiable from data, to be $3/2$, matching that used in simulation model~\hyperlink{S3}{S3}, and convenient in that it simplifies Equation~\ref{eq:matern} to be the product of a polynomial and an exponential
+\begin{equation}
+k(s, s') = \left(1 + \sqrt{3} \lvert s - s' \rvert / l \right) \exp \left(- \sqrt{3} \lvert s - s' \rvert / l \right).
+\end{equation}
+Taking a similar approach to @best1999bayesian, in models~\hyperlink{I5}{I5} and \hyperlink{I7}{I7} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data.
+In models~\hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
+For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51} (though the true number of integration points may be 1-2 above or below 10, as `sf::st_sample` with `type = "hexagonal"` is not guaranteed to be exact).
+We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}.
+For Model~\hyperlink{I6}{I6} although it is feasible to use a much larger number of integration points, since the Gram matrix is only computed once, we chose not to for comparability.
+
+## Inference algorithms
+
+In models~\hyperlink{I1}{I1}-\hyperlink{I5}{I5} choosing a Gaussian distribution for $\beta_0$ ensures that the latent random variables are Gaussian and the model is a LGM.
+This facilitated use of the integrated nested Laplace approximation (INLA) algorithm [@rue2009approximate] for inference, as implemented by the \texttt{R-INLA} package.
+INLA is a deterministic method for approximate Bayesian inference based upon a combination of numerical integration and the Laplace approximation.
+INLA is significantly faster than MCMC and has been shown to have comparable accuracy for LGMs in realistic, pre-asymptotic settings.
+<!-- Although there are particular conditions under which INLA is not accurate [@rue2017bayesian] they do not apply here, as we illustrate in Appendix \ref{app:inlamcmc} where we compare INLA to MCMC using a large number of iterations.
 -->
 
-<!-- \begin{table} -->
-<!-- \small -->
-<!-- \begin{tabularx}{\textwidth}{llX} -->
-<!-- \toprule -->
-<!-- Inferential model & Algorithm & Additional details \\  -->
-<!--  \midrule -->
-<!-- \hypertarget{I1}{I1}. \textit{Constant} & INLA &  -->
-<!-- No spatial random effects \label{row:I1} \\ -->
-<!-- \hypertarget{I2}{I2}. \textit{IID} & INLA &  -->
-<!-- $\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} I_n)$\\ -->
-<!-- \hypertarget{I3}{I3}. \textit{Besag} & INLA &  -->
-<!-- $\bphi \sim \mathcal{N}(0, (\tau_\phi \mathbf{R}^\star)^{-})$ following additional recommendations in~\ref{app:freni}  \\  -->
-<!-- \hypertarget{I4}{I4}. \textit{BYM2} & INLA &  -->
-<!-- $\bphi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\ -->
-<!-- \hypertarget{I5}{I5}. \textit{FCK} & INLA & -->
-<!-- $\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{K})$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\  -->
-<!-- \hypertarget{I6}{I6}. \textit{CK} & NUTS & As in model~\hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\  -->
-<!-- \hypertarget{I7}{I7}.
-\textit{FIK} & INLA & $\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{K})$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\ -->
-<!-- \hypertarget{I8}{I8}. \textit{IK} & NUTS & As in model~\hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\  -->
-<!-- \bottomrule -->
-<!-- \end{tabularx} -->
-<!-- \caption{Each inferential model is implemented as a part of the \textsf{R} package \texttt{bsae} available from \url{github.com/athowes/bsae}.} -->
-<!-- \label{tab:inferentialmodels} -->
-<!-- \small -->
-<!-- \end{table} -->
+## Model assessment
 
-<!-- ### Priors -->
+We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, subject to calibration.
+In the following two sections we describe our approach to model assessment.
+<!-- Analogous approaches may be used to check recovery of the intercept parameter $\beta_0$, precision parameter $\tau_\phi$, and, when relevant, the length-scale $l$. -->
 
-<!-- For the precision, we placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$. -->
-<!-- The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one, neither of which we assumed to reflect reality.
--->
-<!-- A weakly informative $\mathcal{N}(-2, 1)$ prior was placed on $\beta_0$, setting most of the prior probability density for $\rho_i$ within a range $[0, 0.25]$ typical for a disease prevalence.
--->
+### Continuous ranked probability score
 
-<!-- ### Kernel model details -->
+The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts.
+For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, \y)$ and $\omega_i$ be the true value.
+Writing $F$ as the cumulative distribution corresponding to $f$ then
+\begin{equation}
+\text{CRPS}(f, \omega_i) = \int_{-\infty}^{\infty} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i = \int_{0}^{1} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i, \label{eq:crps}
+\end{equation}
+where $\mathbbm{1}$ denotes the indicator function and the second equality follows from $0 \leq \rho_i \leq 1$.
+The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$.
+We use the CRPS because it is a strictly proper scoring rule [@gneiting2007strictly].
+This ensures that the (uniquely) best approach to minimising the score is to truthfully report probabilistic beliefs, which here correspond to the posterior distribution.
+The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, but in line with the findings of @kruger2020predictive, using kernel density estimation we found it difficult to compute satisfactorily.
+In contrast, given a collection of Monte Carlo samples $\{\rho_i^s\}_{s = 1}^S \sim f$, Equation~\ref{eq:crps} may easily be approximated \citep[Equation 37]{sigrist2015stochastic} by
+\begin{equation}
+\text{CRPS}(f, \omega_i) \approx \frac{1}{S} \sum_{s = 1}^S | \rho_i^s - \omega_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | \rho_i^s - \rho_i^l |. \label{eq:crpsapprox}
+\end{equation}
+We obtained these samples using the function \texttt{INLA::inla.posterior.sample} for models fit with \texttt{R-INLA} and directly from the Markov chain using \texttt{rstan::extract} for models fit with Stan.
 
-<!-- Following @stein1999interpolation, for the models~\hyperlink{I5}{I5}-\hyperlink{I8}{I8} we used the Mat\'{e}rn kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by -->
-<!-- \begin{equation} -->
-<!-- k(s, s') =  -->
-<!-- \frac{1}{2^{\nu - 1}\Gamma(\nu)} \left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right)^\nu -->
-<!-- K_\nu\left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right), \label{eq:matern} -->
-<!-- \end{equation} -->
-<!-- where $K_\nu$ is the modified Bessel function of the second kind, $|s - s'|$ is the Euclidean distance between $s$ and $s'$, $\nu$ is the smoothness hyperparameter and $l$ is the length-scale hyperparameter on the latitude-longitude scale.
--->
-<!-- We fixed the smoothness parameter, which is typically unidentifiable from data, to be $3/2$, matching that used in simulation model~\hyperlink{S3}{S3}, and convenient in that it simplifies Equation~\ref{eq:matern} to be the product of a polynomial and an exponential -->
-<!-- \begin{equation} -->
-<!-- k(s, s') = \left(1 + \sqrt{3} \lvert s - s' \rvert / l \right) \exp \left(- \sqrt{3} \lvert s - s' \rvert / l \right). -->
-<!-- \end{equation} -->
-<!-- Taking a similar approach to @best1999bayesian, in models~\hyperlink{I5}{I5} and \hyperlink{I7}{I7} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data.
--->
-<!-- In models~\hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
--->
-<!-- For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51} (though the true number of integration points may be 1-2 above or below 10, as `sf::st_sample` with `type = "hexagonal"` is not guaranteed to be exact).
--->
-<!-- We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}. -->
-<!-- For Model~\hyperlink{I6}{I6} although it is feasible to use a much larger number of integration points, since the Gram matrix is only computed once, we chose not to for comparability. -->
+\subsubsection{Coverage}
 
-<!-- ## Inference algorithms -->
+To assess the coverage of our estimates, we considered probability integral transform (PIT) histograms and empirical cumulative distribution function (ECDF) difference plots [@sailynoja2021graphical].
+We estimated the quantile $q_i = F(\omega_i)$ of the true value within the marginal posterior using numerical integration.
+For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that for any given nominal significance level $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$, should be $1 - \alpha$.
 
-<!-- In models~\hyperlink{I1}{I1}-\hyperlink{I5}{I5} choosing a Gaussian distribution for $\beta_0$ ensures that the latent random variables are Gaussian and the model is a LGM. -->
-<!-- This facilitated use of the integrated nested Laplace approximation (INLA) algorithm [@rue2009approximate] for inference, as implemented by the \texttt{R-INLA} package. -->
-<!-- INLA is a deterministic method for approximate Bayesian inference based upon a combination of numerical integration and the Laplace approximation. -->
-<!-- INLA is significantly faster than MCMC and has been shown to have comparable accuracy for LGMs in realistic, pre-asymptotic settings. -->
-<!-- <!-- Although there are particular conditions under which INLA is not accurate [@rue2017bayesian] they do not apply here, as we illustrate in Appendix \ref{app:inlamcmc} where we compare INLA to MCMC using a large number of iterations.
---> -->
+# Funding {-}
 
-<!-- ## Model assessment -->
+AH, HZ were supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).
+AH, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661).
+YX was supported by the Mary Lister McCammon Summer Research Fellowship 2019.
+HZ was supported by Cervest Limited.
+JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664).
+SF was supported by the EPSRC (EP/V002910/1).
+This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth \& Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union.
 
-<!-- We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, subject to calibration. -->
-<!-- In the following two sections we describe our approach to model assessment. -->
-<!-- % Analogous approaches may be used to check recovery of the intercept parameter $\beta_0$, precision parameter $\tau_\phi$, and, when relevant, the length-scale $l$. -->
+For the purpose of Open Access, the authors have applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission.
 
-<!-- ### Continuous ranked probability score -->
+# Acknowledgements {-}
 
-<!-- The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts.
--->
-<!-- For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, \y)$ and $\omega_i$ be the true value. -->
-<!-- Writing $F$ as the cumulative distribution corresponding to $f$ then -->
-<!-- \begin{equation} -->
-<!-- \text{CRPS}(f, \omega_i) = \int_{-\infty}^{\infty} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i = \int_{0}^{1} (F(\rho_i) - \mathbbm{1} \{\rho_i \geq \omega_i \} )^2 \text{d}\rho_i, \label{eq:crps} -->
-<!-- \end{equation} -->
-<!-- where $\mathbbm{1}$ denotes the indicator function and the second equality follows from $0 \leq \rho_i \leq 1$. -->
-<!-- The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$. -->
-<!-- We use the CRPS because it is a strictly proper scoring rule [@gneiting2007strictly]. -->
-<!-- This ensures that the (uniquely) best approach to minimising the score is to truthfully report probabilistic beliefs, which here correspond to the posterior distribution. -->
-<!-- The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, but in line with the findings of @kruger2020predictive, using kernel density estimation we found it difficult to compute satisfactorily.
--->
-<!-- In contrast, given a collection of Monte Carlo samples $\{\rho_i^s\}_{s = 1}^S \sim f$, Equation~\ref{eq:crps} may easily be approximated \citep[Equation 37]{sigrist2015stochastic} by -->
-<!-- \begin{equation} -->
-<!-- \text{CRPS}(f, \omega_i) \approx \frac{1}{S} \sum_{s = 1}^S | \rho_i^s - \omega_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | \rho_i^s - \rho_i^l |. \label{eq:crpsapprox} -->
-<!-- \end{equation} -->
-<!-- We obtained these samples using the function \texttt{INLA::inla.posterior.sample} for models fit with \texttt{R-INLA} and directly from the Markov chain using \texttt{rstan::extract} for models fit with Stan.
--->
+AH thanks Robert Ashton for help with \textsf{R} package development, and Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for providing feedback on earlier drafts of this manuscript.
 
-<!-- \subsubsection{Coverage} -->
+# Disclaimer {-}
 
-<!-- To assess the coverage of our estimates, we considered probability integral transform (PIT) histograms and empirical cumulative distribution function (ECDF) difference plots [@sailynoja2021graphical].
--->
-<!-- We estimated the quantile $q_i = F(\omega_i)$ of the true value within the marginal posterior using numerical integration. -->
-<!-- For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that for any given nominal significance level $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$, should be $1 - \alpha$.
--->
+The findings and conclusions in this manuscript are those of the authors and do not necessarily represent the official position of the funding agencies.
 
-<!-- ## Funding -->
+\clearpage
 
-<!-- AH, HZ were supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).  -->
-<!-- AH, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661). -->
-<!-- YX was supported by the Mary Lister McCammon Summer Research Fellowship 2019. -->
-<!-- HZ was supported by Cervest Limited. -->
-<!-- JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664).  -->
-<!-- SF was supported by the EPSRC (EP/V002910/1). -->
-<!-- This research was supported by the MRC Centre for Global Infectious Disease Analysis (MR/R015600/1), jointly funded by the UK Medical Research Council (MRC) and the UK Foreign, Commonwealth \& Development Office (FCDO), under the MRC/FCDO Concordat program and is also part of the EDCTP2 programme supported by the European Union.
--->
-
-<!-- For the purpose of Open Access, the authors have applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission. -->
-
-<!-- ## Acknowledgements -->
-
-<!-- AH thanks Robert Ashton for help with \textsf{R} package development, and Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for providing feedback on earlier drafts of this manuscript.
--->
-
-<!-- ## Disclaimer -->
-
-<!-- The findings and conclusions in this manuscript are those of the authors and do not necessarily represent the official position of the funding agencies. -->
-
-<!-- \clearpage -->
-
-<!-- # References {#references .unnumbered} -->
+# References {#references .unnumbered}

### 2022-07-13 make/0_all.R

@@ -4,19 +4,19 @@ source("make/utils.R")
 #' []: not yet run
 #' [/]: tried to run but gave me issues
 
-run_commit_push("0_check_bin-gaussian") #' [x]
-run_commit_push("0_check_ik-compute") #' [x]
-run_commit_push("0_check_ik-converge") #' [x]
+run_commit_push("0_check_bin-gaussian")       #' [x]
+run_commit_push("0_check_ik-compute")         #' [x]
+run_commit_push("0_check_ik-converge")        #' [x]
 run_commit_push("0_check_xbinomial-sampling") #' [x]
-run_commit_push("0_demo_areal-kernels") #' [x]
-run_commit_push("0_explore_hrsl-sampling") #' []
-run_commit_push("0_explore_inla-spde") #' []
-run_commit_push("0_explore_unequal-samples") #' []
-run_commit_push("0_explore_variogram-method") #' []
-run_commit_push("0_explore_wilson-pointless") #' []
-run_commit_push("0_plot_areal-data") #' [x]
-run_commit_push("0_plot_sloo-cv") #' [x]
-run_commit_push("0_plot_geometry-graph") #' [x]
-run_commit_push("0_plot_integration-method") #' [x]
-run_commit_push("0_plot-tikz_besag") #' [/]
-run_commit_push("0_plot-tikz_maup") #' [x]
+run_commit_push("0_demo_areal-kernels")       #' [x]
+run_commit_push("0_explore_hrsl-sampling")    #' [ ]
+run_commit_push("0_explore_inla-spde")        #' [ ]
+run_commit_push("0_explore_unequal-samples")  #' [ ]
+run_commit_push("0_explore_variogram-method") #' [ ]
+run_commit_push("0_explore_wilson-pointless") #' [ ]
+run_commit_push("0_plot_areal-data")          #' [x]
+run_commit_push("0_plot_sloo-cv")             #' [x]
+run_commit_push("0_plot_geometry-graph")      #' [x]
+run_commit_push("0_plot_integration-method")  #' [x]
+run_commit_push("0_plot-tikz_besag")          #' [/]
+run_commit_push("0_plot-tikz_maup")           #' [x]

### 2022-07-13 make/1_assess-marginals.R

@@ -1,6 +1,6 @@
 source("make/utils.R")
 
-run_commit_push("1_assess_rho-marginals") #' [x]
-run_commit_push("1_assess_intercept-marginal") #' [x]
-run_commit_push("1_assess_lengthscale-marginal") #' []
-run_commit_push("1_assess_time-taken") #' [/] (It's not really a marginal)
+run_commit_push("1_assess_rho-marginals")        #' [x]
+run_commit_push("1_assess_intercept-marginal")   #' [x]
+run_commit_push("1_assess_lengthscale-marginal") #' [ ]
+run_commit_push("1_assess_time-taken")           #' [/]

### 2022-07-13 make/1_create-plots.R

@@ -1,7 +1,3 @@
 source("make/utils.R")
 
-run_commit_push("plot_areal-data")
-run_commit_push("plot_geometry-graph")
-run_commit_push("plot-tikz_maup")
-run_commit_push("plot_integration-method")
-run_commit_push("plot_simulation-geometries")
+run_commit_push("1_plot_simulation-geometries") #' [x]

### 2022-07-13 make/1_fit-models.R

@@ -2,15 +2,15 @@ source("make/utils.R")
 
 #' Vignette and realistic together
 run_commit_push("1_fit_constant-inla") #' [x]
-run_commit_push("1_fit_iid-inla") #' [x]
-run_commit_push("1_fit_besag-inla") #' [x]
-run_commit_push("1_fit_bym2-inla") #' [/]
-run_commit_push("1_fit_fck-inla") #' [x]
-run_commit_push("1_fit_fik-inla") #' [x]
-run_commit_push("1_fit_ck-stan") #' []
-run_commit_push("1_fit_ik-stan") #' []
+run_commit_push("1_fit_iid-inla")      #' [x]
+run_commit_push("1_fit_besag-inla")    #' [x]
+run_commit_push("1_fit_bym2-inla")     #' [/]
+run_commit_push("1_fit_fck-inla")      #' [x]
+run_commit_push("1_fit_fik-inla")      #' [x]
+run_commit_push("1_fit_ck-stan")       #' [ ]
+run_commit_push("1_fit_ik-stan")       #' [ ]
 
-#' #' Vignette
+#' #' Vignette alone
 #' run_commit_push("1_fit_constant-inla", param = list(realistic = FALSE))
 #' run_commit_push("1_fit_iid-inla", param = list(realistic = FALSE))
 #' run_commit_push("1_fit_besag-inla", param = list(realistic = FALSE))
@@ -18,7 +18,7 @@ run_commit_push("1_fit_ik-stan") #' []
 #' run_commit_push("1_fit_fck-inla", param = list(realistic = FALSE))
 #' run_commit_push("1_fit_fik-inla", param = list(realistic = FALSE))
 #'
-#' #' Realistic
+#' #' Realistic alone
 #' run_commit_push("1_fit_constant-inla", param = list(vignette = FALSE))
 #' run_commit_push("1_fit_iid-inla", param = list(vignette = FALSE))
 #' run_commit_push("1_fit_besag-inla", param = list(vignette = FALSE))

### 2022-07-13 make/1_post-process.R

@@ -1,7 +1,7 @@
 source("make/utils.R")
 
-run_commit_push("1_process_metric-tables")
-run_commit_push("1_plot_metric-maps") #' [x]
-run_commit_push("1_plot_metric-boxplots") #' [x]
-run_commit_push("1_plot_coverage")
-run_commit_push("1_plot_lengthscale-recovery")
+run_commit_push("1_process_metric-tables")     #' [ ]
+run_commit_push("1_plot_metric-maps")          #' [x]
+run_commit_push("1_plot_metric-boxplots")      #' [x]
+run_commit_push("1_plot_coverage")             #' [x]
+run_commit_push("1_plot_lengthscale-recovery") #' [ ]

### 2022-07-13 make/1_simulate-data.R

@@ -1,5 +1,6 @@
 source("make/utils.R")
 
 run_commit_push("1_draw_vignette-geometries") #' [x]
-run_commit_push("1_sim_vignette-geometries") #' [x]
+run_commit_push("1_sim_vignette-geometries")  #' [x]
 run_commit_push("1_sim_realistic-geometries") #' [x]
+run_commit_push("1_plot_simulated-data")      #' [ ]

### 2022-07-13 src/1_plot_coverage/script.R

@@ -1,6 +1,6 @@
 #' Uncomment and run the two line below to resume development of this script
-# orderly::orderly_develop_start("plot_coverage")
-# setwd("src/plot_coverage")
+# orderly::orderly_develop_start("1_plot_coverage")
+# setwd("src/1_plot_coverage")
 
 df <- readRDS("depends/df_rho.rds") %>%
   mutate(replicate = as.numeric(replicate)) %>%
@@ -9,7 +9,7 @@ df <- readRDS("depends/df_rho.rds") %>%
 
 geometries <- unique(df$geometry)
 
-pdf("histogram-ecdf-no-constant.pdf", h = 8, w = 6.25)
+pdf("histogram-ecdf.pdf", h = 8, w = 6.25)
 
 lapply(geometries, function(x) {
   df <- filter(df, geometry == as.character(x))

### 2022-07-13 src/1_plot_metric-boxplots/functions.R

@@ -18,7 +18,7 @@ dark_palette <- c(darkblue, darkgreen, darkpink)
 
 cbpalette <- c("#56B4E9", "#009E73", "#E69F00", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")
 
-boxplot <- function(df, metric, title = NULL, y_lab = NULL, labs = FALSE, remove_constant = FALSE) {
+boxplot <- function(df, metric, y_lab = NULL, remove_constant = FALSE) {
 
   .calc_boxplot_stat <- function(x) {
     coef <- 1.5
@@ -44,8 +44,8 @@ boxplot <- function(df, metric, title = NULL, y_lab = NULL, labs = FALSE, remove
   ggplot(df, aes(x = inf_model, y = .data[[metric]])) +
     stat_summary(fun.data = .calc_boxplot_stat, geom = "boxplot", width = 0.5, alpha = 0.9, show.legend = FALSE, fill = lightgrey) +
     stat_summary(fun = "mean", geom = "point", position = position_dodge(width = 0.5), alpha = 0.7, show.legend = FALSE) +
-    facet_wrap(geometry ~ sim_model, scales = "free") +
-    labs(x = "Inferential model", y = y_lab, title = title, fill = "Simulation model") +
+    facet_grid(~sim_model, scales = "free") +
+    labs(x = "Inferential model", y = y_lab, title = df$geometry[1], fill = "Simulation model") +
     theme_minimal() +
     theme(
       axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)

### 2022-07-13 src/1_plot_metric-boxplots/script.R

@@ -6,41 +6,41 @@ df_rho <- readRDS("depends/df_rho.rds")
 df_intercept <- readRDS("depends/df_intercept.rds")
 
 #' CRPS boxplot for rho
-pdf("crps-boxplot-rho.pdf", h = 7, w = 6.25)
+pdf("crps-boxplot-rho.pdf", h = 4, w = 6.25)
 
-df_rho %>%
-  bsae::update_naming() %>%
-  boxplot(
-    metric = "crps",
-    y_lab = "CRPS"
-  )
+# df_rho %>%
+#   bsae::update_naming() %>%
+#   split(.$geometry) %>%
+#   lapply(function(x) {
+#     boxplot(x, metric = "crps", y_lab = "CRPS")
+#   })
 
 df_rho %>%
   filter(inf_model != "constant_inla") %>%
   bsae::update_naming() %>%
-  boxplot(
-    metric = "crps",
-    y_lab = "CRPS"
-  )
+  split(.$geometry) %>%
+  lapply(function(x) {
+    boxplot(x, metric = "crps", y_lab = "CRPS")
+  })
 
 dev.off()
 
 #' CRPS boxplot for the intercept
 pdf("crps-boxplot-intercept.pdf", h = 7, w = 6.25)
 
-df_intercept %>%
-  bsae::update_naming() %>%
-  boxplot(
-    metric = "crps",
-    y_lab = "CRPS"
-  )
+# df_intercept %>%
+#   bsae::update_naming() %>%
+#   split(.$geometry) %>%
+#   lapply(function(x) {
+#     boxplot(x, metric = "crps", y_lab = "CRPS")
+#   })
 
 df_intercept %>%
   filter(inf_model != "constant_inla") %>%
   bsae::update_naming() %>%
-  boxplot(
-    metric = "crps",
-    y_lab = "CRPS"
-  )
+  split(.$geometry) %>%
+  lapply(function(x) {
+    boxplot(x, metric = "crps", y_lab = "CRPS")
+  })
 
 dev.off()

### 2022-07-13 src/1_plot_metric-maps/functions.R

@@ -19,7 +19,7 @@ metric_map <- function(df, metric, sf, remove_constant = FALSE) {
     st_as_sf() %>%
     ggplot(aes(fill = .data[[metric_mean]])) +
     facet_grid(sim_model ~ inf_model) +
-    geom_sf() +
+    geom_sf(size = 0.1, colour = scales::alpha("grey", 0.25)) +
     scale_fill_viridis_c(option = "C") +
     labs(fill = toupper(metric)) +
     scale_y_continuous(sec.axis = sec_axis(~ . , name = "SECOND Y AXIS", breaks = NULL, labels = NULL)) +
@@ -71,7 +71,7 @@ metric_map <- function(df, metric, sf, remove_constant = FALSE) {
   ggplotify::as.ggplot(z)
 }
 
-produce_map <- function(arg1, arg2, arg3) {
+produce_maps <- function(arg1, arg2, arg3) {
   df_rho %>%
     filter(geometry == arg1) %>%
     filter(inf_model != "constant_inla") %>%

### 2022-07-14 src/docs_paper/paper.Rmd
Some HIV study text

@@
 where $x$ is the $n$-dimensional latent field, $\theta$ are the $m$-dimensional parameters, $m < n$, and $g$ is a link function.
-Spatial structure is modelled at the level of the latent field, and the observation model, often either binomial or Poisson, is conditionally independent and identically distributed $p(\y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
+Spatial structure is modelled at the level of the latent field, and the observation model, often either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
 As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $x \sim \mathcal{N}(\mu, \Sigma)$ for the latent field.
@@
 The spatial random effects $\phi$ were generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}).
-We then generated synthetic data $\y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via
+We then generated synthetic data $y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via
 \begin{equation}
@@
 The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts.
-For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, \y)$ and $\omega_i$ be the true value.
+For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, y)$ and $\omega_i$ be the true value.
 Writing $F$ as the cumulative distribution corresponding to $f$ then
@@
 
-\subsubsection{Coverage}
+### Coverage
 
@@
 For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that for any given nominal significance level $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$, should be $1 - \alpha$.
+
+# HIV prevalence study \label{sec:hiv}
+
+We compared model performance as, again varying the spatial random effect specification, in estimating HIV prevalence in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire ($n = 33$), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$).
+As with the simulation study, \textsf{R} code for this study is available from \url{https://github.com/athowes/areal-comparison}.
+
+<!-- Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa. -->
+<!-- Epidemic estimates are required to help plan, implement, and evaluate the success of programmes, as well as allocate limited resources [@hallett2016evaluation; @cuadros2017mapping]. -->
+<!-- Household surveys provide nationally representative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level. -->
+<!-- Sometimes auxiliary covariates can be used to aid estimation, but data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, faces similar measurement difficulties as the HIV indicators themselves.
+-->
+<!-- In light of this fact, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}, unlike other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
+-->
+<!-- These circumstances foreground the sharing information between districts via spatial smoothing in HIV mapping. -->
+
+## Household survey data
+
+We used data from the most recent publicly available Population Health Impact Assessment (PHIA) survey in each country.
+These surveys utilise a complex design, where each individual $k$ in area $i$ has an unequal probability $\pi_{ik}$ of being included in the sample.
+<!-- A two-stage design in which enumeration areas are first drawn from a stratified sample and then households are chosen using equal probability systematic sampling from within each enumeration area, is common.
+-->
+To account for the survey design we used sampling weights $\omega_{ik} = 1/\pi_{ik}$ to adjust the raw data in each district, obtaining the Kish effective sample size $m_i^\star = (\sum_k \omega_{ik})^2 / \sum_k \omega_{ik}^2$ and effective number of cases $y_i^\star$.
+These effective values may be thought of as what would have been observed had the survey been a simple random sample.
+
+## Model structure
+
+We used a similar model structure as Section \ref{sec:simulation}, with a minor alteration to the likelihood.
+The effective number of cases and effective sample size may not be integers, so we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
+The working likelihood under this model for $m^\star_i \geq y^\star_i$ is given by
+\begin{equation}
+p(y_i^\star \, | \, m_i^\star, \rho_i) = 
+\frac{\Gamma(m_i^\star + 1)}{\Gamma(y_i^\star + 1) \Gamma(m_i^\star - y_i^\star + 1)}  \rho_i ^{y_i^\star}  (1 - \rho_i)^{(m_i^\star - y_i^\star)}. \label{eq:xbinomial}
+\end{equation}
+As with the simulation study we linked the prevalence $\rho_i \in [0, 1]$ to $x_i$ via
+\begin{equation}
+\log \left( \frac{\rho_i}{1 - \rho_i} \right) = x_i = \beta_0 + \phi_i, \label{eq:linear}
+\end{equation}
+and again considered the range of models for $\bphi$ given by Table \ref{tab:inferentialmodels}.
+
+## Cross-validation \label{sec:cv}
+
+<!-- \begin{figure} -->
+<!--     \centering -->
+<!-- 	\includegraphics[width=\textwidth]{depends/sloo-cv.pdf} -->
+<!-- 	\caption{The $i$th fold in a SLOO-CV in which the model is fitted on the grey (\lightgreyline) areas $A_{-(i, \delta i)}$ with the blue (\lightblueline) areas $A_{\delta i}$ held-out, to be assessed in predicting the green (\lightgreenline) area $A_i$.} -->
+<!-- 	\label{fig:sloo} -->
+<!-- \end{figure} -->
+
+```{r sloo-cv, fig.align="center", fig.cap="SLOO-CV."}
+knitr::include_graphics("depends/sloo-cv.pdf")
+```
+
+We assessed each model using two cross-validation approaches: (1) LOO-CV, a standard leave-one-out cross-validation, in which the prevalence data from a single district were removed, and (2) SLOO-CV, a spatial leave-one-out cross-validation, in which the data from a single district and each of its neighbouring districts were removed.
+We performed cross-validation manually as we found approximations that avoid refitting the model, such as the built-in conditional predictive ordinate statistic in `R-INLA` package [@held2010posterior] and the Pareto smoothed importance sampling [@vehtari2015pareto] approach implemented in the `loo` package [@vehtari2017practical], to be inaccurate in this setting, perhaps because individual observations in spatial random effect models can exert high influence toward local parameters.
+
+\subsubsection{LOO-CV}
+
+In the $i$th LOO-CV fold, the model was fit using the data $y_{-i}$ and assessed according to its prediction on $y_i$.
+We assessed forecasting performance using the CRPS, at the level of the data $y_i$ rather than at the level of the prevalence $\rho_i$.
+To compute the value of CRPS for the $i$th fold we use samples $\{ y_i^s \}_{s = 1}^S$ from the LOO posterior predictive distribution $p(y_i \, | \, y_{-i})$ giving
+\begin{equation}
+\text{LOO-CRPS}_i = \frac{1}{S} \sum_{s = 1}^S | y_i^s - y_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | y_i^s - y_i |. \label{eq:crpsapprox2}
+\end{equation}
+
+\subsubsection{SLOO-CV}
+
+If structural dependencies, like spatial dependence, exist in the data then results from LOO-CV tend to underestimate predictive error [@roberts2017cross].
+To counteract this effect, we modify the spatial leave-one-out (SLOO) approach of @le2014spatial.
+As illustrated by Figure \ref{fig:sloo}, during the $i$th fold of SLOO-CV we leave-out the block $(i, \delta i)$.
+This increases the extent to which the training and validation sets are conditionally independent.
+The result is a SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$ from which, analogously to Equation \ref{eq:crpsapprox2}, we may compute the corresponding value of $\text{SLOO-CRPS}_i$.
 

### 2022-07-14 src/0_plot_sloo-cv/script.R

@@ -22,7 +22,7 @@ plot <- ggplot(df, aes(fill = as.factor(left_out))) +
   scale_fill_manual(
     values = c(sf_lightgrey, lightblue, lightgreen),
     name = "",
-    labels = c("Fit model to", "Left out", "Predicted on")
+    labels = c("Areas the model is fit to", "Areas left out", "Areas predicted on")
   ) +
   theme_void() +
   theme(

### 2022-07-31 src/docs_paper/paper.Rmd
Iterate manuscript forward

@@
 ---
-title: Models for spatial structure in small-area estimation
+title: Modelling spatial structure in small-area estimation
 author:
@@
 abstract: |
-  **Background:**
-
-  **Methods:**
-
-  **Results:**
-
-  **Conclusions:**
+  |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects.
+However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions, which seem particularly dubious for irregular geometries.
+Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data.
+Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
+  |
+  |  **Methods:** In a comprehensive simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications.
+We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities.
+Finally, we compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
+  |
+  |  **Results:**
+  |
+  |  **Conclusions:**
 
@@
 Use of this model is widespread, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
-However, for irregular geometries like the administrative divisions of a country, the assumptions made about space are unrealistic.
+However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
 We hypothesised that incorporating a spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models.
@@
 -->
+
+Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
+Epidemic estimates are required to help plan, implement, and evaluate the success of programmes, as well as allocate limited resources [@hallett2016evaluation; @cuadros2017mapping].
+Household surveys provide nationally representative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
+Sometimes auxiliary covariates can be used to aid estimation, but data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, faces similar measurement difficulties as the HIV indicators themselves.
+In light of this fact, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}, unlike other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
+These circumstances foreground the importance of sharing information between districts via spatial smoothing in HIV mapping.
 
@@
 
-<!-- We also hypothesised that improvements to the Besag model may be transferable to improving convolution models. -->
-<!-- Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
--->
-<!-- Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models. -->
-<!-- Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive. -->
-<!-- These convolutions use a Besag model as the spatially structured component, though any choice of spatial random effect model is possible. -->
-<!-- The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and more broadly both models are of substantive practical interest.
--->
-
-<!-- @bradley2015spatio develop a framework for spatio-temporal change of support (STCOS) modelling, and a review of its implementation in \textsc{R} is provided by @raim2021spatio. -->
-<!-- Lee find that Leroux model of @leroux2000estimation achieves consistently good performance across a range of spatial correlation settings. -->
-<!-- @riebler2016intuitive reach a similar conclusion about both the Leroux and BYM2 model. -->
+Improvements to the Besag model may be transferable to improving convolution models.
+Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
+Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models.
+Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive.
+The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and more broadly both models are of substantive practical interest.
 
@@
 In Section \ref{sec:simulation} we compare models using simulated data, before applying them to mapping HIV prevalence in sub-Saharan Africa in Section \ref{sec:hiv}.
-Finally, we discuss our conclusions and directions for future research in Section \ref{sec:conclusion}.
+Finally, we discuss our conclusions and directions for future research in Section \ref{sec:discussion}.
+
+<!-- @bradley2015spatio develop a framework for spatio-temporal change of support (STCOS) modelling, and a review of its implementation in \textsc{R} is provided by @raim2021spatio. -->
+<!-- Lee find that Leroux model of @leroux2000estimation achieves consistently good performance across a range of spatial correlation settings. -->
+<!-- @riebler2016intuitive reach a similar conclusion about both the Leroux and BYM2 model. -->
 
@@
 where $x$ is the $n$-dimensional latent field, $\theta$ are the $m$-dimensional parameters, $m < n$, and $g$ is a link function.
-Spatial structure is modelled at the level of the latent field, and the observation model, often either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
+Spatial structure is modelled at the level of the latent field, and the observation model, typically either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
 As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $x \sim \mathcal{N}(\mu, \Sigma)$ for the latent field.
@@
 The latent field in a LGM is comprised a constant mean $\mu$, known also as the fixed effects, and (spatial) random effects $\phi \sim \mathcal{N}(0, \Sigma)$.
-The mean may be modelled using a linear predictor $\mu_i = \beta_0 + z_i^\top \beta$, where $(\beta_0, \beta)$ are regression parameters corresponding to areal covariates^[To denote length $p$ vectors of areal and point covariates, we use $z_i = (z_{i,1}, \ldots, z_{i,p})^\top$ and $z(s) = (z_1(s), \ldots, z_p(s))^\top$ respectively.} $(1, \z_i)^\top$].
+The mean may be modelled using a linear predictor $\mu_i = \beta_0 + z_i^\top \beta$, where $(\beta_0, \beta)$ are regression parameters corresponding to areal covariates^[To denote length $p$ vectors of areal and point covariates, we use $z_i = (z_{i,1}, \ldots, z_{i,p})^\top$ and $z(s) = (z_1(s), \ldots, z_p(s))^\top$ respectively.} $(1, z_i)^\top$].
 Here, we focus on the spatial random effects $\phi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
@@
 Since then, it has seen wider use, including in situations where the spatial structure is less regular.
-This leads to a number of concerns about the model's applicability, which we highlight below.
+There are a number of concerns about the model's applicability to this broader setting, which we highlight below.
 Closely linked are the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups.
@@
 
-<!-- ## BYM2 \label{sec:bym2} -->
+## BYM2 \label{sec:bym2}
 
-<!-- Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model. -->
-<!-- The Besag-York-Molli\'{e} (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\phi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
--->
-<!-- The resulting distribution is -->
-<!-- \begin{equation} -->
-<!--     \phi \sim \N(\mathbf{0}, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} R^{-}) \label{eq:bym}. -->
-<!-- \end{equation} -->
-<!-- Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$. -->
-<!-- However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
--->
-<!-- In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale.
--->
-<!-- Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
--->
-<!-- This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal. -->
-<!-- Building on the models of @leroux2000estimation and @dean2001detecting which tackle this the identifiability problem, but do not scale the spatially structured noise, @simpson2017penalising propose a reparameterisation $(\tau_v, \tau_u) \mapsto (\tau_\phi, \pi)$ of the BYM model known as the BYM2 model and given by -->
-<!-- \begin{align} -->
-<!-- \phi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2} -->
-<!-- \end{align} -->
-<!-- where $\tau_\phi$ is the marginal precision of $\phi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $\bu$ with precision matrix given by the scaled structure matrix $R^\star$ (see~\ref{app:freni}).
--->
-<!-- When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model. -->
-<!-- To borrow an analogy, the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately [@rue2020comment].
--->
+Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model.
+The Besag-York-Mollié (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\phi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
+The resulting distribution is
+\begin{equation}
+    \phi \sim \N(\mathbf{0}, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} R^{-}) \label{eq:bym}.
+\end{equation}
+Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$.
+However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
+In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale.
+Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
+This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal.
+Building on the models of @leroux2000estimation and @dean2001detecting which tackle this the identifiability problem, but do not scale the spatially structured noise, @simpson2017penalising propose a reparameterisation $(\tau_v, \tau_u) \mapsto (\tau_\phi, \pi)$ of the BYM model known as the BYM2 model and given by
+\begin{align}
+\phi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2}
+\end{align}
+where $\tau_\phi$ is the marginal precision of $\phi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $\bu$ with precision matrix given by the scaled structure matrix $R^\star$ (see~\ref{app:freni}).
+When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model.
+To borrow an analogy, the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately [@rue2020comment].
 
@@
 Section \ref{sec:ab} reviewed ways to construct spatial random effect precision matrices using an adjacency relation.
-Another approach is to define the covariance matrix using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$, where $\mathcal{P}$ denotes the power set such that $\mathcal{P}(\Sc)$ is the space of subsets of the study region, which gives a measure of similarity between two areas.
+Another approach is to define the covariance matrix using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$ which gives a measure of similarity between two areas, where $\mathcal{P}$ denotes the power set and $\mathcal{P}(\Sc)$ is the space of subsets of the study region.
 If $K$ is positive semi-definite, then we define areal kernel spatial random effects by
@@
 where the $n \times n$ Gram matrix $K$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix.
-We keep $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.
+Here, we place $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.
 
 Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$.
-To construct $K$ we use a simple method, namely averaging the kernel $k$ computed on some collection of points from within each area.
-As such, in the multiple instance learning literature $K$ would be described as a type of kernels on sets [@gartner2002multi].
+We use a simple method to construct $K$ from $k$, namely averaging the kernel $k$ computed on some collection of points from within each area.
+As such, $K$ is described as a kernel on sets in the multiple instance learning literature [@gartner2002multi].
 Although it may be possible to learn the method for combining kernels from data [@gonen2011multiple], we do believe that this is feasible in the small-area estimation setting, and instead hope to construct areal kernels which represent our prior beliefs about the spatial process.
@@
 
-The simplest approach is to use a single point, which one might hope would be representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude.
+The simplest approach is to use a single point, especially one which is representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude.
 This results in the centroid kernel
@@
 \end{equation}
-% More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation.
-Usage of the centroid kernel includes @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both using the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$.
+<!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
+The centroid kernel has been used by @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$.
 The former fix the length-scale $l$ using a semi-variogram and the later place a prior on the inverse length-scale.
-In their simulation experiment, \citet[Section 3]{best2005comparison} used data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas (though it is unsurprising that a stationary covariance structure would struggle to recover non-stationary spatial structure).
+In their simulation experiment, \citet[Section 3]{best2005comparison} used data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas, but it is not suprising that a stationary covariance structure would struggle to recover non-stationary spatial structure).
 More recently, @teh2021efficient used a centroid kernel as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
@@
 
-<!-- Rather than attempting to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest. -->
-<!-- This results in the integrated kernel -->
-<!-- \begin{equation} -->
-<!-- K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s', \label{eq:ik1} -->
-<!-- \end{equation} -->
-<!-- which is equivalent to the covariance structure obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition. -->
-<!-- Various authors [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi] in the machine learning literature have recently contributed to the development of these aggregated Gaussian process models.
--->
+Rather than attempting to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
+This results in the integrated kernel
+\begin{equation}
+K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s', \label{eq:ik1}
+\end{equation}
+which is equivalent to the covariance structure obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition.
+Models of this sort have been studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi; @2021arXiv210501460Z].
 <!-- % For example, @law2018variational develop a variational Bayesian inference method for aggregated observation models and @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region.
 -->
-<!-- Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out.
--->
+Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out.
 
-<!-- Considering the aggregation of point data is particularly natural within the log-Gaussian Cox Process framework [@diggle2013spatial]. -->
-<!-- A Cox process is an inhomogeneous Poisson process with a continuous stochastic intensity function $\{ x(s), s \in \Sc \}$ such that conditional on the realisation of $x(s)$ the number of points in any area $A_i$ follows a Poisson distribution.
--->
-<!-- The rate parameter of this Poisson distribution is explicitly aggregated as follows -->
-<!-- \begin{equation} -->
-<!-- y_i \, | \, x(s) \sim \text{Poisson} \left(\int_{s \in A_i} x(s) \text{d}s \right). -->
-<!-- \end{equation} -->
-<!-- In a LGCP the log intensity $\log x(s) = \eta(s)$ is modelled using a Gaussian process prior $\eta(s) \sim \mathcal{GP}(\mu(s), k(s, s'))$. -->
-<!-- @johnson2019spatially obtain Equation~\ref{eq:ik2} by considering a discrete Poisson log-linear mixed model approximation to a continuous LGCP, whereby $\eta(s)$ is approximated by a piecewise constant $\eta_i = \mu_i + \phi_i$ in each area $A_i$.\footnote{@johnson2019spatially use points generated from a class of inhibition processes which combine sequential inhibition with rejection sampling to evaluate the integral.} -->
-<!-- The $i$th discrete spatial random effect is then $\phi_i = \int_{A_i} w_i(s) \phi(s) \text{d}s$, -->
-<!-- with covariance structure -->
-<!-- \begin{equation} -->
-<!-- \text{Cov} \left( \int_{A_i} w_i(s) \phi(s) \text{d}s, \int_{A_j} w_j(s') \phi(s') \text{d}s' \right) -->
-<!-- = \int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d}s\text{d}s', -->
-<!-- \end{equation} -->
-<!-- corresponding to an areal integrated kernel with a logarithmic link function and Poisson likelihood. -->
+The log-Gaussian Cox Process framework [@diggle2013spatial] also naturally arrives at the integrated kernel formulation.
+A Cox process is an inhomogeneous Poisson process with a continuous stochastic intensity function $\{ x(s), s \in \Sc \}$ such that conditional on the realisation of $x(s)$ the number of points in any area $A_i$ follows a Poisson distribution.
+The rate parameter of this Poisson distribution is explicitly aggregated as follows
+\begin{equation}
+y_i \, | \, x(s) \sim \text{Poisson} \left(\int_{s \in A_i} x(s) \text{d}s \right).
+\end{equation}
+In a LGCP the log intensity $\log x(s) = \eta(s)$ is modelled using a Gaussian process prior $\eta(s) \sim \mathcal{GP}(\mu(s), k(s, s'))$.
+@johnson2019spatially obtain Equation \ref{eq:ik2} by considering a discrete Poisson log-linear mixed model approximation to a continuous LGCP, whereby $\eta(s)$ is approximated by a piecewise constant $\eta_i = \mu_i + \phi_i$ in each area $A_i$^[@johnson2019spatially use points generated from a class of inhibition processes which combine sequential inhibition with rejection sampling to evaluate the integral.].
+The $i$th discrete spatial random effect is then $\phi_i = \int_{A_i} w_i(s) \phi(s) \text{d}s$,
+with covariance structure
+\begin{equation}
+\text{Cov} \left( \int_{A_i} w_i(s) \phi(s) \text{d}s, \int_{A_j} w_j(s') \phi(s') \text{d}s' \right)
+= \int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d}s\text{d}s',
+\end{equation}
+corresponding to an areal integrated kernel with a logarithmic link function and Poisson likelihood.
 
-<!-- One benefit of integration is that additional information accounting for heterogeneity over $A_i$ may be incorperated into the model if it is available. -->
-<!-- This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure. -->
-<!-- The weighted integrated kernel is given by -->
-<!-- \begin{equation} -->
-<!-- K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2} -->
-<!-- \end{equation} -->
-<!-- This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, and utilising a high resolution measure of population density as the weighting distribution would account for this.
--->
+Disaggregation regression (also known as downscaling or interpolation) is another closely related approach which, rather than using the aggregate nature of areal observations as primarily a route towards better area-level estimates, as is our focus, aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
+<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
+<!-- Disaggregation is challenging without auxiliary information. -->
 
-<!-- If areal kernel spatial random effects are created using the integrated kernel, then aggregation occurs at the level of the latent field in the three stage model rather than at the level of the data.
--->
-<!-- This corresponds to a generative model of the form $y_i \sim p(y_i \, | \,  g^{-1}(x_i))$, with $x_i = |A_i|^{-1}\int_{A_i} x(s) \text{d} s$. -->
-<!-- If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data. -->
-<!-- On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model mismatches the proposed data generating process $y_i = |A_i|^{-1}\int_{A_i} y(s) \text{d}s$.
--->
+A benefit of the integration approach is that, if available, additional information accounting for heterogeneity over $A_i$ may be incorperated into the model.
+This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
+\begin{equation}
+K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2}
+\end{equation}
+This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, which could be accounted for by weighting according to a high resolution measure of population density.
 
-<!-- Most of the time we do not expect to be able to calculate the integral in Equation~\ref{eq:ik2} analytically. -->
-<!-- Instead, given $n$ collections of $L_i$ samples $\{s^{(i)}_l \}_{l = 1}^{L_i} \sim \mathcal{U}(A_i)$ drawn uniformly from each area then the integral may be approximated using Monte Carlo integration by the double sum -->
-<!-- \begin{equation} -->
-<!-- K(A_i, A_j) \approx \frac{1}{L_i L_j} \sum_{l = 1}^{L_i} \sum_{m = 1}^{L_j}  -->
-<!-- w_i \left( s^{(i)}_l \right) w_j \left( s^{(j)}_m \right)  -->
-<!-- k \left( s^{(i)}_l, s^{(j)}_m \right). \label{eq:mcintegratedkernel} -->
-<!-- \end{equation} -->
-<!-- Equivalently, samples drawn from $W_i$ may be used unweighted by $w_i(s)$. -->
-<!-- Integration points may also be selected by some deterministic process, and used in Equation~\ref{eq:mcintegratedkernel} to give a numerical integration estimate of the kernel. -->
-<!-- As with learning the way in which the kernels should be combined, in more data rich settings it may be possible to learn the integration points [@campbell2019automated]. -->
-<!-- Taking any of these approaches requires $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$. -->
-<!-- This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or integration points it is feasible to use.
--->
-
-<!-- Another closely linked approach is disaggregation regression (also known as downscaling or interpolation) which, rather than using the aggregate nature of areal observations as just a route towards better area-level estimates as is our focus, aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
--->
-<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates.  -->
-
-<!-- Disaggregation is challenging without auxiliary information. -->
-<!-- Kelsall and Wakefield They use the cubic correlation function and, to reduce the number of Gram matrix constructions and inversions, learn a range parameter by placing a discrete prior. -->
-<!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. -->
+If areal kernel spatial random effects are created using the integrated kernel, then aggregation occurs at the level of the latent field in the three stage model rather than at the level of the data.
+This corresponds to a generative model of the form $y_i \sim p(y_i \, | \,  g^{-1}(x_i))$, with $x_i = |A_i|^{-1}\int_{A_i} x(s) \text{d} s$.
+If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data.
+On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model does not match the proposed data generating process $y_i = |A_i|^{-1}\int_{A_i} y(s) \text{d}s$.
 <!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. -->
@@
 
+Most of the time we do not expect to be able to calculate the integral in Equation \ref{eq:ik2} analytically.
+Instead, given $n$ collections of $L_i$ samples $\{s^{(i)}_l \}_{l = 1}^{L_i} \sim \mathcal{U}(A_i)$ drawn uniformly from each area then the integral may be approximated using Monte Carlo by the double sum
+\begin{equation}
+K(A_i, A_j) \approx \frac{1}{L_i L_j} \sum_{l = 1}^{L_i} \sum_{m = 1}^{L_j}
+w_i \left( s^{(i)}_l \right) w_j \left( s^{(j)}_m \right)
+k \left( s^{(i)}_l, s^{(j)}_m \right). \label{eq:mcintegratedkernel}
+\end{equation}
+Equivalently, samples drawn from $W_i$ may be used without weighting by $w_i(s)$.
+Integration points may also be selected by some deterministic process, and used in Equation \ref{eq:mcintegratedkernel} to give a numerical integration estimate of the kernel.
+As with learning the way in which the kernels should be combined, in more data rich settings it may be possible to learn the integration points [@campbell2019automated].
+Taking any of these approaches requires $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
+This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or integration points it is feasible to use.
+To make inference more feasible, @kelsall2002modeling reduce the number of Gram matrix constructions and inversions required, by using a discrete hyperparameter prior.
+
+<!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. -->
+
 # Simulation study \label{sec:simulation}
 
-```{r simulation-geometries, fig.align="center", fig.cap="Simulation geometries."}
-knitr::include_graphics("depends/simulation-geometries.pdf")
-```
+\begin{figure}
+\centering
+\includegraphics{depends/simulation-geometries.pdf}
+\caption{Simulation geometries.}
+\label{fig:51}
+\end{figure}
 
@@
 We fixed the sample sizes to $m_i = 25$ for all $i$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$.
-We used seven different geometries: the four vignette geometries in Figure \ref{fig:tikz1}, which share an adjacency graph, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of C\^{o}te d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
+We used seven different geometries: the four vignette geometries in Figure \ref{fig:tikz1}, which share an adjacency graph, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of Côte d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
 These more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test how model performance varies by geometry regularity.
@@
 
-We fit eight inferential models, each of which differing in the spatial random effect specification (Table \ref{tab:inferentialmodels}), allowing us to test sensitivity to spatial structure misspecification.
+We fit eight inferential models, each of which differing in the spatial random effect specification (Table \ref{tab:inferentialmodels}).
 
@@
 For the precision, we placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
-The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one, neither of which we assumed to reflect reality.
+The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one, neither of which we assumed to be plausible.
 A weakly informative $\mathcal{N}(-2, 1)$ prior was placed on $\beta_0$, setting most of the prior probability density for $\rho_i$ within a range $[0, 0.25]$ typical for a disease prevalence.
@@
 
-Following @stein1999interpolation, for the models~\hyperlink{I5}{I5}-\hyperlink{I8}{I8} we used the Mat\'{e}rn kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
+Following @stein1999interpolation, for the areal kernel models we used the Matérn kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
 \begin{equation}
@@
 \frac{1}{2^{\nu - 1}\Gamma(\nu)} \left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right)^\nu
-K_\nu\left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right), \label{eq:matern}
+B_\nu\left(\frac{\sqrt{2\nu}\lvert s - s' \rvert}{l}\right), \label{eq:matern}
 \end{equation}
-where $K_\nu$ is the modified Bessel function of the second kind, $|s - s'|$ is the Euclidean distance between $s$ and $s'$, $\nu$ is the smoothness hyperparameter and $l$ is the length-scale hyperparameter on the latitude-longitude scale.
-We fixed the smoothness parameter, which is typically unidentifiable from data, to be $3/2$, matching that used in simulation model~\hyperlink{S3}{S3}, and convenient in that it simplifies Equation~\ref{eq:matern} to be the product of a polynomial and an exponential
+where $B_\nu$ is the modified Bessel function of the second kind, $|s - s'|$ is the Euclidean distance between $s$ and $s'$, $\nu$ is the smoothness hyperparameter and $l$ is the length-scale hyperparameter on the latitude-longitude scale.
+We fixed the smoothness parameter, which is otherwise typically unidentifiable from data, to be $3/2$, matching that used in simulation model \hyperlink{S3}{S3}, and convenient in that it simplifies Equation \ref{eq:matern} to be
 \begin{equation}
@@
 \end{equation}
-Taking a similar approach to @best1999bayesian, in models~\hyperlink{I5}{I5} and \hyperlink{I7}{I7} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data.
-In models~\hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
-For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51} (though the true number of integration points may be 1-2 above or below 10, as `sf::st_sample` with `type = "hexagonal"` is not guaranteed to be exact).
+Taking a similar approach to @best1999bayesian, in models \hyperlink{I5}{I5} and \hyperlink{I7}{I7} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data.
+In models \hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
+For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51}.
+The true number of integration points may be 1-2 above or below 10, as `sf::st_sample` with `type = "hexagonal"` is not guaranteed to be exact.
 We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}.
-For Model~\hyperlink{I6}{I6} although it is feasible to use a much larger number of integration points, since the Gram matrix is only computed once, we chose not to for comparability.
+For Model \hyperlink{I6}{I6} although it is feasible to use a much larger number of integration points, since the Gram matrix is only computed once, we chose not to for comparability.
 
@@
 
-In models~\hyperlink{I1}{I1}-\hyperlink{I5}{I5} choosing a Gaussian distribution for $\beta_0$ ensures that the latent random variables are Gaussian and the model is a LGM.
-This facilitated use of the integrated nested Laplace approximation (INLA) algorithm [@rue2009approximate] for inference, as implemented by the \texttt{R-INLA} package.
+In models \hyperlink{I1}{I1}-\hyperlink{I5}{I5} choosing a Gaussian distribution for $\beta_0$ ensures that the latent random variables are Gaussian and the model is a LGM, facilitating use of the integrated nested Laplace approximation (INLA) algorithm [@rue2009approximate] for inference, as implemented by the \texttt{R-INLA} package.
 INLA is a deterministic method for approximate Bayesian inference based upon a combination of numerical integration and the Laplace approximation.
@@
 We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, subject to calibration.
-In the following two sections we describe our approach to model assessment.
+We describe our approach to model assessment in the following sections.
 <!-- Analogous approaches may be used to check recovery of the intercept parameter $\beta_0$, precision parameter $\tau_\phi$, and, when relevant, the length-scale $l$. -->
@@
 
-The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts.
+The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance, and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts.
 For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, y)$ and $\omega_i$ be the true value.
@@
 The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$.
-We use the CRPS because it is a strictly proper scoring rule [@gneiting2007strictly].
-This ensures that the (uniquely) best approach to minimising the score is to truthfully report probabilistic beliefs, which here correspond to the posterior distribution.
-The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, but in line with the findings of @kruger2020predictive, using kernel density estimation we found it difficult to compute satisfactorily.
+The CRPS is a strictly proper scoring rule [@gneiting2007strictly], which ensures that the (uniquely) best approach to minimising the score is to truthfully report probabilistic beliefs, here corresponding to the posterior distribution.
+The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
 In contrast, given a collection of Monte Carlo samples $\{\rho_i^s\}_{s = 1}^S \sim f$, Equation~\ref{eq:crps} may easily be approximated \citep[Equation 37]{sigrist2015stochastic} by
@@
 As with the simulation study, \textsf{R} code for this study is available from \url{https://github.com/athowes/areal-comparison}.
-
-<!-- Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa. -->
-<!-- Epidemic estimates are required to help plan, implement, and evaluate the success of programmes, as well as allocate limited resources [@hallett2016evaluation; @cuadros2017mapping]. -->
-<!-- Household surveys provide nationally representative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level. -->
-<!-- Sometimes auxiliary covariates can be used to aid estimation, but data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, faces similar measurement difficulties as the HIV indicators themselves.
--->
-<!-- In light of this fact, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}, unlike other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
--->
-<!-- These circumstances foreground the sharing information between districts via spatial smoothing in HIV mapping. -->
 
@@
 
-<!-- \begin{figure} -->
-<!--     \centering -->
-<!-- 	\includegraphics[width=\textwidth]{depends/sloo-cv.pdf} -->
-<!-- 	\caption{The $i$th fold in a SLOO-CV in which the model is fitted on the grey (\lightgreyline) areas $A_{-(i, \delta i)}$ with the blue (\lightblueline) areas $A_{\delta i}$ held-out, to be assessed in predicting the green (\lightgreenline) area $A_i$.} -->
-<!-- 	\label{fig:sloo} -->
-<!-- \end{figure} -->
-
-```{r sloo-cv, fig.align="center", fig.cap="SLOO-CV."}
-knitr::include_graphics("depends/sloo-cv.pdf")
-```
+\begin{figure}
+    \centering
+	\includegraphics[width=\textwidth]{depends/sloo-cv.pdf}
+	\caption{The $i$th fold in a SLOO-CV in which the model is fitted on the grey areas $A_{-(i, \delta i)}$ with the blue areas $A_{\delta i}$ held-out, to be assessed in predicting the green area $A_i$.}
+	\label{fig:sloo}
+\end{figure}
 
@@
 
-\subsubsection{LOO-CV}
+### Leave-one-out cross-validation
 
@@
 
-\subsubsection{SLOO-CV}
+### Spatial leave-one-out cross-validation
 
@@
 The result is a SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$ from which, analogously to Equation \ref{eq:crpsapprox2}, we may compute the corresponding value of $\text{SLOO-CRPS}_i$.
+
+# Discussion \label{sec:discussion}
 

### 2022-07-31 src/docs_paper/preamble.tex
Work on manuscript and simulation study
Iterate manuscript forward

@@
 \RequirePackage{pifont}
+\RequirePackage[margin=1in]{geometry}
 \usepackage{footnote}
@@
 \newcommand{\xmark}{\ding{55}}
+
+\definecolor{lightblue}{HTML}{56B4E9}
+\definecolor{lightgreen}{HTML}{009E73}
+\definecolor{lightorange}{HTML}{E69F00}
+\definecolor{lightgrey}{HTML}{E6E6E6}
+
+\newcommand{\lightblueline}{\raisebox{2pt}{\tikz{\draw[-,lightblue,solid,line width = 3pt](0,0) -- (2mm,0);}}}
+\newcommand{\lightgreenline}{\raisebox{2pt}{\tikz{\draw[-,lightgreen,solid,line width = 3pt](0,0) -- (2mm,0);}}}
+\newcommand{\lightorangeline}{\raisebox{2pt}{\tikz{\draw[-,lightorange,solid,line width = 3pt](0,0) -- (2mm,0);}}}
+\newcommand{\lightgreyline}{\raisebox{2pt}{\tikz{\draw[-,lightgrey,solid,line width = 3pt](0,0) -- (2mm,0);}}}

### 2022-07-31 src/docs_paper/paper.Rmd
Upload manuscript

@@
 
-<!-- \begin{figure} -->
-<!-- \centering -->
-<!-- \includegraphics[width=\textwidth]{depends/integration-method.pdf} -->
-<!-- \caption{Three ways to choose integration points with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample} [@pebesma2018sf].} -->
-<!-- \label{fig:41} -->
-<!-- \end{figure} -->
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{depends/integration-method.pdf}
+\caption{Three ways to choose integration points with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample} [@pebesma2018sf].}
+\label{fig:41}
+\end{figure}
 
@@
 
-We compared model performance as, again varying the spatial random effect specification, in estimating HIV prevalence in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire ($n = 33$), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$).
+We compared model performance in estimating HIV prevalence in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire ($n = 33$), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$), again varying the spatial random effect specification.
 As with the simulation study, \textsf{R} code for this study is available from \url{https://github.com/athowes/areal-comparison}.

### 2022-07-31 make/2_cv-models.R

@@ -1,12 +1,12 @@
 source("make/utils.R")
 
 run_commit_push("2_cv_constant-inla") #' [x]
-run_commit_push("2_cv_iid-inla") #' [x]
-run_commit_push("2_cv_besag-inla") #' [x]
-run_commit_push("2_cv_bym2-inla") #' [x]
-run_commit_push("2_cv_fck-inla") #' [/] See issue 35
-run_commit_push("2_cv_fik-inla") #' [/] See issue 36
-run_commit_push("2_cv_ck-stan") #' []
-run_commit_push("2_cv_ik-stan") #' []
+run_commit_push("2_cv_iid-inla")      #' [x]
+run_commit_push("2_cv_besag-inla")    #' [x]
+run_commit_push("2_cv_bym2-inla")     #' [x]
+run_commit_push("2_cv_fck-inla")      #' [/] See issue 35
+run_commit_push("2_cv_fik-inla")      #' [/] See issue 36
+run_commit_push("2_cv_ck-stan")       #' []
+run_commit_push("2_cv_ik-stan")       #' []
 
 run_commit_push("2_assess_cv") #' [x]

### 2022-07-31 make/2_post-process.R

@@ -1,6 +1,6 @@
 source("make/utils.R")
 
-run_commit_push("2_process_metric-tables") #' [x]
-run_commit_push("2_plot_metric-maps") #' [x]
-run_commit_push("2_plot_prev-ladder") #' [x]
-run_commit_push("2_process_compare-to-manual-cv") #' Probably supplementary material
+run_commit_push("2_process_metric-tables")        #' [x]
+run_commit_push("2_plot_metric-maps")             #' [x]
+run_commit_push("2_plot_prev-ladder")             #' [x]
+run_commit_push("2_process_compare-to-manual-cv") #' [ ] Probably supplementary material

### 2022-07-31 make/2_pull-surveys.R

@@ -6,10 +6,10 @@ source("make/utils.R")
 #' * tza2017phia
 #' * zwe2016phia
 
-orderly::orderly_pull_archive("civ_data_survey", remote = "naomi2") #' [x]
+orderly::orderly_pull_archive("civ_data_survey", remote = "naomi2")            #' [x]
 orderly::orderly_pull_archive("mwi_data_survey-indicators", remote = "malawi") #' [x]
-orderly::orderly_pull_archive("tza_data_survey", remote = "naomi2") #' [x]
-orderly::orderly_pull_archive("zwe_data_survey", remote = "naomi2") #' [x]
+orderly::orderly_pull_archive("tza_data_survey", remote = "naomi2")            #' [x]
+orderly::orderly_pull_archive("zwe_data_survey", remote = "naomi2")            #' [x]
 
 run_commit_push("2_process_hiv-surveys", push = FALSE) #' [x]
 

### 2022-07-31 utils/archive_to_docs.R

@@ -1,0 +1,19 @@
+#' Set the working directory to the project root
+setwd(rprojroot::find_rstudio_root_file())
+
+archive_to_docs <- function(report) {
+  #' Artefacts to be moved
+  filenames <- yaml::read_yaml(file.path(paste0("src/", report, "/orderly.yml")))$artefacts[[1]]$data$filenames
+
+  #' Latest version in archive
+  latest <- orderly::orderly_latest(report)
+
+  #' Copy files over
+  files_from <- paste0("archive/", report, "/", latest, "/", filenames)
+  files_to <- paste0("docs/", filenames)
+
+  file.copy(from = files_from, to = files_to, overwrite = TRUE)
+}
+
+#' Names of the reports to move
+archive_to_docs("docs_paper")

### 2022-08-03 make/1_post-process.R
Import CRPS results for simulation study straight into paper

@@ -1,6 +1,6 @@
 source("make/utils.R")
 
-run_commit_push("1_process_metric-tables")     #' [ ]
+run_commit_push("1_process_metric-tables")     #' [x]
 run_commit_push("1_plot_metric-maps")          #' [x]
 run_commit_push("1_plot_metric-boxplots")      #' [x]
 run_commit_push("1_plot_coverage")             #' [x]

### 2022-08-03 src/1_process_metric-tables/functions.R

@@ -61,8 +61,7 @@ metric_table <- function(df, metric, title = NULL, subtitle = NULL, latex = FALS
     return(
       table %>%
         as_latex() %>%
-        as.character() %>%
-        cat()
+        as.character()
     )
   } else {
     return(table)
@@ -107,8 +106,7 @@ coverage_table <- function(full_df_renamed, latex = FALSE) {
     return(
       table %>%
         as_latex() %>%
-        as.character() %>%
-        cat()
+        as.character()
     )
   } else {
     return(table)

### 2022-08-03 src/1_process_metric-tables/script.R

@@ -7,8 +7,6 @@ df_intercept <- readRDS("depends/df_intercept.rds")
 # df_time <- readRDS("depends/df_time.rds")
 
 #' CRPS table for rho
-sink("crps-table-rho.txt")
-
 df_rho %>%
   group_mean_and_se(group_variables = c("geometry", "sim_model", "inf_model")) %>%
   bsae::update_naming() %>%
@@ -18,13 +16,11 @@ df_rho %>%
     latex = TRUE,
     scale = 10^3,
     figures = 3
-  )
+  ) %>%
+  cat(file = "crps-table-rho.txt")
 
-sink()
 
 #' CRPS table for the intercept
-sink("crps-table-intercept.txt")
-
 df_intercept %>%
   group_mean_and_se(group_variables = c("geometry", "sim_model", "inf_model")) %>%
   bsae::update_naming() %>%
@@ -34,13 +30,10 @@ df_intercept %>%
     latex = TRUE,
     scale = 10^3,
     figures = 3
-  )
-
-sink()
+  ) %>%
+  cat(file = "crps-table-intercept.txt")
 
 #' #' Time table
-#' sink("time-table.txt")
-#'
 #' df_time %>%
 #'   group_by(geometry, sim_model, inf_model) %>%
 #'   bsae::update_naming() %>%
@@ -48,6 +41,5 @@ sink()
 #'   metric_table(
 #'     metric = "t",
 #'     latex = FALSE
-#'   )
-#'
-#' sink()
+#'   ) %>%
+#'   cat(file = "time-table.txt")

### 2022-08-03 src/docs_paper/paper.Rmd
Import CRPS results for simulation study straight into paper
Add metrics table for HIV survey results

@@
 
+## Results
+
+```{r results="asis"}
+#' Some of the lines should not be included
+lines_remove <- c(1, 3, 4, 5)
+cat(readLines("depends/crps-table-rho.txt")[-lines_remove], sep = "\n")
+```
+
 # HIV prevalence study \label{sec:hiv}
@@
 
+## Results
+
+```{r results="asis"}
+#' Some of the lines should not be included
+lines_remove <- c(1)
+cat(readLines("depends/metric-table.txt")[-lines_remove], sep = "\n")
+```
+
 # Discussion \label{sec:discussion}

### 2022-08-03 src/docs_paper/paper.Rmd
Simplify LOO and SLOO-CV paragraph

@@
 We assessed forecasting performance using the CRPS, at the level of the data $y_i$ rather than at the level of the prevalence $\rho_i$.
-To compute the value of CRPS for the $i$th fold we use samples $\{ y_i^s \}_{s = 1}^S$ from the LOO posterior predictive distribution $p(y_i \, | \, y_{-i})$ giving
-\begin{equation}
-\text{LOO-CRPS}_i = \frac{1}{S} \sum_{s = 1}^S | y_i^s - y_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | y_i^s - y_i |. \label{eq:crpsapprox2}
-\end{equation}
+To compute the value of CRPS for the $i$th fold we use samples $\{ y_i^s \}_{s = 1}^S$ from the LOO posterior predictive distribution $p(y_i \, | \, y_{-i})$ giving $\text{LOO-CRPS}_i = \frac{1}{S} \sum_{s = 1}^S | y_i^s - y_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | y_i^s - y_i |.
+\label{eq:crpsapprox2}$
 
@@
 
-If structural dependencies, like spatial dependence, exist in the data then results from LOO-CV tend to underestimate predictive error [@roberts2017cross].
-To counteract this effect, we modify the spatial leave-one-out (SLOO) approach of @le2014spatial.
-As illustrated by Figure \ref{fig:sloo}, during the $i$th fold of SLOO-CV we leave-out the block $(i, \delta i)$.
-This increases the extent to which the training and validation sets are conditionally independent.
-The result is a SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$ from which, analogously to Equation \ref{eq:crpsapprox2}, we may compute the corresponding value of $\text{SLOO-CRPS}_i$.
+When structural dependencies, like spatial dependence, exist in the data then results from LOO-CV tend to underestimate predictive error [@roberts2017cross].
+To counteract this effect, we modified the spatial leave-one-out (SLOO) approach of @le2014spatial by leaving-out the block $(i, \delta i)$ during the $i$th fold of SLOO-CV (Figure \ref{fig:sloo}), thereby increasing the extent to which the training and validation sets are conditionally independent.
+From the SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$  we analogously computed $\text{SLOO-CRPS}_i$.
 

### 2022-08-03 src/docs_paper/paper.Rmd
Some attempt to simplying the writing

@@
 Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
-Epidemic estimates are required to help plan, implement, and evaluate the success of programmes, as well as allocate limited resources [@hallett2016evaluation; @cuadros2017mapping].
-Household surveys provide nationally representative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
-Sometimes auxiliary covariates can be used to aid estimation, but data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, faces similar measurement difficulties as the HIV indicators themselves.
-In light of this fact, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}, unlike other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
+Estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effective used to counter the epidemic [@hallett2016evaluation; @cuadros2017mapping].
+Household surveys provide nationally represenative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
+Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, often face similar measurement difficulties as the HIV indicators themselves.
+Models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}, unlike other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
 These circumstances foreground the importance of sharing information between districts via spatial smoothing in HIV mapping.
@@
 
-Improvements to the Besag model may be transferable to improving convolution models.
 Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
@@
 Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive.
-The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and more broadly both models are of substantive practical interest.
+Improvements to the Besag model are likely transferable to improving these convolution models.
+The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and both models are of substantive practical interest.
 
@@
 
-Usually, areal may be conceptualised as arising from aggregation of point data so that $y_i = \int_{A_i} y(s) \text{d}s$.
+Areal is often aptly conceptualised as arising from aggregation of point data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
 Exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a genuinely discrete spatial structure\footnote{For data of this sort, there is less clear justification for using spatial smoothing.
@@
 Each observation provides less information than it would have had the samples been independent, making it more difficult to estimate global parameters.
-However, spatial correlation can be used to improve local parameter estimates, particularly in parts of the study region where little to no information is available^[We use local parameters to refer to those parameters those indexed by a particular spatial (or spatio-temporal) location.
-These parameters corresponds to the latent field, as defined in Section \ref{sec:lgm}.].
+However, spatial correlation can be used to improve local (indexed by a particular spatial location) parameter estimates particularly in parts of the study region where little to no information is available.
 
@@
 
-The latent field in a LGM is comprised a constant mean $\mu$, known also as the fixed effects, and (spatial) random effects $\phi \sim \mathcal{N}(0, \Sigma)$.
-The mean may be modelled using a linear predictor $\mu_i = \beta_0 + z_i^\top \beta$, where $(\beta_0, \beta)$ are regression parameters corresponding to areal covariates^[To denote length $p$ vectors of areal and point covariates, we use $z_i = (z_{i,1}, \ldots, z_{i,p})^\top$ and $z(s) = (z_1(s), \ldots, z_p(s))^\top$ respectively.} $(1, z_i)^\top$].
-Here, we focus on the spatial random effects $\phi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
+The latent field in a LGM is comprised a constant mean $\mu$, known also as the fixed effects, and spatial random effects $\phi \sim \mathcal{N}(0, \Sigma)$.
+The mean may be modelled using a linear predictor $\mu_i = \beta_0 + z_i^\top \beta$, where $(\beta_0, \beta)$ are regression parameters corresponding to areal covariates $z_i = (z_{i,1}, \ldots, z_{i,p})^\top$.
+We focus on the spatial random effects $\phi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
 If no spatial structure remains after inclusion of covariates in the mean, independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ should be used, where $\tau_\phi$ is a shared precision.
@@
 <!-- TODO: Add calculation of Moran's $I$ to my HIV survey data-sets. -->
-Specifying $\phi$ amounts to specifying the entries of a precision or covariance matrix, as we discuss below.
+Specifying $\phi$ amounts to specifying the entries of a precision or covariance matrix, as we discuss in the following sections.
 
-# Spatial random effects based on adjacency \label{sec:ab}
+# Spatial random effects using adjacency \label{sec:ab}
 
@@
 
-One way to encode spatial structure between areas is by using a symmetric relation, where $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring, as often defined by a shared border^[Any choice of adjacency definition is possible.
-For example, one might chose to allow inclusion of second degree neighbours resulting in so-called higher-order models [@paciorek2013spatial].].
+One way to encode spatial structure between areas is by using a symmetric relation, where $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring.
+Adjacency is often defined by a shared border, though other choices, such as inclusion of second or higher-degree neighbours [@paciorek2013spatial], are also possible.
 The Besag model [@besag1991bayesian] is a type of improper conditional auto-regressive (ICAR) model where the full conditional distribution of the $i$th spatial random effect is given by
@@
 \begin{equation}
-    \phi \sim \mathcal{N}(\mathbf{0}, \tau_\phi^{-1} R^{-}), \label{eq:gmrf}
+    \phi \sim \mathcal{N}(0, \tau_\phi^{-1} R^{-}), \label{eq:gmrf}
 \end{equation}
-where $R^{-}$ is the generalised inverse of the rank-deficient structure matrix\footnote{So-called because it defines the \textit{structure} of the precision matrix $\mathbf{Q} = \tau_\phi R$.} $R$ with entries
+where $R^{-}$ is the generalised inverse of the rank-deficient structure matrix $R$, so-called because it defines the \textit{structure} of the precision matrix, with entries
 \begin{equation}
@@
 The structure matrix $R$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
-Figure \ref{fig:zi} shows the adjacency graph for the districts of Zimbabwe alongside the geometry.
-Shortly, in Section \ref{sec:concerns} we discuss the appropriateness of using the graph rather than the full geometry.
+Figure \ref{fig:geometry-graph-zwe} shows the adjacency graph for the districts of Zimbabwe alongside the geometry.
+Shortly, in Section \ref{sec:concerns} we discuss the appropriateness of using the graph rather than the complete geometry.
 
@@
 \end{equation}
-This density is a function of the pairwise differences $\phi_i - \phi_j$ and so is invariant to the addition of a constant $c$ to each entry $p(\phi) = p(\phi + c\mathbf{1})$, leading to an improper uniform distribution on the average of the $\phi_i$.
-If $\mathcal{G}$ is connected\footnote{In that by traversing the edges, any vertex can be reached from any other vertex.} then there is only one impropriety in the model and $\text{rank}(R) = n - 1$, while if $\mathcal{G}$ is disconnected, and composed of $n_c \geq 2$ connected components with index sets $I_1, \ldots, I_{n_c}$, then the corresponding structure matrix $R$ has rank $n - n_c$ and the density is invariant to the addition of a constant to each of the connected components $p(\phi_{I}) = p(\phi_{I} + c \mathbf{1})$ where $I = I_1, \ldots, I_{n_c}$.
+This density is a function of the pairwise differences $\phi_i - \phi_j$ and so is invariant to the addition of a constant $c$ to each entry $p(\phi) = p(\phi + c1)$, leading to an improper uniform distribution on the average of the $\phi_i$.
+If $\mathcal{G}$ is connected, in that by traversing the edges, any vertex can be reached from any other vertex, then there is only one impropriety in the model and $\text{rank}(R) = n - 1$, while if $\mathcal{G}$ is disconnected, and composed of $n_c \geq 2$ connected components with index sets $I_1, \ldots, I_{n_c}$, then the corresponding structure matrix $R$ has rank $n - n_c$ and the density is invariant to the addition of a constant to each of the connected components $p(\phi_{I}) = p(\phi_{I} + c1)$ where $I = I_1, \ldots, I_{n_c}$.
 
@@
 <!-- The Besag model is a special case of a more general class of (zero-mean) ICAR models in which each neighbour may not be weighted equally. -->
-<!-- These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\bm{\kappa} = (\kappa_1, \ldots, \kappa_n)^\top$ such that $\phi_i \, | \, \phi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$.
+<!-- These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\kappa = (\kappa_1, \ldots, \kappa_n)^\top$ such that $\phi_i \, | \, \phi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$.
 -->
-<!-- The structure matrix $R$ corresponding to the above full conditionals is given by $R = \mathbf{D}_\kappa(\mathbf{I} - \mathbf{B})$, where $\mathbf{B}$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $\mathbf{D}_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$.
+<!-- The structure matrix $R$ corresponding to the above full conditionals is given by $R = D_\kappa(I - B)$, where $B$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $D_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$.
 -->
 <!-- As the structure matrix is symmetric we must have the condition $-b_{ij}\kappa_i = -b_{ji}\kappa_j$. -->
-<!-- To meet this condition, it is often simpler to use an unscaled weights matrix $\mathbf{W} = \mathbf{D}_\kappa \mathbf{B}$ such that $R = \mathbf{D}_\kappa - \mathbf{W}$. -->
-<!-- For example, in the Besag model $\mathbf{W}$ corresponds to the adjacency matrix. -->
+<!-- To meet this condition, it is often simpler to use an unscaled weights matrix $W = D_\kappa B$ such that $R = D_\kappa - W$. -->
+<!-- For example, in the Besag model $W$ corresponds to the adjacency matrix. -->
 <!-- The scaled weights can then be recovered by $b_{ij} = w_{ij} / \kappa_i$ where $\kappa_i = \sum_{k: k \sim i} w_{ik}$. -->
-<!-- A thorough discussion of methods for specifying $\mathbf{W}$ is provided by @duncan2017spatial. -->
+<!-- A thorough discussion of methods for specifying $W$ is provided by @duncan2017spatial. -->
 <!-- Much of the work in this area focuses on the case where the geometry is a lattice. -->
@@
 Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model.
-The Besag-York-Mollié (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\phi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
+The Besag-York-Mollié (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\phi = v + u$ into a sum of an unstructured IID component $v$ and a spatially structured Besag component $u$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
 The resulting distribution is
 \begin{equation}
-    \phi \sim \N(\mathbf{0}, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} R^{-}) \label{eq:bym}.
+    \phi \sim \N(0, \tau_v^{-1} I + \tau_u^{-1} R^{-}) \label{eq:bym}.
 \end{equation}
-Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$.
-However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
-In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale.
-Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
+Including both $v$ and $u$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$.
+However, in this specification, scaling of the Besag precision matrix $Q$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
+In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $Q$ and $I$ have the same scale.
+Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $v$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
 This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal.
@@
 \begin{align}
-\phi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2}
+\phi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, v + \sqrt{\pi} \, u^\star \right), \label{eq:bym2}
 \end{align}
-where $\tau_\phi$ is the marginal precision of $\phi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $\bu$ with precision matrix given by the scaled structure matrix $R^\star$ (see~\ref{app:freni}).
+where $\tau_\phi$ is the marginal precision of $\phi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $u^\star$ is a scaled version of $u$ with precision matrix given by the scaled structure matrix $R^\star$ (see Appendix).
 When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model.
@@
 
-# Areal kernel spatial random effects \label{sec:ks}
+# Spatial random effects using kernels \label{sec:ks}
 
@@
 \begin{equation}
-    \phi \sim \mathcal{N}(\mathbf{0}, \frac{1}{\tau_\phi} K), \label{eq:arealkernel}
+    \phi \sim \mathcal{N}(0, \frac{1}{\tau_\phi} K), \label{eq:arealkernel}
 \end{equation}
@@
 \hypertarget{I4}{I4}. \textit{BYM2} & INLA &
-$\phi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
+$\phi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, v + \sqrt{\pi} \, u^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
 \hypertarget{I5}{I5}. \textit{FCK} & INLA &
@@
 The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$.
-The CRPS is a strictly proper scoring rule [@gneiting2007strictly], which ensures that the (uniquely) best approach to minimising the score is to truthfully report probabilistic beliefs, here corresponding to the posterior distribution.
+The CRPS is a strictly proper scoring rule [@gneiting2007strictly], which ensures that the uniquely best approach to minimising the score is to truthfully report probabilistic beliefs, here corresponding to the posterior distribution.
 The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
@@
 We used data from the most recent publicly available Population Health Impact Assessment (PHIA) survey in each country.
-These surveys utilise a complex design, where each individual $k$ in area $i$ has an unequal probability $\pi_{ik}$ of being included in the sample.
+These surveys utilise a complex design, where each individual $j$ in area $i$ has an unequal probability $\pi_{ij}$ of being included in the sample.
 <!-- A two-stage design in which enumeration areas are first drawn from a stratified sample and then households are chosen using equal probability systematic sampling from within each enumeration area, is common.
 -->
-To account for the survey design we used sampling weights $\omega_{ik} = 1/\pi_{ik}$ to adjust the raw data in each district, obtaining the Kish effective sample size $m_i^\star = (\sum_k \omega_{ik})^2 / \sum_k \omega_{ik}^2$ and effective number of cases $y_i^\star$.
-These effective values may be thought of as what would have been observed had the survey been a simple random sample.
+To account for the survey design we used sampling weights $\omega_{ij} = 1/\pi_{ij}$ to adjust the raw data in each district, obtaining the Kish effective sample size $m_i^\star = (\sum_k \omega_{ij})^2 / \sum_k \omega_{ij}^2$ and effective number of cases $y_i^\star$ which may be thought of as what would have been observed had the survey been a simple random sample.
 
@@
 
-We used a similar model structure as Section \ref{sec:simulation}, with a minor alteration to the likelihood.
+We used a the same models (Table \ref{tab:inferentialmodels}) as Section \ref{sec:simulation}, with one alteration to the likelihood.
 The effective number of cases and effective sample size may not be integers, so we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
@@
 \end{equation}
-As with the simulation study we linked the prevalence $\rho_i \in [0, 1]$ to $x_i$ via
-\begin{equation}
-\log \left( \frac{\rho_i}{1 - \rho_i} \right) = x_i = \beta_0 + \phi_i, \label{eq:linear}
-\end{equation}
-and again considered the range of models for $\bphi$ given by Table \ref{tab:inferentialmodels}.
 

### 2022-08-03 src/docs_paper/paper.Rmd
Update title

@@
 ---
-title: Modelling spatial structure in small-area estimation
+title: Understanding models for spatial structure in small-area estimation
 author:

### 2022-08-03 src/docs_paper/paper.Rmd
Some paper editing

@@
 We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities.
-Finally, we compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
+We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
   |
@@
 We hypothesised that incorporating a spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models.
-Here, we set out to test this hypothesis, and in doing so provide practical recommendations for modelling spatial structure.
+Here, we set out to test this hypothesis, and in doing so provide practical recommendations and further understanding of models for spatial structure.
 <!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)?
@@
 
-<!-- Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related. -->
-<!-- @duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model^[This conclusion was specific to the (often-analysed) Scottish lip cancer example, and like the majority of the literature, based on the deviance information criteria (DIC).
-The DIC is based on summarising the posterior distribution by a point estimate, and is recommended against by @vehtari2017practical.].
--->
-<!-- Another approach is to take into account that areal data is typically produced by aggregating higher resolution data, so as to be a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change] and dating to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
--->
-<!-- Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
--->
-<!-- This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially], `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages.
--->
-<!-- Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].
--->
+Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
+@duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model, though this conclusion was specific to the (often-analysed) Scottish lip cancer example, and based on the recommended against [@vehtari2017practical] deviance information criteria (DIC).
+Another approach is to take into account that areal data is typically produced by aggregating higher resolution data, so as to be a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change] and dating to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
+Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
+This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially], `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages.
+Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].
 
@@
 
-Areal is often aptly conceptualised as arising from aggregation of point data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
-Exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a genuinely discrete spatial structure\footnote{For data of this sort, there is less clear justification for using spatial smoothing.
-Realistically, data relate to many covariates, some number of which are likely have discrete spatial structure and others of which continuous spatial structure, such that the result is often an intermediate.
-It may be possible to model this explicitly, though the result might look similar to what is achieved using e.g.
-BYM2.}.
+Areal data is often aptly conceptualised as arising from aggregation of point data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
+Exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a genuinely discrete spatial structure.
+<!-- Note that, for data of this sort, there is less clear justification for using spatial smoothing. -->
+<!-- Realistically, data relate to many covariates, some number of which are likely have discrete spatial structure and others of which continuous spatial structure, such that the result is often an intermediate.
+-->
+<!-- It may be possible to model this explicitly, though the result might look similar to what is achieved using e.g. BYM2. -->
 If the areal data we observe are indeed an aggregation of point data, why not use the higher resolution point data in our models rather than areal data?
-Explainations include that (a) the underlying point data are unobserved, (b) although point data are collected, they may not be accessible due to privacy constraints, administrative practicality or storage capacity, and (c) the point data are available, but we decide not to use them in the model.
-In the final case (c), researchers may prefer to use area-level models because models for point data are typically more complex, require more data, including covariate data which may not exist, and have higher computational burden.
-Choosing to model the underlying data generating mechanism as either areal or point is often a matter of pragmatism.
+Explanations include that (a) the underlying point data are unobserved, (b) although point data are collected, they may not be accessible due to privacy constraints, administrative practicality or storage capacity, and (c) the point data are available, but we decide not to use them in the model.
+In the final case (c), researchers may prefer to use area-level models as models for point data are typically more complex, require more data, are less immediately applicable to area-level policy making, and have higher computational burden.
+Furthermore, seemly point data may actual be areal, such as data observed at polling stations or health facilities which individuals from the surrounding area travel to.
+As such, choosing to model the underlying data generating mechanism as either areal or point is often a matter of pragmatism.
 
@@
 Smoothing approaches use information from similar units to "borrow strength" from one parameter to another, with determining precisely what is meant by "similar" a central challenge.
-@wakefield2020small review design-based approaches and model-based approaches to spatial analysis of health-indicators, including both area and unit-level analysis.
+For a recent review of both design-based and model-based approaches to spatial analysis of health-indicators, including both area and unit-level analysis, see @wakefield2020small.
 
@@
 
-A striking feature of Equation~\ref{eq:besag} is that the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.
+A striking feature of Equation \ref{eq:besag} is that the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.
 It follow that as $n_{\delta i} \to \infty$ then $\text{Var}(\phi_i) \to 0$.
@@
 
-<!-- ## Weighted ICAR \label{sec:icar} -->
+## Weighted ICAR \label{sec:icar}
 
-<!-- The Besag model is a special case of a more general class of (zero-mean) ICAR models in which each neighbour may not be weighted equally. -->
-<!-- These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\kappa = (\kappa_1, \ldots, \kappa_n)^\top$ such that $\phi_i \, | \, \phi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$.
--->
-<!-- The structure matrix $R$ corresponding to the above full conditionals is given by $R = D_\kappa(I - B)$, where $B$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $D_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$.
--->
-<!-- As the structure matrix is symmetric we must have the condition $-b_{ij}\kappa_i = -b_{ji}\kappa_j$. -->
-<!-- To meet this condition, it is often simpler to use an unscaled weights matrix $W = D_\kappa B$ such that $R = D_\kappa - W$. -->
-<!-- For example, in the Besag model $W$ corresponds to the adjacency matrix. -->
-<!-- The scaled weights can then be recovered by $b_{ij} = w_{ij} / \kappa_i$ where $\kappa_i = \sum_{k: k \sim i} w_{ik}$. -->
-<!-- A thorough discussion of methods for specifying $W$ is provided by @duncan2017spatial. -->
-<!-- Much of the work in this area focuses on the case where the geometry is a lattice. -->
+The Besag model is a special case of a more general class of (zero-mean) ICAR models in which each neighbour may not be weighted equally.
+These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\kappa = (\kappa_1, \ldots, \kappa_n)^\top$ such that $\phi_i \, | \, \phi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$.
+The structure matrix $R$ corresponding to the above full conditionals is given by $R = D_\kappa(I - B)$, where $B$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $D_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$.
+As the structure matrix is symmetric we must have the condition $-b_{ij}\kappa_i = -b_{ji}\kappa_j$.
+To meet this condition, it is often simpler to use an unscaled weights matrix $W = D_\kappa B$ such that $R = D_\kappa - W$.
+For example, in the Besag model $W$ corresponds to the adjacency matrix.
+The scaled weights can then be recovered by $b_{ij} = w_{ij} / \kappa_i$ where $\kappa_i = \sum_{k: k \sim i} w_{ik}$.
+A thorough discussion of methods for specifying $W$ is provided by @duncan2017spatial.
+Much of the work in this area focuses on the case where the geometry is a lattice.
 
@@
 <!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
-The centroid kernel has been used by @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$.
-The former fix the length-scale $l$ using a semi-variogram and the later place a prior on the inverse length-scale.
-In their simulation experiment, \citet[Section 3]{best2005comparison} used data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas, but it is not suprising that a stationary covariance structure would struggle to recover non-stationary spatial structure).
-More recently, @teh2021efficient used a centroid kernel as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
+The centroid kernel has been used by @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$; as well as more recently by @teh2021efficient as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
+@wakefield1999spatial fix the length-scale $l$ using a semi-variogram and @best2005comparison place a prior on the inverse length-scale.
+\citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas, though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.
 
@@
 \includegraphics[width=\textwidth]{depends/integration-method.pdf}
-\caption{Three ways to choose integration points with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample} [@pebesma2018sf].}
+\caption{Three ways to choose integration points with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample}.}
 \label{fig:41}
@@
 \includegraphics{depends/simulation-geometries.pdf}
-\caption{Simulation geometries.}
+\caption{The seven simulation geometries that we considered.}
 \label{fig:51}
@@
 \hypertarget{S3}{S3}.
-\textit{IK} & $\phi \sim \mathcal{N}(0, K^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation~\ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
+\textit{IK} & $\phi \sim \mathcal{N}(0, K^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation \ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
 Like the Besag model, we scaled the Gram matrix to have generalised variance one \\
@@
 
-We fit eight inferential models, each of which differing in the spatial random effect specification (Table \ref{tab:inferentialmodels}).
+We fit eight inferential models, each of which differing in its spatial random effect specification (Table \ref{tab:inferentialmodels}).
 
@@
 \hypertarget{I3}{I3}. \textit{Besag} & INLA &
-$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations in~\ref{app:freni}  \\
+$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations in \ref{app:freni}  \\
 \hypertarget{I4}{I4}. \textit{BYM2} & INLA &
@@
 $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
-\hypertarget{I6}{I6}. \textit{CK} & NUTS & As in model~\hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
+\hypertarget{I6}{I6}. \textit{CK} & NUTS & As in model \hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
 \hypertarget{I7}{I7}.
 \textit{FIK} & INLA & $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
-\hypertarget{I8}{I8}. \textit{IK} & NUTS & As in model~\hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
+\hypertarget{I8}{I8}. \textit{IK} & NUTS & As in model \hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
 \bottomrule
@@
 
-For the precision, we placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
+We placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
 The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one, neither of which we assumed to be plausible.
@@
 In models \hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
-For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51}.
-The true number of integration points may be 1-2 above or below 10, as `sf::st_sample` with `type = "hexagonal"` is not guaranteed to be exact.
+For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51} -- though `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].
 We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}.
@@
 The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$.
-The CRPS is a strictly proper scoring rule [@gneiting2007strictly], which ensures that the uniquely best approach to minimising the score is to truthfully report probabilistic beliefs, here corresponding to the posterior distribution.
+The CRPS is a strictly proper scoring rule [@gneiting2007strictly], ensuring that the uniquely best approach to minimising the score is for the model to report its true probabilistic beliefs about the quantity in question.
 The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
-In contrast, given a collection of Monte Carlo samples $\{\rho_i^s\}_{s = 1}^S \sim f$, Equation~\ref{eq:crps} may easily be approximated \citep[Equation 37]{sigrist2015stochastic} by
 \begin{equation}
@@
 
-To assess the coverage of our estimates, we considered probability integral transform (PIT) histograms and empirical cumulative distribution function (ECDF) difference plots [@sailynoja2021graphical].
+To assess the coverage of our estimates, we considered probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2021graphical].
 We estimated the quantile $q_i = F(\omega_i)$ of the true value within the marginal posterior using numerical integration.
-For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that for any given nominal significance level $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$, should be $1 - \alpha$.
+For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that for any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$, in other words the empirical coverage, should also be $1 - \alpha$.
 
@@
 
-We used a the same models (Table \ref{tab:inferentialmodels}) as Section \ref{sec:simulation}, with one alteration to the likelihood.
-The effective number of cases and effective sample size may not be integers, so we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
+We used a the same models as in Section \ref{sec:simulation} (Table \ref{tab:inferentialmodels}), with one alteration to the likelihood.
+Since the effective number of cases and effective sample size may not be integers, we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
 The working likelihood under this model for $m^\star_i \geq y^\star_i$ is given by
@@
 
-We assessed each model using two cross-validation approaches: (1) LOO-CV, a standard leave-one-out cross-validation, in which the prevalence data from a single district were removed, and (2) SLOO-CV, a spatial leave-one-out cross-validation, in which the data from a single district and each of its neighbouring districts were removed.
-We performed cross-validation manually as we found approximations that avoid refitting the model, such as the built-in conditional predictive ordinate statistic in `R-INLA` package [@held2010posterior] and the Pareto smoothed importance sampling [@vehtari2015pareto] approach implemented in the `loo` package [@vehtari2017practical], to be inaccurate in this setting, perhaps because individual observations in spatial random effect models can exert high influence toward local parameters.
+We assessed each model using two cross-validation approaches: (1) a standard leave-one-out cross-validation (LOO-CV), in which at each fold the data from a single district were held-out, and (2) a spatial leave-one-out cross-validation (SLOO-CV), in which at each fold the data from a single district and each of its neighbouring districts were removed.
+We performed both types of cross-validation manually, as we found approximations that avoid refitting the model, such as the built-in conditional predictive ordinate statistic in `R-INLA` package [@held2010posterior] and the Pareto smoothed importance sampling [@vehtari2015pareto] approach implemented in the `loo` package [@vehtari2017practical], were inaccurate in this setting.
+This might be because individual observations in spatial random effect models can exert high influence toward local parameters.
 
@@
 
-When structural dependencies, like spatial dependence, exist in the data then results from LOO-CV tend to underestimate predictive error [@roberts2017cross].
-To counteract this effect, we modified the spatial leave-one-out (SLOO) approach of @le2014spatial by leaving-out the block $(i, \delta i)$ during the $i$th fold of SLOO-CV (Figure \ref{fig:sloo}), thereby increasing the extent to which the training and validation sets are conditionally independent.
+In the presense of structural dependencies, like spatial dependence, LOO-CV tends to underestimate predictive error [@roberts2017cross].
+We counteracted this effect by modifying the spatial leave-one-out (SLOO) approach of @le2014spatial.
+During the $i$th fold of SLOO-CV, we leave-out the block $(i, \delta i)$ (Figure \ref{fig:sloo}), increasing the extent to which the training and validation sets are conditionally independent.
 From the SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$  we analogously computed $\text{SLOO-CRPS}_i$.
@@
 lines_remove <- c(1)
-cat(readLines("depends/metric-table.txt")[-lines_remove], sep = "\n")
+
+#' An obnoxious number of escape characters
+readLines("depends/metric-table.txt")[-lines_remove] %>%
+  stringr::str_replace(pattern = "\\(\\\\\\$n\\\\\\$", replacement = "($n$") %>%
+  cat(sep = "\n")
 ```

### 2022-08-03 src/0_check_ik-compute/ik-compute.Rmd
ik-compute to HTML

---
title: "Check integrated kernel computation"
author: "Adam Howes"
output: html_document
---

```{r}
#' Checking that the Stan code to calculate the integrated kernel is correct

sf <- mw
L <- 10
type = "random"

n <- nrow(sf)
samples <- sf::st_sample(sf, type = type, exact = TRUE, size = rep(L, n))
S <- sf::st_distance(samples, samples)
sample_index <- sf::st_intersects(sf, samples)
sample_lengths <- lengths(sample_index)
start_index <- sapply(sample_index, function(x) x[1])

#' First calculate it directly using R only
D <- sf::st_distance(samples, samples)
kD <- matern(D, l = 1)
K <- matrix(nrow = n, ncol = n)

for(i in 1:n) {
  K[i, i] <- mean(kD[sample_index[[i]], sample_index[[i]]])
}

for(i in 1:(n - 1)) {
  for(j in (i + 1):n) {
    K[i, j] <- mean(kD[sample_index[[i]], sample_index[[j]]]) #' Fill the upper triangle
    K[j, i] <- K[i, j] #' Fill the lower triangle
  }
}

cov_r <- K

#' Now calculate it in Stan via two different methods
#' rstan::expose_stan_functions puts the C++ functions into the R local workspace to use
rstan::expose_stan_functions("cov_sample_average.stan")

#' 1. Newer method
cov_stan <- cov_sample_average(
  S = S,
  l = 1,
  n = n,
  start_index = start_index,
  sample_lengths = sample_lengths,
  total_samples = sum(sample_lengths)
)

#' 2. Old method
cov_stan_old <- cov_sample_average_old(
  n = n,
  L = 10,
  l = 1,
  S = S
)

#' Comparison (these statements should all be TRUE)
all.equal(cov_r, cov_stan)
all.equal(cov_r, cov_stan_old)
all.equal(cov_stan, cov_stan_old)

#' Hard to tell from an image if something is identical but anyway
cowplot::plot_grid(
  plot_matrix(cov_r) + labs(title = "R"),
  plot_matrix(cov_stan_old) + labs(title = "Stan old"),
  plot_matrix(cov_stan) + labs(title = "Stan new"),
  ncol = 2
)
```

### 2022-08-03 utils/archive_to_docs.R

@@ -17,3 +17,4 @@ archive_to_docs <- function(report) {
 
 #' Names of the reports to move
 archive_to_docs("docs_paper")
+archive_to_docs("0_check_ik-compute")

### 2022-08-03 utils/archive_to_docs.R
INLA-SPDE notes to HTML

@@ -18,3 +18,4 @@ archive_to_docs <- function(report) {
 #' Names of the reports to move
 archive_to_docs("docs_paper")
 archive_to_docs("0_check_ik-compute")
+archive_to_docs("0_explore_inla-spde")

### 2022-08-06 src/0_check_bin-gaussian/bin-gaussian.Rmd
bin-gaussian to HTMl

---
title: "Binomial of sum of Gaussians versus sum of binomial of Gaussians"
author: "Adam Howes"
output: html_document
---

```{r}
#' Binomial of sum of Gaussians versus sum of binomial of Gaussians
#' IID latent field
#' Without aggregation:
#' `y = \sum_{i = 1}^n y_i$, $y_i \sim \text{Bin}(m, \rho_i)$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`
#' With aggregation:
#' `y \sim \text{Bin}(nm, \bar \rho)$, $\bar \rho = \frac{1}{n} \sum_{i = 1}^n \rho_i$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`

experiment <- function(nsim, n, phi_mean, phi_sd, y_sample_size) {
  phi <- rnorm(n, phi_mean, phi_sd)
  rho <- plogis(phi)
  m <- rep(y_sample_size, n)

  #' Without aggregation
  y <- replicate(nsim, rbinom(n, m, rho), simplify = FALSE)
  y <- lapply(y, sum)

  #' With aggregation (at the level of the latent field)
  y_agg <- replicate(nsim, rbinom(1, sum(m), weighted.mean(rho)), simplify = FALSE)

  return(list(y = unlist(y), y_agg = unlist(y_agg), phi = phi, rho = rho))
}

plot_hist <- function(result) {
  data.frame(
    type = c(rep("No aggregation (truth)", nsim), rep("Aggregation (approximation)", nsim)),
    y = c(result$y, result$y_agg)
  ) %>%
    ggplot() +
    geom_histogram(aes(x = y, fill = type), alpha = 0.5, position = "identity") +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )
}

nsim <- 500
n <- 36

#' Low prevalence setting (disease)
result <- experiment(nsim, n, -2.5, 0.5, 30)
plot(result$rho) #' This is what the underlying latent field looks like for these settings

plot_hist(result) #' And then the difference between the truth and the aggregated approximation

sd(result$y)
sd(result$y_agg)
sd(result$y_agg) / sd(result$y) #' Ratio

#' Probability approximately 1/2 setting (elections)
#' As before but with different `phi_mean` settings:
result_half <- experiment(nsim, n, 0, 0.5, 30)
plot(result_half$rho)

plot_hist(result_half)

sd(result_half$y)
sd(result_half$y_agg)
sd(result_half$y_agg) / sd(result_half$y) #' Ratio

#' Conclusions, things to add:
#' * Standard deviation of aggregation tends to overestimate true standard deviation,
#' * Effect is larger in the small `\rho` setting than the large `\rho` setting
#' * Should test latent field with spatial structure e.g. ICAR latent field and so on
#' * What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?
```

### 2022-08-06 src/0_check_bin-gaussian/functions.R

@@ -1,27 +1,0 @@
-experiment <- function(nsim, n, phi_mean, phi_sd, y_sample_size) {
-  phi <- rnorm(n, phi_mean, phi_sd)
-  rho <- plogis(phi)
-  m <- rep(y_sample_size, n)
-
-  #' Without aggregation
-  y <- replicate(nsim, rbinom(n, m, rho), simplify = FALSE)
-  y <- lapply(y, sum)
-
-  #' With aggregation (at the level of the latent field)
-  y_agg <- replicate(nsim, rbinom(1, sum(m), weighted.mean(rho)), simplify = FALSE)
-
-  return(list(y = unlist(y), y_agg = unlist(y_agg), phi = phi, rho = rho))
-}
-
-plot_hist <- function(result) {
-  data.frame(
-    type = c(rep("No aggregation (truth)", nsim), rep("Aggregation (approximation)", nsim)),
-    y = c(result$y, result$y_agg)
-  ) %>%
-    ggplot() +
-    geom_histogram(aes(x = y, fill = type), alpha = 0.5, position = "identity") +
-    theme_minimal() +
-    theme(
-      legend.position = "bottom"
-    )
-}

### 2022-08-06 src/0_check_bin-gaussian/script.R

@@ -1,47 +1,4 @@
 # orderly::orderly_develop_start("0_check_bin-gaussian")
 # setwd("src/0_check_bin-gaussian")
 
-#' Binomial of sum of Gaussians versus sum of binomial of Gaussians
-#' IID latent field
-#' Without aggregation:
-#' `y = \sum_{i = 1}^n y_i$, $y_i \sim \text{Bin}(m, \rho_i)$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`
-#' With aggregation:
-#' `y \sim \text{Bin}(nm, \bar \rho)$, $\bar \rho = \frac{1}{n} \sum_{i = 1}^n \rho_i$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`
-
-nsim <- 500
-n <- 36
-
-#' Low prevalence setting (disease)
-result <- experiment(nsim, n, -2.5, 0.5, 30)
-plot(result$rho) #' This is what the underlying latent field looks like for these settings
-
-pdf("histogram-disease.pdf", h = 4, w = 6.25)
-
-plot_hist(result) #' And then the difference between the truth and the aggregated approximation
-
-dev.off()
-
-sd(result$y)
-sd(result$y_agg)
-sd(result$y_agg) / sd(result$y) #' Ratio
-
-#' Probability approximately 1/2 setting (elections)
-#' As before but with different `phi_mean` settings:
-result_half <- experiment(nsim, n, 0, 0.5, 30)
-plot(result_half$rho)
-
-pdf("histogram-election.pdf", h = 4, w = 6.25)
-
-plot_hist(result_half)
-
-dev.off()
-
-sd(result_half$y)
-sd(result_half$y_agg)
-sd(result_half$y_agg) / sd(result_half$y) #' Ratio
-
-#' Conclusions, things to add:
-#' * Standard deviation of aggregation tends to overestimate true standard deviation,
-#' * Effect is larger in the small `\rho` setting than the large `\rho` setting
-#' * Should test latent field with spatial structure e.g. ICAR latent field and so on
-#' * What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?
+rmarkdown::render("bin-gaussian.Rmd")

### 2022-08-06 src/0_check_ik-converge/ik-converge.Rmd
ik-converge to HTML

---
title: "Check integrated kernel convergence"
author: "Adam Howes"
output: html_document
---

```{r}
#' Checking the convergence of the integrated kernel as a function of number of samples per area

type <- "random"
L <- 200
grid <- readRDS("depends/grid.rds")
n <- nrow(grid)
samples <- sf::st_sample(grid, type = type, exact = TRUE, size = rep(L, n))
length(samples) # n * L
sf::st_crs(samples) <- NA

#' Add id column
samples <- st_sf(id = rep(seq_along(1:L), n), geom = samples)

sample_sets <- lapply(seq(10, L, by = 10), function(i) subset(samples, id <= i))

# Checking that the samples look OK
cowplot::plot_grid(
  plot_samples(sample_sets[[1]]),
  plot_samples(sample_sets[[10]]),
  plot_samples(sample_sets[[20]]),
  ncol = 3
)

#' Calculate the Gram matrices
covs <- lapply(sample_sets, adapted_integrated_covariance)

cowplot::plot_grid(
  plot_matrix(covs[[1]]),
  plot_matrix(covs[[10]]),
  plot_matrix(covs[[20]]),
  ncol = 3
)

matrix_comparison <- outer(covs, covs, Vectorize(frobenius))
plot_matrix(matrix_comparison)

#' Experiment: 10, 20 and 50 points versus ground truth 300 points
L_settings <- list(10, 20, 50)
gt_samples <- sf::st_sample(grid, type = type, exact = TRUE, size = rep(300, n))
gt_cov <- adapted_integrated_covariance(gt_samples)
result <- lapply(L_settings, int_convergence_experiment)

names(result) <- c("L_10", "L_20", "L_50")

result <- list_to_df(result) %>%
  pivot_longer(
    cols = 1:3,
    names_to = "L",
    names_prefix= "L_",
    values_to = "value"
  ) %>%
  mutate(L = as.numeric(L))

ggplot(result, aes(x = value)) +
  geom_histogram() +
  facet_grid(~L) +
  labs(x = "Frobenius norm", y = "") +
  theme_minimal()
```

### 2022-08-06 src/0_check_ik-converge/script.R

@@ -1,78 +1,4 @@
 # orderly::orderly_develop_start("0_check_ik-converge")
 # setwd("src/0_check_ik-converge")
 
-#' Checking the convergence of the integrated kernel as a function of number of samples per area
-
-type <- "random"
-L <- 200
-grid <- readRDS("depends/grid.rds")
-n <- nrow(grid)
-samples <- sf::st_sample(grid, type = type, exact = TRUE, size = rep(L, n))
-length(samples) # n * L
-sf::st_crs(samples) <- NA
-
-#' Add id column
-samples <- st_sf(id = rep(seq_along(1:L), n), geom = samples)
-
-sample_sets <- lapply(seq(10, L, by = 10), function(i) subset(samples, id <= i))
-
-# Checking that the samples look OK
-pdf("sample-check.pdf", h = 3, w = 6.25)
-
-cowplot::plot_grid(
-  plot_samples(sample_sets[[1]]),
-  plot_samples(sample_sets[[10]]),
-  plot_samples(sample_sets[[20]]),
-  ncol = 3
-)
-
-dev.off()
-
-#' Calculate the Gram matrices
-covs <- lapply(sample_sets, adapted_integrated_covariance)
-
-pdf("gram-matrices.pdf", h = 3, w = 6.25)
-
-cowplot::plot_grid(
-  plot_matrix(covs[[1]]),
-  plot_matrix(covs[[10]]),
-  plot_matrix(covs[[20]]),
-  ncol = 3
-)
-
-dev.off()
-
-matrix_comparison <- outer(covs, covs, Vectorize(frobenius))
-
-pdf("matrix-comparison.pdf", h = 5, w = 6.25)
-
-plot_matrix(matrix_comparison)
-
-dev.off()
-
-#' Experiment: 10, 20 and 50 points versus ground truth 300 points
-L_settings <- list(10, 20, 50)
-gt_samples <- sf::st_sample(grid, type = type, exact = TRUE, size = rep(300, n))
-gt_cov <- adapted_integrated_covariance(gt_samples)
-result <- lapply(L_settings, int_convergence_experiment)
-
-names(result) <- c("L_10", "L_20", "L_50")
-
-result <- list_to_df(result) %>%
-  pivot_longer(
-    cols = 1:3,
-    names_to = "L",
-    names_prefix= "L_",
-    values_to = "value"
-  ) %>%
-  mutate(L = as.numeric(L))
-
[…]

### 2022-08-06 src/0_demo_areal-kernels/areal-kernels.Rmd
areal-kernels to HTML

@@
-# Comparison
+---
+title: "Areal kernel demonstration"
+author: "Adam Howes"
+output: html_document
+---
+
+# Code to fit centroid kernel
 
 ```{r}
-fit_centroid <- readRDS("centroid.rds")
-fit_integrated <- readRDS("integrated.rds")
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+sf <- mw
+
+centroid_distance <- function(sf) {
+  cent <- sf::st_centroid(sf)
+  D <- sf::st_distance(cent, cent)
+  return(D)
+}
+
+D <- centroid_distance(sf)
+
+dat <- list(n = nrow(sf),
+            y = sf$y,
+            m = sf$n_obs,
+            mu = rep(0, nrow(sf)),
+            D = D)
+
+nsim_warm <- 100
+nsim_iter <- 1000
+
+fit_centroid <- rstan::stan(
+  "centroid.stan",
+  data = dat,
+  warmup = nsim_warm,
+  iter = nsim_iter
+)
 ```
+
+# Code to fit integrated kernel
+
+```{r}
+sf <- mw
+
+# The number of integration points within each area
+L <- 10
+
+# The method for selecting integration points
+type <- "hexagonal"
+
+# The number of areas
+n <- nrow(sf)
+
+# Draw L samples from each area according to method type
+samples <- sf::st_sample(sf, type = type, size = rep(L, n))
+
+plot_samples <- function(samples){
+  ggplot(sf) +
+    geom_sf(fill = "lightgrey") +
+    geom_sf(data = samples, alpha = 0.5, shape = 4) +
+    labs(x = "Longitude", y = "Latitude") +
+    theme_minimal() +
+    labs(fill = "") +
+    theme(axis.title = element_blank(),
+          axis.text = element_blank(),
+          axis.ticks = element_blank())
+}
+
+plot_samples(samples)
+
+# Construct an (L * n) * (L * n) matrix containing the Euclidean distance between each sample
+# Note that this ^ will not be exact for some settings of type (hexagonal, regular)
+S <- sf::st_distance(samples, samples)
+dim(S)
+
+# Data structure for unequal number of points in each area
+sample_index <- sf::st_intersects(sf, samples)
+sample_lengths <- lengths(sample_index)
+start_index <- sapply(sample_index, function(x) x[1])
+
+dat <- list(n = nrow(sf),
+            y = sf$y,
+            m = sf$n_obs,
+            mu = rep(0, nrow(sf)),
+            sample_lengths = sample_lengths,
+            total_samples = sum(sample_lengths),
+            start_index = start_index,
+            S = S)
+
+fit_integrated <- rstan::stan(
+  "integrated.stan",
+  data = dat,
+  warmup = nsim_warm,
+  iter = nsim_iter
+)
+```
+
+# Comparison
 

### 2022-08-06 src/0_demo_areal-kernels/centroid.R

@@ -1,28 +1,0 @@
-rstan_options(auto_write = TRUE)
-options(mc.cores = parallel::detectCores())
-
-sf <- mw
-
-centroid_distance <- function(sf) {
-  cent <- sf::st_centroid(sf)
-  D <- sf::st_distance(cent, cent)
-  return(D)
-}
-
-D <- centroid_distance(sf)
-
-dat <- list(n = nrow(sf),
-            y = sf$y,
-            m = sf$n_obs,
-            mu = rep(0, nrow(sf)),
-            D = D)
-
-nsim_warm <- 100
-nsim_iter <- 1000
-
-fit <- rstan::stan("centroid.stan",
-                   data = dat,
-                   warmup = nsim_warm,
-                   iter = nsim_iter)
-
-saveRDS(fit, file = "centroid.rds")

### 2022-08-06 src/0_demo_areal-kernels/script.R

@@ -1,9 +1,4 @@
 # orderly::orderly_develop_start("0_demo_areal-kernels")
 # setwd("src/0_demo_areal-kernels")
 
-#' Run models
-source("centroid.R")
-source("integrated.R")
-
-#' Create report
 rmarkdown::render("areal-kernels.Rmd")

### 2022-08-06 utils/archive_to_docs.R

@@ -20,4 +20,5 @@ archive_to_docs("docs_paper")
 archive_to_docs("0_check_bin-gaussian")
 archive_to_docs("0_check_ik-compute")
 archive_to_docs("0_check_ik-converge")
+archive_to_docs("0_demo_areal-kernels")
 archive_to_docs("0_explore_inla-spde")

### 2022-09-30 docs/resources.Rmd
Add resources from Notion

---
title: "Resources"
author: "Adam Howes"
date: "2022-09-30"
output: html_document
---

# GMRFs

* [Lecture 4: Numerical methods for sparse matrices](http://www.math.chalmers.se/~bodavid/GMRF2015/Lectures/F4slides.pdf)

# Cross-validation

* [Spatial leave-one-out blog post](https://davidrroberts.wordpress.com/2016/03/11/spatial-leave-one-out-sloo-cross-validation/)
* [mlr3spatiotempcv package](https://github.com/mlr-org/mlr3spatiotempcv)
* [https://github.com/paul-buerkner/brms/issues/637](https://github.com/paul-buerkner/brms/issues/637)
* [https://github.com/jayverhoef/spAReco](https://github.com/jayverhoef/spAReco)
* [Efficient LOO-CV, the `loo` package](http://mc-stan.org/loo/reference/loo-package.html)
* [Approximate cross‐validatory predictive checks in disease mapping models](https://onlinelibrary.wiley.com/doi/abs/10.1002/sim.1403?casa_token=G0o6vee_DZQAAAAA:WB3XhagKTyVQ0FQBtfzT8LOVvwZB0o4vywZhB-ulmgx5zCQ2-AOa2qdkgyA23m_7KnVacuoNC6XoMQ)
* [When LOO and other cross-validation approaches are valid](https://statmodeling.stat.columbia.edu/2018/08/03/loo-cross-validation-approaches-valid/)
* [I am the supercargo](https://statmodeling.stat.columbia.edu/2018/06/21/i-am-the-supercargo/)
* [Comments on Limitations of Bayesian Leave-One-Out Cross-Validation for Model Selection](https://statmodeling.stat.columbia.edu/2018/06/05/comments-limitations-bayesian-leave-one-cross-validation-model-selection/)
* [Bayesian leave-one-out cross-validation for large data](https://arxiv.org/pdf/1904.10679.pdf)

# INLA

* [Bayesian inference with INLA](https://becarioprecario.bitbucket.io/inla-gitbook/index.html)
* [The R-INLA package, David Bolin](https://sites.stat.washington.edu/peter/PASI/pasi_practical_intro.pdf)
* [Haakon Bakka](https://haakonbakkagit.github.io/index.html)
* Posterior predictive
    * [http://noserose.net/e/mmm3-blog/2018/03/17/posterior-predictive-in-inla/](http://noserose.net/e/mmm3-blog/2018/03/17/posterior-predictive-in-inla/)
    * [Julian Faraway](https://people.bath.ac.uk/jjf23/)
    * [Obtain summary of posterior predictive distribution, R-INLA Google group](https://groups.google.com/g/r-inla-discussion-group/c/2FYu8lAqid4)
* [Defining a latent model in R: The rgeneric model](https://raw.githubusercontent.com/inbo/INLA/master/vignettes/rgeneric.pdf)

# Stan

* [https://mc-stan.org/users/documentation/case-studies/mbjoseph-CARStan.html](https://mc-stan.org/users/documentation/case-studies/mbjoseph-CARStan.html)
* [https://mc-stan.org/users/documentation/case-studies/icar_stan.html](https://mc-stan.org/users/documentation/case-studies/icar_stan.html)
* [https://insp.memberclicks.net/assets/docs/Local_Events/2018/Paris/Stanforpmx2018_MitziMorris.pdf](https://insp.memberclicks.net/assets/docs/Local_Events/2018/Paris/Stanforpmx2018_MitziMorris.pdf)
* [https://statmodeling.stat.columbia.edu/2016/09/02/two-weird-tricks-for-fast-conditional-autoregressive-models-in-stan/](https://statmodeling.stat.columbia.edu/2016/09/02/two-weird-tricks-for-fast-conditional-autoregressive-models-in-stan/)
* [https://github.com/stan-dev/rstanarm/issues/207](https://github.com/stan-dev/rstanarm/issues/207)
* [https://mc-stan.org/loo/articles/loo2-with-rstan.html](https://mc-stan.org/loo/articles/loo2-with-rstan.html)

# Hilbert space methods

* [http://gpss.cc/gpa15/assets/sarkka.pdf](http://gpss.cc/gpa15/assets/sarkka.pdf)
* [https://discourse.mc-stan.org/t/practical-hilbert-space-approximate-bayesian-gaussian-processes-for-probabilistic-programming/14633](https://discourse.mc-stan.org/t/practical-hilbert-space-approximate-bayesian-gaussian-processes-for-probabilistic-programming/14633)
* [https://github.com/gabriuma/basis_functions_approach_to_GP/blob/master/uni_dimensional/simulated_dataset/stancode_BF_1dim.stan](https://github.com/gabriuma/basis_functions_approach_to_GP/blob/master/uni_dimensional/simulated_dataset/stancode_BF_1dim.stan)
* [https://arxiv.org/pdf/2004.11408.pdf](https://arxiv.org/pdf/2004.11408.pdf)

# Gaussian processes in Stan

* [https://github.com/paul-buerkner/brms/issues/234](https://github.com/paul-buerkner/brms/issues/234)
* [https://discourse.mc-stan.org/t/multilevel-gaussian-proccess-with-spectral-mixture-kernel/13335](https://discourse.mc-stan.org/t/multilevel-gaussian-proccess-with-spectral-mixture-kernel/13335)
* [https://github.com/paul-buerkner/brms/issues/540](https://github.com/paul-buerkner/brms/issues/540)
* [https://discourse.mc-stan.org/t/adding-gaussian-process-covariance-functions/237/12](https://discourse.mc-stan.org/t/adding-gaussian-process-covariance-functions/237/12)
* [https://discourse.mc-stan.org/t/approximate-gps-with-spectral-stuff/1041](https://discourse.mc-stan.org/t/approximate-gps-with-spectral-stuff/1041)
* [https://discourse.mc-stan.org/t/approximate-gp-prediction-question/11519](https://discourse.mc-stan.org/t/approximate-gp-prediction-question/11519)

# Non-stationarity

* [https://cran.r-project.org/web/packages/BayesNSGP/index.html](https://cran.r-project.org/web/packages/BayesNSGP/index.html)
* [https://arxiv.org/pdf/1910.14101.pdf](https://arxiv.org/pdf/1910.14101.pdf)

# R packages

* Both TMB and Stan working in an R package
    * [https://stackoverflow.com/questions/48627069/guidelines-for-including-tmb-c-code-in-an-r-package](https://stackoverflow.com/questions/48627069/guidelines-for-including-tmb-c-code-in-an-r-package)
    * [https://rtbecard.gitlab.io/2018/02/11/Distributing-TMB-in-R-packages.html](https://rtbecard.gitlab.io/2018/02/11/Distributing-TMB-in-R-packages.html)
    * [https://stackoverflow.com/questions/52537897/tmb-c-code-in-r-package-and-cran-checks](https://stackoverflow.com/questions/52537897/tmb-c-code-in-r-package-and-cran-checks)
    * [https://github.com/kaskr/adcomp/issues/43#issuecomment-204719483](https://github.com/kaskr/adcomp/issues/43#issuecomment-204719483)
* Formatting Rd files
    * [https://cran.r-project.org/web/packages/roxygen2/vignettes/rd-formatting.html#lists](https://cran.r-project.org/web/packages/roxygen2/vignettes/rd-formatting.html#lists)
* Getting a Linux R package to work on Windows
    * [https://stackoverflow.com/questions/1685181/making-r-package-work-in-both-windows-and-linux](https://stackoverflow.com/questions/1685181/making-r-package-work-in-both-windows-and-linux)
    * [https://win-builder.r-project.org/](https://win-builder.r-project.org/)
    * [https://seandavi.github.io/post/build-linux-r-binary-packages/](https://seandavi.github.io/post/build-linux-r-binary-packages/)
* [https://r-pkgs.org/description.html](https://r-pkgs.org/description.html)

# Weights

* [https://www.sciencedirect.com/science/article/pii/S0047259X12000589](https://www.sciencedirect.com/science/article/pii/S0047259X12000589)
* [Evaluating the effect of neighbourhood weight matrices on smoothing properties of Conditional Autoregressive (CAR) models](https://ij-healthgeographics.biomedcentral.com/articles/10.1186/1476-072X-6-54)
* [Spatial smoothing in Bayesian models: a comparison of weights matrix specifications and their impact on inference](https://ij-healthgeographics.biomedcentral.com/track/pdf/10.1186/s12942-017-0120-x.pdf)

# Gerrymandering

* [https://warwick.ac.uk/fac/sci/statistics/staff/academic-research/pollock/projects/m_gerrymandering.pdf](https://warwick.ac.uk/fac/sci/statistics/staff/academic-research/pollock/projects/m_gerrymandering.pdf)

# Model comparison

* [https://avehtari.github.io/modelselection/](https://avehtari.github.io/modelselection/)
* [https://radfordneal.wordpress.com/2009/03/07/does-coverage-matter/](https://radfordneal.wordpress.com/2009/03/07/does-coverage-matter/)
* [Simulation-based calibration](https://arxiv.org/abs/1804.06788)
* A random interval $(L(X), U(X))$ has $100(1 * \alpha)$% frequentist coverage for $\theta$ if, before the data are gathered $P(L(X) < \theta < U(X) | \theta) = 1 * \alpha$
* DIC
    * [https://www.mrc-bsu.cam.ac.uk/software/bugs/the-bugs-project-dic](https://www.mrc-bsu.cam.ac.uk/software/bugs/the-bugs-project-dic)

# Coresets

* [https://people.csail.mit.edu/tbroderick/nips_2016_spotlight_coresets_bayes.html](https://people.csail.mit.edu/tbroderick/nips_2016_spotlight_coresets_bayes.html)
* [Coreset (Wikipedia)](https://en.wikipedia.org/wiki/Coreset)
* [Automated Scalable Bayesian Inference via Hilbert Coresets](https://arxiv.org/abs/1710.05053)

# Modifiable areal unit problem

* [Flowerdew slides](https://www.ed.ac.uk/files/imports/fileManager/RFlowerdew_Slides.pdf)
* [The modifiable areal unit problem (MAUP) in physical geography](https://journals.sagepub.com/doi/pdf/10.1177/0309133307083294)
* [Neighbourhood effects on health: does it matter where you draw the boundaries?](https://pubmed.ncbi.nlm.nih.gov/18177988/)
* [Spatio-Temporal Change of Support Modeling with R](https://arxiv.org/pdf/1904.12092.pdf)

# Log scoring

* [Predictive Inference Based on Markov Chain Monte Carlo Output](https://arxiv.org/abs/1608.06802)

# Variograms

* [Optimal interpolation and isarithmic mapping of soil properties: I The semi‐variogram and punctual kriging](https://onlinelibrary.wiley.com/doi/epdf/10.1111/ejss.12784?saml_referrer)
* [Quantifying the scale effect in geospatial big data using semi-variograms](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0225139)
* `[gstat` package](https://cran.r-project.org/web/packages/gstat/vignettes/gstat.pdf)
* [Implementing variograms in R](https://www.aspexit.com/en/implementing-variograms-in-r/)
* [http://gsp.humboldt.edu/OLM/R/04_01_Variograms.html](http://gsp.humboldt.edu/OLM/R/04_01_Variograms.html)

"Some software for geostatistical analysis uses an empirical variogram to fit the parameters of the correlation function. However, this option does not make any assumptions about a likelihood function for the data or uses a multivariate Gaussian distribution for the spatially structured random effect. A good description of these techniques is available in Cressie (1993)."

# Chosing sample sizes

* [http://finzi.psych.upenn.edu/library/mcmcse/html/minESS.html](http://finzi.psych.upenn.edu/library/mcmcse/html/minESS.html)

# `xbinomial`

* Non-integer counts in `hiv_surveys$y` due to the use of survey weights
* Currently using `round(hiv_surveys$y)` but this is an unsatisfying solution
* In `stan` and `TMB` a custom continuous version of the binomial likelihood could be implemented
* In `R-INLA` there is the `xbinomial` function for non-integer counts
    * Be warned that this may make the marginal likelihood not interpretable as it is not clear what to do with the normalising constant
* Keywords to investigate
    * Bayesian pseudo likelihood
* [https://stats.stackexchange.com/questions/310676/continuous-generalization-of-the-negative-binomial-distribution](https://stats.stackexchange.com/questions/310676/continuous-generalization-of-the-negative-binomial-distribution)
* `R-INLA` Google group links
    * [https://groups.google.com/forum/#!searchin/r-inla-discussion-group/Jeff$20Eaton|sort:date/r-inla-discussion-group/FLpdv5jB90w/kSwlgxb9AgAJ](https://groups.google.com/forum/#!searchin/r-inla-discussion-group/Jeff$20Eaton%7Csort:date/r-inla-discussion-group/FLpdv5jB90w/kSwlgxb9AgAJ)
    * [https://groups.google.com/forum/#!topic/r-inla-discussion-group/k5KbmV3HKPU](https://groups.google.com/forum/#!topic/r-inla-discussion-group/k5KbmV3HKPU)

# Compact support kernels

* Empty so far!

# Multiple instance learning

* [Multiple Instance Learning with Bag Dissimilarities](https://arxiv.org/pdf/1309.5643.pdf)
* [Gaussian Processes Multiple Instance Learning](https://icml.cc/Conferences/2010/papers/412.pdf)
* [Variational Bayesian Multiple Instance Learning with Gaussian Processes](https://openaccess.thecvf.com/content_cvpr_2017/papers/Haussmann_Variational_Bayesian_Multiple_CVPR_2017_paper.pdf)

# Graph Laplacians

* https://csustan.csustan.edu/~tom/Clustering/GraphLaplacian-tutorial.pdf

# To categorise

* [Developing a Cancer Atlas using Bayesian Methods: A Practical Guide for Application and Interpretation](https://atlas.cancer.org.au/developing-a-cancer-atlas/Chapter_10.html#modelling)
    * [https://cancerqld.blob.core.windows.net/content/docs/Investigation-of-Bayesian-spatial-models.pdf](https://cancerqld.blob.core.windows.net/content/docs/Investigation-of-Bayesian-spatial-models.pdf)
* [A quick refresher on the coordinate reference system](https://s3.amazonaws.com/assets.datacamp.com/production/course_4422/slides/chapter2.pdf)
* [Variational Learning on Aggregate Outputs with Gaussian Processes](https://arxiv.org/pdf/1805.08463.pdf)
* [Multi-resolution Multi-task Gaussian Processes](https://arxiv.org/pdf/1906.08344.pdf)
* [A Case Study Competition Among Methods for Analyzing Large Spatial Data](https://imperiallondon-my.sharepoint.com/personal/sflaxman_ic_ac_uk/Documents/Microsoft%20Teams%20Chat%20Files/Heaton2019_Article_ACaseStudyCompetitionAmongMeth%20(1).pdf)
* [A spatial regression model for the disaggregation of areal unit based data to high-resolution grids with application to vaccination coverage mapping](https://journals.sagepub.com/doi/pdf/10.1177/0962280218797362)
* [Generalized spatial fusion model framework for joint analysis of point and areal data](https://www.sciencedirect.com/science/article/pii/S2211675317302531)
* [https://cran.r-project.org/web/packages/SUMMER/vignettes/summer-vignette.html](https://cran.r-project.org/web/packages/SUMMER/vignettes/summer-vignette.html)
* [Generalised inverse](https://www.math.wustl.edu/~sawyer/handouts/GenrlInv.pdf)
* [Spatial change-of-support and misalignment problems](https://sites.stat.washington.edu/peter/PASI/Craigmile_misalignment_problems.pdf)
* [https://docs.displayr.com/wiki/Design_Effects_and_Effective_Sample_Size](https://docs.displayr.com/wiki/Design_Effects_and_Effective_Sample_Size)
* [Hierarchical Bayesian species distribution models with the hSDMR Package](https://bioscenemada.cirad.fr/FileTransfer/hSDM-vignette.pdf)
* [[R-sig-Geo] Choosing the right nb2listw-style](https://stat.ethz.ch/pipermail/r-sig-geo/2010-August/009059.html)
* [Spatial weights objects as sparse matrices and graphs](https://cran.r-project.org/web/packages/spdep/vignettes/nb_igraph.html)
* [The Effect of Bias, Variance Estimation, Skewness and Kurtosis of the Empirical Logit on Weighted Least Squares Analyses on JSTOR](https://www.jstor.org/stable/2336348?casa_token=VEgq0pe9p8cAAAAA:Uo9-8vSQpPaw0KGjL7J4kEs7aBjIIuTNRFYiZPoGcDVtSQ7khYsHVwMbrqy5SIGBvz1s7DD2DPriO_ff8VkPoHiLeGHVk0Y36jA9BunjS1qcekIdPobO&seq=1)
* [Comparing Bayesian spatial models:Goodness-of-smoothing criteria for assessing under-and over-smoothin](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7239453/pdf/pone.0233019.pdf)g
* [change of support / moving average /kernel smoothing](https://groups.google.com/g/r-inla-discussion-group/c/SL5McDqCRyI/m/j1veDTtJaDQJ) INLA Google group

### 2022-10-07 src/0_check_bin-gaussian/bin-gaussian.Rmd
Iterate report

@@
 title: "Binomial of sum of Gaussians versus sum of binomial of Gaussians"
-author: "Adam Howes"
-output: html_document
+author:
+- name: Adam Howes
+output:
+  html_document:
+    toc: yes
+    toc_float: yes
+    toc_collapsed: yes
+    df_print: paged
+    code_folding: show
+    theme: lumen
+abstract: |
+  **Background**
+
+  **Task**
 ---
 
+## Set-up
+
+Consider an IID latent field.
+Without aggregation:
+$$
+\begin{align*}
+y &= \sum_{i = 1}^n y_i, \\
+y_i &\sim \text{Bin}(m, \rho_i), \\
+\rho_i &\sim \text{Logitnormal}(\mu, \sigma^2)
+\end{align*}
+$$
+With aggregation:
+$$
+\begin{align*}
+y &\sim \text{Bin}(nm, \bar \rho), \\
+\bar \rho &= \frac{1}{n} \sum_{i = 1}^n \rho_i, \\
+\rho_i &\sim \text{Logitnormal}(\mu, \sigma^2)
+\end{align*}
+$$
+
 ```{r}
-#' Binomial of sum of Gaussians versus sum of binomial of Gaussians
-#' IID latent field
-#' Without aggregation:
-#' `y = \sum_{i = 1}^n y_i$, $y_i \sim \text{Bin}(m, \rho_i)$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`
-#' With aggregation:
-#' `y \sim \text{Bin}(nm, \bar \rho)$, $\bar \rho = \frac{1}{n} \sum_{i = 1}^n \rho_i$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`
-
 experiment <- function(nsim, n, phi_mean, phi_sd, y_sample_size) {
@@
 n <- 36
+```
 
-#' Low prevalence setting (disease)
+## Low prevalence setting (disease)
+
+```{r}
 result <- experiment(nsim, n, -2.5, 0.5, 30)
@@
 sd(result$y_agg) / sd(result$y) #' Ratio
+```
 
-#' Probability approximately 1/2 setting (elections)
-#' As before but with different `phi_mean` settings:
+## Probability approximately 1/2 setting (elections)
+
+As before but with different `phi_mean` settings:
+
+```{r}
 result_half <- experiment(nsim, n, 0, 0.5, 30)
@@
 sd(result_half$y_agg) / sd(result_half$y) #' Ratio
+```
 
-#' Conclusions, things to add:
-#' * Standard deviation of aggregation tends to overestimate true standard deviation,
-#' * Effect is larger in the small `\rho` setting than the large `\rho` setting
-#' * Should test latent field with spatial structure e.g. ICAR latent field and so on
-#' * What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?
-```
+## Conclusions, things to add
+
+* Standard deviation of aggregation tends to overestimate true standard deviation,
+* Effect is larger in the small `rho` setting than the large `rho` setting
+* Should test latent field with spatial structure e.g. ICAR latent field and so on
+* What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?

### 2022-10-14 make/0_all.R
Create TikZ annotated Besag equation

@@ -18,5 +18,5 @@ run_commit_push("0_plot_areal-data")          #' [x]
 run_commit_push("0_plot_sloo-cv")             #' [x]
 run_commit_push("0_plot_geometry-graph")      #' [x]
 run_commit_push("0_plot_integration-method")  #' [x]
-run_commit_push("0_plot-tikz_besag")          #' [/]
+run_commit_push("0_plot-tikz_besag")          #' [x]
 run_commit_push("0_plot-tikz_maup")           #' [x]

### 2022-10-14 src/0_plot-tikz_besag/script.R

@@ -3,3 +3,4 @@
 # setwd("src/0_plot-tikz_besag")
 
 tools::texi2dvi("besag.tex", pdf = TRUE, clean = TRUE)
+knitr::plot_crop("besag.pdf")

### 2022-10-27 docs/resources.Rmd
Update resources

@@
 ---
-title: "Resources"
-author: "Adam Howes"
-date: "2022-09-30"
-output: html_document
+title: "Resources on spatial models"
+author:
+- name: Adam Howes
+output:
+  html_document:
+    toc: yes
+    toc_float: yes
+    toc_collapsed: yes
+    df_print: paged
+    code_folding: show
+    theme: lumen
+abstract: |
+  A compliation of links to resources which I've found useful while working on `areal-comparison`.
 ---
@@
 * [mlr3spatiotempcv package](https://github.com/mlr-org/mlr3spatiotempcv)
-* [https://github.com/paul-buerkner/brms/issues/637](https://github.com/paul-buerkner/brms/issues/637)
-* [https://github.com/jayverhoef/spAReco](https://github.com/jayverhoef/spAReco)
+* [Cross-validation of CAR models cannot be done with stratified folds](https://github.com/paul-buerkner/brms/issues/637)
+* [An R package in support of publication, "Spatial Autoregressive Models for Statistical Inference from Ecological Data."](https://github.com/jayverhoef/spAReco)
 * [Efficient LOO-CV, the `loo` package](http://mc-stan.org/loo/reference/loo-package.html)
@@
 * Posterior predictive
-    * [http://noserose.net/e/mmm3-blog/2018/03/17/posterior-predictive-in-inla/](http://noserose.net/e/mmm3-blog/2018/03/17/posterior-predictive-in-inla/)
+    * [Posterior Predictive in INLA](http://noserose.net/e/mmm3-blog/2018/03/17/posterior-predictive-in-inla/)
     * [Julian Faraway](https://people.bath.ac.uk/jjf23/)
@@
 
+# Jeff's notebooks
+
+* [INLA sandbox](https://athowes.github.io/inla-sandbox/)
+
 # Stan
 
-* [https://mc-stan.org/users/documentation/case-studies/mbjoseph-CARStan.html](https://mc-stan.org/users/documentation/case-studies/mbjoseph-CARStan.html)
-* [https://mc-stan.org/users/documentation/case-studies/icar_stan.html](https://mc-stan.org/users/documentation/case-studies/icar_stan.html)
-* [https://insp.memberclicks.net/assets/docs/Local_Events/2018/Paris/Stanforpmx2018_MitziMorris.pdf](https://insp.memberclicks.net/assets/docs/Local_Events/2018/Paris/Stanforpmx2018_MitziMorris.pdf)
-* [https://statmodeling.stat.columbia.edu/2016/09/02/two-weird-tricks-for-fast-conditional-autoregressive-models-in-stan/](https://statmodeling.stat.columbia.edu/2016/09/02/two-weird-tricks-for-fast-conditional-autoregressive-models-in-stan/)
-* [https://github.com/stan-dev/rstanarm/issues/207](https://github.com/stan-dev/rstanarm/issues/207)
-* [https://mc-stan.org/loo/articles/loo2-with-rstan.html](https://mc-stan.org/loo/articles/loo2-with-rstan.html)
+* [Exact sparse CAR models in Stan](https://mc-stan.org/users/documentation/case-studies/mbjoseph-CARStan.html)
+* [Spatial Models in Stan: Intrinsic Auto-Regressive Models for Areal Data](https://mc-stan.org/users/documentation/case-studies/icar_stan.html)
+* [Spatial Smoothing in Stan: Conditional Auto-Regressive Models](https://insp.memberclicks.net/assets/docs/Local_Events/2018/Paris/Stanforpmx2018_MitziMorris.pdf)
+* [Fast CAR: Two weird tricks for fast conditional autoregressive models in Stan](https://statmodeling.stat.columbia.edu/2016/09/02/two-weird-tricks-for-fast-conditional-autoregressive-models-in-stan/)
+* [feature/car in `rstanarm``](https://github.com/stan-dev/rstanarm/issues/207)
+* [Writing Stan programs for use with the loo package](https://mc-stan.org/loo/articles/loo2-with-rstan.html)
+
+## Gaussian processes in Stan
+
+* [Add more Gaussian process kernels in `brms`](https://github.com/paul-buerkner/brms/issues/234)
+* [Multilevel Gaussian Proccess with spectral mixture kernel](https://discourse.mc-stan.org/t/multilevel-gaussian-proccess-with-spectral-mixture-kernel/13335)
+* [Basis function approximation to Gaussian processes in `brms`](https://github.com/paul-buerkner/brms/issues/540)
+* [Adding Gaussian Process Covariance Functions](https://discourse.mc-stan.org/t/adding-gaussian-process-covariance-functions/237/12)
+* [Approximate GPs with Spectral Stuff](https://discourse.mc-stan.org/t/approximate-gps-with-spectral-stuff/1041)
+* [Approximate GP prediction question](https://discourse.mc-stan.org/t/approximate-gp-prediction-question/11519)
 
@@
 
-* [http://gpss.cc/gpa15/assets/sarkka.pdf](http://gpss.cc/gpa15/assets/sarkka.pdf)
-* [https://discourse.mc-stan.org/t/practical-hilbert-space-approximate-bayesian-gaussian-processes-for-probabilistic-programming/14633](https://discourse.mc-stan.org/t/practical-hilbert-space-approximate-bayesian-gaussian-processes-for-probabilistic-programming/14633)
-* [https://github.com/gabriuma/basis_functions_approach_to_GP/blob/master/uni_dimensional/simulated_dataset/stancode_BF_1dim.stan](https://github.com/gabriuma/basis_functions_approach_to_GP/blob/master/uni_dimensional/simulated_dataset/stancode_BF_1dim.stan)
-* [https://arxiv.org/pdf/2004.11408.pdf](https://arxiv.org/pdf/2004.11408.pdf)
-
-# Gaussian processes in Stan
-
-* [https://github.com/paul-buerkner/brms/issues/234](https://github.com/paul-buerkner/brms/issues/234)
-* [https://discourse.mc-stan.org/t/multilevel-gaussian-proccess-with-spectral-mixture-kernel/13335](https://discourse.mc-stan.org/t/multilevel-gaussian-proccess-with-spectral-mixture-kernel/13335)
-* [https://github.com/paul-buerkner/brms/issues/540](https://github.com/paul-buerkner/brms/issues/540)
-* [https://discourse.mc-stan.org/t/adding-gaussian-process-covariance-functions/237/12](https://discourse.mc-stan.org/t/adding-gaussian-process-covariance-functions/237/12)
-* [https://discourse.mc-stan.org/t/approximate-gps-with-spectral-stuff/1041](https://discourse.mc-stan.org/t/approximate-gps-with-spectral-stuff/1041)
-* [https://discourse.mc-stan.org/t/approximate-gp-prediction-question/11519](https://discourse.mc-stan.org/t/approximate-gp-prediction-question/11519)
+* [Hilbert Space Methods for Reduced-Rank Gaussian Process Regression](http://gpss.cc/gpa15/assets/sarkka.pdf)
+* [Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming](https://discourse.mc-stan.org/t/practical-hilbert-space-approximate-bayesian-gaussian-processes-for-probabilistic-programming/14633)
+* [`stancode_BF_1dim.stan`](https://github.com/gabriuma/basis_functions_approach_to_GP/blob/master/uni_dimensional/simulated_dataset/stancode_BF_1dim.stan)
 
@@
 
-* [https://cran.r-project.org/web/packages/BayesNSGP/index.html](https://cran.r-project.org/web/packages/BayesNSGP/index.html)
-* [https://arxiv.org/pdf/1910.14101.pdf](https://arxiv.org/pdf/1910.14101.pdf)
+* [BayesNSGP: Bayesian Analysis of Non-Stationary Gaussian Process Models](https://cran.r-project.org/web/packages/BayesNSGP/index.html)
+* [Bayesian inference for high-dimensional nonstationary Gaussian processes](https://arxiv.org/pdf/1910.14101.pdf)
 
@@
 * Both TMB and Stan working in an R package
-    * [https://stackoverflow.com/questions/48627069/guidelines-for-including-tmb-c-code-in-an-r-package](https://stackoverflow.com/questions/48627069/guidelines-for-including-tmb-c-code-in-an-r-package)
-    * [https://rtbecard.gitlab.io/2018/02/11/Distributing-TMB-in-R-packages.html](https://rtbecard.gitlab.io/2018/02/11/Distributing-TMB-in-R-packages.html)
-    * [https://stackoverflow.com/questions/52537897/tmb-c-code-in-r-package-and-cran-checks](https://stackoverflow.com/questions/52537897/tmb-c-code-in-r-package-and-cran-checks)
-    * [https://github.com/kaskr/adcomp/issues/43#issuecomment-204719483](https://github.com/kaskr/adcomp/issues/43#issuecomment-204719483)
+    * [Guidelines for including TMB c++ code in an R package](https://stackoverflow.com/questions/48627069/guidelines-for-including-tmb-c-code-in-an-r-package)
+    * [TMB C++ code in R package and CRAN checks](https://stackoverflow.com/questions/52537897/tmb-c-code-in-r-package-and-cran-checks)
+    * [Best practices for distributing TMB and R code](https://github.com/kaskr/adcomp/issues/43#issuecomment-204719483)
 * Formatting Rd files
-    * [https://cran.r-project.org/web/packages/roxygen2/vignettes/rd-formatting.html#lists](https://cran.r-project.org/web/packages/roxygen2/vignettes/rd-formatting.html#lists)
+    * [(R)Markdown support](https://cran.r-project.org/web/packages/roxygen2/vignettes/rd-formatting.html#lists)
 * Getting a Linux R package to work on Windows
-    * [https://stackoverflow.com/questions/1685181/making-r-package-work-in-both-windows-and-linux](https://stackoverflow.com/questions/1685181/making-r-package-work-in-both-windows-and-linux)
-    * [https://win-builder.r-project.org/](https://win-builder.r-project.org/)
-    * [https://seandavi.github.io/post/build-linux-r-binary-packages/](https://seandavi.github.io/post/build-linux-r-binary-packages/)
-* [https://r-pkgs.org/description.html](https://r-pkgs.org/description.html)
+    * [Making R package work in both Windows and Linux](https://stackoverflow.com/questions/1685181/making-r-package-work-in-both-windows-and-linux)
+    * [Building and checking R source packages for Windows](https://win-builder.r-project.org/)
+    * [Building R Binary Packages for Linux](https://seandavi.github.io/post/build-linux-r-binary-packages/)
+* [R Packages (2e)](https://r-pkgs.org/)
 
-# Weights
+# Neighbourhood weights
 
-* [https://www.sciencedirect.com/science/article/pii/S0047259X12000589](https://www.sciencedirect.com/science/article/pii/S0047259X12000589)
+* [Bayesian spatial models with a mixture neighborhood structure](https://www.sciencedirect.com/science/article/pii/S0047259X12000589)
 * [Evaluating the effect of neighbourhood weight matrices on smoothing properties of Conditional Autoregressive (CAR) models](https://ij-healthgeographics.biomedcentral.com/articles/10.1186/1476-072X-6-54)
 * [Spatial smoothing in Bayesian models: a comparison of weights matrix specifications and their impact on inference](https://ij-healthgeographics.biomedcentral.com/track/pdf/10.1186/s12942-017-0120-x.pdf)
-
-# Gerrymandering
-
-* [https://warwick.ac.uk/fac/sci/statistics/staff/academic-research/pollock/projects/m_gerrymandering.pdf](https://warwick.ac.uk/fac/sci/statistics/staff/academic-research/pollock/projects/m_gerrymandering.pdf)
+* [Spatial weights objects as sparse matrices and graphs](https://cran.r-project.org/web/packages/spdep/vignettes/nb_igraph.html)
 
@@
 
-* [https://avehtari.github.io/modelselection/](https://avehtari.github.io/modelselection/)
-* [https://radfordneal.wordpress.com/2009/03/07/does-coverage-matter/](https://radfordneal.wordpress.com/2009/03/07/does-coverage-matter/)
+* [Model selection tutorials and talks](https://avehtari.github.io/modelselection/)
+* [Does coverage matter?](https://radfordneal.wordpress.com/2009/03/07/does-coverage-matter/)
 * [Simulation-based calibration](https://arxiv.org/abs/1804.06788)
 * A random interval $(L(X), U(X))$ has $100(1 * \alpha)$% frequentist coverage for $\theta$ if, before the data are gathered $P(L(X) < \theta < U(X) | \theta) = 1 * \alpha$
-* DIC
-    * [https://www.mrc-bsu.cam.ac.uk/software/bugs/the-bugs-project-dic](https://www.mrc-bsu.cam.ac.uk/software/bugs/the-bugs-project-dic)
+* [DIC: Deviance Information Criteria](https://www.mrc-bsu.cam.ac.uk/software/bugs/the-bugs-project-dic)
 
@@
 
-* [https://people.csail.mit.edu/tbroderick/nips_2016_spotlight_coresets_bayes.html](https://people.csail.mit.edu/tbroderick/nips_2016_spotlight_coresets_bayes.html)
+* [Coresets for automated, scalable Bayesian inference](https://tamarabroderick.com/files/broderick_2018_coresets.pdf)
 * [Coreset (Wikipedia)](https://en.wikipedia.org/wiki/Coreset)
@@
 * [Spatio-Temporal Change of Support Modeling with R](https://arxiv.org/pdf/1904.12092.pdf)
+* [Spatial change-of-support and misalignment problems](https://sites.stat.washington.edu/peter/PASI/Craigmile_misalignment_problems.pdf)
+* [Change of support / moving average /kernel smoothing](https://groups.google.com/g/r-inla-discussion-group/c/SL5McDqCRyI/m/j1veDTtJaDQJ) INLA Google group
 
@@
 * [Quantifying the scale effect in geospatial big data using semi-variograms](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0225139)
-* `[gstat` package](https://cran.r-project.org/web/packages/gstat/vignettes/gstat.pdf)
+* [`gstat` package](https://cran.r-project.org/web/packages/gstat/vignettes/gstat.pdf)
 * [Implementing variograms in R](https://www.aspexit.com/en/implementing-variograms-in-r/)
-* [http://gsp.humboldt.edu/OLM/R/04_01_Variograms.html](http://gsp.humboldt.edu/OLM/R/04_01_Variograms.html)
+* [R Variograms & Kriging](http://gsp.humboldt.edu/OLM/R/04_01_Variograms.html)
 
@@
 
-# Chosing sample sizes
+# Chosing MCMC sample sizes
 
-* [http://finzi.psych.upenn.edu/library/mcmcse/html/minESS.html](http://finzi.psych.upenn.edu/library/mcmcse/html/minESS.html)
+* [MCMC Output Analysis with R package mcmcse](https://warwick.ac.uk/fac/sci/wdsi/events/wrug/resources/mcmcse.pdf)
+  * See the function `mcmcse::minESS`
 
-# `xbinomial`
+# Survey weights
 
-* Non-integer counts in `hiv_surveys$y` due to the use of survey weights
-* Currently using `round(hiv_surveys$y)` but this is an unsatisfying solution
-* In `stan` and `TMB` a custom continuous version of the binomial likelihood could be implemented
-* In `R-INLA` there is the `xbinomial` function for non-integer counts
-    * Be warned that this may make the marginal likelihood not interpretable as it is not clear what to do with the normalising constant
+* [Design Effects and Effective Sample Size](https://docs.displayr.com/wiki/Design_Effects_and_Effective_Sample_Size)
+
+## Non-integer pseduo-likelihoods
+
+* Non-integer counts due to the use of survey weights. Could use `round` but this is an unsatisfying solution
+* In `stan` and `TMB` a custom continuous version of the likelihood could be implemented
+* In `R-INLA` there are `xbinomial` and `xpoisson` functions for non-integer counts
+    * Be warned that these may make the marginal likelihood not interpretable as it is not clear what to do with the normalising constant
 * Keywords to investigate
     * Bayesian pseudo likelihood
-* [https://stats.stackexchange.com/questions/310676/continuous-generalization-of-the-negative-binomial-distribution](https://stats.stackexchange.com/questions/310676/continuous-generalization-of-the-negative-binomial-distribution)
+* [Continuous generalization of the negative binomial distribution](https://stats.stackexchange.com/questions/310676/continuous-generalization-of-the-negative-binomial-distribution)
 * `R-INLA` Google group links
-    * [https://groups.google.com/forum/#!searchin/r-inla-discussion-group/Jeff$20Eaton|sort:date/r-inla-discussion-group/FLpdv5jB90w/kSwlgxb9AgAJ](https://groups.google.com/forum/#!searchin/r-inla-discussion-group/Jeff$20Eaton%7Csort:date/r-inla-discussion-group/FLpdv5jB90w/kSwlgxb9AgAJ)
-    * [https://groups.google.com/forum/#!topic/r-inla-discussion-group/k5KbmV3HKPU](https://groups.google.com/forum/#!topic/r-inla-discussion-group/k5KbmV3HKPU)
+    * [Non-integer counts for binomial model](https://groups.google.com/forum/#!searchin/r-inla-discussion-group/Jeff$20Eaton%7Csort:date/r-inla-discussion-group/FLpdv5jB90w/kSwlgxb9AgAJ)
+    * [what is binomial distribution doing with non-integer data](https://groups.google.com/forum/#!topic/r-inla-discussion-group/k5KbmV3HKPU)
 
@@
 
-* https://csustan.csustan.edu/~tom/Clustering/GraphLaplacian-tutorial.pdf
+* [A Short Tutorial on Graph Laplacians, Laplacian Embedding, and Spectral Clustering](https://csustan.csustan.edu/~tom/Clustering/GraphLaplacian-tutorial.pdf)
 
-# To categorise
+# Aggregated Gaussian processes
+
+* [Variational Learning on Aggregate Outputs with Gaussian Processes](https://arxiv.org/pdf/1805.08463.pdf)
+* [Multi-resolution Multi-task Gaussian Processes](https://arxiv.org/pdf/1906.08344.pdf)
+
+# Misc.
 
 * [Developing a Cancer Atlas using Bayesian Methods: A Practical Guide for Application and Interpretation](https://atlas.cancer.org.au/developing-a-cancer-atlas/Chapter_10.html#modelling)
-    * [https://cancerqld.blob.core.windows.net/content/docs/Investigation-of-Bayesian-spatial-models.pdf](https://cancerqld.blob.core.windows.net/content/docs/Investigation-of-Bayesian-spatial-models.pdf)
+  * [Investigation of Bayesian spatial models](https://cancerqld.blob.core.windows.net/content/docs/Investigation-of-Bayesian-spatial-models.pdf)
 * [A quick refresher on the coordinate reference system](https://s3.amazonaws.com/assets.datacamp.com/production/course_4422/slides/chapter2.pdf)
-* [Variational Learning on Aggregate Outputs with Gaussian Processes](https://arxiv.org/pdf/1805.08463.pdf)
-* [Multi-resolution Multi-task Gaussian Processes](https://arxiv.org/pdf/1906.08344.pdf)
 * [A Case Study Competition Among Methods for Analyzing Large Spatial Data](https://imperiallondon-my.sharepoint.com/personal/sflaxman_ic_ac_uk/Documents/Microsoft%20Teams%20Chat%20Files/Heaton2019_Article_ACaseStudyCompetitionAmongMeth%20(1).pdf)
@@
 * [Generalized spatial fusion model framework for joint analysis of point and areal data](https://www.sciencedirect.com/science/article/pii/S2211675317302531)
-* [https://cran.r-project.org/web/packages/SUMMER/vignettes/summer-vignette.html](https://cran.r-project.org/web/packages/SUMMER/vignettes/summer-vignette.html)
+* [SUMMER: Bayesian Small Area Estimation using Complex Survey Data](https://cran.r-project.org/web/packages/SUMMER/vignettes/summer-vignette.html)
 * [Generalised inverse](https://www.math.wustl.edu/~sawyer/handouts/GenrlInv.pdf)
-* [Spatial change-of-support and misalignment problems](https://sites.stat.washington.edu/peter/PASI/Craigmile_misalignment_problems.pdf)
-* [https://docs.displayr.com/wiki/Design_Effects_and_Effective_Sample_Size](https://docs.displayr.com/wiki/Design_Effects_and_Effective_Sample_Size)
 * [Hierarchical Bayesian species distribution models with the hSDMR Package](https://bioscenemada.cirad.fr/FileTransfer/hSDM-vignette.pdf)
 * [[R-sig-Geo] Choosing the right nb2listw-style](https://stat.ethz.ch/pipermail/r-sig-geo/2010-August/009059.html)
-* [Spatial weights objects as sparse matrices and graphs](https://cran.r-project.org/web/packages/spdep/vignettes/nb_igraph.html)
 * [The Effect of Bias, Variance Estimation, Skewness and Kurtosis of the Empirical Logit on Weighted Least Squares Analyses on JSTOR](https://www.jstor.org/stable/2336348?casa_token=VEgq0pe9p8cAAAAA:Uo9-8vSQpPaw0KGjL7J4kEs7aBjIIuTNRFYiZPoGcDVtSQ7khYsHVwMbrqy5SIGBvz1s7DD2DPriO_ff8VkPoHiLeGHVk0Y36jA9BunjS1qcekIdPobO&seq=1)
-* [Comparing Bayesian spatial models:Goodness-of-smoothing criteria for assessing under-and over-smoothin](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7239453/pdf/pone.0233019.pdf)g
-* [change of support / moving average /kernel smoothing](https://groups.google.com/g/r-inla-discussion-group/c/SL5McDqCRyI/m/j1veDTtJaDQJ) INLA Google group
+* [Comparing Bayesian spatial models: Goodness-of-smoothing criteria for assessing under-and over-smoothing](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7239453/pdf/pone.0233019.pdf)g

### 2022-10-27 docs/resources.Rmd
Typo

@@
 * [Fast CAR: Two weird tricks for fast conditional autoregressive models in Stan](https://statmodeling.stat.columbia.edu/2016/09/02/two-weird-tricks-for-fast-conditional-autoregressive-models-in-stan/)
-* [feature/car in `rstanarm``](https://github.com/stan-dev/rstanarm/issues/207)
+* [feature/car in `rstanarm`](https://github.com/stan-dev/rstanarm/issues/207)
 * [Writing Stan programs for use with the loo package](https://mc-stan.org/loo/articles/loo2-with-rstan.html)

### 2022-10-27 src/0_check_ik-compute/ik-compute.Rmd
Update bin-gaussian and ik-compute

@@
 title: "Check integrated kernel computation"
-author: "Adam Howes"
-output: html_document
+author:
+- name: Adam Howes
+output:
+  html_document:
+    toc: yes
+    toc_float: yes
+    toc_collapsed: yes
+    df_print: paged
+    code_folding: show
+    theme: lumen
+abstract: |
+  **Task** We check that the Stan code to calculate the integrated kernel is correct.
 ---
@@
 ```{r}
-#' Checking that the Stan code to calculate the integrated kernel is correct
-
 sf <- mw
@@
 start_index <- sapply(sample_index, function(x) x[1])
+```
 
-#' First calculate it directly using R only
+First, calculate the integrated kernel directly using R only:
+
+```{r}
 D <- sf::st_distance(samples, samples)
@@
 cov_r <- K
+```
 
-#' Now calculate it in Stan via two different methods
-#' rstan::expose_stan_functions puts the C++ functions into the R local workspace to use
+Now, calculate it in Stan via two different methods.
+We use `rstan::expose_stan_functions` to put the C++ functions into the R local workspace to use:
+
+```{r}
 rstan::expose_stan_functions("cov_sample_average.stan")
+```
 
-#' 1. Newer method
+Newer method (using `cov_sample_average`):
+
+```{r}
 cov_stan <- cov_sample_average(
@@
 )
+```
 
-#' 2. Old method
+Old method (using `cov_sample_average_old`):
+
+```{r}
 cov_stan_old <- cov_sample_average_old(
@@
 )
+```
 
-#' Comparison (these statements should all be TRUE)
+Comparison (these statements should all be TRUE):
+
+```{r}
 all.equal(cov_r, cov_stan)
@@
 all.equal(cov_stan, cov_stan_old)
+```
 
-#' Hard to tell from an image if something is identical but anyway
+And here are the images (it's hard to tell from an image if something is identical, but anyway):
+
+```{r}
+
 cowplot::plot_grid(
@@
 ```
+
+# `cov_sample_average.stan`
+
+```{cpp, echo=TRUE, eval=FALSE, output.var="ex", code=readLines('cov_sample_average.stan')}
+```
+
+# Original computing environment
+
+```{r}
+sessionInfo()
+```

### 2022-10-27 src/0_check_ik-converge/ik-converge.Rmd
Polish ik-convergence

@@
 title: "Check integrated kernel convergence"
-author: "Adam Howes"
-output: html_document
+author:
+- name: Adam Howes
+output:
+  html_document:
+    toc: yes
+    toc_float: yes
+    toc_collapsed: yes
+    df_print: paged
+    code_folding: show
+    theme: lumen
+abstract: |
+  **Task** We check how the convergence of the integrated kernel varies as a function of number of samples per area.
 ---
 
+# Performing the check
+
 ```{r}
-#' Checking the convergence of the integrated kernel as a function of number of samples per area
-
 type <- "random"
@@
 sf::st_crs(samples) <- NA
+```
 
-#' Add id column
+Add an `id` column:
+
+```{r}
 samples <- st_sf(id = rep(seq_along(1:L), n), geom = samples)
@@
 sample_sets <- lapply(seq(10, L, by = 10), function(i) subset(samples, id <= i))
+```
 
-# Checking that the samples look OK
+Check that the samples look OK:
+
+```{r}
 cowplot::plot_grid(
@@
 )
+```
 
-#' Calculate the Gram matrices
+Calculate the Gram matrices:
+
+```{r}
 covs <- lapply(sample_sets, adapted_integrated_covariance)
@@
 plot_matrix(matrix_comparison)
+```
 
-#' Experiment: 10, 20 and 50 points versus ground truth 300 points
+Perform an experiment with 10, 20 and 50 points versus ground truth 300 points:
+
+```{r}
 L_settings <- list(10, 20, 50)
@@
 ```
+
+# `functions.R`
+
+```{r, echo=TRUE, eval=FALSE, code=readLines('functions.R')}
+```
+
+# Original computing environment
+
+```{r}
+sessionInfo()
+```

### 2022-10-27 src/0_check_xbinomial-sampling/xbinom.Rmd
Not verbose please!

---
title: "`xbinomial` sampling"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Task** We test sampling from the custom likelihoods `xbinomial_lpdf` and `xbinomial_logit_lpdf`.
---

# `xbinomial_lpdf`

```{r}
temp1 <- rstan::stan(file = "xbinom.stan", data = list(m = 10, rho = 0.5), warmup = 500, iter = 1000, refresh = 0)
rstan::summary(temp1)$summary
out1 <- rstan::extract(temp1)
plot(out1$y)
hist(out1$y)
```

# `xbinomial_logit_lpdf`

```{r}
temp2 <- rstan::stan(file = "xbinom_logit.stan", data = list(m = 10, eta = 0), warmup = 500, iter = 1000, refresh = 0)
rstan::summary(temp2)$summary
out2 <- rstan::extract(temp2)
plot(out2$y)
hist(out2$y)
```

# Comparison

```{r}
df1 <- data.frame(x = out1$y)
df2 <- data.frame(x = out2$y)

ks.test(out1$y, out2$y)

ggplot() +
  stat_ecdf(data = df1, aes(x = x), geom = "step") +
  stat_ecdf(data = df2, aes(x = x), geom = "step", col = "red")
```

# Stan code

## `xbinom.stan`

```{cpp, echo=TRUE, eval=FALSE, output.var="ex", code=readLines('xbinom.stan')}
```

## `xbinom_logit.stan`

```{cpp, echo=TRUE, eval=FALSE, output.var="ex", code=readLines('xbinom_logit.stan')}
```

# Original computing environment

```{r}
sessionInfo()
```

### 2022-10-27 src/0_demo_areal-kernels/areal-kernels.Rmd
Improve areal-kernels

@@
 title: "Areal kernel demonstration"
-author: "Adam Howes"
-output: html_document
+author:
+- name: Adam Howes
+output:
+  html_document:
+    toc: yes
+    toc_float: yes
+    toc_collapsed: yes
+    df_print: paged
+    code_folding: show
+    theme: lumen
+abstract: |
+  **Task** We briefly demonstrate the fitting of areal kernel models in Stan.
 ---
@@
 sf <- mw
+```
 
-# The number of integration points within each area
+The number of integration points within each area:
+
+```{r}
 L <- 10
+```
 
-# The method for selecting integration points
+The method for selecting integration points:
+
+```{r}
 type <- "hexagonal"
+```
 
-# The number of areas
+The number of areas:
+
+```{r}
 n <- nrow(sf)
+```
 
-# Draw L samples from each area according to method type
+Draw $L$ samples from each area according to method type:
+
+```{r}
 samples <- sf::st_sample(sf, type = type, size = rep(L, n))
@@
 plot_samples(samples)
+```
 
-# Construct an (L * n) * (L * n) matrix containing the Euclidean distance between each sample
-# Note that this ^ will not be exact for some settings of type (hexagonal, regular)
+Construct an $nL \times nL$ matrix containing the Euclidean distance between each sample.
+Note that this the $L$ actually obtained will not be exact for some settings of type (`hexagonal`, `regular`):
+
+```{r}
 S <- sf::st_distance(samples, samples)
 dim(S)
+```
 
-# Data structure for unequal number of points in each area
+Data structure for unequal number of points in each area:
+
+```{r}
 sample_index <- sf::st_intersects(sf, samples)
@@
 ```
+
+# Stan code
+
+## `centroid.stan`
+
+```{cpp, echo=TRUE, eval=FALSE, output.var="ex", code=readLines('centroid.stan')}
+```
+
+## `integrated.stan`
+
+```{cpp, echo=TRUE, eval=FALSE, output.var="ex", code=readLines('integrated.stan')}
+```
+
+# Original computing environment
+
+```{r}
+sessionInfo()
+```

### 2022-10-27 src/0_demo_areal-kernels/areal-kernels.Rmd
Move unequal-samples development to a note in areal-kernels

@@
 
-Data structure for unequal number of points in each area:
+We use a start index data structure for unequal number of points in each area.
+Note that other possible ways to do this include a padded data structure (`matrix(unlist(padded_sample_index), nrow = n, ncol = max_length, byrow = TRUE)`) and a database structure `group_db <- data.frame(i = unlist(sample_index), group_id = rep(seq(length(sample_index)), lengths(sample_index)))`).
 

### 2022-10-28 src/0_explore_wilson-pointless/pointless.Rmd
Add pointless

---
title: "Pointless spatial modeling"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Task** We explore the code from the paper "Pointless spatial modeling" by Katie Wilson and Jon Wakefield.
---

# Code

```{r}
set.seed(10)
load("kenya.rdata")

#' Read in population data
#' Follow instructions in pointless-spatial-modeling README as to how to get this
pop.dat <- raster("gpw_v4_population_count_rev11_2015_2pt5_min.tif")
pop.dat <- setMinMax(pop.dat)
kenya.extent <- extent(33.9, 42, -4.7, 5.1)
pop.dat.kenya <- crop(pop.dat, kenya.extent)

kenya.df1 <- fortify(adm0) %>%
  mutate(id = as.numeric(id))

kenya.df47 <- fortify(adm1) %>%
  mutate(id = as.numeric(id))

kenya.df8 <- fortify(adm0.5) %>%
  mutate(id = as.numeric(id) + 1)

#' Set-up variables for simulation
#' * Create mesh for first 5 scenarios: "mesh.true"
#' * Create D matrix used for scenario 2 and 3: "D" (47 x m)
#' * Create D matrix used for scenario 4 and 5: "D8" (see note above)
#' * Add columns to kenya.data indicating which polygons the EAs reside in

# Create internal mesh points
points.kenya <- spsample(adm0, 2500, "regular")

boundary.kenya <- inla.nonconvex.hull(
  points.kenya@coords,
  convex = 0.1,
  resolution = c(120, 120)
)

#' Determine population near these mesh points
pop.at.meshX <- raster::extract(pop.dat.kenya, points.kenya@coords)
pop.at.meshX[is.na(pop.at.meshX)] <- 0
pop.at.mesh <- data.frame(points.kenya@coords)
pop.at.mesh$pop <- pop.at.meshX
names(pop.at.mesh)[1:2] <- c("mlong", "mlat")

#' Make the mesh
mesh.true <- inla.mesh.2d(
  pop.at.mesh[, 1:2],
  offset = c(1, 2),
  max.edge = 1 * c(1, 5)
)

plot(mesh.true)

#' adm1
#' Determine which polys the mesh points live in
meshgrid <- data.frame(
  Longitude = mesh.true$loc[,1],
  Latitude = mesh.true$loc[,2]
)

coordinates(meshgrid) <- ~ Longitude + Latitude
proj4string(meshgrid) <- proj4string(adm1)
meshlocs <- over(meshgrid, adm1)[, 1]
tapply(rep(1, length(meshlocs)), meshlocs, sum)

mesh.df <- data.frame(
  mlong = mesh.true$loc[, 1],
  mlat = mesh.true$loc[, 2],
  loc = meshlocs,
  id = 1:length(meshlocs)
)

#' adm0.5
# Determine which polys the mesh points live in

#' There is slight misalignment of the boundaries, which should completely
#' overlap. Thus, we manually enter the provinces based on which of the 47
#' districts the mesh point is in.
meshlocs8 <- over(meshgrid, adm0.5)[, 1]
mesh.df$loc8X <- meshlocs8
mapping47to8 <- as.vector(apply(table(mesh.df[, c("loc", "loc8X")]), 1, which.max))
mesh.df$loc8 <- mapping47to8[mesh.df$loc]

#' Population at mesh including polygon location
pop.at.mesh <- dplyr::right_join(mesh.df, pop.at.mesh)

#' Create D matrix

#' adm1
D <- inla.spde.make.A(
  mesh = mesh.true,
  loc = as.matrix(pop.at.mesh[,1:2]),
  block = pop.at.mesh$loc,
  weights = pop.at.mesh$pop
)

D.tmp <- list()
D.tmp$D <- D / rowSums(D)
D.tmp$mesh.weights <- colSums(D)
mesh.df$weight.unscaled <- D.tmp$mesh.weights
D <- D.tmp$D

mesh.df$weight.scaled <- colSums(D)
nmesh.area <- 1 / tapply(rep(1, nrow(mesh.df)), mesh.df$loc, sum)
mesh.df$weight.comp <- nmesh.area[mesh.df$loc] #' Comparison weight

#' adm0.5

D8 <- inla.spde.make.A(
  mesh = mesh.true,
  loc = as.matrix(pop.at.mesh[,1:2]),
  block = pop.at.mesh$loc8,
  weights = pop.at.mesh$pop
)

D.tmp8 <- list()
D.tmp8$D <- D8 / rowSums(D8)
D.tmp8$mesh.weights <- colSums(D8)

mesh.df$weight.unscaled8 <- D.tmp8$mesh.weights
D8 <- D.tmp8$D
mesh.df$weight.scaled8 <- colSums(D8)
nmesh.area8 <- 1 / tapply(rep(1, nrow(mesh.df)), mesh.df$loc8, sum)
mesh.df$weight.comp8 <- nmesh.area8[mesh.df$loc8] #' Comparison weight

#' Set-up parameters
alpha <- 0
sigma2 <- 0.25
lambda2 <- 1 - sigma2
erange <- sqrt(8) / exp(0.5)
theta2 <- log(sqrt(8) / erange)
kappa <- exp(theta2)
tau2 <- 1 / (4 * pi * kappa^2 * lambda2)
theta1 <- log(sqrt(tau2))

#' Set-up field
spde <- inla.spde2.matern(mesh.true)
Q.true <- inla.spde.precision(spde, c(theta1, theta2))
field.true <- as.numeric(inla.qsample(1, Q = Q.true, seed = 50L))

#' Connecting field to data
proj.survey <- inla.mesh.projector(
  mesh.true,
  loc = as.matrix(kenya.data[, c("LONGNUM", "LATNUM")])
)
field.survey <- inla.mesh.project(proj.survey, field = field.true)
kenya.data$field <- field.survey
kenya.data$mu <- alpha + kenya.data$field

#' Make matrix that will be used to obtain predictions of the surface
pred.info <- MakePredictionAMatrix(mesh.true, adm0, dims=c(110, 110))
A.pred <- pred.info$A.pred
stkgrid <- inla.stack(
  data = list(y = NA, N = NA),
  A = list(A.pred, 1),
  effects = list(i = 1:spde$n.spde, alpha = rep(1, nrow(A.pred))), tag = 'prd.gr'
)
rel.points.clip <- pred.info$rel.points.clip

#' Simulate data

#' Census
#' Data simulated on 1km x 1km grid
pop.data.kenya.points <- data.frame(rasterToPoints(pop.dat.kenya, spatial = F))
names(pop.data.kenya.points) <- c("long", "lat", "pop")
coordinates(pop.data.kenya.points) <- ~ long + lat
pop.data.kenya.points@proj4string <- adm0@proj4string
pop.pt.insideX <- pop.data.kenya.points[adm0, ]
pop.pt.inside <- as.data.frame(pop.pt.insideX)
coordinates(pop.pt.inside) <- ~ long + lat
pop.pt.inside@proj4string <- adm0@proj4string
```

# `functions.R`

```{r, echo=TRUE, eval=FALSE, output.var="ex", code=readLines('functions.R')}
```

# Original computing environment

```{r}
sessionInfo()
```

### 2022-10-28 make/0_all.R

@@ -13,7 +13,7 @@ run_commit_push("0_explore_hrsl-sampling")    #' [ ]
 run_commit_push("0_explore_inla-spde")        #' [ ]
 run_commit_push("0_explore_unequal-samples")  #' [ ]
 run_commit_push("0_explore_variogram-method") #' [ ]
-run_commit_push("0_explore_wilson-pointless") #' [ ]
+run_commit_push("0_explore_wilson-pointless") #' [x]
 run_commit_push("0_plot_areal-data")          #' [x]
 run_commit_push("0_plot_sloo-cv")             #' [x]
 run_commit_push("0_plot_geometry-graph")      #' [x]

### 2022-10-28 src/0_explore_wilson-pointless/script.R

@@ -1,172 +1,4 @@
-# orderly::orderly_develop_start("explore_wilson-pointless")
-# setwd("src/explore_wilson-pointless")
+# orderly::orderly_develop_start("0_explore_wilson-pointless")
+# setwd("src/0_explore_wilson-pointless")
 
-set.seed(10)
-
-load("kenya.rdata")
-
-#' Read in population data
-#' Follow instructions in pointless-spatial-modeling README as to how to get this
-pop.dat <- raster("gpw_v4_population_count_rev11_2015_2pt5_min.tif")
-pop.dat <- setMinMax(pop.dat)
-kenya.extent <- extent(33.9, 42, -4.7, 5.1)
-pop.dat.kenya <- crop(pop.dat, kenya.extent)
-
-kenya.df1 <- fortify(adm0) %>%
-  mutate(id = as.numeric(id))
-
-kenya.df47 <- fortify(adm1) %>%
-  mutate(id = as.numeric(id))
-
-kenya.df8 <- fortify(adm0.5) %>%
-  mutate(id = as.numeric(id) + 1)
-
-#' Set-up variables for simulation
-#' * Create mesh for first 5 scenarios: "mesh.true"
-#' * Create D matrix used for scenario 2 and 3: "D" (47 x m)
-#' * Create D matrix used for scenario 4 and 5: "D8" (see note above)
-#' * Add columns to kenya.data indicating which polygons the EAs reside in
-
-# Create internal mesh points
-points.kenya <- spsample(adm0, 2500, "regular")
-
-boundary.kenya <- inla.nonconvex.hull(
-  points.kenya@coords,
-  convex = 0.1,
-  resolution = c(120, 120)
-)
-
-#' Determine population near these mesh points
-pop.at.meshX <- raster::extract(pop.dat.kenya, points.kenya@coords)
-pop.at.meshX[is.na(pop.at.meshX)] <- 0
-pop.at.mesh <- data.frame(points.kenya@coords)
-pop.at.mesh$pop <- pop.at.meshX
-names(pop.at.mesh)[1:2] <- c("mlong", "mlat")
-
-#' Make the mesh
-mesh.true <- inla.mesh.2d(
-  pop.at.mesh[, 1:2],
-  offset = c(1, 2),
-  max.edge = 1 * c(1, 5)
-)
-
-plot(mesh.true)
-
[…]

### 2022-10-28 utils/archive_to_docs.R

@@ -23,3 +23,4 @@ archive_to_docs("0_check_ik-converge")
 archive_to_docs("0_check_xbinomial-sampling")
 archive_to_docs("0_demo_areal-kernels")
 archive_to_docs("0_explore_inla-spde")
+archive_to_docs("0_explore_wilson-pointless")

### 2022-11-10 src/0_explore_inla-spde/inla-spde.Rmd
INLA-SPDE notes to HTML
Update explore_inla-spde

@@
 title: "SPDE approach notes"
-author: "Adam Howes"
-output: pdf_document
+author:
+- name: Adam Howes
+output:
+  html_document:
+    toc: yes
+    toc_float: yes
+    toc_collapsed: yes
+    df_print: paged
+    code_folding: show
+    theme: lumen
 bibliography: citations.bib
+abstract: |
+  **Background** To write.
 ---
 
-## List of possible sources
+# Resources
 
@@
 
-## Background
+# Background
 
@@
 
-## @miller2020understanding explanation
+# @miller2020understanding explanation
 
@@
 
-### Connection to basis-penalty smoothing
+## Connection to basis-penalty smoothing
 
@@
 
-## A short introduction on how to fit a SPDE model with INLA (Krainski and Rue)
+# A short introduction on how to fit a SPDE model with INLA (Krainski and Rue)
 
@@
 
-## @krainski2018advanced notes
+# @krainski2018advanced notes
 
-### First result
+## First result
 
@@
 
-### Second result
+## Second result
 
@@
 
-### Toy example
+## Toy example
 
@@
 
-## Bibliography {-}
+# Original computing environment
+
+```{r}
+sessionInfo()
+```
+
+# Bibliography {-}

### 2022-11-10 make/0_all.R

@@ -10,7 +10,7 @@ run_commit_push("0_check_ik-converge")        #' [x]
 run_commit_push("0_check_xbinomial-sampling") #' [x]
 run_commit_push("0_demo_areal-kernels")       #' [x]
 run_commit_push("0_explore_hrsl-sampling")    #' [ ]
-run_commit_push("0_explore_inla-spde")        #' [ ]
+run_commit_push("0_explore_inla-spde")        #' [x]
 run_commit_push("0_explore_unequal-samples")  #' [ ]
 run_commit_push("0_explore_variogram-method") #' [ ]
 run_commit_push("0_explore_wilson-pointless") #' [x]

### 2022-11-10 utils/archive_to_docs.R

@@ -24,3 +24,4 @@ archive_to_docs("0_check_xbinomial-sampling")
 archive_to_docs("0_demo_areal-kernels")
 archive_to_docs("0_explore_inla-spde")
 archive_to_docs("0_explore_wilson-pointless")
+archive_to_docs("0_explore_inla-spde")

### 2022-12-12 src/0_checks/bin-gaussian.Rmd
Add xbinom to 0_checks

--- 2022-10-07 src/0_check_bin-gaussian/bin-gaussian.Rmd
@@
 abstract: |
-  **Background**
+  **Background** Suppose that you take a binomial sample from the average of Gaussians. How different is that to if you aggregated binomial samples from each Gaussian?
 
-  **Task**
+  **Task** We simulate data from both models in a setting where $\rho$ is low (disease prevalence) and when $\rho \approx 0.5$ (elections) and compare.
 ---
 
-## Set-up
+# Set-up
 
@@
 
-## Low prevalence setting (disease)
+# Low probability setting (disease prevalence)
 
@@
 
-## Probability approximately 1/2 setting (elections)
+# Probability approximately 1/2 setting (elections)
 
@@
 
-## Conclusions, things to add
+# Conclusions, things to add
 
@@
 * What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?
+
+# Original computing environment
+
+```{r}
+sessionInfo()
+```

### 2022-12-12 make/0_all.R

@@ -5,7 +5,6 @@ source("make/utils.R")
 #' [/]: tried to run but gave me issues
 
 run_commit_push("checks")                     #' [x]
-run_commit_push("0_check_xbinomial-sampling") #' [x]
 run_commit_push("0_demo_areal-kernels")       #' [x]
 run_commit_push("0_explore_hrsl-sampling")    #' [ ]
 run_commit_push("0_explore_inla-spde")        #' [x]

### 2022-12-12 utils/archive_to_docs.R

@@ -21,7 +21,7 @@ archive_to_docs("docs_paper")
 archive_to_docs("checks", i = 1)
 archive_to_docs("checks", i = 2)
 archive_to_docs("checks", i = 3)
-archive_to_docs("0_check_xbinomial-sampling")
+archive_to_docs("checks", i = 4)
 archive_to_docs("0_demo_areal-kernels")
 archive_to_docs("0_explore_inla-spde")
 archive_to_docs("0_explore_wilson-pointless")

### 2022-12-12 src/docs_paper/paper.Rmd
Small paper edits

@@
 However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
-We hypothesised that incorporating a spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models.
-Here, we set out to test this hypothesis, and in doing so provide practical recommendations and further understanding of models for spatial structure.
+In this work, we set out to test the hypothesis that incorporating spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models, and in doing so provide practical recommendations and further understanding of models for spatial structure.
 <!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)?
@@
 Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
-Estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effective used to counter the epidemic [@hallett2016evaluation; @cuadros2017mapping].
+Estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effectively used to respond to the epidemic [@hallett2016evaluation; @cuadros2017mapping].
 Household surveys provide nationally represenative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
 Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, often face similar measurement difficulties as the HIV indicators themselves.
-Models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}, unlike other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
+Models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
+This stands in contrast to other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
 These circumstances foreground the importance of sharing information between districts via spatial smoothing in HIV mapping.
@@
 Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
-@duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model, though this conclusion was specific to the (often-analysed) Scottish lip cancer example, and based on the recommended against [@vehtari2017practical] deviance information criteria (DIC).
-Another approach is to take into account that areal data is typically produced by aggregating higher resolution data, so as to be a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change] and dating to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
+@duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model, though this conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
+Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
+The problem is then a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change], which dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
 Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
-This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially], `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages.
+This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially].
+<!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
 Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].
@@
 \end{equation}
-We obtained these samples using the function \texttt{INLA::inla.posterior.sample} for models fit with \texttt{R-INLA} and directly from the Markov chain using \texttt{rstan::extract} for models fit with Stan.
 
-### Coverage
+### Posterior predictive check for coverage
 
-To assess the coverage of our estimates, we considered probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2021graphical].
-We estimated the quantile $q_i = F(\omega_i)$ of the true value within the marginal posterior using numerical integration.
-For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that for any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$, in other words the empirical coverage, should also be $1 - \alpha$.
+We assessed the coverage of our estimates using posterior predictive checks based on the quantile $q_i = F(\omega_i)$ of the true value within each marginal posterior predictive distribution, computed using numerical integration.
+For calibrated models, over repeated simulations, then $q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also be $1 - \alpha$.
+To check uniformity we used probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2021graphical].
 
@@
 
-We compared model performance in estimating HIV prevalence in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire ($n = 33$), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$), again varying the spatial random effect specification.
+We compared model performance in estimating HIV prevalence in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire (which has $n = 33$ districts), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$), again varying the spatial random effect specification.
 As with the simulation study, \textsf{R} code for this study is available from \url{https://github.com/athowes/areal-comparison}.

### 2022-12-12 src/2_checks/cv.Rmd
Create 2_checks from the CV code check, and move that to a notebook

---
title: "Cross-validation code check"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background**

  **Task**
---

```{r}
#' Test of cross-validation approach using missing data in Stan

loo_test_sets <- create_folds(mw, type = "LOO")
sloo_test_sets <- create_folds(mw, type = "SLOO")

#' What do the test sets look like?
cowplot::plot_grid(
  plotlist = lapply(loo_test_sets, function(x) {
    x$data %>%
      ggplot(aes(fill = y)) +
        geom_sf() +
        theme_void() +
        theme(
          legend.position = "none"
        )
  }),
  ncol = 7
)

cowplot::plot_grid(
  plotlist = lapply(sloo_test_sets, function(x) {
    x$data %>%
      ggplot(aes(fill = y)) +
      geom_sf() +
      theme_void() +
      theme(
        legend.position = "none"
      )
  }),
  ncol = 7
)

#' Try fitting one of them in Stan
sf <- loo_test_sets[[1]]$data

#' Index of observations which are not missing
ii_obs <- which(!is.na(sf$y))

#' Index of missing observations
ii_mis <- which(is.na(sf$y))

#' Number of not missing observations
n_obs <- length(ii_obs)

#' Number of missing observations
n_mis <- length(ii_mis)

dat <- list(
  n_obs = n_obs,
  n_mis = n_mis,
  ii_obs = array(ii_obs),
  ii_mis = array(ii_mis),
  n = nrow(sf),
  y_obs = sf$y[ii_obs],
  m = sf$n_obs
)

temp <- rstan::stan(file = "constant.stan", data = dat, warmup = 100, iter = 900)
rstan::summary(temp)$summary
out <- rstan::extract(temp)

#' Could add truth to this plot and remake in ggplot2
plot(out$y_mis)
```

### 2022-12-12 src/2_checks/script.R

@@ -1,0 +1,5 @@
+#' Uncomment and run the two line below to resume development of this script
+# orderly::orderly_develop_start("2_checks")
+# setwd("src/2_checks")
+
+rmarkdown::render("cv.Rmd")

### 2022-12-13 src/1_checks/inla-mcmc.Rmd
Consolidate 1_checks

---
title: "Check R-INLA against MCMC"
author:
  - name: "Adam Howes"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Task**
---

```{r}
#' Checking that R-INLA and Stan produce similar results
#' Would using Stan (slower) for the models below rather than R-INLA (faster) change anything?

#' Example data, should repeat this for others
data <- readRDS("depends/data_iid_grid.rds")
sf <- data[[1]]$sf

#' Constant
fit_inla <- constant_inla(sf)
fit_stan <- constant_stan(sf)
get_time(fit_inla); get_time(fit_stan)

plot(marginal_intervals(fit_inla)$mean, marginal_intervals(fit_stan, parameter = "rho")$mean)

#' IID
fit_inla <- iid_inla(sf)
fit_stan <- iid_stan(sf)
get_time(fit_inla); get_time(fit_stan)

plot(marginal_intervals(fit_inla)$mean, marginal_intervals(fit_stan, parameter = "rho")$mean)

#' Besag
fit_inla <- besag_inla(sf)
fit_stan <- besag_stan(sf)
get_time(fit_inla); get_time(fit_stan)

plot(marginal_intervals(fit_inla)$mean, marginal_intervals(fit_stan, parameter = "rho")$mean)

#' BYM2
fit_inla <- bym2_inla(sf)
fit_stan <- bym2_stan(sf)
get_time(fit_inla); get_time(fit_stan)

plot(marginal_intervals(fit_inla)$mean, marginal_intervals(fit_stan, parameter = "rho")$mean)

#' TODO: These plots could be improved, ggplot2, a line through x = y
#' Could also check how things look for more than one dataset
```

### 2022-12-13 src/1_checks/lengthscale-recovery.Rmd
Consolidate 1_checks

---
title: "Can the lengthscale parameter be properly recovered?"
author:
  - name: "Adam Howes"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Task**
---

```{r}
#' Simulating data from l = 1 or 3 and seeing if it can be recovered

grid <- readRDS("depends/grid.rds")

#' l = 1
ck_one <- sim_ck(grid, nsim = 10, l = 1, save = FALSE)

fits_one <- lapply(ck_one, FUN = function(x) bsae::ck_stan(x$sf))
lengthscale_summary_one <- lapply(fits_one, FUN = function(fit) summary(fit)$summary["l", ])
df_lengthscale_one <- list_to_df(lengthscale_summary_one)

ggplot(df_lengthscale_one, aes(x = replicate, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
  geom_hline(yintercept = 1, size = 0.75, col = "grey", alpha = 0.75) +
  labs(x = "Replicate", y = "Posterior CrI", title = "Is l = 1 recovered?") +
  theme_minimal()

#' Ratio of lengthscale to noise
sigma_summary_one <- lapply(fits_one, FUN = function(fit) summary(fit)$summary["sigma_phi", ])
df_sigma_one <- list_to_df(sigma_summary_one)

ggplot(df_sigma_one, aes(x = replicate, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
  geom_hline(yintercept = 1, size = 0.75, col = "grey", alpha = 0.75) +
  labs(x = "Replicate", y = "Posterior CrI", title = "Is sigma = 1 recovered?") +
  theme_minimal()

plot(df_sigma_one$mean, df_lengthscale_one$mean)

lengthscale_draws_one <- lapply(fits_one, FUN = function(fit) as.data.frame(rstan::extract(fit)$l))
sigma_draws_one <- lapply(fits_one, FUN = function(fit) as.data.frame(rstan::extract(fit)$sigma_phi))

#' Contour plots of lengthscale and sigma

cbind(bind_rows(lengthscale_draws_one, .id = "replicate"), bind_rows(sigma_draws_one)) %>%
  rename("l" = "rstan::extract(fit)$l", "sigma_phi" = "rstan::extract(fit)$sigma_phi") %>%
  ggplot(aes(x = l, y = sigma_phi, col = as.factor(replicate))) +
  geom_density2d() +
  labs(x = "Length-scale",
       y = "Spatial random effect standard deviation",
       col = "Replicate",
       subtitle = "Tend to observe positive correlation in the samples") +
  theme_minimal()

#' l = 3
#' TODO: Replicate this analysis for l = 3

# ck_three <- sim_ck(grid, nsim = 30, l = 3, save = FALSE)
#
# fits_three <- lapply(ck_three, FUN = function(x) bsae::ck_stan(x$sf))
# lengthscale_summary_three <- lapply(fits_three, FUN = function(fit) summary(fit)$summary["l", ])
# df_lengthscale_three <- list_to_df(lengthscale_summary_three)
#
# ggplot(df_lengthscale_three, aes(x = replicate, y = mean)) +
#   geom_point() +
#   geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
#   geom_hline(yintercept = 3, col = lightblue, size = 1.5)
#
# hist(df_lengthscale_three$mean - 3, breaks = 15)
# mean(df_lengthscale_three$mean - 3)

#' What does the prior look like, Gamma(1, 1)
ggplot(data = data.frame(x = c(0, 10)), aes(x)) +
  stat_function(fun = dgamma, n = 101, args = list(1, 1))

#' Note: Nyquist-Shannon sampling theorem
#' https://en.wikipedia.org/wiki/Nyquist%E2%80%93Shannon_sampling_theorem
```

### 2022-12-13 src/1_checks/plausible-lengthscale.Rmd
Consolidate 1_checks

---
title: "Average distances between centroids for plausible lengthscale choice"
author:
  - name: "Adam Howes"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Task**
---

```{r}
grid <- readRDS("depends/geometry_grid.rds")
civ <- readRDS("depends/geometry_civ.rds")
tex <- readRDS("depends/geometry_tex.rds")

grid_distances <- c(centroid_distance(grid))
min(grid_distances[grid_distances > 0])

civ_distances <- c(centroid_distance(civ))
min(civ_distances[ci_distances > 0])

tex_distances <- c(centroid_distance(tex))
min(tex_distances[tex_distances > 0])

cowplot::plot_grid(
  graph_length_plot(grid),
  ggplot() + geom_histogram(data = data.frame(x = grid_distances), aes(x = x))
)

cowplot::plot_grid(
  graph_length_plot(civ),
  ggplot() + geom_histogram(data = data.frame(x = civ_distances), aes(x = x))
)

cowplot::plot_grid(
  graph_length_plot(tex),
  ggplot() + geom_histogram(data = data.frame(x = tex_distances), aes(x = x))
)
```

### 2022-12-13 src/1_checks/prior-sensitivity.Rmd
Consolidate 1_checks

---
title: "Sensitivity of the length-scale posterior to prior specification"
author:
  - name: "Adam Howes"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Task**
---

```{r}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

data <- readRDS("depends/data_iid_grid.rds")
sf <- data[[1]]$sf

D <- centroid_distance(sf)
dat <- list(n = nrow(sf), y = sf$y, m = sf$n_obs, mu = rep(0, nrow(sf)), D = D)

# No prior on l (uniform): this shouldn't be a good idea
fit_non <- rstan::stan("prior-sensitivity/non-informative.stan", data = dat, warmup = 100, iter = 1000)
# Nice visual checking tool--a lot of between chain posterior variability would be a bad sign
bayesplot::mcmc_hist_by_chain(fit_non, pars = "l")

# l ~ Gamma(1, 1)
fit_gamma <- rstan::stan("prior-sensitivity/gamma_.stan", data = dat, warmup = 100, iter = 1000)
bayesplot::mcmc_hist_by_chain(fit_non, pars = "l")

# l ~ N+(0, (ub - lb) / 3): Half-normal with standard deviation based on range of distances
# https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html
dist <- as.vector(D)
lb <- min(dist) # Could do dist[dist > 0] here but results in lb = 1
lb <- 0.1
ub <- max(dist)
(ub - lb) / 3 # Approximately 2.3
fit_normal_data <- rstan::stan("prior-sensitivity/normal-data.stan", data = dat, warmup = 100, iter = 1000)
bayesplot::mcmc_hist_by_chain(fit_normal_data, pars = "l")

# l ~ N(1, 0.5): Informative prior at the true value
fit_normal_inform <- rstan::stan("prior-sensitivity/informative.stan", data = dat, warmup = 100, iter = 1000)
bayesplot::mcmc_hist_by_chain(fit_normal_inform, pars = "l")

# l ~ Inverse-Gamma(a, b):
# https://warwick.ac.uk/fac/sci/statistics/crism/workshops/masterclassapril/presentation4.pdf

# Copied from brms (https://github.com/paul-buerkner/brms/commit/524e738aeeb82e49c5338839c3e10113763b6de1)
# plb: prior probability of being lower than minimum length-scale
# pub: prior probability of being higher than maximum length-scale
opt_fun <- function(x, lb, ub) {
  # optimize parameters on the log-scale to make them positive only
  x <- exp(x)
  y1 <- invgamma::pinvgamma(lb, x[1], x[2], log.p = TRUE)
  y2 <- invgamma::pinvgamma(ub, x[1], x[2], lower.tail = FALSE, log.p = TRUE)
  c(y1 - log(plb), y2 - log(pub))
}

opt_fun(c(0, 0), lb = lb, ub = ub)

plb <- 0.01
pub <- 0.01

opt_res <- nleqslv::nleqslv(
  c(0, 0), opt_fun, lb = lb, ub = ub,
  control = list(allowSingular = TRUE)
)

exp(opt_res$x) # Approximately a = 1.7, b = 0.6

fit_invgamma_data <- rstan::stan("prior-sensitivity/invgamma-data.stan", data = dat, warmup = 100, iter = 1000)
bayesplot::mcmc_hist_by_chain(fit_invgamma_data, pars = "l")

cbpalette <- c("#56B4E9", "#009E73", "#E69F00", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

# Prior comparison

ggplot(data = data.frame(x = c(0, 10)), aes(x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 2.3), col = cbpalette[1]) +
  stat_function(fun = dgamma, n = 101, args = list(1, 1), col = cbpalette[2]) +
  stat_function(fun = invgamma::dinvgamma, n = 101, args = list(1.7, 0.6), col = cbpalette[3]) +
  stat_function(fun = dnorm, n = 101, args = list(1, 0.5), col = cbpalette[4]) +
  stat_function(fun = dunif, n = 101, args = list(0, 100), col = cbpalette[5]) +
  labs(x = "l", y = "p(l)") +
  theme_minimal()

# Posterior comparison

models <- list(
  "Non informative" = fit_non,
  "Gamma" = fit_gamma,
  "Data informed normal" = fit_normal_data,
  "Informative normal" = fit_normal_inform,
  "Informative inverse gamma" = fit_invgamma_data
)

draws <- lapply(models, FUN = function(model) rstan::extract(model)$l)

df <- do.call(
  "rbind",
  lapply(
    seq_along(draws),
    FUN = function(i) {
      data.frame(l = draws[[i]], type = names(draws)[i], mean = mean(draws[[i]]), truth = 1)
    }
  )
)

ggplot(df, aes(x = l, fill = type)) +
  geom_histogram(aes(y = ..density..), alpha = 0.8) +
  geom_vline(aes(xintercept = mean, fill = type)) +
  geom_vline(aes(xintercept = truth), linetype = "dashed") +
  facet_wrap(~ type, ncol = 1) +
  scale_fill_manual(values = cbpalette) +
  labs(y = "Density", x = "Lengthscale", fill = "Prior type") +
  theme_minimal()
```

### 2022-12-13 src/1_check_plausible-lengthscale/functions.R

@@ -1,18 +1,0 @@
-graph_length_plot <- function(g) {
-  nb <- sf_to_nb(g)
-
-  nb_sf <- spdep::nb2lines(nb, coords = sp::coordinates(as(g, "Spatial"))) %>%
-    as("sf") %>%
-    st_set_crs(st_crs(g))
-
-  line_sf <- nb_sf %>%
-    mutate(length = round(sf::st_length(nb_sf), 2),
-           geometry = sf::st_centroid(geometry))
-
-  ggplot(g) +
-    geom_sf() +
-    geom_sf(data = nb_sf) +
-    geom_sf_label(data = line_sf, aes(label = length)) +
-    theme_minimal() +
-    theme_void()
-}

### 2022-12-13 src/1_checks/script.R

@@ -1,0 +1,5 @@
+# orderly::orderly_develop_start("1_checks")
+# setwd("src/1_checks")
+
+rmarkdown::render("inla-mcmc.Rmd")
+rmarkdown::render("lengthscale-recovery.Rmd")

### 2023-02-16 src/1_checks/lengthscale-recovery.Rmd
arealutils not bsae, and other updates

@@
 
-fits_one <- lapply(ck_one, FUN = function(x) bsae::ck_stan(x$sf))
+fits_one <- lapply(ck_one, FUN = function(x) arealutils::ck_stan(x$sf))
 lengthscale_summary_one <- lapply(fits_one, FUN = function(fit) summary(fit)$summary["l", ])
@@
 #
-# fits_three <- lapply(ck_three, FUN = function(x) bsae::ck_stan(x$sf))
+# fits_three <- lapply(ck_three, FUN = function(x) arealutils::ck_stan(x$sf))
 # lengthscale_summary_three <- lapply(fits_three, FUN = function(fit) summary(fit)$summary["l", ])

### 2023-02-16 src/2_check_assess-code/assess.Rmd
arealutils not bsae, and other updates

---
title: "Marginal assessment code check"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background**

  **Task**
---

Does the code to assess each of the posterior marginals work as intended?
There is also a possibility to switch some of the assessment code to use `INLA::` inbuilt functions if they are more efficient.
It would also be helpful to know what proportion of the computation is assessing models.

Fit a quick constant model on example data:

```{r}
data <- readRDS("depends/data_iid_grid.rds")
sf <- data[[1]]$sf
fit <- arealutils::constant_inla(sf)
```

Some made-up values for the truth:

```{r}
rho <- rep(0.14, 28)
```

Draws from the first posterior marginal:

```{r}
marginal <- ith_marginal_rho(fit, 1)
plot(marginal$samples)
```

Check that the mean and mode are above the minimum sample and below the maximum sample:

```{r}
stopifnot(min(marginal$samples) < marginal$mean)
stopifnot(max(marginal$samples) > marginal$mean)
stopifnot(min(marginal$samples) < marginal$mode)
stopifnot(max(marginal$samples) > marginal$mode)
```

A list of the evaluation metrics:

```{r}
metrics <- assess_ith_marginal_rho(rho, fit, 1)

metric_names <- c(
  "obs", "mean", "mode", "lower", "upper", "mse", "mae", "mse_mean",
  "mae_mean", "mse_mode", "mae_mode", "crps", "lds", "quantile"
)

stopifnot(names(metrics) == metric_names)
```

`df` with columns giving evaluation metrics with first row as above.
Note that the model fitted is a constant so the metrics should all be equal.

```{r}
df <- assess_marginals_rho(rho, fit)

plot(df$mse)
plot(df$mae)
plot(df$lds)
plot(df$crps)
plot(df$quantile)
```

Test that `assess_replicates` works as intended:

```{r}
x <- assess_replicates(list(list(rho = rho)), list(fit))
```

This is for the lengthscale test:

```{r}
fit2 <- arealutils::ck_stan(mw)
x2 <- assess_replicates(list(list(rho = rho)), list(fit2))
```

Does `inla.posterior.sample` match `$fitted.values`?

```{r}
samples_list <- INLA::inla.posterior.sample(n = 100000, fit, selection = list(Predictor = 1))
eta_samples = sapply(samples_list, function(x) x$latent)
samples <- plogis(eta_samples)
samples_df <- data.frame(x = samples)

cbpalette <- c("#56B4E9","#009E73", "#E69F00", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

ggplot(marginal$df) +
  geom_histogram(data = samples_df, aes(x = x, y = ..density..), bins = 100, alpha = 0.5) +
  geom_line(aes(x = x, y = y), col = cbpalette[1])

f <- approxfun(marginal$df$x, marginal$df$y, yleft = 0, yright = 0)

#' Are samples generated outside the range of f (from $fitted.values)
log(f(max(samples))) #' Not -Inf
log(f(min(samples)))

#' Kernel density approach on samples
kde <- density(samples, bw = 0.1, n = 256, from = 0, to = 1)
df2 <- data.frame(x = kde$x, y = kde$y)
plot(df2$x, df2$y) #' What is going on here?
plot(marginal$df$x, marginal$df$y)
f2 <- approxfun(df2$x, df2$y, yleft = 0, yright = 0)
log(f2(max(samples))) #' Not -Inf
log(f2(min(samples)))

#' inla. function testing

#' Should be approximately the same
INLA::inla.dmarginal(0.14, marginal$df, log = TRUE)
log(f(0.14))

#' inla.qmarginals gives the value corresponding to a quantile q
#' inla.pmarginal gives the cumulative distribution function at p
#' Should be approxiamtely the same
INLA::inla.pmarginal(0.14, marginal$df)
cubature::cubintegrate(f, lower = 0, upper = 0.14, method = "pcubature")$integral

#' Do these functions work for df created outside of INLA (I don't see why not)
#' Yes they do, so is it worth switching to them?
fit_stan <- arealutils::constant_stan(sf)
marginal_stan <- ith_marginal_rho(fit_stan, 1)
plot(marginal_stan$df$x, marginal_stan$df$y)
plot(df2$x, df2$y)

#' Looks the same as that based upon inla.posterior.sample
INLA::inla.dmarginal(0.14, marginal_stan$df, log = TRUE)
INLA::inla.dmarginal(0.14, df2, log = TRUE)
INLA::inla.pmarginal(0.14, marginal_stan$df)
INLA::inla.pmarginal(0.14, df2)
```

### 2023-02-16 src/docs_paper/paper.Rmd
arealutils not bsae, and other updates

@@
     corresponding: ath19@ic.ac.uk
-  - name: Harrison Zhu
-    affiliation: A
-  - name: Yidan Xu
-    affiliation: C
-  - name: Samir Bhatt
-    affiliation: D
   - name: Jeffrey W. Eaton
@@
   - name: Seth R. Flaxman
-    affiliation: E
+    affiliation: C
 address:
@@
   - code: C
-    address: Department of Statistics, University of Michigan
-  - code: D
-    address: Department of Public Health, University of Copenhagen
-  - code: E
     address: Department of Computer Science, University of Oxford
@@
   |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects.
-However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions, which seem particularly dubious for irregular geometries.
+However, for some geometries, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions.
 Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data.
@@
   |
-  |  **Methods:** In a comprehensive simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications.
+  |  **Methods:** In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications.
 We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities.
@@
 \end{tabularx}
-\caption{Each inferential model is implemented as a part of the \textsf{R} package \texttt{bsae} available from \url{github.com/athowes/bsae}.}
+\caption{Each inferential model is implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}.}
 \label{tab:inferentialmodels}
@@
 In models \hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
-For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51} -- though `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].
+For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51}.
+Though note `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].
 We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}.
@@
 
-We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, subject to calibration.
-We describe our approach to model assessment in the following sections.
+We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, as well as the calibration of the model's estimates.
+<!-- We describe our approach to model assessment in the following sections. -->
 <!-- Analogous approaches may be used to check recovery of the intercept parameter $\beta_0$, precision parameter $\tau_\phi$, and, when relevant, the length-scale $l$. -->
@@
 For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, y)$ and $\omega_i$ be the true value.
-Writing $F$ as the cumulative distribution corresponding to $f$ then
+Writing $F$ as the cumulative distribution function corresponding to $f$ then
 \begin{equation}
@@
 The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$.
-The CRPS is a strictly proper scoring rule [@gneiting2007strictly], ensuring that the uniquely best approach to minimising the score is for the model to report its true probabilistic beliefs about the quantity in question.
+The CRPS is a strictly proper scoring rule such that the uniquely best approach to minimising the score is for the model to report its true probabilistic beliefs about the quantity in question [@gneiting2007strictly].
 The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
+In contrast, the CRPS can be evaluated directly using samples $\{\rho_i^s\}_{s = 1}^S \sim p(\rho_i \, | \, y)$ as
 \begin{equation}
@@
 We assessed the coverage of our estimates using posterior predictive checks based on the quantile $q_i = F(\omega_i)$ of the true value within each marginal posterior predictive distribution, computed using numerical integration.
-For calibrated models, over repeated simulations, then $q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also be $1 - \alpha$.
+For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also be $1 - \alpha$.
 To check uniformity we used probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2021graphical].
@@
 These surveys utilise a complex design, where each individual $j$ in area $i$ has an unequal probability $\pi_{ij}$ of being included in the sample.
-<!-- A two-stage design in which enumeration areas are first drawn from a stratified sample and then households are chosen using equal probability systematic sampling from within each enumeration area, is common.
--->
-To account for the survey design we used sampling weights $\omega_{ij} = 1/\pi_{ij}$ to adjust the raw data in each district, obtaining the Kish effective sample size $m_i^\star = (\sum_k \omega_{ij})^2 / \sum_k \omega_{ij}^2$ and effective number of cases $y_i^\star$ which may be thought of as what would have been observed had the survey been a simple random sample.
+A two-stage design in which enumeration areas are first drawn from a stratified sample and then households are chosen using equal probability systematic sampling from within each enumeration area, is common.
+To account for the survey design we used sampling weights $w_{ij} = 1 / \pi_{ij}$ to adjust the raw data in each district, obtaining the Kish effective sample size $m_i^\star = (\sum_k w_{ij})^2 / \sum_k w_{ij}^2$ and effective number of cases $y_i^\star$ which may be thought of as what would have been observed had the survey been a simple random sample.
 
@@
 
-We used a the same models as in Section \ref{sec:simulation} (Table \ref{tab:inferentialmodels}), with one alteration to the likelihood.
+We used the same eight inferential models (Section \ref{sec:simulation}, Table \ref{tab:inferentialmodels}), with a small alteration to the likelihood.
 Since the effective number of cases and effective sample size may not be integers, we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
@@
 We assessed each model using two cross-validation approaches: (1) a standard leave-one-out cross-validation (LOO-CV), in which at each fold the data from a single district were held-out, and (2) a spatial leave-one-out cross-validation (SLOO-CV), in which at each fold the data from a single district and each of its neighbouring districts were removed.
-We performed both types of cross-validation manually, as we found approximations that avoid refitting the model, such as the built-in conditional predictive ordinate statistic in `R-INLA` package [@held2010posterior] and the Pareto smoothed importance sampling [@vehtari2015pareto] approach implemented in the `loo` package [@vehtari2017practical], were inaccurate in this setting.
-This might be because individual observations in spatial random effect models can exert high influence toward local parameters.
+<!-- We performed both types of cross-validation manually, as we found approximations that avoid refitting the model, such as the built-in conditional predictive ordinate statistic in `R-INLA` package [@held2010posterior] and the Pareto smoothed importance sampling [@vehtari2015pareto] approach implemented in the `loo` package [@vehtari2017practical], were inaccurate in this setting.
+-->
+<!-- This might be because individual observations in spatial random effect models can exert high influence toward local parameters. -->
 
@@
 In the presense of structural dependencies, like spatial dependence, LOO-CV tends to underestimate predictive error [@roberts2017cross].
-We counteracted this effect by modifying the spatial leave-one-out (SLOO) approach of @le2014spatial.
-During the $i$th fold of SLOO-CV, we leave-out the block $(i, \delta i)$ (Figure \ref{fig:sloo}), increasing the extent to which the training and validation sets are conditionally independent.
+To counteract this effect we modified the spatial leave-one-out (SLOO) approach of @le2014spatial by leaving-out the block $(i, \delta i)$ (Figure \ref{fig:sloo}) during the $i$th fold, increasing the extent to which the training and validation sets are conditionally independent.
 From the SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$  we analogously computed $\text{SLOO-CRPS}_i$.
@@
 
-AH, HZ were supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).
+AH was supported by the EPSRC Centre for Doctoral Training in Modern Statistics and Statistical Machine Learning (EP/S023151/1).
 AH, JWE were supported by the Bill and Melinda Gates Foundation (OPP1190661).
-YX was supported by the Mary Lister McCammon Summer Research Fellowship 2019.
-HZ was supported by Cervest Limited.
 JWE was supported by UNAIDS and National Institute of Allergy and Infectious Disease of the National Institutes of Health (R01AI136664).
@@
 
-AH thanks Robert Ashton for help with \textsf{R} package development, and Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for providing feedback on earlier drafts of this manuscript.
+Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
+AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations.
 

### 2023-02-16 make/1_all.R

@@ -11,7 +11,7 @@ run_commit_push("1_plot_simulation-geometries") #' [x]
 #' 3. Fit models
 
 #' Vignette and realistic together
-run_commit_push("1_fit", param = list(f = "constant_inla")) #' [ ]
+run_commit_push("1_fit", param = list(f = "constant_inla")) #' [x]
 run_commit_push("1_fit", param = list(f = "iid_inla"))      #' [ ]
 run_commit_push("1_fit", param = list(f = "besag_inla"))    #' [ ]
 run_commit_push("1_fit", param = list(f = "bym2_inla"))     #' [ ]
@@ -21,24 +21,24 @@ run_commit_push("1_fit", param = list(f = "ck_stan"))       #' [ ]
 run_commit_push("1_fit", param = list(f = "ik_stan"))       #' [ ]
 
 #' Vignette alone
-run_commit_push("1_fit", param = list(f = "constant_inla", realistic = FALSE))
-run_commit_push("1_fit", param = list(f = "iid_inla", realistic = FALSE))
-run_commit_push("1_fit", param = list(f = "besag_inla", realistic = FALSE))
-run_commit_push("1_fit", param = list(f = "bym2_inla", realistic = FALSE))
-run_commit_push("1_fit", param = list(f = "fck_inla", realistic = FALSE))
-run_commit_push("1_fit", param = list(f = "fik_inla", realistic = FALSE))
-run_commit_push("1_fit", param = list(f = "ck_stan", realistic = FALSE))
-run_commit_push("1_fit", param = list(f = "ik_stan", realistic = FALSE))
+# run_commit_push("1_fit", param = list(f = "constant_inla", realistic = FALSE))
+# run_commit_push("1_fit", param = list(f = "iid_inla", realistic = FALSE))
+# run_commit_push("1_fit", param = list(f = "besag_inla", realistic = FALSE))
+# run_commit_push("1_fit", param = list(f = "bym2_inla", realistic = FALSE))
+# run_commit_push("1_fit", param = list(f = "fck_inla", realistic = FALSE))
+# run_commit_push("1_fit", param = list(f = "fik_inla", realistic = FALSE))
+# run_commit_push("1_fit", param = list(f = "ck_stan", realistic = FALSE))
+# run_commit_push("1_fit", param = list(f = "ik_stan", realistic = FALSE))
 
 #' Realistic alone
-run_commit_push("1_fit", param = list(f = "constant_inla", vignette = FALSE))
[…]

### 2023-02-16 src/1_assess_intercept-marginal/functions.R

@@ -48,7 +48,7 @@ assess_marginal_intercept <- function(intercept = -2, fit) {
     upper = marginal$upper,
     mse = mean(error_samples^2),
     mae = mean(abs(error_samples)),
-    crps = bsae::crps(marginal$samples, obs),
+    crps = arealutils::crps(marginal$samples, obs),
     lds = log(f(obs)),
     quantile = cubature::cubintegrate(f, lower = -Inf, upper = obs, method = "pcubature")$integral
   ))

### 2023-02-16 src/1_assess_intercept-marginal/script.R

@@ -31,7 +31,7 @@ process_fits <- function(geometry, sim_model, inf_model) {
 
   #' Assess the intercept marginal
   df <- lapply(fits, assess_marginal_intercept, intercept = -2) %>%
-    bsae::list_to_df() %>%
+    arealutils::list_to_df() %>%
     #' Add columns for meta data
     mutate(
       geometry = geometry,

### 2023-02-16 src/1_assess_lengthscale-marginal/script.R

@@ -30,7 +30,7 @@ process_fits <- function(geometry, sim_model, inf_model) {
 
   #' Assess the lengthscale marginal
   df <- lapply(fits, assess_marginal_lengthscale, lengthscale = 5/2) %>%
-    bsae::list_to_df() %>%
+    arealutils::list_to_df() %>%
     #' Add columns for meta data
     mutate(
       geometry = geometry,

### 2023-02-16 src/1_assess_rho-marginals/functions.R

@@ -67,7 +67,7 @@ assess_ith_marginal_rho <- function(rho, fit, i) {
     upper = marginal$upper,
     mse = mean(error_samples^2),
     mae = mean(abs(error_samples)),
-    crps = bsae::crps(marginal$samples, obs),
+    crps = arealutils::crps(marginal$samples, obs),
     lds = lds,
     quantile = quantile
   ))

### 2023-02-16 src/1_assess_time-taken/script.R

@@ -28,7 +28,7 @@ process_fits <- function(geometry, sim_model, inf_model) {
 
   #' Assess the intercept marginal
   df <- lapply(fits, function(fit) list(t = get_time(fit))) %>%
-    bsae::list_to_df() %>%
+    arealutils::list_to_df() %>%
     #' Add columns for meta data
     mutate(
       geometry = geometry,

### 2023-02-16 src/1_plot_lengthscale-recovery/script.R

@@ -27,7 +27,7 @@ df_lengthscale %>%
     lengthscale_mean,
     by = c("inf_model", "geometry")
   ) %>%
-  bsae::update_naming() %>%
+  arealutils::update_naming() %>%
   split(.$geometry) %>%
   lapply(function(x) {
     x_max <- max(as.numeric(x$replicate))

### 2023-02-16 src/1_plot_metric-maps/functions.R

@@ -75,7 +75,7 @@ produce_maps <- function(arg1, arg2, arg3) {
   df_rho %>%
     filter(geometry == arg1) %>%
     filter(inf_model != "constant_inla") %>%
-    bsae::update_naming() %>%
+    arealutils::update_naming() %>%
     group_mean_and_se(c("geometry", "sim_model", "inf_model", "id")) %>%
     metric_map(metric = arg3, sf = arg2)
     # %>% ggsave(filename = paste0(arg3, "-map-rho-", tolower(arg1), "-no-constant.pdf"), width = 6.25, height = 4)

### 2023-02-16 src/1_process_metric-tables/script.R

@@ -9,7 +9,7 @@ df_intercept <- readRDS("depends/df_intercept.rds")
 #' CRPS table for rho
 df_rho %>%
   group_mean_and_se(group_variables = c("geometry", "sim_model", "inf_model")) %>%
-  bsae::update_naming() %>%
+  arealutils::update_naming() %>%
   metric_table(
     metric = "crps",
     title = "Continuous Ranked Probability Score",
@@ -23,7 +23,7 @@ df_rho %>%
 #' CRPS table for the intercept
 df_intercept %>%
   group_mean_and_se(group_variables = c("geometry", "sim_model", "inf_model")) %>%
-  bsae::update_naming() %>%
+  arealutils::update_naming() %>%
   metric_table(
     metric = "crps",
     title = "Continuous Ranked Probability Score",
@@ -36,7 +36,7 @@ df_intercept %>%
 #' #' Time table
 #' df_time %>%
 #'   group_by(geometry, sim_model, inf_model) %>%
-#'   bsae::update_naming() %>%
+#'   arealutils::update_naming() %>%
 #'   summarise(n = n(), across(t, list(mean = mean, se = ~ sd(.x) / sqrt(length(.x))))) %>%
 #'   metric_table(
 #'     metric = "t",

### 2023-02-16 src/2_check_assess-code/script.R

@@ -1,109 +1,5 @@
 #' Uncomment and run the two line below to resume development of this script
-# orderly::orderly_develop_start("check_assess-code")
-# setwd("src/check_assess-code")
+# orderly::orderly_develop_start("2_check_assess-code")
+# setwd("src/2_check_assess-code")
 
-#' Does the code to assess each of the posterior marginals work as intended?
-#' There is also a possibility to switch some of the assessment code to use INLA:: inbuilt functions
-#' Presumably this would be more efficient? Perhaps make an issue to look into this
-#' Might want to know what proportion of the computation is assessing models (if it's not very much then matters less)
-
-#' Just fit a quick constant model on example data
-data <- readRDS("depends/data_iid_grid.rds")
-sf <- data[[1]]$sf
-fit <- bsae::constant_inla(sf)
-
-#' Make up some values for the truth
-rho <- rep(0.14, 28)
-
-marginal <- ith_marginal_rho(fit, 1)
-plot(marginal$samples)
-
-#' Draws from the posterior marginal
-stopifnot(min(marginal$samples) < marginal$mean)
-stopifnot(max(marginal$samples) > marginal$mean)
-stopifnot(min(marginal$samples) < marginal$mode)
-stopifnot(max(marginal$samples) > marginal$mode)
-
-#' A list of the evaluation metrics
-metrics <- assess_ith_marginal_rho(rho, fit, 1)
-
-metric_names <- c(
-  "obs", "mean", "mode", "lower", "upper", "mse", "mae", "mse_mean",
-  "mae_mean", "mse_mode", "mae_mode", "crps", "lds", "quantile"
-)
-
-stopifnot(names(metrics) == metric_names)
-
-
-#' df with columns giving evaluation metrics with first row as above
-#' The model fitted is a constant so the metrics should all be equal
-df <- assess_marginals_rho(rho, fit)
-
-plot(df$mse)
-plot(df$mae)
-plot(df$lds)
-plot(df$crps)
-plot(df$quantile)
-
-#' Test that assess_replicates works as intended
-x <- assess_replicates(list(list(rho = rho)), list(fit))
-
-#' This is for the lengthscale test
-fit2 <- bsae::ck_stan(mw)
-x2 <- assess_replicates(list(list(rho = rho)), list(fit2))
-
[…]

### 2023-02-16 src/2_cv/script.R

@@ -4,9 +4,9 @@
 
 surveys <- list("civ2017phia", "mwi2016phia", "tza2017phia", "zwe2016phia")
 types <- list("loo", "sloo")
-fs <- list(get(f, envir = asNamespace("bsae")))
+fs <- list(get(f, envir = asNamespace("arealutils")))
 pars <- expand.grid("survey" = surveys, "f" = fs)
 cv_pars <- expand.grid("survey" = surveys, "type" = types, "f" = fs)
 
-purrr::pmap(cv_pars, bsae:::run_cv_models)
-purrr::pmap(pars, bsae:::run_models)
+purrr::pmap(cv_pars, arealutils:::run_cv_models)
+purrr::pmap(pars, arealutils:::run_models)

### 2023-02-16 src/2_process_hiv-surveys/script.R

@@ -1,6 +1,6 @@
 #' Uncomment and run the two line below to resume development of this script
-# orderly::orderly_develop_start("process_hiv-surveys")
-# setwd("src/process_hiv-surveys")
+# orderly::orderly_develop_start("2_process_hiv-surveys")
+# setwd("src/2_process_hiv-surveys")
 
 analysis_level <- c(
   "CIV" = 1,
@@ -17,6 +17,8 @@ survey_name <- c(
 )
 
 iso3 <- c("civ", "mwi", "tza", "zwe")
+
+#' This function writes the survey to file as well as outputing
 surveys <- lapply(iso3, extract_survey)
 
 pdf("hiv-surveys.pdf", h = 5, w = 6.25)

### 2023-02-16 utils/archive_to_docs.R

@@ -27,6 +27,7 @@ archive_to_docs("0_explore_inla-spde")
 archive_to_docs("0_explore_wilson-pointless")
 archive_to_docs("0_explore_inla-spde")
 
+archive_to_docs("1_plot_simulation-geometries")
 archive_to_docs("1_plot_coverage")
 archive_to_docs("1_plot_metric-boxplots")
 

### 2023-05-30 src/docs_paper/paper.Rmd
Update paper

@@
     email: ath19@ic.ac.uk
-    affiliation: A
+    affiliation: A,B
     corresponding: ath19@ic.ac.uk
   - name: Jeffrey W. Eaton
-    affiliation: B
+    affiliation: C,B
   - name: Seth R. Flaxman
-    affiliation: C
+    affiliation: D
 address:
@@
   - code: C
+    address: Center for Communicable Disease Dynamics, Harvard T.H. Chan School of Public Health, Harvard University
+  - code: D
     address: Department of Computer Science, University of Oxford
 abstract: |
-  |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects.
-However, for some geometries, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions.
+  |  Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects.
+However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions when the geometry is not a grid.
 Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data.
@@
   |
-  |  **Methods:** In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications.
-We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities.
+  |  In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications.
+The simulations were performed under model misspecification and across a range of vignette and realistic geometries with different regularities.
 We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
-  |
-  |  **Results:**
-  |
-  |  **Conclusions:**
 
@@
 bibliography: citations.bib
+classoption: preprint, authoryear
 output:
@@
       in_header: preamble.tex
-# plos.csl can be download and used locally
-csl: http://www.zotero.org/styles/plos
+csl: elsarticle.cls
 ---
@@
 Gaussian Markov random fields (GMRFs) [@rue2005gaussian], which combine a Gaussian distribution with Markov conditional independence assumptions between areas, are the most common class of models used to specify spatial random effects.
-Using a GMRF typically reflects the assumption that observations made in areas close together are correlated, and more distant relationships may be ignored.
-The simplest GMRF model is arguably that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
-This model requires minimal additional modelling choices and is widely implemented, including as a part of the `R-INLA` package [@rue2017bayesian], popular within spatial statistics.
-Use of this model is widespread, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
+Under these Markov assumptions, observations made in areas close together are correlated, and more distant relationships are ignored.
+Arguably the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
+This model is attractive as it requires minimal additional choices and is widely implemented.
+As such, it has seen widespread use, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
 However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
-In this work, we set out to test the hypothesis that incorporating spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models, and in doing so provide practical recommendations and further understanding of models for spatial structure.
+As such, we set out to test the hypothesis that using more realistic assumptions about spatial structure would improve the performance of small-area estimation models.
+In doing so, we provide practical recommendations and further understanding of models for spatial structure.
 <!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)?
@@
 Household surveys provide nationally represenative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
-Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, often face similar measurement difficulties as the HIV indicators themselves.
-Models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
+Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour [@howes2023spatio] or the prevalence of male circumcision, often face analogous measurement difficulties.
+Possibly as a result, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
 This stands in contrast to other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
-These circumstances foreground the importance of sharing information between districts via spatial smoothing in HIV mapping.
+These circumstances foreground the importance of models for spatial structure in HIV mapping.
 
 Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
-@duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model, though this conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
+@duncan2017spatial compared seventeen methods for specifying these weights and did not find any which outperform the Besag model, though this conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
 Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
-The problem is then a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change], which dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
+Obtaining inference about a variable at a different resolution to that it was observed at is known in geostatistics as the change-of-support problem [@gelfand2001change], and dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
 Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
@@
 
+\newpage
+
 # Background \label{sec:background}
@@
 Point data are another type of spatial data where instead observations $y(s)$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.
-
 Areal data is often aptly conceptualised as arising from aggregation of point data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
@@
 <!-- It may be possible to model this explicitly, though the result might look similar to what is achieved using e.g. BYM2. -->
-If the areal data we observe are indeed an aggregation of point data, why not use the higher resolution point data in our models rather than areal data?
-Explanations include that (a) the underlying point data are unobserved, (b) although point data are collected, they may not be accessible due to privacy constraints, administrative practicality or storage capacity, and (c) the point data are available, but we decide not to use them in the model.
+
+If the areal data we observe indeed arise from aggregation, we may not use the underlying point data because (a) it is unobserved, (b) it may be inaccessible due to privacy constraints, administrative practicality or storage capacity, and (c) the point data are available, but we decide not to use them in the model.
 In the final case (c), researchers may prefer to use area-level models as models for point data are typically more complex, require more data, are less immediately applicable to area-level policy making, and have higher computational burden.
@@
 
-Bayesian hierarchical models are an attractive framework for small-area estimation, whereby areal data $y$ corresponding to the areas $\{A_i\}_{i = 1}^n$, are modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
+Bayesian hierarchical models are an attractive framework for small-area estimation, whereby areal data $\y = (y_1, \ldots, y_n)$ corresponding to the areas $\{A_i\}_{i = 1}^n$, are modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
 \begin{alignat}{2}
-&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, g^{-1}(x_i), \theta), \quad i = 1, \ldots, n, \label{eq:data} \\
-&\text{(Latent field)}  &        x &\sim p(x \, | \, \theta), \label{eq:process} \\
-&\text{(Parameters)}    & \qquad \theta &\sim p(\theta), \label{eq:parameters}
+&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, \x, \btheta), \quad i = 1, \ldots, n, \label{eq:data} \\
+&\text{(Latent field)}  &        \x &\sim p(\x \, | \, \btheta), \label{eq:process} \\
+&\text{(Parameters)}    & \qquad \btheta &\sim p(\btheta), \label{eq:parameters}
 \end{alignat}
-where $x$ is the $n$-dimensional latent field, $\theta$ are the $m$-dimensional parameters, $m < n$, and $g$ is a link function.
+where $\x$ is a $N$-dimensional latent field, $\btheta$ are the $m$-dimensional parameters, $m < N$.
 Spatial structure is modelled at the level of the latent field, and the observation model, typically either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
-As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $x \sim \mathcal{N}(\mu, \Sigma)$ for the latent field.
+As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $\x \sim \mathcal{N}(\bmu, \bSigma)$ for the latent field.
 Three-layer models with Gaussian latent fields comprise a wide array of popular models, including generalised linear mixed models (GLMMs), and have been collectively studied under the title latent Gaussian models (LGMs) [@rue2009approximate].
@@
 
+\newpage
+
 # Spatial random effects using adjacency \label{sec:ab}
+
+## Besag \label{sec:besag}
 
@@
 
-## Besag \label{sec:besag}
-
-One way to encode spatial structure between areas is by using a symmetric relation, where $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring.
+Spatial structure can be encoded using a symmetric relation between areas, where $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring.
 Adjacency is often defined by a shared border, though other choices, such as inclusion of second or higher-degree neighbours [@paciorek2013spatial], are also possible.
@@
 
+\newpage
+
 # Spatial random effects using kernels \label{sec:ks}
@@
 
-Section \ref{sec:ab} reviewed ways to construct spatial random effect precision matrices using an adjacency relation.
-Another approach is to define the covariance matrix using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$ which gives a measure of similarity between two areas, where $\mathcal{P}$ denotes the power set and $\mathcal{P}(\Sc)$ is the space of subsets of the study region.
+In the previous section we reviewed ways to construct spatial random effect precision matrices using an adjacency relation.
+Another approach is to define the covariance matrix using an areal kernel function which gives a measure of similarity between two areas $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$, where $\mathcal{P}$ denotes the power set such that $\mathcal{P}(\Sc)$ is the space of subsets of the study region.
 If $K$ is positive semi-definite, then we define areal kernel spatial random effects by
 \begin{equation}
-    \phi \sim \mathcal{N}(0, \frac{1}{\tau_\phi} K), \label{eq:arealkernel}
+    \phi \sim \mathcal{N} \left( 0, \frac{1}{\tau_\phi} \mathbf{K} \right), \label{eq:arealkernel}
 \end{equation}
-where the $n \times n$ Gram matrix $K$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix.
-Here, we place $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.
-
+where the $n \times n$ Gram matrix $\mathbf{K}$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix.
+We place $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.
 Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$.
-We use a simple method to construct $K$ from $k$, namely averaging the kernel $k$ computed on some collection of points from within each area.
-As such, $K$ is described as a kernel on sets in the multiple instance learning literature [@gartner2002multi].
-Although it may be possible to learn the method for combining kernels from data [@gonen2011multiple], we do believe that this is feasible in the small-area estimation setting, and instead hope to construct areal kernels which represent our prior beliefs about the spatial process.
+We consider a simple method to construct $K$ from $k$: namely averaging the kernel $k$ computed on some collection of points from within each area.
 
@@
 
-The simplest approach is to use a single point, especially one which is representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude.
+The simplest approach is to use a single point.
+A natural choice is the centroid $c_i \in A_i$, given by the arithmetic mean of the latitude and longitude, which we might hope to be representative of the area.
 This results in the centroid kernel
@@
 <!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
-The centroid kernel has been used by @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$; as well as more recently by @teh2021efficient as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
+The centroid kernel has been used in an environmental epidemiology setting [@wakefield1999spatial] and in model comparison study [@best2005comparison], both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$; as well as more recently by @teh2021efficient as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
 @wakefield1999spatial fix the length-scale $l$ using a semi-variogram and @best2005comparison place a prior on the inverse length-scale.
@@
 
-Rather than attempting to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
+Rather than looking to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
 This results in the integrated kernel
 \begin{equation}
-K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s', \label{eq:ik1}
+K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s'. \label{eq:ik1}
 \end{equation}
-which is equivalent to the covariance structure obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition.
-Models of this sort have been studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi; @2021arXiv210501460Z].
-<!-- % For example, @law2018variational develop a variational Bayesian inference method for aggregated observation models and @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region.
+This results in equivalent covariance structure to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].å
+<!-- @law2018variational develop a variational Bayesian inference method for aggregated observation models. -->
+<!-- @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region.
 -->
@@
 
-The log-Gaussian Cox Process framework [@diggle2013spatial] also naturally arrives at the integrated kernel formulation.
-A Cox process is an inhomogeneous Poisson process with a continuous stochastic intensity function $\{ x(s), s \in \Sc \}$ such that conditional on the realisation of $x(s)$ the number of points in any area $A_i$ follows a Poisson distribution.
-The rate parameter of this Poisson distribution is explicitly aggregated as follows
-\begin{equation}
-y_i \, | \, x(s) \sim \text{Poisson} \left(\int_{s \in A_i} x(s) \text{d}s \right).
-\end{equation}
-In a LGCP the log intensity $\log x(s) = \eta(s)$ is modelled using a Gaussian process prior $\eta(s) \sim \mathcal{GP}(\mu(s), k(s, s'))$.
-@johnson2019spatially obtain Equation \ref{eq:ik2} by considering a discrete Poisson log-linear mixed model approximation to a continuous LGCP, whereby $\eta(s)$ is approximated by a piecewise constant $\eta_i = \mu_i + \phi_i$ in each area $A_i$^[@johnson2019spatially use points generated from a class of inhibition processes which combine sequential inhibition with rejection sampling to evaluate the integral.].
-The $i$th discrete spatial random effect is then $\phi_i = \int_{A_i} w_i(s) \phi(s) \text{d}s$,
-with covariance structure
-\begin{equation}
-\text{Cov} \left( \int_{A_i} w_i(s) \phi(s) \text{d}s, \int_{A_j} w_j(s') \phi(s') \text{d}s' \right)
-= \int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d}s\text{d}s',
-\end{equation}
-corresponding to an areal integrated kernel with a logarithmic link function and Poisson likelihood.
-
-Disaggregation regression (also known as downscaling or interpolation) is another closely related approach which, rather than using the aggregate nature of areal observations as primarily a route towards better area-level estimates, as is our focus, aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
-<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
-<!-- Disaggregation is challenging without auxiliary information. -->
-
-A benefit of the integration approach is that, if available, additional information accounting for heterogeneity over $A_i$ may be incorperated into the model.
+If available, additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel.
 This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
@@
 Integration points may also be selected by some deterministic process, and used in Equation \ref{eq:mcintegratedkernel} to give a numerical integration estimate of the kernel.
-As with learning the way in which the kernels should be combined, in more data rich settings it may be possible to learn the integration points [@campbell2019automated].
 Taking any of these approaches requires $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
@@
 
+### Connection to log-Gaussian Cox processes
+
+The log-Gaussian Cox Process framework [@diggle2013spatial] arrives naturally at the integrated kernel formulation.
+A Cox process is an inhomogeneous Poisson process with a continuous stochastic intensity function $\{ x(s), s \in \Sc \}$ such that conditional on the realisation of $x(s)$ the number of points in any area $A_i$ follows a Poisson distribution.
+The rate parameter of this Poisson distribution is explicitly aggregated as follows
+\begin{equation}
+y_i \, | \, x(s) \sim \text{Poisson} \left(\int_{s \in A_i} x(s) \text{d}s \right).
+\end{equation}
+In a LGCP the log intensity $\log x(s) = \eta(s)$ is modelled using a Gaussian process prior $\eta(s) \sim \mathcal{GP}(\mu(s), k(s, s'))$.
+@johnson2019spatially obtain Equation \ref{eq:ik2} by considering a discrete Poisson log-linear mixed model approximation to a continuous LGCP, whereby $\eta(s)$ is approximated by a piecewise constant $\eta_i = \mu_i + \phi_i$ in each area $A_i$.
+The $i$th discrete spatial random effect is then $\phi_i = \int_{A_i} w_i(s) \phi(s) \text{d}s$,
+with covariance structure
+\begin{equation}
+\text{Cov} \left( \int_{A_i} w_i(s) \phi(s) \text{d}s, \int_{A_j} w_j(s') \phi(s') \text{d}s' \right)
+= \int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d}s\text{d}s',
+\end{equation}
+corresponding to an areal integrated kernel with a logarithmic link function and Poisson likelihood.
+
+### Connection to disaggregation regression
+
+Disaggregation regression, also known as downscaling or interpolation, is another closely related approach.
+Rather than focusing on the aggregate nature of areal observations as primarily a route towards better area-level estimates, disaggregation regression aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
+<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
+<!-- Disaggregation is challenging without auxiliary information. -->
 <!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. -->
+
+\newpage
 
@@
 We tested the ability of inferential models with varying spatial random effect specifications to accurately recover small-area quantities.
-The data and modelling choices were designed with a spatial epidemiology setting in mind, and are similar to those used in Section \ref{sec:hiv} where we considered models for HIV prevalence using data from national household surveys in sub-Saharan Africa.
+The data and modelling choices were designed with a spatial epidemiology setting in mind.
 The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`.
@@
 \small
-\begin{tabularx}{\textwidth}{@{}lX@{}}
+\begin{tabularx}{\textwidth}{llX}
 \toprule
-Simulation model & Additional details \\
+ & Simulation model & Additional details \\
 \midrule
-\hypertarget{S1}{S1}. \textit{IID} & $\phi \sim \mathcal{N}(0, I_n)$ \\
-\hypertarget{S2}{S2}. \textit{Besag} & $\phi \sim \mathcal{N}(0, (R^\star)^{-})$ as described in Section \ref{sec:besag} where $R^\star$ is the scaled structure matrix \\
-\hypertarget{S3}{S3}.
-\textit{IK} & $\phi \sim \mathcal{N}(0, K^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation \ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
+\hypertarget{S1}{\texttt{S1}} & IID & $\phi \sim \mathcal{N}(0, I_n)$ \\
+\hypertarget{S2}{\texttt{S2}} & Besag & $\phi \sim \mathcal{N}(0, (R^\star)^{-})$ as described in Section \ref{sec:besag} where $R^\star$ is the scaled structure matrix \\
+\hypertarget{S3}{\texttt{S3}} & IK & $\phi \sim \mathcal{N}(0, K^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation \ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
 Like the Besag model, we scaled the Gram matrix to have generalised variance one \\
@@
 
-We studied robustness to spatial misspecification, by simulating synthetic data-sets from three known data generating processes.
-The spatial random effects $\phi$ were generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}).
-We then generated synthetic data $y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via
+To study robustness to spatial misspecification we simulated synthetic data-sets from three known data generating processes, using spatial random effects $\bphi$ generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}).
+We then generated synthetic data $\y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via
 \begin{equation}
@@
 \end{equation}
-We fixed the sample sizes to $m_i = 25$ for all $i$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$.
-We used seven different geometries: the four vignette geometries in Figure \ref{fig:tikz1}, which share an adjacency graph, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of Côte d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
-These more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test how model performance varies by geometry regularity.
+The sample sizes were fixed as $m_i = 25$ for all $i$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$.
+We used seven different geometries: the four vignette geometries sharing an adjacency graph in Figure \ref{fig:tikz1}, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of Côte d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
+The more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test how model performance varies by geometry regularity.
 For each combination of spatial random effect and geometry we replicated the simulation process above 200 times to generate a total of 1400 synthetic data-sets.
@@
 \small
-\begin{tabularx}{\textwidth}{llX}
+\begin{tabularx}{\textwidth}{lllX}
 \toprule
-Inferential model & Algorithm & Additional details \\
+& Inferential model & Algorithm & Additional details \\
  \midrule
-\hypertarget{I1}{I1}. \textit{Constant} & INLA &
+\hypertarget{I1}{\texttt{I1}} & Constant & AGHQ &
 No spatial random effects \label{row:I1} \\
-\hypertarget{I2}{I2}. \textit{IID} & INLA &
+\hypertarget{I2}{\texttt{I2}} & IID & AGHQ &
 $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} I_n)$\\
-\hypertarget{I3}{I3}. \textit{Besag} & INLA &
-$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations in \ref{app:freni}  \\
-\hypertarget{I4}{I4}. \textit{BYM2} & INLA &
+\hypertarget{I3}{\texttt{I3}} & Besag & AGHQ &
+$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations of \citet{freni2018note}  \\
+\hypertarget{I4}{\texttt{I4}} & BYM2 & AGHQ &
 $\phi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, v + \sqrt{\pi} \, u^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
-\hypertarget{I5}{I5}. \textit{FCK} & INLA &
+\hypertarget{I5}{\texttt{I5}} & FCK & AGHQ &
 $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
-\hypertarget{I6}{I6}. \textit{CK} & NUTS & As in model \hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
-\hypertarget{I7}{I7}.
-\textit{FIK} & INLA & $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
-\hypertarget{I8}{I8}. \textit{IK} & NUTS & As in model \hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
+\hypertarget{I6}{\texttt{I6}} & CK & AGHQ & As in model \hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
+\hypertarget{I7}{\texttt{I7}} & FIK & AGHQ & $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
+\hypertarget{I8}{\texttt{I8}} & IK & AGHQ & As in model \hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
 \bottomrule
 \end{tabularx}
-\caption{Each inferential model is implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}.}
+\caption{All inferential models are implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}.}
 \label{tab:inferentialmodels}
@@
 
-### Kernel model details
+### Kernels
 
-Following @stein1999interpolation, for the areal kernel models we used the Matérn kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
+For the areal kernel models we used the Matérn [@stein1999interpolation] kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
 \begin{equation}
@@
 where $B_\nu$ is the modified Bessel function of the second kind, $|s - s'|$ is the Euclidean distance between $s$ and $s'$, $\nu$ is the smoothness hyperparameter and $l$ is the length-scale hyperparameter on the latitude-longitude scale.
-We fixed the smoothness parameter, which is otherwise typically unidentifiable from data, to be $3/2$, matching that used in simulation model \hyperlink{S3}{S3}, and convenient in that it simplifies Equation \ref{eq:matern} to be
+We fixed the smoothness parameter, which is otherwise typically unidentifiable from data, to be $3/2$, matching that used in simulation model \hyperlink{S3}{\texttt{S3}}, and convenient in that it simplifies Equation \ref{eq:matern} to be
 \begin{equation}
@@
 \end{equation}
-Taking a similar approach to @best1999bayesian, in models \hyperlink{I5}{I5} and \hyperlink{I7}{I7} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data.
-In models \hyperlink{I6}{I6} and \hyperlink{I8}{I8}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
+In models \hyperlink{I5}{\texttt{I5}} and \hyperlink{I7}{\texttt{I7}} we fixed the length-scale such that points an average distance apart in the model have 1\% correlation  a priori, independent of the data [@best1999bayesian].
+In models \hyperlink{I6}{\texttt{I6}} and \hyperlink{I8}{\texttt{I8}}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
 For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51}.
-Though note `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].
+Note that `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].
 We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}.
@@
 
-## Inference algorithms
+## Inference
 
-In models \hyperlink{I1}{I1}-\hyperlink{I5}{I5} choosing a Gaussian distribution for $\beta_0$ ensures that the latent random variables are Gaussian and the model is a LGM, facilitating use of the integrated nested Laplace approximation (INLA) algorithm [@rue2009approximate] for inference, as implemented by the \texttt{R-INLA} package.
-INLA is a deterministic method for approximate Bayesian inference based upon a combination of numerical integration and the Laplace approximation.
-INLA is significantly faster than MCMC and has been shown to have comparable accuracy for LGMs in realistic, pre-asymptotic settings.
-<!-- Although there are particular conditions under which INLA is not accurate [@rue2017bayesian] they do not apply here, as we illustrate in Appendix \ref{app:inlamcmc} where we compare INLA to MCMC using a large number of iterations.
--->
+We used adaptive Gauss-Hermite quadrature (AGHQ) with $k = 3$ quadrature points per hyperparamter dimension to perform approximate Bayesian inference.
 
@@
 
+\newpage
+
 # HIV prevalence study \label{sec:hiv}
 
-We compared model performance in estimating HIV prevalence in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire (which has $n = 33$ districts), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$), again varying the spatial random effect specification.
-As with the simulation study, \textsf{R} code for this study is available from \url{https://github.com/athowes/areal-comparison}.
+We compared model performance in estimating HIV prevalence in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire (which has $n = 33$ districts), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$).
+As before, we varyied the spatial random effect specification.
+The \textsf{R} code for this study is available from \url{https://github.com/athowes/areal-comparison}.
 
@@
 
-We used the same eight inferential models (Section \ref{sec:simulation}, Table \ref{tab:inferentialmodels}), with a small alteration to the likelihood.
-Since the effective number of cases and effective sample size may not be integers, we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
+We used the eight inferential models from Table \ref{tab:inferentialmodels}, with a small alteration to the likelihood.
+The effective number of cases and effective sample size may not be integers, so we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
 The working likelihood under this model for $m^\star_i \geq y^\star_i$ is given by
@@
 
-In the $i$th LOO-CV fold, the model was fit using the data $y_{-i}$ and assessed according to its prediction on $y_i$.
-We assessed forecasting performance using the CRPS, at the level of the data $y_i$ rather than at the level of the prevalence $\rho_i$.
-To compute the value of CRPS for the $i$th fold we use samples $\{ y_i^s \}_{s = 1}^S$ from the LOO posterior predictive distribution $p(y_i \, | \, y_{-i})$ giving $\text{LOO-CRPS}_i = \frac{1}{S} \sum_{s = 1}^S | y_i^s - y_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | y_i^s - y_i |.
+In the $i$th LOO-CV fold, the model was fit using the data $\y_{-i}$ and assessed according to its prediction on $y_i$.
+We assessed forecasting performance using the root mean square error (RMSE), mean absolute error, and CRPS, at the level of the data $y_i$ rather than at the level of the prevalence $\rho_i$.
+To compute the value of CRPS for the $i$th fold we use samples $\{ y_i^s \}_{s = 1}^S$ from the LOO posterior predictive distribution $p(y_i \, | \, \y_{-i})$ giving $\text{LOO-CRPS}_i = \frac{1}{S} \sum_{s = 1}^S | y_i^s - y_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | y_i^s - y_i |.
 \label{eq:crpsapprox2}$
@@
 
-In the presense of structural dependencies, like spatial dependence, LOO-CV tends to underestimate predictive error [@roberts2017cross].
-To counteract this effect we modified the spatial leave-one-out (SLOO) approach of @le2014spatial by leaving-out the block $(i, \delta i)$ (Figure \ref{fig:sloo}) during the $i$th fold, increasing the extent to which the training and validation sets are conditionally independent.
-From the SLOO posterior predictive distribution $p(y_i \, | \, y_{-(i, \delta i)})$  we analogously computed $\text{SLOO-CRPS}_i$.
+In the presense of structural dependencies, like spatial dependence, LOO-CV tends to underestimate predictive error [@le2014spatial; @roberts2017cross].
+To counteract this effect during the $i$th fold we left out the block $(i, \delta i)$ (Figure \ref{fig:sloo}), increasing the extent to which the training and validation sets are conditionally independent.
+From the SLOO posterior predictive distribution $p(y_i \, | \, \y_{-(i, \delta i)})$  we analogously computed $\text{SLOO-CRPS}_i$.
 
@@
 
+\newpage
+
 # Discussion \label{sec:discussion}
+
+Areal kernels can be thought of as kernels on sets [@gartner2002multi].
+Though it may be possible to learn more sophisticated methods for combining kernels using data [@gonen2011multiple], we do believe that this is feasible in the small-area estimation setting, and instead focused on constructing areal kernels which represent our prior beliefs about spatial processes.
+In more data rich settings it may be possible to learn the integration points used within the integrated kernel [@campbell2019automated].
 

### 2023-05-31 src/docs_paper/paper.Rmd
Paper writing and update template

@@
 title: Understanding models for spatial structure in small-area estimation
+classoption: ba
 author:
-  - name: Adam Howes
+  - firstname: Adam 
+    lastname: Howes
     email: ath19@ic.ac.uk
-    affiliation: A,B
-    corresponding: ath19@ic.ac.uk
-  - name: Jeffrey W. Eaton
-    affiliation: C,B
-  - name: Seth R. Flaxman
-    affiliation: D
-address:
-  - code: A
-    address: Department of Mathematics, Imperial College London
-  - code: B
-    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
-  - code: C
-    address: Center for Communicable Disease Dynamics, Harvard T.H. Chan School of Public Health, Harvard University
-  - code: D
-    address: Department of Computer Science, University of Oxford
+    affiliationref: A,B
+    footnoterefs: 
+      - t1
+  - firstname: Jeffrey W.
+    lastname: Eaton
+    affiliationref: C,B
+  - firstname: Seth R.
+    lastname: Flaxman
+    affiliationref: D
+affiliations:
+  - ref: A
+    name: Imperial College London 
+    address: Department of Mathematics
+  - ref: B
+    name: Imperial College London
+    address: MRC Centre for Global Infectious Disease Analysis
+  - ref: C
+    name: Harvard University
+    address: Center for Communicable Disease Dynamics, Harvard T.H. Chan School of Public Health
+  - ref: D
+    name: University of Oxford
+    address: Department of Computer Science 
+footnotes:
+  - ref: t1
+    text: "Corresponding Author"
+supplements:
+  - title: Title of Supplement A
+    description: Short description of Supplement A.
+  - title: Title of Supplement B
+    description: Short description of Supplement B.
+acknowledgements: |
+  | Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
+AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations.
 abstract: |
-  |  Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects.
-However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions when the geometry is not a grid.
-Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data.
+  |  Spatial correlation between small-areas is typically accounted for using spatial random effects.
+However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions for non-grid geometries.
+Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous geostatistical data.
 Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
@@
 We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
-
-author_summary: |
-  Author summary
-
+date: "`r Sys.Date()`"
 bibliography: citations.bib
-classoption: preprint, authoryear
+#linenumbers: true
+#numbersections: true
 output: 
-  rticles::elsevier_article:
-    number_sections: true
+  bookdown::pdf_book: 
+    base_format: rticles::isba_article
+    md_extensions: -autolink_bare_uris #keep these off or \printead{} will fail on email addresses
     includes:
       in_header: preamble.tex
-csl: elsarticle.cls
 ---
@@
 Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
-Gaussian Markov random fields (GMRFs) [@rue2005gaussian], which combine a Gaussian distribution with Markov conditional independence assumptions between areas, are the most common class of models used to specify spatial random effects.
-Under these Markov assumptions, observations made in areas close together are correlated, and more distant relationships are ignored.
-Arguably the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
+The most common class of models used to specify spatial random effects are Gaussian Markov random fields (GMRFs), which combine a Gaussian distribution with Markov conditional independence assumptions between areas [@rue2005gaussian].
+Under the Markov assumptions, observations made in areas close together are correlated, and more distant relationships are ignored.
+Perhaps the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
 This model is attractive as it requires minimal additional choices and is widely implemented.
@@
 However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
-As such, we set out to test the hypothesis that using more realistic assumptions about spatial structure would improve the performance of small-area estimation models.
+As such, in this work we test the hypothesis that using more realistic assumptions about spatial structure improves the performance of small-area estimation models.
 In doing so, we provide practical recommendations and further understanding of models for spatial structure.
@@
 
-## Areal and point spatial data
+## Areal and geostatistical spatial data
 
-Let $\Sc \subset \mathbb{R}^2$ be the study region, and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$.
-Areal data are a type of spatial data where observations $y_i$ are associated to areas $A_i$.
-Examples include the colour of a pixel, the minimum wage in a state, and the number of disease cases in a region.
-Point data are another type of spatial data where instead observations $y(s)$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.
-Areal data is often aptly conceptualised as arising from aggregation of point data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
-Exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a genuinely discrete spatial structure.
+Let $\Sc \subset \mathbb{R}^2$ be the study region, and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$ such that $A_i \cap A_j = \emptyset$ and $\cup_{i = 1}^n A_i = \Sc$.
+Areal data are a type of spatial data where observations $y_i \in \mathcal{Y}$ are associated to areas $A_i$.
+Examples of areal data include the colour of a pixel, the minimum wage in a state, and the number of disease cases in a region.
+Geostatistical data are another type of spatial data where instead observations $y(s) \in \mathcal{Y}$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.
+Examples of geostatistical spatial data include the temperature at a monitoring station, the response to a household survey, and the quality of a soil sample.
+
+Areal data may often be conceptualised as arising from aggregation of geostatistical data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
+Notable exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a truly discrete spatial structure.
 <!-- Note that, for data of this sort, there is less clear justification for using spatial smoothing. -->
@@
 -->
-<!-- It may be possible to model this explicitly, though the result might look similar to what is achieved using e.g. BYM2. -->
 
-If the areal data we observe indeed arise from aggregation, we may not use the underlying point data because (a) it is unobserved, (b) it may be inaccessible due to privacy constraints, administrative practicality or storage capacity, and (c) the point data are available, but we decide not to use them in the model.
-In the final case (c), researchers may prefer to use area-level models as models for point data are typically more complex, require more data, are less immediately applicable to area-level policy making, and have higher computational burden.
-Furthermore, seemly point data may actual be areal, such as data observed at polling stations or health facilities which individuals from the surrounding area travel to.
-As such, choosing to model the underlying data generating mechanism as either areal or point is often a matter of pragmatism.
+In some situations it may be preferable to work using aggregated areal data rather than the original geostatistical data.
+For one, the geostatistical data may be unavailabile or inaccessible due to privacy constraints, administrative practicality or limitations in storage capacity.
+Alternatively, opting to model at the area-level can be advantageous as models for geostatistical data tend to be more complex, require more data, have higher computational burden, and offer less immediate applicablility to area-level policy-making.
+In addition, in some circumstances, seemingly geostatistical data may truly be areal, such as data observed at polling stations or health facilities which individuals from the surrounding area travel to.
+Ultimately, the decision to model the underlying data generating mechanism as either areal or geostatistical is often a matter of practical considerations and pragmatism.
 
@@
 Auto-correlation is an important property of most spatial processes.
-Usually, outcomes for locations close together in space tend to be more similar than those far apart.
-This fact is known as "Tobler's first law of geography" [@tobler1970computer] and, from a statistical point of view, is both a challenge and a benefit of working with spatial data.
+Usually, outcomes for locations close together in space tend to be more similar than those far apart, known as "Tobler's first law of geography" [@tobler1970computer].
+This property both presents a challenge and an opportunity.
 Each observation provides less information than it would have had the samples been independent, making it more difficult to estimate global parameters.
-However, spatial correlation can be used to improve local (indexed by a particular spatial location) parameter estimates particularly in parts of the study region where little to no information is available.
+On the other hand, spatial correlation can be used to improve local parameter estimates, indexed by a particular spatial location, especially in parts of the study region where little to no information is available.
 
 The latter benefit is basis for the statistical task of small-area estimation, which aims to produce reliable local estimates where small sample sizes lead to noisy data [@pfeffermann2013new].
-In a spatial setting, this is often in small geographic areas, though the phrase "small-area" is not restricted to geographic areas and may be interpreted more broadly to mean any area where data are insufficient to make accurate local parameter inferences.
-In the context of multilevel regression and post-stratification [@gelman1997poststratification], small-areas are generated by the intersection of demographic variables like age, gender and race, alongside geographic variables.
-Due to the cost of gathering samples, a survey may be designed to give reliable estimates at at an aggregated level but not at a small-area level.
+In a spatial setting, this is often in small geographic areas. 
+However, the phrase "small-area" is not restricted to geographic areas and may be interpreted more broadly to mean any area where data are insufficient to make accurate local parameter inferences.
+For example, in the context of multilevel regression and post-stratification [@gelman1997poststratification], small-areas are generated by the intersection of demographic variables like age, gender and race, alongside geographic variables.
+Because of the cost of gathering samples, a survey may be designed to give reliable estimates at at an aggregate level but not at a small-area level.
 Although direct estimators of local parameters are unbiased, when data are sparse the total error may be reduced by accepting some bias in exchange for reduced variance using so-called indirect estimators.
 Smoothing approaches use information from similar units to "borrow strength" from one parameter to another, with determining precisely what is meant by "similar" a central challenge.
-For a recent review of both design-based and model-based approaches to spatial analysis of health-indicators, including both area and unit-level analysis, see @wakefield2020small.
+See @wakefield2020small for a recent review of both design-based and model-based approaches to small-area estimation of health-indicators, including both area and unit-level analysis.
 
@@
 
-Bayesian hierarchical models are an attractive framework for small-area estimation, whereby areal data $\y = (y_1, \ldots, y_n)$ corresponding to the areas $\{A_i\}_{i = 1}^n$, are modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
+Bayesian hierarchical models are an attractive framework for small-area estimation.
+Let $[n]$ denote $1, \ldots, n$.
+Areal data $\y = (y_i)_{i \in [n]}$ may be modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
 \begin{alignat}{2}
-&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, \x, \btheta), \quad i = 1, \ldots, n, \label{eq:data} \\
-&\text{(Latent field)}  &        \x &\sim p(\x \, | \, \btheta), \label{eq:process} \\
-&\text{(Parameters)}    & \qquad \btheta &\sim p(\btheta), \label{eq:parameters}
+&\text{(Observations)}  &        \y &\sim p(\y \, | \, \x, \btheta_1), \label{eq:data} \\
+&\text{(Latent field)}  &        \x &\sim p(\x \, | \, \btheta_2), \label{eq:process} \\
+&\text{(Parameters)}    & \qquad (\btheta_1, \btheta_2) &= \btheta \sim p(\btheta), \label{eq:parameters}
 \end{alignat}
-where $\x$ is a $N$-dimensional latent field, $\btheta$ are the $m$-dimensional parameters, $m < N$.
-Spatial structure is modelled at the level of the latent field, and the observation model, typically either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
-As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $\x \sim \mathcal{N}(\bmu, \bSigma)$ for the latent field.
-Three-layer models with Gaussian latent fields comprise a wide array of popular models, including generalised linear mixed models (GLMMs), and have been collectively studied under the title latent Gaussian models (LGMs) [@rue2009approximate].
+where $\x = (x_i)_{i \in [N]}$ is a $N$-dimensional latent field, and $\btheta = (\theta_i)_{i \in [m]}$ are the $m$-dimensional hyperparameters, with $m < N$.
+Small-area estimation models are usually descriptive rather than mechanistic, so a natural choice is to model the latent field with a flexible and computationally efficient Gaussian distribution $\x \sim \mathcal{N}(\bmu, \mathbf{Q})$ with mean vector $\bmu$ and precision matrix $\mathbf{Q}$.
+Three-layer models with Gaussian latent fields have been collectively studied as latent Gaussian models (LGMs) [@rue2009approximate] and comprise a wide array of popular models, including generalised linear mixed models.
 
@@
 
-The latent field in a LGM is comprised a constant mean $\mu$, known also as the fixed effects, and spatial random effects $\phi \sim \mathcal{N}(0, \Sigma)$.
-The mean may be modelled using a linear predictor $\mu_i = \beta_0 + z_i^\top \beta$, where $(\beta_0, \beta)$ are regression parameters corresponding to areal covariates $z_i = (z_{i,1}, \ldots, z_{i,p})^\top$.
-We focus on the spatial random effects $\phi$, whose role is to capture the spatial structure between areas not already accounted for in the mean.
-If no spatial structure remains after inclusion of covariates in the mean, independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ should be used, where $\tau_\phi$ is a shared precision.
+The observations in an LGM are conditionally independent and identically distributed given linear predictors $\bmeta = (\eta)_{i \in [n]}$ such that
+\begin{equation*}
+p(\y \, | \, \, \x, \btheta_1) = p(\y \, | \, \, \bmeta, \btheta_1) = \prod_{i = 1}^n p(y_i \, | \, \eta_i, \btheta_1).
+\end{equation*}
+Each linear predictor $\eta_i$ is constructed from the Gaussian latent field such that
+\begin{equation*}
+\eta_i = \beta_0 + \sum_{l = 1}^{p} \beta_j z_{ji} + \sum_{k = 1}^{r} f_k(u_{ki}),
+\end{equation*}
+where the intercept $\beta_0$, linear effects $\beta_j$ of the covariates $z_{ji}$, and functions $f_k(\cdot)$ of the covariates $u_{ki}$ are each Gaussian and may be collected into $\x$.
+The conditional mean $\mathbb{E}(y_i \, | \, \eta_i)$ is given by $g(\eta_i)$ where $g: \mathbb{R} \to \mathcal{Y}$ is an inverse link function
+
+Spatial random effects $\phi(\cdot)$ are one Gaussian function which may be included in the model, whose role is to capture the spatial structure between areas not already accounted for by $\beta_0 + \sum_{l = 1}^{p} \beta_j z_{ji}$.
+Should no spatial structure remain, then independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ may be appropriate, where $\tau_\phi$ is a shared precision.
 Exploratory data analysis techniques such as visual inspection or Moran's $I$ coefficient [@cliff1981spatial] may be used to determine if there remains any spatial auto-correlation in the data.
 <!-- TODO: Add calculation of Moran's $I$ to my HIV survey data-sets. -->
-Specifying $\phi$ amounts to specifying the entries of a precision or covariance matrix, as we discuss in the following sections.
+Specifying $\bphi = (\phi_i)_{i \in [n]}$ amounts to specifying the entries of a precision or covariance matrix, as we discuss in the following sections.
 
@@
 
-Spatial structure can be encoded using a symmetric relation between areas, where $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring.
-Adjacency is often defined by a shared border, though other choices, such as inclusion of second or higher-degree neighbours [@paciorek2013spatial], are also possible.
-The Besag model [@besag1991bayesian] is a type of improper conditional auto-regressive (ICAR) model where the full conditional distribution of the $i$th spatial random effect is given by
+Spatial structure can be encoded using a symmetric relation between areas.
+Let $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring.
+Adjacency is often defined by a shared border, though other choices are also possible [@paciorek2013spatial].
+The Besag model [@besag1991bayesian] is an improper conditional auto-regressive (ICAR) model where the full conditional distribution of the $i$th spatial random effect is given by
 \begin{equation}
-    \phi_i \, | \, \phi_{-i} \sim \N \left(\frac{1}{n_{\delta i}} \sum_{j: j \sim i} \phi_j, \frac{1}{n_{\delta i}\tau_\phi}\right), \label{eq:besag}
+    \phi_i \, | \, \bphi_{-i} \sim \N \left(\frac{1}{n_{\delta i}} \sum_{j: j \sim i} \phi_j, \frac{1}{n_{\delta i}\tau_\phi}\right), \label{eq:besag}
 \end{equation}
-where $\delta i$ is the set of neighbours of $A_i$ with cardinality $n_{\delta i} = |\delta i|$ and $\phi_{-i} = (\phi_1, \ldots, \phi_{i - 1}, \phi_{i + 1}, \ldots, \phi_n)^\top$ is the vector of spatial random effects with the $i$th entry removed.
+where $\delta i$ is the set of neighbours of $A_i$ with cardinality $n_{\delta i} = |\delta i|$ and $\bphi_{-i}$ is the vector of spatial random effects with the $i$th entry removed.
 The conditional mean of the random effect $\phi_i$ is the average of its neighbours $\{\phi_j\}_{j \sim i}$ and the precision $n_{\delta i}\tau_\phi$ is proportional to the number of neighbours $n_{\delta i}$.
@@
 \begin{equation}
-    \phi \sim \mathcal{N}(0, \tau_\phi^{-1} R^{-}), \label{eq:gmrf}
+    \bphi \sim \mathcal{N}(\mathbf{0}, \tau_\phi^{-1} \bR^{-}), \label{eq:gmrf}
 \end{equation}
-where $R^{-}$ is the generalised inverse of the rank-deficient structure matrix $R$, so-called because it defines the \textit{structure} of the precision matrix, with entries
+where $\bR^{-}$ is the generalised inverse of the rank-deficient structure matrix $\bR$, so-called because it defines the structure of the precision matrix, with entries
 \begin{equation}
@@
 \end{equation}
-The Markov property arises due to the conditional independence structure $p(\phi_i \, | \, \phi_{-i}) = p(\phi_i \, | \, \phi_{\delta i})$ whereby each area only depends on its neighbours.
-This is reflected in the sparsity of $R$ whereby $\phi_i \perp \phi_j \, | \, \phi_{-ij}$ if and only if $R_{ij} = 0$.
-The structure matrix $R$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
+The Markov property arises due to the conditional independence structure $p(\phi_i \, | \, \bphi_{-i}) = p(\phi_i \, | \, \bphi_{\delta i})$ whereby each area only depends on its neighbours.
+This is reflected in the sparsity of $\bR$ whereby $\phi_i \perp \phi_j \, | \, \bphi_{-ij}$ if and only if $R_{ij} = 0$.
+The structure matrix $\bR$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
 Figure \ref{fig:geometry-graph-zwe} shows the adjacency graph for the districts of Zimbabwe alongside the geometry.
-Shortly, in Section \ref{sec:concerns} we discuss the appropriateness of using the graph rather than the complete geometry.
+In Section \ref{sec:concerns} we will discuss the appropriateness of using the graph rather than the complete geometry.
 
-Rewriting Equation \ref{eq:gmrf}, the probability density function of $\phi$ is
+Rewriting Equation \ref{eq:gmrf}, the probability density function of $\bphi$ is
 \begin{equation}
-    p(\phi)
-    \propto \exp \left( -\frac{\tau_\phi}{2} \phi^\top R \phi \right)
+    p(\bphi)
+    \propto \exp \left( -\frac{\tau_\phi}{2} \bphi^\top \bR \bphi \right)
     \propto \exp \left( -\frac{\tau_\phi}{2} \sum_{i \sim j} (\phi_i - \phi_j)^2 \right). \label{eq:pdfphi}
 \end{equation}
-This density is a function of the pairwise differences $\phi_i - \phi_j$ and so is invariant to the addition of a constant $c$ to each entry $p(\phi) = p(\phi + c1)$, leading to an improper uniform distribution on the average of the $\phi_i$.
-If $\mathcal{G}$ is connected, in that by traversing the edges, any vertex can be reached from any other vertex, then there is only one impropriety in the model and $\text{rank}(R) = n - 1$, while if $\mathcal{G}$ is disconnected, and composed of $n_c \geq 2$ connected components with index sets $I_1, \ldots, I_{n_c}$, then the corresponding structure matrix $R$ has rank $n - n_c$ and the density is invariant to the addition of a constant to each of the connected components $p(\phi_{I}) = p(\phi_{I} + c1)$ where $I = I_1, \ldots, I_{n_c}$.
+This density is a function of the pairwise differences $\phi_i - \phi_j$ and so is invariant to the addition of a constant $c$ to each entry $p(\bphi) = p(\bphi + c\mathbf{1})$, leading to an improper uniform distribution on the average of the $\phi_i$.
+If $\mathcal{G}$ is connected, in that by traversing the edges, any vertex can be reached from any other vertex, then there is only one impropriety in the model and $\text{rank}(\bR) = n - 1$, while if $\mathcal{G}$ is disconnected, and composed of $n_c \geq 2$ connected components with index sets $I_1, \ldots, I_{n_c}$, then the corresponding structure matrix $\bR$ has rank $n - n_c$ and the density is invariant to the addition of a constant to each of the connected components $p(\bphi_{I}) = p(\bphi_{I} + c\mathbf{1})$ where $I = I_1, \ldots, I_{n_c}$.
 
@@
 \includegraphics{depends/maup1.pdf}
-\caption{Each of these four geometries corresponds to the same adjacency graph.}
+\caption{Each of these four geometries have the same adjacency graph.}
 \label{fig:tikz1}
@@
 Since then, it has seen wider use, including in situations where the spatial structure is less regular.
-There are a number of concerns about the model's applicability to this broader setting, which we highlight below.
-Closely linked are the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups.
+We highlight a number of concerns about the model's applicability to this broader setting.
+This discussion is closely linked to the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups.
 <!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. -->
@@
 Many geometries share the same adjacency graph, and therefore the same probability model, as illustrated by Figure \ref{fig:tikz1}.
-This fact is not concerning in itself, though it prompts consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.\footnote{The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be.
-In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.}
-It is intuitive that the more regular the spatial structure the less information is lost in compression to an adjacency graph.
+This fact is not concerning in itself, but it prompts consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.^[The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be.
+In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.]
+Intuitively the more regular the spatial structure, the less information is lost in compression to an adjacency graph.
 In image analysis, very little spatial information is lost in compression a lattice structure to an adjacency graph.
@@
 
-A feature specific to the Besag model is that all adjacent areas count equally.
+In the Besag model is all adjacent areas count equally.
 This assumption is unsatisfying: for most geometries, we expect some areas to be more highly correlated to some neighbours than to others.
-Figure \ref{fig:tikz1} illustrates a number of heuristic features for neighbour importance, including length of shared border in geometry (iii), and the proximity of centers of mass in geometry (iv).
+Figure \ref{fig:tikz1} illustrates a number of heuristic features for neighbour importance, including length of shared border, and the proximity of centers of mass.
 This may be remedied using more general weighted ICAR models, as we outline in Section \ref{sec:icar}.
@@
 
-A striking feature of Equation \ref{eq:besag} is that the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.
+In Equation \ref{eq:besag} the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.
 It follow that as $n_{\delta i} \to \infty$ then $\text{Var}(\phi_i) \to 0$.
-This is illustrated by Figure \ref{fig:tikz2} where the area on the right is repeatedly divided such that its number of neighbours increases by one at each step.
+This is illustrated by Figure \ref{fig:tikz2} where the area on the right is repeatedly divided such that its number of neighbours increases.
 This property is a consequence of averaging the conditional mean over a greater number of areas, which, in certain situations, can correspond to a greater amount of information.
-However, if the amount of information in the shaded area remains fixed, it is inappropriate that $\text{Var}(\phi_1)$ should tend to zero just as a result of drawing additional, arbitrary, boundaries.
+However, if the amount of information in the shaded area remains fixed, it seems inappropriate that $\text{Var}(\phi_1)$ should tend to zero simply as a result of drawing additional, arbitrary, boundaries.
 In the image analysis setting this modelling assumption is reasonable: each pixel represents a fixed amount of information and a higher pixel density represents a greater amount of information.
-On the other hand, in public health and epidemiology, drawing boundaries to create additional regional areas cannot be expected to correspond to a greater amount of information and this assumption is not appropriate.
+On the other hand, in public health and epidemiology, drawing boundaries to create additional areas is not expected to correspond to a greater amount of information.
 
@@
 The Besag model is a special case of a more general class of (zero-mean) ICAR models in which each neighbour may not be weighted equally.
-These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\kappa = (\kappa_1, \ldots, \kappa_n)^\top$ such that $\phi_i \, | \, \phi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$.
-The structure matrix $R$ corresponding to the above full conditionals is given by $R = D_\kappa(I - B)$, where $B$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $D_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$.
+These models can be specified in terms of scaled weights $\{b_{ij}\}_{j \sim i}$ and a precision vector $\bkappa = (\kappa_i)_{i \in [n]}$ such that $\phi_i \, | \, \bphi_{-i} \sim \N ( \sum_{j: j \sim i} b_{ij} \phi_j, (\kappa_i \tau_\phi)^{-1} )$.
+The structure matrix $\bR$ corresponding to the above full conditionals is given by $\bR = \mathbf{D}_\kappa(\mathbf{I} - \mathbf{B})$, where $\mathbf{B}$ has the entries $b_{ij}$ for $i \sim j$, diagonal entries $b_{ii} = 0$ and $b_{ij} = 0$ for $i \nsim j$ and the matrix $\mathbf{D}_\kappa$ is given by $\text{diag}(\kappa_1, \ldots, \kappa_n)$.
 As the structure matrix is symmetric we must have the condition $-b_{ij}\kappa_i = -b_{ji}\kappa_j$.
-To meet this condition, it is often simpler to use an unscaled weights matrix $W = D_\kappa B$ such that $R = D_\kappa - W$.
+To meet this condition, it is often simpler to use an unscaled weights matrix $\mathbf{W} = \mathbf{D}_\kappa \mathbf{B}$ such that $\mathbf{R} = \mathbf{D}_\kappa - \mathbf{W}$.
 For example, in the Besag model $W$ corresponds to the adjacency matrix.
 The scaled weights can then be recovered by $b_{ij} = w_{ij} / \kappa_i$ where $\kappa_i = \sum_{k: k \sim i} w_{ik}$.
-A thorough discussion of methods for specifying $W$ is provided by @duncan2017spatial.
+A thorough discussion of methods for specifying $\mathbf{W}$ is provided by @duncan2017spatial.
 Much of the work in this area focuses on the case where the geometry is a lattice.
@@
 Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model.
-The Besag-York-Mollié (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\phi = v + u$ into a sum of an unstructured IID component $v$ and a spatially structured Besag component $u$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
+The Besag-York-Mollié (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\bphi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
 The resulting distribution is
 \begin{equation}
-    \phi \sim \N(0, \tau_v^{-1} I + \tau_u^{-1} R^{-}) \label{eq:bym}.
+    \bphi \sim \N(0, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} \mathbf{R}^{-}) \label{eq:bym}.
 \end{equation}
-Including both $v$ and $u$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$.
-However, in this specification, scaling of the Besag precision matrix $Q$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
-In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $Q$ and $I$ have the same scale.
-Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $v$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
+Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$.
+However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
+In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale.
+Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
 This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal.
@@
 \begin{align}
-\phi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, v + \sqrt{\pi} \, u^\star \right), \label{eq:bym2}
+\bphi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2}
 \end{align}
-where $\tau_\phi$ is the marginal precision of $\phi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $u^\star$ is a scaled version of $u$ with precision matrix given by the scaled structure matrix $R^\star$ (see Appendix).
+where $\tau_\phi$ is the marginal precision of $\bphi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $u$ with precision matrix given by the scaled structure matrix $\mathbf{R}^\star$ (see Appendix).
 When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model.
@@
 <!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
-The centroid kernel has been used in an environmental epidemiology setting [@wakefield1999spatial] and in model comparison study [@best2005comparison], both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$; as well as more recently by @teh2021efficient as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
-@wakefield1999spatial fix the length-scale $l$ using a semi-variogram and @best2005comparison place a prior on the inverse length-scale.
-\citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas, though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.
+The centroid kernel has been used in environmental epidemiology [@wakefield1999spatial] and to model the reproduction number of COVID-19 [@teh2021efficient].
+In a model comparison study \citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas (though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure).
 
@@
 \end{equation}
-This results in equivalent covariance structure to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].å
+This covariance structure is equivalent to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].
 <!-- @law2018variational develop a variational Bayesian inference method for aggregated observation models. -->
@@
 
-# Acknowledgements {-}
-
-Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
-AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations.
-
 # Disclaimer {-}

### 2023-05-31 src/docs_paper/appendix.Rmd
Update paper
Will Adam ever finish writing this paper? Stay tuned to find out (I hope so)

@@
 ---
-title: Appendix to "Models for spatial structure in small-area estimation"
-subtitle: |
-  Corresponding author: Adam Howes (`ath19@ic.ac.uk`)
+title: Appendix to "Understanding models for spatial structure in small-area estimation"
+author:
+  - Adam Howes^[Department of Mathematics, Imperial College London]
+  - Jeffrey W. Eaton^[Department of Epidemiology, Harvard University]
+  - Seth R. Flaxman^[Department of Computer Science, Oxford University]
 output:
@@
 
+```{r echo = FALSE}
+options(scipen = 100)
+knitr::opts_chunk$set(
+  echo = FALSE,
+  warning = FALSE,
+  message = FALSE,
+  dpi = 320,
+  cache = TRUE,
+  out.width = "95%",
+  fig.align = 'center'
+)
+
+comma <- function(x) formatC(x, format = "d", big.mark = ",")
+```
+
 <!-- So that the tables and figures in this appendix are prefixed by "A." -->
@@
 \renewcommand{\thefigure}{A.\arabic{figure}}
+
+\clearpage
+
+# Comparison of AGHQ to other inference methods \label{app:aghq-nuts}
 
@@
 
-1. \textit{Scaling}
-The structure matrix $R$ should be rescaled to have generalised variance, defined by the geometric mean of the diagonal elements of its generalised inverse
+## Scaling
+
+The structure matrix $\mathbf{R}$ should be rescaled to have generalised variance, defined by the geometric mean of the diagonal elements of its generalised inverse
 \begin{equation}
-    \sigma^2_{\text{GV}}(R) = \prod_{i = 1}^n (R^-_{ii})^{1/n} = \exp \left( \frac{1}{n} \sum_{i = 1}^n \log (R^-_{ii}) \right),
+    \sigma^2_{\text{GV}}(\mathbf{R}) = \prod_{i = 1}^n (\mathbf{R}^-_{ii})^{1/n} = \exp \left( \frac{1}{n} \sum_{i = 1}^n \log (\mathbf{R}^-_{ii}) \right),
 \end{equation}
-equal to one, by replacing $R$ with $R^\star = R / \sigma^2_{\text{GV}}(R)$.
+equal to one, by replacing $\mathbf{R}$ with $\mathbf{R}^\star = \mathbf{R} / \sigma^2_{\text{GV}}(\mathbf{R})$.
 As the diagonal elements $R^-_{ii}$ correspond to marginal variances, the generalised variance gives a measure of the average marginal variance.
@@
 Allowing the variance to be controlled by $\tau_\phi$ alone is important as it allows for consistent, interpretable prior selection.
-When the adjacency graph is disconnected it is not appropriate to scale the structure matrix $R$ uniformly since for a given precision $\tau_\phi$, local smoothing operates on each connected component independently.
-As such, each connected component should be scaled independently to have generalised variance one giving $R^\star_I = R_I / \sigma^2_{\text{GV}}(R_I)$ where $R_I$ is the sub-matrix of the structure matrix corresponding to index set $I$.
+When the adjacency graph is disconnected it is not appropriate to scale the structure matrix $\mathbf{R}$ uniformly since for a given precision $\tau_\phi$, local smoothing operates on each connected component independently.
+As such, each connected component should be scaled independently to have generalised variance one giving $\mathbf{R}^\star_I = R_I / \sigma^2_{\text{GV}}(R_I)$ where $R_I$ is the sub-matrix of the structure matrix corresponding to index set $I$.
 
-2.
-\textit{Singletons}
+## Singletons
+
 When one of the connected components is a single area, known either as a singleton or an island, the probability density $\exp \left( -\frac{\tau_\phi}{2} \sum_{i \sim j} (\phi_i - \phi_j)^2 \right)$ has no dependence on $\phi_i$.
@@
 
-3.
-\textit{Constraints}
+## Constraints
+
 To avoid confounding of the spatial random effects with the intercept, it is recommended to place a sum-to-zero constraint on each non-singleton connected component.
@@
 
-# Simulation study
+# Further results for the simulation study
+
+```{r results="asis"}
+#' Some of the lines should not be included
+lines_remove <- c(1, 3, 4, 5)
+cat(readLines("depends/crps-table-rho.txt")[-lines_remove], sep = "\n")
+```
 
@@
 
-# HIV study
+# Further results for the HIV study
 

### 2023-05-31 src/docs_paper/paper.Rmd
Will Adam ever finish writing this paper? Stay tuned to find out (I hope so)

@@
 supplements:
-  - title: Title of Supplement A
-    description: Short description of Supplement A.
-  - title: Title of Supplement B
-    description: Short description of Supplement B.
+  - title: Appendix
+    description: All further materials for this study.
 acknowledgements: |
-  | Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
-AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations.
+  | AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Tim Wolock, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations.
+Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
 abstract: |
@@
 Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
-@duncan2017spatial compared seventeen methods for specifying these weights and did not find any which outperform the Besag model, though this conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
+@duncan2017spatial compared seventeen methods for specifying these weights, but did not find any which outperform the Besag model.
+This conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
 Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
 Obtaining inference about a variable at a different resolution to that it was observed at is known in geostatistics as the change-of-support problem [@gelfand2001change], and dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
-Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
+@kelsall2002modeling applied ideas from change-of-support modelling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
 This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially].
 <!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
-Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].
+Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2016tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].
 
-Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
-Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models.
+Spatial heterogeneity commonly reflects a combination of both spatially correlated processes and specific circumstances at a given area.
+It is usually therefore recommended to use a spatial random effect specification which includes both structured and unstructured components, such as the BYM2 model [@simpson2017penalising] or earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models.
 Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive.
-Improvements to the Besag model are likely transferable to improving these convolution models.
-The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and both models are of substantive practical interest.
+Any improvement to the Besag model is likely directly transferable to improving these convolution models, and both are of substantive practical interest.
+For example, the Naomi HIV small-area estimation model [@eaton2021naomi] uses both Besag and BYM2 random effects.
 
@@
 In Section \ref{sec:ab}, we review developments in specifying spatial random effects based on adjacency, before presenting an alternative approach based on kernels in Section \ref{sec:ks}.
-In Section \ref{sec:simulation} we compare models using simulated data, before applying them to mapping HIV prevalence in sub-Saharan Africa in Section \ref{sec:hiv}.
+In Section \ref{sec:simulation} we present a simulated case-study, before moving on to mapping HIV prevalence in sub-Saharan Africa in Section \ref{sec:hiv}.
 Finally, we discuss our conclusions and directions for future research in Section \ref{sec:discussion}.
@@
 <!-- @riebler2016intuitive reach a similar conclusion about both the Leroux and BYM2 model. -->
-
-\newpage
 
@@
 
-\newpage
-
-# Spatial random effects using adjacency \label{sec:ab}
+# Spatial random effects defined using adjacency \label{sec:ab}
 
@@
 When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model.
-To borrow an analogy, the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately [@rue2020comment].
+To borrow an analogy [@rue2020comment] the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately.
 
-\newpage
-
-# Spatial random effects using kernels \label{sec:ks}
+# Spatial random effects defined using kernels \label{sec:ks}
 
@@
 \includegraphics[width=\textwidth]{depends/integration-method.pdf}
-\caption{Three ways to choose integration points with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample}.}
+\caption{Three ways to choose nodes with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample}.}
 \label{fig:41}
@@
 
+### Accounting for heterogeneity
+
 If available, additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel.
@@
 
-If areal kernel spatial random effects are created using the integrated kernel, then aggregation occurs at the level of the latent field in the three stage model rather than at the level of the data.
-This corresponds to a generative model of the form $y_i \sim p(y_i \, | \,  g^{-1}(x_i))$, with $x_i = |A_i|^{-1}\int_{A_i} x(s) \text{d} s$.
-If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data.
-On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model does not match the proposed data generating process $y_i = |A_i|^{-1}\int_{A_i} y(s) \text{d}s$.
-<!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. -->
-<!-- Explain under what circumstances we believe this approximation to be acceptable. -->
+### Computation
 
-Most of the time we do not expect to be able to calculate the integral in Equation \ref{eq:ik2} analytically.
+Most of the time we do not expect to be able to calculate Equation \ref{eq:ik2} analytically.
 Instead, given $n$ collections of $L_i$ samples $\{s^{(i)}_l \}_{l = 1}^{L_i} \sim \mathcal{U}(A_i)$ drawn uniformly from each area then the integral may be approximated using Monte Carlo by the double sum
@@
 Equivalently, samples drawn from $W_i$ may be used without weighting by $w_i(s)$.
-Integration points may also be selected by some deterministic process, and used in Equation \ref{eq:mcintegratedkernel} to give a numerical integration estimate of the kernel.
-Taking any of these approaches requires $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
-This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or integration points it is feasible to use.
-To make inference more feasible, @kelsall2002modeling reduce the number of Gram matrix constructions and inversions required, by using a discrete hyperparameter prior.
+Nodes may also be selected by some deterministic process to give a numerical quadrature estimate of the kernel.
+These approaches require $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
+This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or nodes it is feasible to use.
+@kelsall2002modeling make inference more feasible by using a discrete hyperparameter prior to reduce the number of Gram matrix constructions and inversions required.
+
+### Mismatch to data generating process
+
+Aggregation via the integrated kernel occurs at the level of the latent field rather than at the level of the data.
+<!-- This corresponds to a generative model of the form $y_i \sim p(y_i \, | \,  g^{-1}(x_i))$, with $x_i = |A_i|^{-1}\int_{A_i} x(s) \text{d} s$. -->
+If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data.
+On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model does not match the proposed data generating process.
+<!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. -->
+<!-- Explain under what circumstances we believe this approximation to be acceptable. -->
 
@@
 Disaggregation regression, also known as downscaling or interpolation, is another closely related approach.
-Rather than focusing on the aggregate nature of areal observations as primarily a route towards better area-level estimates, disaggregation regression aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
+Rather than focusing on the aggregate nature of areal observations as primarily a route towards better area-level estimates, disaggregation regression aims to produce high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
 <!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
@@
 <!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. -->
-
-\newpage
 
@@
 \centering
-\includegraphics{depends/simulation-geometries.pdf}
+\includegraphics[width=\textwidth]{depends/simulation-geometries.pdf}
 \caption{The seven simulation geometries that we considered.}
@@
 The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`.
-We used the `orderly` package [@orderly] for reproducible research.
+We used `orderly` [@orderly] for reproducible research, `ggplot2` for data visualisation [@wickham2016ggplot2] and `rticles` [@allaire2022rticles] for reporting via `rmarkdown` [@allaire2022rmarkdown].
 
@@
 \midrule
-\hypertarget{S1}{\texttt{S1}} & IID & $\phi \sim \mathcal{N}(0, I_n)$ \\
-\hypertarget{S2}{\texttt{S2}} & Besag & $\phi \sim \mathcal{N}(0, (R^\star)^{-})$ as described in Section \ref{sec:besag} where $R^\star$ is the scaled structure matrix \\
-\hypertarget{S3}{\texttt{S3}} & IK & $\phi \sim \mathcal{N}(0, K^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation \ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
+\hypertarget{S1}{\texttt{S1}} & IID & $\bphi \sim \mathcal{N}(0, \mathbf{I}_n)$ \\
+\hypertarget{S2}{\texttt{S2}} & Besag & $\bphi \sim \mathcal{N}(0, (\mathbf{R}^\star)^{-})$ as described in Section \ref{sec:besag} where $\mathbf{R}^\star$ is the scaled structure matrix \\
+\hypertarget{S3}{\texttt{S3}} & IK & $\bphi \sim \mathcal{N}(0, \mathbf{K}^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation \ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
 Like the Besag model, we scaled the Gram matrix to have generalised variance one \\
@@
 \end{tabularx}
-\caption{The three simulation models from which we generate $\phi$.}
+\caption{The three simulation models from which we generate $\bphi$.}
 \label{tab:simulationmodels}
@@
 To study robustness to spatial misspecification we simulated synthetic data-sets from three known data generating processes, using spatial random effects $\bphi$ generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}).
-We then generated synthetic data $\y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via
+We then generated synthetic data $\y = (y_i)_{i \in [n]}$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the linear predictors $\eta_i \in \mathbb{R}$ via
 \begin{equation}
-\log \left( \frac{\rho_i}{1 - \rho_i} \right) = x_i = \beta_0 + \phi_i, \quad i = 1, \ldots, n.
+\log \left( \frac{\rho_i}{1 - \rho_i} \right) = \eta_i = \beta_0 + \phi_i, \quad i \in [n].
 \end{equation}
-The sample sizes were fixed as $m_i = 25$ for all $i$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$.
-We used seven different geometries: the four vignette geometries sharing an adjacency graph in Figure \ref{fig:tikz1}, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of Côte d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
+The sample sizes were fixed as $m_i = 25$ for all $i \in [n]$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$.
+We used seven different geometries including the four vignette geometries sharing an adjacency graph (Figure \ref{fig:tikz1}), as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of Côte d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
 The more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test how model performance varies by geometry regularity.
@@
 \small
-\begin{tabularx}{\textwidth}{lllX}
+\begin{tabularx}{\textwidth}{llX}
 \toprule
-& Inferential model & Algorithm & Additional details \\
+& Inferential model & Additional details \\
  \midrule
-\hypertarget{I1}{\texttt{I1}} & Constant & AGHQ &
+\hypertarget{I1}{\texttt{I1}} & Constant &
 No spatial random effects \label{row:I1} \\
-\hypertarget{I2}{\texttt{I2}} & IID & AGHQ &
-$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} I_n)$\\
-\hypertarget{I3}{\texttt{I3}} & Besag & AGHQ &
-$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations of \citet{freni2018note}  \\
-\hypertarget{I4}{\texttt{I4}} & BYM2 & AGHQ &
-$\phi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, v + \sqrt{\pi} \, u^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
-\hypertarget{I5}{\texttt{I5}} & FCK & AGHQ &
-$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
-\hypertarget{I6}{\texttt{I6}} & CK & AGHQ & As in model \hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
-\hypertarget{I7}{\texttt{I7}} & FIK & AGHQ & $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
-\hypertarget{I8}{\texttt{I8}} & IK & AGHQ & As in model \hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
+\hypertarget{I2}{\texttt{I2}} & IID &
+$\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{I}_n)$\\
+\hypertarget{I3}{\texttt{I3}} & Besag &
+$\bphi \sim \mathcal{N}(0, (\tau_\phi \mathbf{R}^\star)^{-})$ following additional recommendations of \citet{freni2018note}  \\
+\hypertarget{I4}{\texttt{I4}} & BYM2 &
+$\bphi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
+\hypertarget{I5}{\texttt{I5}} & FCK &
+$\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{K})$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
+\hypertarget{I6}{\texttt{I6}} & CK & As \hypertarget{I5}{\texttt{I5}}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
+\hypertarget{I7}{\texttt{I7}} & FIK & $\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{K})$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
+\hypertarget{I8}{\texttt{I8}} & IK & As \hypertarget{I7}{\texttt{I7}}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
 \bottomrule
 \end{tabularx}
-\caption{All inferential models are implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}.}
+\caption{All inferential models are implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}.
+The AGHQ algorithm was used for approximate Bayesian inference of all inferential models.}
 \label{tab:inferentialmodels}
@@
 We placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
-The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one, neither of which we assumed to be plausible.
+The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one.
 A weakly informative $\mathcal{N}(-2, 1)$ prior was placed on $\beta_0$, setting most of the prior probability density for $\rho_i$ within a range $[0, 0.25]$ typical for a disease prevalence.
@@
 
-For the areal kernel models we used the Matérn [@stein1999interpolation] kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
+We used the Matérn [@stein1999interpolation] kernel $k: \Sc \times \Sc \to \mathbb{R}$ given by
 \begin{equation}
@@
 In models \hyperlink{I6}{\texttt{I6}} and \hyperlink{I8}{\texttt{I8}}, we used a length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$, with parameters $a$ and $b$ chosen for each model such that 1\% of the prior mass is below 0.1 and 1\% of the prior mass is above the maximum distance between points in the model [@betancourt2017robust].
-For the integration points we set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51}.
-Note that `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].
-We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{I8}.
-For Model \hyperlink{I6}{I6} although it is feasible to use a much larger number of integration points, since the Gram matrix is only computed once, we chose not to for comparability.
+
+We set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51}.^[Note that `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].]
+We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{\texttt{I8}}.
+For Model \hyperlink{I6}{\texttt{I6}} the Gram matrix is only computed once and it is feasible to use a much larger number of nodes, but we chose not to for comparability.
 
@@
 
-We used adaptive Gauss-Hermite quadrature (AGHQ) with $k = 3$ quadrature points per hyperparamter dimension to perform approximate Bayesian inference.
+We used adaptive Gauss-Hermite quadrature [AGHQ; @stringer2022fast] with $k = 3$ quadrature points per hyperparameter dimension to perform approximate Bayesian inference.
+Each model is implemented using a Template Model Builder \textsc{C++} template for the log-posterior via the `TMB` package [@kristensen2016tmb]
+We are then able to conduct inference using both AGHQ via the `aghq` package [@stringer2021implementing] and the No-U-Turn Sampling (NUTS) Hamiltonian Monte Carlo (HMC) algorithm using Stan [@carpenter2017stan] via the `tmbstan` package [@monnahan2018no].
+In Appendix we demonstrate that the inference results from AGHQ are comparable to those obtained running NUTS over a longer time.
+We also compare to empirical Bayes, and where possible integrated nested Laplace approximation via `R-INLA` [@martins2013bayesian].
 
@@
 
-The continuous ranked probability score (CRPS) [@matheson1976scoring] gives an overall measure of forecasting performance, and can be seen as a generalisation of the Brier score [@brier1950verification] to distributional forecasts.
-For a given fitted model and area $A_i$ let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, y)$ and $\omega_i$ be the true value.
+The continuous ranked probability score [CRPS; @matheson1976scoring] generalises the Brier score [@brier1950verification] to distributional forecasts.
+Let $\rho_i$ have posterior marginal $f(\rho_i) = p(\rho_i \, | \, y)$ and $\omega_i$ be the true value.
 Writing $F$ as the cumulative distribution function corresponding to $f$ then
@@
 where $\mathbbm{1}$ denotes the indicator function and the second equality follows from $0 \leq \rho_i \leq 1$.
-The minimum possible value of the CRPS is zero, corresponding to a a point mass correctly forecast at the true value $f(\rho_i) = \delta_{\omega_i}$.
-The CRPS is a strictly proper scoring rule such that the uniquely best approach to minimising the score is for the model to report its true probabilistic beliefs about the quantity in question [@gneiting2007strictly].
-The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
-In contrast, the CRPS can be evaluated directly using samples $\{\rho_i^s\}_{s = 1}^S \sim p(\rho_i \, | \, y)$ as
+$\text{CRPS}(f, \omega_i)$ if and only if $f(\rho_i) = \delta_{\omega_i}$.
+The CRPS is a strictly proper scoring rule [@gneiting2007strictly] and can be evaluated directly using samples $\{\rho_i^s\}_{s = 1}^S$ as
 \begin{equation}
@@
 \end{equation}
+<!-- The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
+-->
 
@@
 
-We assessed the coverage of our estimates using posterior predictive checks based on the quantile $q_i = F(\omega_i)$ of the true value within each marginal posterior predictive distribution, computed using numerical integration.
-For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also be $1 - \alpha$.
-To check uniformity we used probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2021graphical].
+For calibrated models, over repeated simulations, $F(\omega_i) = q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also $1 - \alpha$.
+We checked uniformity using probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2022graphical].
 
@@
 
-```{r results="asis"}
-#' Some of the lines should not be included
-lines_remove <- c(1, 3, 4, 5)
-cat(readLines("depends/crps-table-rho.txt")[-lines_remove], sep = "\n")
-```
+See Appendix.
 
@@
 
-We compared model performance in estimating HIV prevalence in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire (which has $n = 33$ districts), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$).
-As before, we varyied the spatial random effect specification.
+We compared model performance in estimating HIV prevalence $\rho_i \in [0, 1]$ in adults aged $15-49$ across four countries in sub-Saharan Africa: Côte d'Ivoire (which has $n = 33$ districts), Malawi ($n = 28$), Tanzania ($n = 159$) and Zimbabwe ($n = 60$).
+As before, we varied the spatial random effect specification.
 The \textsf{R} code for this study is available from \url{https://github.com/athowes/areal-comparison}.
@@
 
-We used data from the most recent publicly available Population Health Impact Assessment (PHIA) survey in each country.
-These surveys utilise a complex design, where each individual $j$ in area $i$ has an unequal probability $\pi_{ij}$ of being included in the sample.
-A two-stage design in which enumeration areas are first drawn from a stratified sample and then households are chosen using equal probability systematic sampling from within each enumeration area, is common.
-To account for the survey design we used sampling weights $w_{ij} = 1 / \pi_{ij}$ to adjust the raw data in each district, obtaining the Kish effective sample size $m_i^\star = (\sum_k w_{ij})^2 / \sum_k w_{ij}^2$ and effective number of cases $y_i^\star$ which may be thought of as what would have been observed had the survey been a simple random sample.
+In each country we used data from the most recent publicly available Population Health Impact Assessment (PHIA) survey.
+These surveys utilise a complex design, where each individual $j$ in area $i$ has an unequal probability $\pi_{ij}$ of being included in the sample: a two-stage design in which enumeration areas are first drawn from a stratified sample and then households are chosen using equal probability systematic sampling from within each enumeration area is common.
+We used sampling weights $w_{ij} = 1 / \pi_{ij}$ to account for the survey design by adjusting the raw data in each district, obtaining the Kish effective sample size [@kish1965survey] $m_i^\star = (\sum_k w_{ij})^2 / \sum_k w_{ij}^2$ and effective number of cases $y_i^\star$ which may be thought of as what would have been observed had the survey been a simple random sample.
 
@@
 
-We used the eight inferential models from Table \ref{tab:inferentialmodels}, with a small alteration to the likelihood.
-The effective number of cases and effective sample size may not be integers, so we modelled the effective number of cases $y_i^\star \in \mathbb{R}$ according to a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$, where $m_i^\star \in \mathbb{R}$ is the effective sample size and $\rho_i \in [0, 1]$ is the adult (15-49) HIV prevalence.
+We make a slight alteration to the inferential models from Table \ref{tab:inferentialmodels}.
+As the effective number of cases $y_i^\star \in \mathbb{R}$ and effective sample size $m_i^\star \in \mathbb{R}$ may not be integers, we use a generalised binomial distribution $y_i^\star \sim \text{xBin}(m_i^\star, \rho_i)$.
 The working likelihood under this model for $m^\star_i \geq y^\star_i$ is given by
@@
 
-We assessed each model using two cross-validation approaches: (1) a standard leave-one-out cross-validation (LOO-CV), in which at each fold the data from a single district were held-out, and (2) a spatial leave-one-out cross-validation (SLOO-CV), in which at each fold the data from a single district and each of its neighbouring districts were removed.
+We assessed each model using two approaches: (1) a standard leave-one-out cross-validation (LOO-CV), and (2) a spatial leave-one-out cross-validation (SLOO-CV).
 <!-- We performed both types of cross-validation manually, as we found approximations that avoid refitting the model, such as the built-in conditional predictive ordinate statistic in `R-INLA` package [@held2010posterior] and the Pareto smoothed importance sampling [@vehtari2015pareto] approach implemented in the `loo` package [@vehtari2017practical], were inaccurate in this setting.
@@
 In the $i$th LOO-CV fold, the model was fit using the data $\y_{-i}$ and assessed according to its prediction on $y_i$.
-We assessed forecasting performance using the root mean square error (RMSE), mean absolute error, and CRPS, at the level of the data $y_i$ rather than at the level of the prevalence $\rho_i$.
-To compute the value of CRPS for the $i$th fold we use samples $\{ y_i^s \}_{s = 1}^S$ from the LOO posterior predictive distribution $p(y_i \, | \, \y_{-i})$ giving $\text{LOO-CRPS}_i = \frac{1}{S} \sum_{s = 1}^S | y_i^s - y_i | - \frac{1}{2S^2} \sum_{s = 1}^S \sum_{l = 1}^S | y_i^s - y_i |.
-\label{eq:crpsapprox2}$
+We assessed forecasting performance using the root mean square error (RMSE), mean absolute error (MAE), and CRPS, at the level of the data $y_i$ rather than at the level of the prevalence $\rho_i$.
 
@@
 
-In the presense of structural dependencies, like spatial dependence, LOO-CV tends to underestimate predictive error [@le2014spatial; @roberts2017cross].
-To counteract this effect during the $i$th fold we left out the block $(i, \delta i)$ (Figure \ref{fig:sloo}), increasing the extent to which the training and validation sets are conditionally independent.
-From the SLOO posterior predictive distribution $p(y_i \, | \, \y_{-(i, \delta i)})$  we analogously computed $\text{SLOO-CRPS}_i$.
+In the presence of structural dependencies, LOO-CV tends to underestimate predictive error [@le2014spatial; @roberts2017cross].
+To counteract this effect, during fold $i$ we left out the block $(i, \delta i)$, increasing the extent to which the training and validation sets are conditionally independent, and predicted on $y_i$ (Figure \ref{fig:sloo}).
 
@@
 
-\newpage
-
 # Discussion \label{sec:discussion}
@@
 Though it may be possible to learn more sophisticated methods for combining kernels using data [@gonen2011multiple], we do believe that this is feasible in the small-area estimation setting, and instead focused on constructing areal kernels which represent our prior beliefs about spatial processes.
-In more data rich settings it may be possible to learn the integration points used within the integrated kernel [@campbell2019automated].
+In more data rich settings it may be possible to learn the nodes used within the integrated kernel [@campbell2019automated].
 

### 2023-05-31 src/docs_paper/paper.Rmd
Fixes to bibliography captials, some paper small improvements

@@
 
-```{r geometry-graph-zwe, fig.align="center", fig.cap="A map of the districts of Zimbabwe together with the corresponding adjacency graph structure $\\mathcal{G}$. \\label{fig:geometry-graph-zwe}"}
+```{r geometry-graph-zwe, fig.align="center", fig.cap="A map of the districts of Zimbabwe together with the corresponding adjacency graph structure $\\mathcal{G}$.
+The Besag model operates on the adjacency graph, rather than the geometry.
+\\label{fig:geometry-graph-zwe}"}
 knitr::include_graphics("depends/geometry-graph-zwe.pdf")
@@
 Let $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring.
-Adjacency is often defined by a shared border, though other choices are also possible [@paciorek2013spatial].
+Adjacency is often defined by a shared border, though other choices are possible [@paciorek2013spatial].
 The Besag model [@besag1991bayesian] is an improper conditional auto-regressive (ICAR) model where the full conditional distribution of the $i$th spatial random effect is given by
@@
 The Markov property arises due to the conditional independence structure $p(\phi_i \, | \, \bphi_{-i}) = p(\phi_i \, | \, \bphi_{\delta i})$ whereby each area only depends on its neighbours.
-This is reflected in the sparsity of $\bR$ whereby $\phi_i \perp \phi_j \, | \, \bphi_{-ij}$ if and only if $R_{ij} = 0$.
+This is reflected in the sparsity of $\bR$ such that $\phi_i \perp \phi_j \, | \, \bphi_{-ij}$ if and only if $R_{ij} = 0$.
 The structure matrix $\bR$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
@@
 We highlight a number of concerns about the model's applicability to this broader setting.
-This discussion is closely linked to the modifiable areal unit problem [@openshow1979million] whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, and the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation], which applies particularly to inference about individuals and groups.
+This discussion is closely linked to the modifiable areal unit problem [@openshow1979million], whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, as well as the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation].
 <!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. -->
@@
 
-Summarising all spatial information with an adjacency graph represents a loss of information.
+Summarising a geometry by an adjacency graph represents a loss of information.
 Many geometries share the same adjacency graph, and therefore the same probability model, as illustrated by Figure \ref{fig:tikz1}.
-This fact is not concerning in itself, but it prompts consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.^[The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be.
+This fact is not concerning in itself, but does prompt consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.^[The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be.
 In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.]
 Intuitively the more regular the spatial structure, the less information is lost in compression to an adjacency graph.
-In image analysis, very little spatial information is lost in compression a lattice structure to an adjacency graph.
+In image analysis, very little spatial information is lost in compression of a lattice structure to an adjacency graph.
 On the other hand, the regions of a country, determined by political and geographic forces, tend to display greater irregularity.
@@
 In the Besag model is all adjacent areas count equally.
-This assumption is unsatisfying: for most geometries, we expect some areas to be more highly correlated to some neighbours than to others.
+This assumption is unsatisfying: for most geometries, we expect different amounts of correlation between neighbours.
 Figure \ref{fig:tikz1} illustrates a number of heuristic features for neighbour importance, including length of shared border, and the proximity of centers of mass.
-This may be remedied using more general weighted ICAR models, as we outline in Section \ref{sec:icar}.
+More general weighted ICAR models remedy this problem, as we outline in Section \ref{sec:icar}.
 
@@
 The simplest approach is to use a single point.
-A natural choice is the centroid $c_i \in A_i$, given by the arithmetic mean of the latitude and longitude, which we might hope to be representative of the area.
+A natural choice is the centroid $c_i \in A_i$, given by the arithmetic mean of the latitude and longitude, which may be representative of the area.^[Though note that it is not guaranteed for the centroid to (even) lie within the area.]
 This results in the centroid kernel
@@
 The centroid kernel has been used in environmental epidemiology [@wakefield1999spatial] and to model the reproduction number of COVID-19 [@teh2021efficient].
-In a model comparison study \citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas (though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure).
+In a model comparison study \citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas.^[Though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.]
 
@@
 \end{equation}
-This covariance structure is equivalent to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].
+This covariance structure is equivalent to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, and has been studied in the machine learning literature under the title aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].
 <!-- @law2018variational develop a variational Bayesian inference method for aggregated observation models. -->
@@
 
-If available, additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel.
+Additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel.
 This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
@@
 
-### Connection to log-Gaussian Cox processes
+### Log-Gaussian Cox processes
 
@@
 
-### Connection to disaggregation regression
+<!-- Discuss Gary's paper here? [@konstantinoudis2020discrete] -->
+
+### Disaggregation regression
 
 Disaggregation regression, also known as downscaling or interpolation, is another closely related approach.
-Rather than focusing on the aggregate nature of areal observations as primarily a route towards better area-level estimates, disaggregation regression aims to produce high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
+Rather than focusing on the aggregate nature of areal observations as primarily a route towards better area-level estimates, disaggregation regression aims to produce high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2023disaggregation].
 <!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
@@
 $\text{CRPS}(f, \omega_i)$ if and only if $f(\rho_i) = \delta_{\omega_i}$.
-The CRPS is a strictly proper scoring rule [@gneiting2007strictly] and can be evaluated directly using samples $\{\rho_i^s\}_{s = 1}^S$ as
+The CRPS is a strictly proper scoring rule^[As is the logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.] [@gneiting2007strictly] and can be evaluated directly using samples $\{\rho_i^s\}_{s = 1}^S$ as
 \begin{equation}
@@
 \end{equation}
-<!-- The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another popular strictly proper scoring rule, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.
--->
 

### 2023-06-23 docs/resources.Rmd
Update to new repo name

@@
 abstract: |
-  A compliation of links to resources which I've found useful while working on `areal-comparison`.
+  A compliation of links to resources which I've found useful while working on `beyond-borders`.
 ---

### 2023-06-23 src/1_checks/lengthscale-recovery.Rmd
Update to new repo name

@@
 #' l = 1
-ck_one <- sim_ck(grid, nsim = 10, l = 1, save = FALSE)
+ck_one <- sim_ck(grid, nsim = 10, l = 1)
 
@@
   geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
-  geom_hline(yintercept = 1, size = 0.75, col = "grey", alpha = 0.75) +
+  geom_hline(yintercept = 1, linewidth = 0.75, col = "grey", alpha = 0.75) +
   labs(x = "Replicate", y = "Posterior CrI", title = "Is l = 1 recovered?") +
@@
   geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
-  geom_hline(yintercept = 1, size = 0.75, col = "grey", alpha = 0.75) +
+  geom_hline(yintercept = 1, linewidth = 0.75, col = "grey", alpha = 0.75) +
   labs(x = "Replicate", y = "Posterior CrI", title = "Is sigma = 1 recovered?") +
@@
 
-plot(df_sigma_one$mean, df_lengthscale_one$mean)
+cor(df_sigma_one$mean, df_lengthscale_one$mean)
 
@@
 #' Contour plots of lengthscale and sigma
-
 cbind(bind_rows(lengthscale_draws_one, .id = "replicate"), bind_rows(sigma_draws_one)) %>%
@@
 
-# ck_three <- sim_ck(grid, nsim = 30, l = 3, save = FALSE)
+# ck_three <- sim_ck(grid, nsim = 10, l = 3, save = FALSE)
 #
@@
 #   geom_errorbar(aes(ymin = X2.5., ymax = X97.5.)) +
-#   geom_hline(yintercept = 3, col = lightblue, size = 1.5)
+#   geom_hline(yintercept = 3, col = lightblue, linewidth = 1.5)
 #
@@
 ggplot(data = data.frame(x = c(0, 10)), aes(x)) +
-  stat_function(fun = dgamma, n = 101, args = list(1, 1))
+  stat_function(fun = dgamma, n = 101, args = list(1, 1)) +
+  theme_minimal()
 

### 2023-06-23 src/docs_paper/appendix.Rmd
Update to new repo name

@@
 ---
-title: Appendix to "Understanding models for spatial structure in small-area estimation"
+title: |
+  | Appendix to "Beyond borders: evaluating the suitability of spatial adjacency for small-area estimation"
 author:
@@
 
-# Comparison of AGHQ to other inference methods \label{app:aghq-nuts}
+# Comparison of AGHQ to NUTS \label{app:aghq-nuts}
 

### 2023-06-23 src/docs_paper/paper.Rmd
Update to new repo name

@@
 ---
-title: Understanding models for spatial structure in small-area estimation
+title: |
+  | Beyond borders: evaluating the suitability of spatial adjacency for small-area estimation
 classoption: ba
@@
   |  Spatial correlation between small-areas is typically accounted for using spatial random effects.
-However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions for non-grid geometries.
-Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous geostatistical data.
+However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic spatial assumptions for geometries which are not a grid.
+Furthermore, they do not take into account areal data is commonly generated by aggregating continuous geostatistical data.
 Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
@@
 We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
+  |
+  |  The results of this analysis are forthcoming, and will be described here.
 date: "`r Sys.Date()`"
@@
 Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
-The most common class of models used to specify spatial random effects are Gaussian Markov random fields (GMRFs), which combine a Gaussian distribution with Markov conditional independence assumptions between areas [@rue2005gaussian].
-Under the Markov assumptions, observations made in areas close together are correlated, and more distant relationships are ignored.
+The most common class of models used to specify spatial random effects are Gaussian Markov random fields (GMRFs).
+These models combine a Gaussian distribution with Markov conditional independence assumptions between areas [@rue2005gaussian].
+It is assumed that observations made in areas close together are correlated, and more distant relationships are ignored.
 Perhaps the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
-This model is attractive as it requires minimal additional choices and is widely implemented.
+This model is attractive as it requires minimal additional modelling choices and is widely implemented.
 As such, it has seen widespread use, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
 However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
-As such, in this work we test the hypothesis that using more realistic assumptions about spatial structure improves the performance of small-area estimation models.
+We test the hypothesis that using more realistic assumptions about spatial structure improves the performance of small-area estimation models.
 In doing so, we provide practical recommendations and further understanding of models for spatial structure.
@@
 Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
-Estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effectively used to respond to the epidemic [@hallett2016evaluation; @cuadros2017mapping].
-Household surveys provide nationally represenative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
-Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour [@howes2023spatio] or the prevalence of male circumcision, often face analogous measurement difficulties.
-Possibly as a result, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
-This stands in contrast to other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
+These estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effectively used to respond to the epidemic [@hallett2016evaluation; @cuadros2017mapping].
+Household surveys provide nationally represenative data about the general population, but are expensive to carry out, and as such have small sample sizes at a district level.
+Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour [@howes2023spatio] or the prevalence of male circumcision, often face similar measurement difficulties.
+Potentially as a result, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
+In contrast, covariates do see substantial use for mapping other infectious diseases, like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
 These circumstances foreground the importance of models for spatial structure in HIV mapping.
@@
 @duncan2017spatial compared seventeen methods for specifying these weights, but did not find any which outperform the Besag model.
-This conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
+This conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) recommended against by @vehtari2017practical.
 Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
@@
 <!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
-Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2016tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].
+Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2016tmb], and made available as a part of the `disaggregation` package [@nandi2023disaggregation].
 
@@
 Examples of geostatistical spatial data include the temperature at a monitoring station, the response to a household survey, and the quality of a soil sample.
-
 Areal data may often be conceptualised as arising from aggregation of geostatistical data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
-Notable exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a truly discrete spatial structure.
+Notable exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures.
 <!-- Note that, for data of this sort, there is less clear justification for using spatial smoothing. -->
@@
 
-In some situations it may be preferable to work using aggregated areal data rather than the original geostatistical data.
+In some situations it may be preferable to work using aggregated areal data rather than the underlying geostatistical data.
 For one, the geostatistical data may be unavailabile or inaccessible due to privacy constraints, administrative practicality or limitations in storage capacity.
-Alternatively, opting to model at the area-level can be advantageous as models for geostatistical data tend to be more complex, require more data, have higher computational burden, and offer less immediate applicablility to area-level policy-making.
-In addition, in some circumstances, seemingly geostatistical data may truly be areal, such as data observed at polling stations or health facilities which individuals from the surrounding area travel to.
-Ultimately, the decision to model the underlying data generating mechanism as either areal or geostatistical is often a matter of practical considerations and pragmatism.
+Alternatively, opting to model at the area-level can be advantageous: models for geostatistical data tend to be more complex, require more data, have higher computational burden, and offer less immediate applicability to policy.
+Under some circumstances, seemingly geostatistical data may truly be areal, as with data observed at polling stations or health facilities which individuals from the surrounding area travel to.
+Ultimately, the decision to model the data generating mechanism as either areal or geostatistical is often a matter of practical considerations and pragmatism.
+This is reflected in the variety of approaches taken by modellers.
 
@@
 Usually, outcomes for locations close together in space tend to be more similar than those far apart, known as "Tobler's first law of geography" [@tobler1970computer].
-This property both presents a challenge and an opportunity.
+This property presents both a challenge and an opportunity.
 Each observation provides less information than it would have had the samples been independent, making it more difficult to estimate global parameters.
@@
 However, the phrase "small-area" is not restricted to geographic areas and may be interpreted more broadly to mean any area where data are insufficient to make accurate local parameter inferences.
-For example, in the context of multilevel regression and post-stratification [@gelman1997poststratification], small-areas are generated by the intersection of demographic variables like age, gender and race, alongside geographic variables.
+For example, in the context of multilevel regression and post-stratification [@gelman1997poststratification], small-areas arise by the intersection of demographic variables like age, gender and race, alongside geographic variables.
 Because of the cost of gathering samples, a survey may be designed to give reliable estimates at at an aggregate level but not at a small-area level.
 Although direct estimators of local parameters are unbiased, when data are sparse the total error may be reduced by accepting some bias in exchange for reduced variance using so-called indirect estimators.
-Smoothing approaches use information from similar units to "borrow strength" from one parameter to another, with determining precisely what is meant by "similar" a central challenge.
+Smoothing approaches use information from similar units to "borrow strength" from one parameter to another. 
+Determining precisely what is meant by "similar" a central challenge.
 See @wakefield2020small for a recent review of both design-based and model-based approaches to small-area estimation of health-indicators, including both area and unit-level analysis.
@@
 where $\x = (x_i)_{i \in [N]}$ is a $N$-dimensional latent field, and $\btheta = (\theta_i)_{i \in [m]}$ are the $m$-dimensional hyperparameters, with $m < N$.
-Small-area estimation models are usually descriptive rather than mechanistic, so a natural choice is to model the latent field with a flexible and computationally efficient Gaussian distribution $\x \sim \mathcal{N}(\bmu, \mathbf{Q})$ with mean vector $\bmu$ and precision matrix $\mathbf{Q}$.
+Small-area estimation models are usually descriptive rather than mechanistic, so a natural choice is to model the large latent field with a flexible and computationally efficient Gaussian distribution $\x \sim \mathcal{N}(\bmu, \mathbf{Q})$ with mean vector $\bmu$ and precision matrix $\mathbf{Q}$.
 Three-layer models with Gaussian latent fields have been collectively studied as latent Gaussian models (LGMs) [@rue2009approximate] and comprise a wide array of popular models, including generalised linear mixed models.
@@
 
-Spatial random effects $\phi(\cdot)$ are one Gaussian function which may be included in the model, whose role is to capture the spatial structure between areas not already accounted for by $\beta_0 + \sum_{l = 1}^{p} \beta_j z_{ji}$.
+Spatial random effects $\phi(\cdot)$ are a Gaussian function, whose role is to capture the spatial structure between areas not already accounted for by the intercept and linear effects of the covariates $\beta_0 + \sum_{l = 1}^{p} \beta_j z_{ji}$.
 Should no spatial structure remain, then independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ may be appropriate, where $\tau_\phi$ is a shared precision.
@@
 The Besag model was originally proposed for use in image analysis, where areas correspond to pixels arranged in a regular lattice structure.
-Since then, it has seen wider use, including in situations where the spatial structure is less regular.
-We highlight a number of concerns about the model's applicability to this broader setting.
+Since then, it has seen wider use, including in situations, like our work mapping HIV, where the spatial structure is less regular.
+We have a number of concerns about the model's applicability to this broader setting.
 This discussion is closely linked to the modifiable areal unit problem [@openshow1979million], whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, as well as the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation].
@@
 Summarising a geometry by an adjacency graph represents a loss of information.
-Many geometries share the same adjacency graph, and therefore the same probability model, as illustrated by Figure \ref{fig:tikz1}.
-This fact is not concerning in itself, but does prompt consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.^[The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be.
+Many geometries share the same adjacency graph, and are as such isomorphic identical under the Besag model (Figure \ref{fig:tikz1}).
+This is not in itself a problem, but does prompt consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.^[The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be.
 In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.]
-Intuitively the more regular the spatial structure, the less information is lost in compression to an adjacency graph.
+Intuitively, the more regular the spatial structure, the less information is lost in compression to an adjacency graph.
 In image analysis, very little spatial information is lost in compression of a lattice structure to an adjacency graph.
 On the other hand, the regions of a country, determined by political and geographic forces, tend to display greater irregularity.
-As such, the appropriateness of adjacency compression varies by the type of geometry common to the application setting.
+The appropriateness of adjacency compression therefore varies by the type of geometry common to the application setting.
 
@@
 This property is a consequence of averaging the conditional mean over a greater number of areas, which, in certain situations, can correspond to a greater amount of information.
-However, if the amount of information in the shaded area remains fixed, it seems inappropriate that $\text{Var}(\phi_1)$ should tend to zero simply as a result of drawing additional, arbitrary, boundaries.
+However, if the amount of information in the shaded area remains fixed, it is inappropriate that $\text{Var}(\phi_1)$ should tend to zero as a result of drawing additional, arbitrary, boundaries.
 In the image analysis setting this modelling assumption is reasonable: each pixel represents a fixed amount of information and a higher pixel density represents a greater amount of information.
@@
 
-The simplest approach is to use a single point.
-A natural choice is the centroid $c_i \in A_i$, given by the arithmetic mean of the latitude and longitude, which may be representative of the area.^[Though note that it is not guaranteed for the centroid to (even) lie within the area.]
+The simplest approach is to use a single point such that $K(A_i, A_j) = k(p_i, p_j)$.
+A natural choice is the centroid $p_i = c_i$, given by the arithmetic mean of the latitude and longitude, which may be representative of the area.^[Note it is not guaranteed for the centroid to (even) lie within the area i.e.
+we may have $c_i \notin A_i$.]
 This results in the centroid kernel
@@
 \end{equation}
-<!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
 The centroid kernel has been used in environmental epidemiology [@wakefield1999spatial] and to model the reproduction number of COVID-19 [@teh2021efficient].
-In a model comparison study \citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas.^[Though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.]
+<!-- In a model comparison study \citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas.^[Though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.] -->
 
@@
 
-Rather than looking to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
+Rather than choosing a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
 This results in the integrated kernel
@@
 \end{equation}
-This covariance structure is equivalent to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, and has been studied in the machine learning literature under the title aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].
+This covariance structure is equivalent to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, and has been studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].
 <!-- @law2018variational develop a variational Bayesian inference method for aggregated observation models. -->
@@
 
-Additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel.
+Additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel^[Analogously, weighted centroids could also be used in the centroid kernel.]
 This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
@@
 Equivalently, samples drawn from $W_i$ may be used without weighting by $w_i(s)$.
-Nodes may also be selected by some deterministic process to give a numerical quadrature estimate of the kernel.
+Nodes may also be selected deterministically to give a numerical quadrature estimate of the kernel.
 These approaches require $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
@@
 If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data.
-On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model does not match the proposed data generating process.
+On the other hand, for non-linear link functions $g$ such as the commonly used exponential or logistic, the generative model does not match the proposed data generating process.
 <!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. -->
@@
 The data and modelling choices were designed with a spatial epidemiology setting in mind.
-The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`.
+The \textsf{R} [@R2021] code used is available from `github.com/athowes/beyond-borders`.
 We used `orderly` [@orderly] for reproducible research, `ggplot2` for data visualisation [@wickham2016ggplot2] and `rticles` [@allaire2022rticles] for reporting via `rmarkdown` [@allaire2022rmarkdown].
@@
 We used seven different geometries including the four vignette geometries sharing an adjacency graph (Figure \ref{fig:tikz1}), as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of Côte d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
-The more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test how model performance varies by geometry regularity.
+The realistic geometries were chosen to represent plausible variation over spatial regularity for the small-area estimation setting, allowing us to test how model performance varies by geometry regularity.
 For each combination of spatial random effect and geometry we replicated the simulation process above 200 times to generate a total of 1400 synthetic data-sets.
@@
 
-We fit eight inferential models, each of which differing in its spatial random effect specification (Table \ref{tab:inferentialmodels}).
+We fit eight inferential models \ref{tab:inferentialmodels}).
+Only the spatial random effect specification differed between models.
 
@@
 \end{tabularx}
-\caption{All inferential models are implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}.
-The AGHQ algorithm was used for approximate Bayesian inference of all inferential models.}
+\caption{All inferential models are implemented as a part of the \textsf{R} package \texttt{arealutils} \citep{arealutils}.
+The AGHQ algorithm was used for approximate Bayesian inference of all inferential models, via the \texttt{aghq} \textsf{R} package \citep{stringer2021implementing}}
 \label{tab:inferentialmodels}
@@
 
-We placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
-The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one.
+We placed a weakly informative half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
+The value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one.
 A weakly informative $\mathcal{N}(-2, 1)$ prior was placed on $\beta_0$, setting most of the prior probability density for $\rho_i$ within a range $[0, 0.25]$ typical for a disease prevalence.
@@
 
-We set $L_i = 10$ using a hexagonal spacing structure, as with the central panel of Figure \ref{fig:51}.^[Note that `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].]
+We set $L_i = 10$ using a hexagonal spacing structure (Figure \ref{fig:51}).^[Note that `sf::st_sample` with `type = "hexagonal"` does not guarantee exactly the specified number of samples are returned [@pebesma2018sf].]
 We are limited to this relatively small number by the computational costs of Model \hyperlink{I8}{\texttt{I8}}.
@@
 We used adaptive Gauss-Hermite quadrature [AGHQ; @stringer2022fast] with $k = 3$ quadrature points per hyperparameter dimension to perform approximate Bayesian inference.
-Each model is implemented using a Template Model Builder \textsc{C++} template for the log-posterior via the `TMB` package [@kristensen2016tmb]
-We are then able to conduct inference using both AGHQ via the `aghq` package [@stringer2021implementing] and the No-U-Turn Sampling (NUTS) Hamiltonian Monte Carlo (HMC) algorithm using Stan [@carpenter2017stan] via the `tmbstan` package [@monnahan2018no].
+Each model was implemented using a Template Model Builder \textsc{C++} template for the log-posterior via the `TMB` package [@kristensen2016tmb]
+We were then able to conduct inference using both AGHQ via the `aghq` package [@stringer2021implementing] and the No-U-Turn Sampling (NUTS) Hamiltonian Monte Carlo (HMC) algorithm using Stan [@carpenter2017stan] via the `tmbstan` package [@monnahan2018no].
 In Appendix we demonstrate that the inference results from AGHQ are comparable to those obtained running NUTS over a longer time.
@@
 As before, we varied the spatial random effect specification.
-The \textsf{R} code for this study is available from \url{https://github.com/athowes/areal-comparison}.
+The \textsf{R} code for this study is available from \url{https://github.com/athowes/beyond-borders}.
 

### 2023-06-23 make/1_all.R

@@ -13,7 +13,7 @@ run_commit_push("1_plot_simulation-geometries") #' [x]
 #' 3. Fit models
 
 #' Vignette and realistic together
-run_commit_push("1_fit", param = list(f = "constant_inla")) #' [x]
+run_commit_push("1_fit", param = list(f = "constant_aghq")) #' [x]
 run_commit_push("1_fit", param = list(f = "iid_inla"))      #' [ ]
 run_commit_push("1_fit", param = list(f = "besag_inla"))    #' [ ]
 run_commit_push("1_fit", param = list(f = "bym2_inla"))     #' [ ]

### 2023-06-23 src/0_plots/geometry-graph.R

@@ -1,11 +1,51 @@
 pdf("geometry-graph-mwi.pdf", h = 2.5, w = 6.25)
 
-arealutils::plot_graph(mw, add_geography = TRUE)
+sf <- mw
+
+nb <- sf_to_nb(sf)
+
+nb_sf <- spdep::nb2lines(nb, coords = sp::coordinates(as(sf, "Spatial"))) %>%
+  as("sf") %>%
+  sf::st_set_crs(sf::st_crs(sf))
+
+b <- ggplot(sf) +
+  geom_sf(data = nb_sf) +
+  theme_minimal() +
+  labs(subtitle = "Graph") +
+  theme_void()
+
+a <- ggplot(sf) +
+  geom_sf() +
+  theme_minimal() +
+  labs(subtitle = "Geography") +
+  theme_void()
+
+a + b
 
 dev.off()
 
 pdf("geometry-graph-zwe.pdf", h = 2.5, w = 6.25)
 
-arealutils::plot_graph(zw, add_geography = TRUE)
+sf <- zw
+
+nb <- sf_to_nb(sf)
+
+nb_sf <- spdep::nb2lines(nb, coords = sp::coordinates(as(sf, "Spatial"))) %>%
+  as("sf") %>%
+  sf::st_set_crs(sf::st_crs(sf))
+
+b <- ggplot(sf) +
+  geom_sf(data = nb_sf) +
+  theme_minimal() +
+  labs(subtitle = "Graph") +
+  theme_void()
+
+a <- ggplot(sf) +
+  geom_sf() +
+  theme_minimal() +
+  labs(subtitle = "Geography") +
+  theme_void()
+
+a + b
 
 dev.off()

### 2023-06-23 src/0_plots/sloo-cv.R

@@ -18,7 +18,7 @@ plot <- ggplot(df, aes(fill = as.factor(left_out))) +
   scale_fill_manual(
     values = c(sf_lightgrey, lightblue, lightgreen),
     name = "",
-    labels = c("Areas the model is fit to", "Areas left out", "Areas predicted on")
+    labels = c("Fit to", "Left out", "Predicted on")
   ) +
   theme_void() +
   theme(

### 2023-06-23 src/1_draw_vignette-geometries/script.R

@@ -48,7 +48,7 @@ geometry_4 <- lapply(list(area1, area2, area3), function(area) list(area) %>% sf
 
 saveRDS(geometry_4, "geometry-4.rds")
 
-pdf("vignette-geometries.pdf", h = 4, w = 6.25)
+pdf("vignette-geometries.pdf", h = 4.5, w = 5)
 
 cowplot::plot_grid(
   ggplot() +

### 2023-11-10 src/docs_paper/paper.Rmd
small typo
Remaining paper and appendix changes

@@
   | Beyond borders: evaluating the suitability of spatial adjacency for small-area estimation
+runtitle: "Evaluating the suitability of spatial adjacency for small-area estimation"
 classoption: ba
@@
 bibliography: citations.bib
-#linenumbers: true
+linenumbers: true
 #numbersections: true
@@
     base_format: rticles::isba_article
-    md_extensions: -autolink_bare_uris #keep these off or \printead{} will fail on email addresses
+    md_extensions: -autolink_bare_uris
     includes:
@@
 <!-- @riebler2016intuitive reach a similar conclusion about both the Leroux and BYM2 model. -->
+
+\newpage
 
@@
 Specifying $\bphi = (\phi_i)_{i \in [n]}$ amounts to specifying the entries of a precision or covariance matrix, as we discuss in the following sections.
+
+\newpage
 
@@
 \end{align}
-where $\tau_\phi$ is the marginal precision of $\bphi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $u$ with precision matrix given by the scaled structure matrix $\mathbf{R}^\star$ (see Appendix).
+where $\tau_\phi$ is the marginal precision of $\bphi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $u$ with precision matrix given by the scaled structure matrix $\mathbf{R}^\star$ (see Appendix S1).
 When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model.
 To borrow an analogy [@rue2020comment] the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately.
+
+\newpage
 
@@
 
+\newpage
+
 # Simulation study \label{sec:simulation}
@@
 \includegraphics[width=\textwidth]{depends/simulation-geometries.pdf}
-\caption{The seven simulation geometries that we considered.}
+\caption{The seven simulation geometries that we considered. The top row contains the four vignette geometries from \ref{fig:tikz1}, and the bottom row contains three more realistic geometries.}
 \label{fig:51}
@@
 \hypertarget{I3}{\texttt{I3}} & Besag &
-$\bphi \sim \mathcal{N}(0, (\tau_\phi \mathbf{R}^\star)^{-})$ following additional recommendations of \citet{freni2018note}  \\
+$\bphi \sim \mathcal{N}(0, (\tau_\phi \mathbf{R}^\star)^{-})$ following additional recommendations of \citet{freni2018note} (Appendix S1)  \\
 \hypertarget{I4}{\texttt{I4}} & BYM2 &
@@
 \caption{All inferential models are implemented as a part of the \textsf{R} package \texttt{arealutils} \citep{arealutils}.
-The AGHQ algorithm was used for approximate Bayesian inference of all inferential models, via the \texttt{aghq} \textsf{R} package \citep{stringer2021implementing}}
+We used adaptive Gauss-Hermite quadrature for approximate Bayesian inference, implemented via the \texttt{aghq} \textsf{R} package \citep{stringer2021implementing}}
 \label{tab:inferentialmodels}
@@
 We used adaptive Gauss-Hermite quadrature [AGHQ; @stringer2022fast] with $k = 3$ quadrature points per hyperparameter dimension to perform approximate Bayesian inference.
-Each model was implemented using a Template Model Builder \textsc{C++} template for the log-posterior via the `TMB` package [@kristensen2016tmb]
+Each model was implemented using a Template Model Builder \textsc{C++} template for the log-posterior via the `TMB` package [@kristensen2016tmb].
 We were then able to conduct inference using both AGHQ via the `aghq` package [@stringer2021implementing] and the No-U-Turn Sampling (NUTS) Hamiltonian Monte Carlo (HMC) algorithm using Stan [@carpenter2017stan] via the `tmbstan` package [@monnahan2018no].
-In Appendix we demonstrate that the inference results from AGHQ are comparable to those obtained running NUTS over a longer time.
-We also compare to empirical Bayes, and where possible integrated nested Laplace approximation via `R-INLA` [@martins2013bayesian].
+In Appendix S2 we demonstrate that the inference results from AGHQ are comparable to those obtained running NUTS to convergence.
+<!-- We also compare to empirical Bayes, and where possible integrated nested Laplace approximation via `R-INLA` [@martins2013bayesian]. -->
 
@@
 
-We assessed each fitted inferential model according to its ability to recover the true underlying value of the probabilities $\rho_i$ at each area in the study region, as well as the calibration of the model's estimates.
+We assessed each fitted inferential model according to its ability to recover the true underlying value of $\rho_i$ at each area in the study region, as well as the calibration of the model's estimates.
 <!-- We describe our approach to model assessment in the following sections. -->
@@
 $\text{CRPS}(f, \omega_i)$ if and only if $f(\rho_i) = \delta_{\omega_i}$.
-The CRPS is a strictly proper scoring rule^[As is the logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$, though in line with the findings of @kruger2020predictive, we found it difficult to compute satisfactorily using kernel density estimation.] [@gneiting2007strictly] and can be evaluated directly using samples $\{\rho_i^s\}_{s = 1}^S$ as
+The CRPS is a strictly proper scoring rule and can be evaluated directly using samples^[The logarithmic score [@good1952rational] given by $\text{LogS}(f, \omega_i) = \log f(\omega_i)$ is another strictly proper scoring rule, however we found it difficult to compute satisfactorily using kernel density estimation, in line with the findings of @kruger2020predictive.] [@gneiting2007strictly] $\{\rho_i^s\}_{s = 1}^S$ as
 \begin{equation}
@@
 
-See Appendix.
+See Appendix S3.
 
@@
 
+See Appendix S4.
+
 ```{r results="asis"}
@@
 ```
+
+\newpage
 

### 2023-11-17 make/0_all.R
Shorten names

@@ -11,5 +11,5 @@ run_commit_push("0_explore_inla-spde")        #' [ ]
 run_commit_push("0_explore_variogram-method") #' [ ]
 run_commit_push("0_explore_wilson-pointless") #' [ ]
 run_commit_push("0_plots")                    #' [ ]
-run_commit_push("0_plot-tikz_besag")          #' [ ]
-run_commit_push("0_plot-tikz_maup")           #' [ ]
+run_commit_push("0_tikz_besag")               #' [ ]
+run_commit_push("0_tikz_maup")                #' [ ]
